### 2018-07-02 manuscript_draft.Rmd
manuscript template created

---
title: "Title"
author: "$^{1}$Petr Capek*, $^{1}$Katherine Todd-Brown, $^{1}$Natalie Sadler, $^{1}$Nancy Hess, $^{1}$Kirsten Hofmockel"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontfamily: mathpazo
fontsize: 11pt
#bibliography: mint.bib
output: 
  pdf_document
header-includes:
    - \usepackage{setspace}
    - \doublespacing
    - \usepackage{lineno}
    - \linenumbers

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

$^{1}$Pacific Northwest National Laboratory, Environmental Molecular Sciences Laboratory, Richland, WA USA

## Abstract

## Introduction

## Material and Methods

## Results

## Discussion

### 2018-07-02 data_analysis.R

@@ -3774,8 +3774,8 @@ mmem_substrates<-function(data){
   }
   
   for(i in 1:3){
-    obs_E<-append(obs_E, cost_E(pars=summary(res[[i]])[6,], data=dat[dat$id==i, ])$residuals$obs)
-    mod_E<-append(mod_E, cost_E(pars=summary(res[[i]])[6,], data=dat[dat$id==i, ])$residuals$mod)
+    obs_E<-append(obs_E, cost_E(pars=summary(res[[i]])[1,], data=dat[dat$id==i, ])$residuals$obs)
+    mod_E<-append(mod_E, cost_E(pars=summary(res[[i]])[1,], data=dat[dat$id==i, ])$residuals$mod)
   }
   
   OvP_E<-data.frame(obs_E, mod_E)
@@ -3808,7 +3808,7 @@ mmem_substrates<-function(data){
   return(al)
 }
 
-#no_cors<-detectCores()
+#no_cors<-detectCores()-1
 #cl<-makeCluster(no_cors)
 registerDoParallel(cl)
 
@@ -4357,7 +4357,7 @@ monod_substrates_fixed<-function(data){
     
     res<-modMCMC(f=cost_function, p=c(Vmax=0.1, Km=3, CUE=0.8),
                  lower=c(Vmax=1e-4, Km=1e-4, CUE=1e-2),
-                 upper=c(Vmax=1e4, Km=1e4, CUE=0.999),niter=10000)
+                 upper=c(Vmax=1e4, Km=1e4, CUE=0.999),niter=100000)
     
     res$Substrate<-data[1, "Substrate"]
     
@@ -4596,7 +4596,7 @@ mem_substrates_fixed<-function(data){
     
     res<-modMCMC(f=cost_function, p=c(Vmax=0.1, Km=3, CUE=0.8),
                  lower=c(Vmax=1e-4, Km=1e-4, CUE=1e-2),
-                 upper=c(Vmax=1e4, Km=1e4, CUE=0.999),niter=10000)
+                 upper=c(Vmax=1e4, Km=1e4, CUE=0.999),niter=100000)
     res$Structure<-data[1, "Substrate"]
     
     return(res)
@@ -4913,7 +4913,7 @@ mmem_substrates_fixed<-function(data){
     
     res<-modMCMC(f=cost_function, p=c(Vmax=0.1, Km=3, Yatp=1),
                  lower=c(Vmax=1e-4, Km=1e-4, Yatp=0.01),
[…]

### 2018-07-02 manuscript_draft.Rmd
MM started

@@
 ---
-title: "Title"
+title: "**Title**"
 author: "$^{1}$Petr Capek*, $^{1}$Katherine Todd-Brown, $^{1}$Natalie Sadler, $^{1}$Nancy Hess, $^{1}$Kirsten Hofmockel"
@@
     - \linenumbers
+    - \usepackage{titling}
+    - \pretitle{\begin{flushleft}}
+    - \posttitle{\end{flushleft}}
+    - \preauthor{\begin{flushleft}\large}
+    - \postauthor{\par\end{flushleft}}
+    - \predate{\begin{flushleft}}
+    - \postdate{\par\end{flushleft}\vskip 0.5em}
 
@@
 
-$^{1}$Pacific Northwest National Laboratory, Environmental Molecular Sciences Laboratory, Richland, WA USA
+_$^{1}$Pacific Northwest National Laboratory, Environmental Molecular Sciences Laboratory, Richland, WA USA_
 
@@
 ## Material and Methods
+###Experimental design
+Microbial culture obtained from the soil (see later) was incubated in liquid batch culture microcosms at **_??24$^{\circ}C$??_** in the dark for five days under nine different experimental treatments.
+These treatments include three different organic substrates as a sole source of carbon for microbial culture and three different levels of microcosm structural complexity in fully factorial design.
+The three organic substrates were glucose, cellobiose and mixture of glucose, cellobiose and cellulose.
+The three levels of structural complexity were represented by **_??34 ml??_** incubation vials with liquid medium only (denoted as **"BROTH"** further in the text), with liquid medium and mixture of 2.7, 0.1 and 0.1 mm glass beads (5g of 2.7mm Biospec glass beads, 2g of 1.0mm Biospec glass beads and 1g of 0.1mm Next-Advanced glass beads; denoted as **"GLASS"** further in the text) and with liquid medium and glass wool (0.2g of Pyrex fiber glass wool cutted to $\sim0.25cm$ pieces; denoted as **"WOOL"** further in the text).
+All experimental treatments were replicated four times.
+
+###Microbial culture
+Microbial culture used in the experiment was obtained from the **_??prosser soil??
+(is it correct)_**, _short characteristic of soil_ ...
+.One gram of air dried soil was supplemented by the 10 ml of sterile Cellulose Degrader M9 Media with 6.25 ml of LB media (see later).
+The soil with medium was incubated seven days at 24$^{\circ}C$ in dark.
+
+###Microcosm
 

### 2018-07-05 manuscript_draft.Rmd
MM continue

--- 2018-07-02 manuscript_draft.Rmd
@@
 fontsize: 11pt
-#bibliography: mint.bib
+bibliography: mint.bib
 output: 
@@
 ## Material and Methods
+###Experimental design
+Microbial culture obtained from the soil (see later) was incubated in liquid batch culture microcosms at 24$^{\circ}C$ in the dark for five days under nine different experimental treatments.
+These treatments include three different organic substrates as a sole source of carbon for microbial culture and three different levels of microcosm structural complexity in fully factorial design.
+The three organic substrates were glucose, cellobiose and mixture of glucose, cellobiose and cellulose.
+The three levels of structural complexity were represented by **_??34 ml??_** incubation vials with liquid medium only (denoted as **"BROTH"** further in the text), with liquid medium and mixture of 2.7, 0.1 and 0.1 mm glass beads (5 g of 2.7 mm Biospec glass beads, 2 g of 1.0 mm Biospec glass beads and 1 g of 0.1 mm Next-Advanced glass beads; denoted as **"GLASS"** further in the text) and with liquid medium and glass wool (0.2 g of Pyrex fiber glass wool cutted to ~0.25 cm pieces; denoted as **"WOOL"** further in the text).
+All experimental treatments were replicated four times **_but this is probably not true since the sampling was destructive_**.
+
+###Microbial culture
+Microbial culture used in the experiment was obtained by imbibement of the **_??prosser soil??
+(is it correct)_**, _short characteristic of soil_ ...
+with the growth medium.
+One gram of air dried soil was supplemented by the 10 ml of sterile Cellulose Degrader M9 Media (see later) with 6.25 ml of lysogeny broth.
+The final organic carbon concentration of added medium was ~40 mmol $l^{-1}$ (0.4 mmol $g(DW)^{-1}$).
+The soil with medium was incubated seven days at 24$^{\circ}C$ in dark.
+At the end of incubation, soil-medium suspension was shaken with ~ten 2.7 mm glass beats for ten minutes.
+The suspension was transferred to sterile flask and soil particles were allowed to settle down.
+The liquid phase was further transferred to new sterile flask and used to inoculate a microcosms.
+
+###Microcosm
+**_??34 ml??_** incubation vials with glass beads, glass wool or empty were filled with 3 ml of sterile medium and 1 ml of microbial culture.
+The medium was composed of Cellulose Degrader M9 Media and organic substrate (the final concentration 1 g of organic substrate per one liter of medium).
+Cellulose Degrader M9 Media was prepared by mixing 300 ml of M9 Minimal Media 10X stock solution (59.623 g anhydrous $NaH_{2}PO_{4}$, 29.938 g $KH_{2}PO_{4}$, 4.967 g NaCl, 10.003 g $NH_{4}Cl$, 1.204 g anhydrous $MgSO_{4}$ and 0.140 g $CaCl_{2}\cdot2H_{2}0$ in 1 l of ultra pure water), 10 ml of Hutner's Trace Element Solution (10 g $C_{6}H_{9}NO_{6}$, 7.3 g KOH, 14.45 g $MgSO_{4}$, 3.335 g $CaCl_{2}\cdot2H_{2}0$, 0.00925 g $(NH_{4})_{6}Mo_{7}O_{24}\cdot4H_{2}0$, 0.099 g $FeSO_{4}\cdot7H_{2}0$ and 50 mL of Hutner's Stock Salt Solution in 1 l of ultra pure water), 1 ml of Biotin (1 g $l^{-1}$), 1 ml of Thiamin (1 g $l^{-1}$) and ultra pure water to 1 l.
+Accounting for the dilution by the inoculum with assumed organic carbon concentration nearly zero, initial organic carbon concentration of the microcosms was 25 mmol $l^{-1}$.
+Initial microbial biomass concentration of the microocosm was measured (see later) to be 7 mmol of microbial carbon per liter.
+Incubation vials were kept covered by Breathe-Easy sealing membrane secured with the aluminium cap at all times the respiration rate was not measured.
+**_(short decription of the incubator)_**.
+
+###Respiration rate
+Microbial respiration rate was measured at the beginning of the experiment and each following day (six times in total).
+Approximately two hours before the headspace $CO_{2}$ concentration measurement, Breathe-Easy sealing membranes were removed from the incubation vials and replaced by the ruber septa secured with the alluminium cap.
+Vials headspace was exhcanged for the $CO_{2}$-free air and incubated.
+Headspace was sampled using 5 ml gas tight syringe.
+1.5 ml of well mixed headspace air was sampled and directly injected to Li-Cor Li 7000 (LI-COR, Inc., Lincoln, Nebraska, USA).
+$CO_{2}$ concentration was calculated against the calibartion standard gas of the $CO_{2}$ concentration 2000 ppm.
+$CO_{2}$ concentration was corrected for the $CO_{2}$ dissolution in liquid media according to Sparling and West [-@Sparling1990].
+Respiration rate was further calculated as corrected $CO_{2}$ concentration divided by the time since the vials were closed.
+
+###DNA isolation and quantification
+DNA was isolated before the start of the experiment, first, second and last day of incubation to quantify microbial biomass in the microcosms.
+Two ml of well mixed liquid solution was sampled and thorougly mixed with two ml of phosphate buffer.
+The final solution was centrifuged at 4,700 g for ten minutes at 4$^{\circ}C$ and pelet was separated from supernatant.
+Supernatant was anaylzed for protein concentration (see later).
+Pellet was resuspended in 0.5 ml and kept deep frozen (-80$^{\circ}C$) until the assay was conducted.
+Cell pelet was bead beated in ethanol solution for 5 minutes to lyse the cells.
+To isolate DNA from the cell lysate, DNeasy PowerSoil Kit (QIAGEN, Venlo, Netherland) was used.
+DNA concentration was measured by a Nanodrop meter **_(specification)_**.
+DNA concentration [@Christensen1993]
 
@@
 ## Discussion
+
+##References

### 2018-07-05 data_analysis.R
data recalculated

@@ -1,7 +1,178 @@
 ##################################MinT#################################################
 ##data 
+#respiration rates in umol/ml/h
+resp<-read.xlsx(xlsxFile = c("C:/Users/cape159/Documents/pracovni/data_statistika/minT/data_checked_respiration.xlsx"),
+                sheet = 1)
 
-dat<-subset(mint, Substrate!="Carbon Free" & Substrate!="Avicel")[,c(1:6,9:12, 18)]
+#arrange the data
+resp.ordered<-resp[order(resp$Sample, resp$Day), ]
+
+#biological data
+biology<-read.xlsx(xlsxFile = c("C:/Users/cape159/Documents/pracovni/data_statistika/minT/data_checked_biology.xlsx"),
+                   sheet = 1)
+
+#arrange the data in the same way
+biology.ordered<-biology[order(biology$Sample, biology$Day), -c(2,3)]
+biology.ordered$Sample<-paste0("CSub3_", biology.ordered$Sample)
+
+#merging both data sets
+m0<-merge(resp.ordered, biology.ordered, by.x=c("Sample", "Day"), by.y = c("Sample", "Day"), all=T)
+
+####################################proxy based Biomass calculations##################################
+#There are plenty of possible conversions
+#I will create a new data frame to store them all at one place for later use
+
+conversions<-data.frame(Sample=m0[,1])
+
+#1. protein based conversions
+#Henriksen et al., 1996
+conversions$Cmic<-m0$Prot.in/12.01/4
+conversions$Proxy<-c("Cellular protein")
+conversions$Reference<-c("Henriksen et al., 1996")
+conversions$Organism<-c("Penicillium chrysogenum")
+
+#Hanegraaf and Muller 2001 presented different data
+#Paracoccus denitrificans
+conversions<-rbind(conversions, data.frame(Sample=m0[,1],
+                                           Cmic=m0$Prot.in/0.57*0.45/12.01/4,
+                                           Proxy=rep("Cellular protein"),
+                                           Reference=rep("Hanegraaf and Muller, 2001"),
+                                           Organism=rep("Paracoccus denitrificans")))
+
+#E. Coli
+conversions<-rbind(conversions, data.frame(Sample=m0[,1],
+                                           Cmic=m0$Prot.in/0.82*0.45/12.01/4,
+                                           Proxy=rep("Cellular protein"),
+                                           Reference=rep("Hanegraaf and Muller, 2001"),
[…]

### 2018-07-10 data_analysis.R
models defined in external files

@@ -161,7 +161,7 @@ ggplot(conversions, aes(Reference, Cmic))+geom_boxplot(cex=0.8, aes(colour=Organ
 ######################################################################################################
 #carefull here, there is slightly different recalculation
 m0$Cmic.dna.init<-m0$DNA.init/0.0494*0.44/12.01/3*0.25
-
+######################################################################################################
 
 #extracellular protein to extracellular protein carbon (46% of carbon in protein - Vrede et al., 2004)
 m0$E<-m0$Prot.out*0.46/12.01/4
@@ -186,6 +186,7 @@ library(foreach)
 library(doParallel)
 library(data.table)
 library(openxlsx)
+library(plyr)
 
 ##ggplot theme
 theme_min<-theme(axis.text.x=element_text(vjust=0.2, size=14, colour="black"),
@@ -210,4214 +211,250 @@ theme_min<-theme(axis.text.x=element_text(vjust=0.2, size=14, colour="black"),
 #######################################################################################
 ##############################First order decay########################################
 #######################################################################################
+source("../first_order_decay_function.R")
 
-
-#different substrates used in the experiment have slightly different concentrations at t=0
-#but I would like to fit one function across substrates to estimate only one parameter
-#I will try to create a function, which could do that
-#the function use one model (first order decay) and apply it to 3 different substrates with
-#3 different initial carbon concentrations. It combines the output into one data frame. 
-#This data frame can be used with the modCost function, which can be subsequently used to
-#estimate parameters across all data
-
-
-#this is the function
-first_order<-function(X, pars, t){
-  
-  #first order decay function
-  deriv<-function(time, state, pars){
-    
-    with(as.list(c(state, pars)),{
-      
-      dC<--k*C
-      
-      return(list(c(dC), r=k*C))
-      
-    })
-  }
-  
[…]

### 2018-07-10 first_order_decay_function.R

@@ -1,0 +1,318 @@
+###first order decay
+first_order_function<-function(data, SUB, FACT, Vars, Niter){
+  
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-setDT(data)[, id := .GRP, by = .(Substrate)]
+  
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-setDT(data)[, id := .GRP, by = .(Structure)]
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-setDT(data)[, id := .GRP, by = .(Substrate,Structure)]
+      }else{
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  #############################################################################################
+  #if parameters are estimated across different substrates, different initial C conc
+  #has to be aknowledged
+  
+  if(SUB==TRUE){
+    
+    #function I need
+    first_order<-function(X, pars, t){
+      
+      #first order decay function
+      deriv<-function(time, state, pars){
+        
+        with(as.list(c(state, pars)),{
+          
+          dC<--k*C
+          
+          return(list(c(dC), r=k*C))
+          
+        })
+      }
+      
+      #function to apply acros 3 different initial substrate concentrations
+      base_function<-function(Ci){
+        
+        
+        out<-ode(y=c(C=Ci), parms=pars, times=t, func=deriv)
+        return(as.data.frame(out))}
+      
+      #results
+      res<-lapply(X=X, FUN = "base_function") %>% bind_cols()
+      
+      return(res)
+      
+    }
+    
+    #parameter estimation function
+    estim<-function(data){
+      
+      #Obs_decay<-dat[,c("Substrate", "r", "Time")]
+      Obs_decay<-select(data, Vars)
+      
+      #m1<-merge(Obs_decay[Obs_decay$Substrate=="Glucose", c("r", "Time")], 
+                #Obs_decay[Obs_decay$Substrate=="Cellobiose", c("r", "Time")],all = T, by="Time")
+      
+      m1<-merge(select(filter(Obs_decay, Substrate=="Glucose"), Vars[-1]),
[…]

### 2018-07-10 mem_function.R

@@ -1,0 +1,588 @@
+mem_function<-function(data, SUB, FACT, Vars, ColM, ColV, VarsCmic, Cmic, Niter){
+  
+  #SUB - needs to be calibrated in matrix
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  #Vars = names of variables to calibrated the model against
+  #ColM = colnames for these varaibles in matrix
+  #ColV = colnames for these varaibles in normal setting
+  #VarsCmic = which biomass to use for likelihood calculation (it should follow the same notation as Vars)
+  #Cmic = initial microbial biomass C
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-setDT(data)[, id := .GRP, by = .(Substrate)]
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-setDT(data)[, id := .GRP, by = .(Structure)]
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-setDT(data)[, id := .GRP, by = .(Substrate,Structure)]
+      }else{
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  ############################################################################################
+  #if parameters are estimated across different substrates, different initial C conc
+  #has to be acknowledged
+  
+  if(SUB==TRUE){
+    
+    #this is the mem model fitted across different substrates 
+    #with different initial carbon concentrations
+    mem<-function(X, pars, t){
+      
+      #mem model
+      deriv<-function(time, state, pars){
+        
+        with(as.list(c(state, pars)),{
+          
+          dCmic<--kmic*Cmic+CUE*Vmax*Cmic*C/(Km+C)-pe*Cmic
+          dC<-kmic*Cmic-Vmax*Cmic*C/(Km+C)+ke*E
+          dE<-pe*Cmic-ke*E
+          
+          return(list(c(dCmic, dC, dE), r=(1-CUE)*Vmax*Cmic*C/(Km+C)))
+          
+        })
+      }
+      
+      #function to apply acros 3 different initial substrate concentrations
+      base_function<-function(Ci){
+        
+        
[…]

### 2018-07-10 mend_function.R

@@ -1,0 +1,618 @@
+mend_function<-function(data, SUB, FACT, Vars, ColM, ColV, VarsCmic, Cmic, Niter){
+  
+  #SUB - needs to be calibrated in matrix
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  #Vars = names of variables to calibrated the model against
+  #ColM = colnames for these varaibles in matrix
+  #ColV = colnames for these varaibles in normal setting
+  #VarsCmic = which biomass to use for likelihood calculation (it should follow the same notation as Vars)
+  #Cmic = initial microbial biomass C
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-setDT(data)[, id := .GRP, by = .(Substrate)]
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-setDT(data)[, id := .GRP, by = .(Structure)]
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-setDT(data)[, id := .GRP, by = .(Substrate,Structure)]
+      }else{
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  ############################################################################################
+  #if parameters are estimated across different substrates, different initial C conc
+  #has to be acknowledged
+  
+  if(SUB==TRUE){
+    
+    #this is the mend model fitted across different substrates 
+    #with different initial carbon concentrations
+    mend<-function(X, pars, t){
+      
+      #mend model
+      deriv<-function(time, state, pars){
+        
+        with(as.list(c(state, pars)),{
+          
+          #equations
+          #C uptake
+          F1=1/CUE*(Vmax+mr)*Cmic*C/(Km+C)
+          #growth respiration
+          F4=(1/CUE-1)*Vmax*Cmic*C/(Km+C)
+          #maintnance respiration
+          F5=(1/CUE-1)*mr*Cmic*C/(Km+C)
+          #microbial mortality
+          F8=(1-pe)*mr*Cmic
+          #enzyme production
+          F9=pe*mr*Cmic
+          #enzyme decay
+          F10=ke*E
+          
+          #states
[…]

### 2018-07-10 mmem_function.R

@@ -1,0 +1,624 @@
+mmem_function<-function(data, SUB, FACT, Vars, ColM, ColV, VarsCmic, Cmic, Niter){
+  
+  #SUB - needs to be calibrated in matrix
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  #Vars = names of variables to calibrated the model against
+  #ColM = colnames for these varaibles in matrix
+  #ColV = colnames for these varaibles in normal setting
+  #VarsCmic = which biomass to use for likelihood calculation (it should follow the same notation as Vars)
+  #Cmic = initial microbial biomass C
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-setDT(data)[, id := .GRP, by = .(Substrate)]
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-setDT(data)[, id := .GRP, by = .(Structure)]
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-setDT(data)[, id := .GRP, by = .(Substrate,Structure)]
+      }else{
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  ############################################################################################
+  #if parameters are estimated across different substrates, different initial C conc
+  #has to be acknowledged
+  
+  if(SUB==TRUE){
+    
+    #this is the mmem model fitted across different substrates 
+    #with different initial carbon concentrations
+    mmem<-function(X, pars, t){
+      
+      #mmem model
+      deriv<-function(time, state, pars){
+        
+        with(as.list(c(state, pars)),{
+          
+          #Carbon uptake
+          
+          Cu=Vmax*C*Cmic/(Km+C)
+          
+          #defining stochiometric coefficients
+          #PO ratio
+          #PO=1.5*(1-((Cu/Cmic)/((Cu/Cmic)+Sover)))
+          y=2*1.5*Yatp
+          
+          #x
+          #x=(Cu/Cmic)/((Cu/Cmic)+Sprod)
+          
+          #coefficients
+          psi.r=1/((y*x)+(y*0.64*(1-x))+1)
[…]

### 2018-07-10 monod_growth_function.R

@@ -1,0 +1,484 @@
+###monod growth function
+monod_growth_function<-function(data, SUB, FACT, Vars, ColM, ColV, VarsCmic, Cmic, Niter){
+  
+  #SUB - needs to be calibrated in matrix
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  #Vars = names of variables to calibrated the model against
+  #ColM = colnames for these varaibles in matrix
+  #ColV = colnames for these varaibles in normal setting
+  #VarsCmic = which biomass to use for likelihood calculation (it should follow the same notation as Vars)
+  #Cmic = initial microbial biomass C
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-setDT(data)[, id := .GRP, by = .(Substrate)]
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-setDT(data)[, id := .GRP, by = .(Structure)]
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-setDT(data)[, id := .GRP, by = .(Substrate,Structure)]
+      }else{
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  ############################################################################################
+  #if parameters are estimated across different substrates, different initial C conc
+  #has to be aknowledged
+  
+  if(SUB==TRUE){
+    
+    #this is the monod growth function fitted across different substrates 
+    #with different initial carbon concentrations
+    monod<-function(X, pars, t){
+      
+      #monod growth function
+      deriv<-function(time, state, pars){
+        
+        with(as.list(c(state, pars)),{
+          
+          dCmic<--k*Cmic+CUE*Vmax*Cmic*C/(Km+C)
+          dC<-k*Cmic-Vmax*Cmic*C/(Km+C)
+          
+          return(list(c(dCmic, dC), r=(1-CUE)*Vmax*Cmic*C/(Km+C)))
+          
+        })
+      }
+      
+      #function to apply acros 3 different initial substrate concentrations
+      base_function<-function(Ci){
+        
+        
+        out<-ode(y=c(Cmic=Cmic, C=Ci), parms=pars, times=t, func=deriv)
[…]

### 2018-07-10 manuscript_draft.Rmd
first results in the manuscript

@@
 knitr::opts_chunk$set(echo = TRUE)
+library(knitr)
+library(ggplot2)
+library(grid)
+
+theme_min<-theme(axis.text.x=element_text(vjust=0.2, size=14, colour="black"),
+                 axis.text.y=element_text(hjust=0.2, size=14, colour="black"),
+                 axis.title=element_text(size=14, colour="black"),
+                 axis.line=element_line(size=0.5, colour="black"),
+                 strip.text=element_text(size=16, face="bold"),
+                 axis.ticks=element_line(size=1, colour="black"),
+                 axis.ticks.length=unit(-0.05, "cm"),
+                 panel.background=element_rect(colour="black", fill="white"),
+                 panel.grid=element_line(linetype=0),
+                 legend.text=element_text(size=14, colour="black"),
+                 legend.title=element_text(size=14, colour="black"),
+                 legend.position=c("right"),
+                 legend.key.size=unit(1, "cm"),
+                 strip.background=element_rect(fill="grey98", colour="black"),
+                 legend.key=element_rect(fill="white", size=1.2),
+                 legend.spacing=unit(0.5, "cm"),
+                 plot.title=element_text(size=28, face="bold", hjust=-0.05))
+
 ```
@@
 DNA concentration was measured by a Nanodrop meter **_(specification)_**.
-DNA concentration [@Christensen1993]
+DNA concentration was conversed to microbial biomass carbon using several conversion factors published in literature (see Fig.
+1)
 
@@
 
+##Tables
+```{r Tab. 1, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, results='asis'}
+#tab1<-rbind(decay_all_results$likelihood, monod_all_results$ll_r, mem_all_results$ll_r,
+            #mend_all_results$ll_r, mmem_all_results$ll_r)
+#colnames(tab1)<-c("_Log Likelihood_", "_N_", "_R^{2}_", "_AIC_")
+#rownames(tab1)<-c("First order decay", "Monod growth", "MEM", "MEND", "MMEM")
+#write.csv(tab1, "./tab1.csv")
+
+tab1<-read.csv("tab1.csv")
+colnames(tab1)<-c(" ", "_Log Likelihood_", "_N_", "_R\\textsuperscript{2}_", "_AIC_")
+
+kable(tab1, align='c', digits=c(0, 1, 0, 2, 1), caption = "_**Table 1:** Correspondence between measured resiration rate and respiration rate modeled by five different models (First order decay, Monod growth, two different Microbial Enzyme Models (MEM and MEND) and Metabolic Microbial Enzyme Model (MMEM)).
+Log Likelihood, coefficient of determination ($R^{2}$), Akaike Information Criterion (AIC) and the number of estimated parameters are reported.
+All models are calibrated against measured respiration rate only._")
+```
+
+##Figures
+```{r Scheme 1, eval=FALSE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
+
+ggplot(data.frame(a=0, b=0), aes(a,b))+geom_blank()+
+  theme(axis.text.x = element_blank(), axis.title.x = element_blank(),
+        axis.text.y = element_blank(), axis.title.y = element_blank(),
+        panel.background = element_blank(),
+        axis.line.x = element_blank(), axis.line.y = element_blank(),
+        axis.ticks.x = element_blank(), axis.ticks.y = element_blank())+
+  geom_rect(aes(xmin=0.8, xmax=1.2, ymin=0.65, ymax=0.95), fill="white", colour="black", lwd=1.2)+
+  geom_text(aes(x=1, y=0.8, label="DOC"), size=14)+ylim(0.5,1.5)+xlim(0.5,1.5)+
+  geom_text(aes(x=1, y=1.3, label="CO[2]"), size=14, parse = T)+
+  geom_text(aes(x=1, y=1.5, label=c("First order decay")), size=11, parse = F, fontface="italic")+
+  geom_segment(aes(x=1, y=0.95, xend=1, yend=1.25), lwd=1.2, arrow = arrow(length = unit(0.5, "cm")))
+
+ggplot(data.frame(a=0, b=0), aes(a,b))+geom_blank()+
+  theme(axis.text.x = element_blank(), axis.title.x = element_blank(),
+        axis.text.y = element_blank(), axis.title.y = element_blank(),
+        panel.background = element_blank(),
+        axis.line.x = element_blank(), axis.line.y = element_blank(),
+        axis.ticks.x = element_blank(), axis.ticks.y = element_blank())+
+  geom_rect(aes(xmin=1.1, xmax=1.5, ymin=0.65, ymax=0.95), fill="white", colour="black", lwd=1.2)+
+  geom_text(aes(x=1.3, y=0.8, label="DOC"), size=14)+ylim(0.5,1.5)+xlim(0.5,1.5)+
+  geom_text(aes(x=1.3, y=1.3, label="CO[2]"), size=14, parse = T)+
+  geom_text(aes(x=1, y=1.5, label=c("Monod growth")), size=11, parse = F, fontface="italic")+
+  geom_segment(aes(x=1.3, y=0.95, xend=1.3, yend=1.25), lwd=1.2, arrow = arrow(length = unit(0.5, "cm")))+
+  geom_rect(aes(xmin=0.6, xmax=0.8, ymin=0.65, ymax=0.95), fill="white", colour="black", lwd=1.2)
+
+```
+
+```{r Fig.
+1, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=16, fig.width=9, fig.cap="_**Fig.
+1:** Distribution of microbial biomass carbon data as directly measured by the DNA or cellular protein concentration and calculated by different conversion factors published in the literature [@Baart2008; @Beck2018; @VanDuuren2013; @Marstorp2000, @Christensen1993; Christensen1993; @Yokoyama2017; @VanPutten1995; @Henriksen1996; @Hanegraaf2001].
+Middle line corresonds to the median, lower and upper hinges to first and third quartile respectively, lower and upper whisker extends from lower and upper hinge to 1.5 time the interquartile range.
+Note that the x axis of both plots have different scales._"}
+
+#write.csv(conversions, "./fig1.csv")
+fig1<-read.csv("fig1.csv")
+
+ggplot(fig1, aes(Reference, Cmic))+geom_boxplot(cex=0.8, aes(colour=Organism), show.legend = F)+
+  facet_wrap(~Proxy, scales="free", nrow=2)+coord_flip()+theme_min+theme(axis.title.y = element_blank())+
+  ylab(expression(paste("Microbial biomass carbon (", mu, "mol ", ml^{-1},")")))
+```
+
 ##References

### 2018-07-10 data_analysis.R

@@ -163,6 +163,7 @@ ggplot(conversions, aes(Reference, Cmic))+geom_boxplot(cex=0.8, aes(colour=Organ
 m0$Cmic.dna.init<-m0$DNA.init/0.0494*0.44/12.01/3*0.25
 ######################################################################################################
 
+
 #extracellular protein to extracellular protein carbon (46% of carbon in protein - Vrede et al., 2004)
 m0$E<-m0$Prot.out*0.46/12.01/4
 
@@ -216,7 +217,7 @@ source("../first_order_decay_function.R")
 #across all
 decay_all_results<-first_order_function(data = dat, SUB = TRUE, FACT = 4, Niter = 10000, Vars = c("Substrate", "r", "Time"))
 
-#no_cors<-detectCores()-1
+#no_cors<-detectCores()
 #cl<-makeCluster(no_cors)
 registerDoParallel(cl)
 
@@ -452,34 +453,97 @@ mmem_structures_results$ll_r
 mmem_substrates_results$ll_r
 mmem_unique_results$ll_r
 
+plot(obs_E~mod_E, mmem_unique_results$OvP_E)
 ####################################################################################################
 ####################################################################################################
 ####################################################################################################
+#Monod, MEM, MEND and MMEM are further calibrated against respiration and microbial biomass as well
+#since several conversion factors between DNA or protein content and micrbial biomass C exist, models
+#are calibrated against 4 different datasets with 4 different conversion factors used:
+#highest and lowest DNA based microbial biomass and highest and lowest protein based microbial biomass
 
+#1. Highest DNA
+#measured microbial biomass
+dat$Cmic.dna_H<-m0[(m0$Substrate!="Free" & m0$Substrate!="Celluloze" & !is.na(m0$Substrate)), "DNA"]/0.004*0.45/12.01/4
+#measured initial microbial biomass
+Cmic.dna.init_H<-mean(m0$DNA.init/0.004*0.45/12.01/3*0.25, na.rm=T)
 
-##############Do again
-#comparing model fits based on AIC and log likelihood
[…]

### 2018-07-16 manuscript_draft.Rmd
further analyses

--- 2018-07-05 manuscript_draft.Rmd
@@
 output: 
-  pdf_document
+  pdf_document:
+    fig_caption: yes
+    keep_tex: yes
 header-includes:
@@
     - \postdate{\par\end{flushleft}\vskip 0.5em}
+    - \usepackage{float}
+    - \floatplacement{figure}{H}
 
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, tidy.opts=list(width.cutoff=40),tidy=TRUE)
+
+library(knitr)
+library(ggplot2)
+library(grid)
+library(DiagrammeR)
+library(DiagrammeRsvg)
+library(rsvg)
+
+theme_min<-theme(axis.text.x=element_text(vjust=0.2, size=14, colour="black"),
+                 axis.text.y=element_text(hjust=0.2, size=14, colour="black"),
+                 axis.title=element_text(size=14, colour="black"),
+                 axis.line=element_line(size=0.5, colour="black"),
+                 strip.text=element_text(size=16, face="bold"),
+                 axis.ticks=element_line(size=1, colour="black"),
+                 axis.ticks.length=unit(-0.05, "cm"),
+                 panel.background=element_rect(colour="black", fill="white"),
+                 panel.grid=element_line(linetype=0),
+                 legend.text=element_text(size=14, colour="black"),
+                 legend.title=element_text(size=14, colour="black"),
+                 legend.position=c("right"),
+                 legend.key.size=unit(1, "cm"),
+                 strip.background=element_rect(fill="grey98", colour="black"),
+                 legend.key=element_rect(fill="white", size=1.2),
+                 legend.spacing=unit(0.5, "cm"),
+                 plot.title=element_text(size=28, face="bold", hjust=-0.05))
+
 ```
@@
 
+##Tables
+```{r Tab. 1, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE, results='asis'}
+#tab1<-rbind(decay_all_results$likelihood, monod_all_results$ll_r, mem_all_results$ll_r,
+            #mend_all_results$ll_r, mmem_all_results$ll_r)
+#colnames(tab1)<-c("_Log Likelihood_", "_N_", "_R^{2}_", "_AIC_")
+#rownames(tab1)<-c("First order decay", "Monod growth", "MEM", "MEND", "MMEM")
+#write.csv(tab1, "./tab1.csv")
+
+tab1<-read.csv("tab1.csv")
+colnames(tab1)<-c(" ", "_Log Likelihood_", "_N_", "_R\\textsuperscript{2}_", "_AIC_")
+
+kable(tab1, align='c', digits=c(0, 1, 0, 2, 1), caption = "_Correspondence between measured resiration rate and respiration rate modeled by five different models (First order decay, Monod growth, two different Microbial Enzyme Models (MEM and MEND) and Metabolic Microbial Enzyme Model (MMEM)).
+Log Likelihood, coefficient of determination ($R^{2}$), Akaike Information Criterion (AIC) and the number of estimated parameters are reported.
+All models are calibrated against measured respiration rate only._")
+```
+
+##Figures
+
+```{r Scheme 1, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+scheme<-"digraph all{
+
+graph [overlap = false, fontsize = 10, layout = dot, rankdir = LR, penwidth = 0.001, root = cluster1];
+
+subgraph cluster0{
+
+graph [overlap = false, fontsize = 10, layout = dot, rankdir = LR];
+
+#nodes
+
+node [shape = circle];
+f [label = 'C@_{S}'] ;
+g [label = 'C@_{MB}'];
+h [label = 'C@_{Enz}'];
+
+node [shape = plaintext];
+i [label = 'CO@_{2}'];
+
+#edges
+f -> g   [label = 'uptake', fontsize = 10];
+g -> f [label = 'death', fontsize = 10];
+g -> h [label = 'enzyme production', fontsize = 10];
+h -> f  [label = 'enzyme decay', fontsize = 10];
+g -> i [label = 'respiration', fontsize = 10];
+
+label='Microbial-enzyme model';
+
+}
+
+subgraph cluster1 {
+
+graph [overlap = false, fontsize = 10, layout = dot, rankdir = LR];
+
+#nodes
+
+node [shape = circle];
+c [label = 'C@_{S}'] ;
+d [label = 'C@_{MB}'];
+
+node [shape = plaintext];
+e [label = 'CO@_{2}'];
+
+#edges
+c -> d [label = 'uptake', fontsize = 10];
+d -> c [label = 'death', fontsize = 10];
+d -> e [label = 'respiration', fontsize = 10];
+
+label='Monod growth';
+}
+
+subgraph cluster2 {
+
+graph [overlap = false, fontsize = 10, layout = dot, rankdir = LR];
+
+#nodes
+node [shape = circle];
+a [label = 'C@_{S}'];
+
+node [shape = plaintext];
+b [label = 'CO@_{2}'];
+
+#edges
+a -> b [label = 'decay', fontsize = 10];
+label = 'First order decay' ;
+
+}
+}"
+
+grViz(scheme) %>% export_svg() %>% charToRaw %>% rsvg_pdf("scheme.pdf")
+
+include_graphics("scheme.pdf")
+
+```
+
+```{r Fig.
+1, echo=FALSE, resultsw = 'asis', message=FALSE, warning=FALSE, fig.height=14, fig.width=9, fig.cap="_Distribution of microbial biomass carbon data as directly measured by the DNA or cellular protein concentration and calculated by different conversion factors published in the literature [@Baart2008; @Beck2018; @VanDuuren2013; @Marstorp2000, @Christensen1993; Christensen1993; @Yokoyama2017; @VanPutten1995; @Henriksen1996; @Hanegraaf2001].
+Middle line corresonds to the median, lower and upper hinges to first and third quartile respectively, lower and upper whisker extends from lower and upper hinge to 1.5 time the interquartile range.
+Note that the x axis of both plots have different scales._"}
+
+#write.csv(conversions, "./fig1.csv")
+fig1<-read.csv("fig1.csv")
+
+ggplot(fig1, aes(Reference, Cmic))+geom_boxplot(cex=0.8, aes(colour=Organism), show.legend = F)+
+  facet_wrap(~Proxy, scales="free", nrow=2)+coord_flip()+theme_min+theme(axis.title.y = element_blank())+
+  ylab(expression(paste("Microbial biomass carbon (", mu, "mol ", ml^{-1},")")))
+```
+
+\
+
 ##References

### 2018-07-16 data_analysis.R

@@ -1,4 +1,38 @@
-##################################MinT#################################################
+#################################################MinT#################################################
+##libraries
+library(deSolve)
+library(dplyr)
+library(FME)
+library(reshape)
+library(reshape2)
+library(ggplot2)
+library(foreach)
+library(doParallel)
+library(data.table)
+library(openxlsx)
+library(plyr)
+
+##ggplot theme
+theme_min<-theme(axis.text.x=element_text(vjust=0.2, size=14, colour="black"),
+                 axis.text.y=element_text(hjust=0.2, size=14, colour="black"),
+                 axis.title=element_text(size=14, colour="black"),
+                 axis.line=element_line(size=0.5, colour="black"),
+                 strip.text=element_text(size=16, face="bold"),
+                 axis.ticks=element_line(size=1, colour="black"),
+                 axis.ticks.length=unit(-0.05, "cm"),
+                 panel.background=element_rect(colour="black", fill="white"),
+                 panel.grid=element_line(linetype=0),
+                 legend.text=element_text(size=14, colour="black"),
+                 legend.title=element_text(size=14, colour="black"),
+                 legend.position=c("right"),
+                 legend.key.size=unit(1, "cm"),
+                 strip.background=element_rect(fill="grey98", colour="black"),
+                 legend.key=element_rect(fill="white", size=1.2),
+                 legend.spacing=unit(0.5, "cm"),
+                 plot.title=element_text(size=28, face="bold", hjust=-0.05))
+
+#######################################################################################################
+
 ##data 
 #respiration rates in umol/ml/h
 resp<-read.xlsx(xlsxFile = c("C:/Users/cape159/Documents/pracovni/data_statistika/minT/data_checked_respiration.xlsx"),
@@ -175,39 +209,6 @@ mint<-mint[!is.na(mint$Substrate), ]
 
[…]

### 2018-07-20 manuscript_draft.Rmd
to word

@@
 output: 
-  pdf_document:
-    fig_caption: yes
-    keep_tex: yes
-header-includes:
-    - \usepackage{setspace}
-    - \doublespacing
-    - \usepackage{lineno}
-    - \linenumbers
-    - \usepackage{titling}
-    - \pretitle{\begin{flushleft}}
-    - \posttitle{\end{flushleft}}
-    - \preauthor{\begin{flushleft}\large}
-    - \postauthor{\par\end{flushleft}}
-    - \predate{\begin{flushleft}}
-    - \postdate{\par\end{flushleft}\vskip 0.5em}
-    - \usepackage{float}
-    - \floatplacement{figure}{H}
+  word_document:
+    fig_caption: true
 
@@
 library(rsvg)
+library(kableExtra)
+library(magick)
 
@@
 ##Tables
-```{r Tab. 1, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE, results='asis'}
+```{r Tab. 1, echo=FALSE, fig.width=130, message=FALSE, warning=FALSE, results='asis'}
 #tab1<-rbind(decay_all_results$likelihood, monod_all_results$ll_r, mem_all_results$ll_r,
@@
 tab1<-read.csv("tab1.csv")
-colnames(tab1)<-c(" ", "_Log Likelihood_", "_N_", "_R\\textsuperscript{2}_", "_AIC_")
 
-kable(tab1, align='c', digits=c(0, 1, 0, 2, 1), caption = "_Correspondence between measured resiration rate and respiration rate modeled by five different models (First order decay, Monod growth, two different Microbial Enzyme Models (MEM and MEND) and Metabolic Microbial Enzyme Model (MMEM)).
+#colnames(tab1)<-c(" ", "Log Likelihood", "N", "$R^{2}$", "AIC")
+
+kable(tab1, align='c', digits=c(0, 1, 0, 2, 1), "latex", booktabs=T, col.names = c(" ", "Log Likelihood", "N", "R^2^", "AIC"), caption = "Correspondence between measured resiration rate and respiration rate modeled by five different models (First order decay, Monod growth, two different Microbial Enzyme Models (MEM and MEND) and Metabolic Microbial Enzyme Model (MMEM)).
 Log Likelihood, coefficient of determination ($R^{2}$), Akaike Information Criterion (AIC) and the number of estimated parameters are reported.
-All models are calibrated against measured respiration rate only._")
+All models are calibrated against measured respiration rate only.") %>%
+  kable_as_image()
+
 ```
@@
 
-```{r Scheme 1, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r Scheme 1, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width = 20, fig.cap="_Schematic representation of three mathematical models, which are commonly used to predict the respiration rate.
+$C_{S}$ - carbon substrate, $C_{MB}$ - microbial biomass carbon, $C_{Enz}$ - enzymes carbon._"}
 
@@
 f [label = 'C@_{S}'] ;
+h [label = 'C@_{Enz}'];
 g [label = 'C@_{MB}'];
-h [label = 'C@_{Enz}'];
 
@@
 
-grViz(scheme) %>% export_svg() %>% charToRaw %>% rsvg_pdf("scheme.pdf")
+grViz(scheme) %>% export_svg() %>% charToRaw %>% rsvg_png("scheme.png")
 
-include_graphics("scheme.pdf")
+include_graphics("scheme.png")
 

### 2018-07-20 manuscript_draft.Rmd
methods finished

@@
 title: "**Title**"
-author: "$^{1}$Petr Capek*, $^{1}$Katherine Todd-Brown, $^{1}$Natalie Sadler, $^{1}$Nancy Hess, $^{1}$Kirsten Hofmockel"
+author: "^1^Petr Capek*, ^1^Katherine Todd-Brown, ^1^Natalie Sadler, ^1^Nancy Hess, ^1^Kirsten Hofmockel"
 date: "`r format(Sys.time(), '%B %d, %Y')`"
@@
 
-_$^{1}$Pacific Northwest National Laboratory, Environmental Molecular Sciences Laboratory, Richland, WA USA_
+_^1^Pacific Northwest National Laboratory, Environmental Molecular Sciences Laboratory, Richland, WA USA_
 
@@
 ###Experimental design
-Microbial culture obtained from the soil (see later) was incubated in liquid batch culture microcosms at 24$^{\circ}C$ in the dark for five days under nine different experimental treatments.
+Microbial culture obtained from the soil (see later) was incubated in liquid batch culture microcosms at 24&deg;C in the dark for five days under nine different experimental treatments.
 These treatments include three different organic substrates as a sole source of carbon for microbial culture and three different levels of microcosm structural complexity in fully factorial design.
@@
 One gram of air dried soil was supplemented by the 10 ml of sterile Cellulose Degrader M9 Media (see later) with 6.25 ml of lysogeny broth.
-The final organic carbon concentration of added medium was ~40 mmol $l^{-1}$ (0.4 mmol $g(DW)^{-1}$).
-The soil with medium was incubated seven days at 24$^{\circ}C$ in dark.
+The final organic carbon concentration of added medium was ~40 mmol l^-1^ (0.4 mmol g(DW)^-1^).
+The soil with medium was incubated seven days at 24&deg;C in dark.
 At the end of incubation, soil-medium suspension was shaken with ~ten 2.7 mm glass beats for ten minutes.
@@
 ###DNA isolation and quantification
-DNA was isolated before the start of the experiment, first, second and last day of incubation to quantify microbial biomass in the microcosms.
-Two ml of well mixed liquid solution was sampled and thorougly mixed with two ml of phosphate buffer.
-The final solution was centrifuged at 4,700 g for ten minutes at 4$^{\circ}C$ and pelet was separated from supernatant.
+To quantify microbial biomass in the microcosms, DNA was isolated from samples before the start of the experiment, first, second and last day of incubation.
+The sampling was destructive.
+The whole volume of microcosm containing microbial cells was quantitatively transfered to 15 ml falcon tubes.
+Cell scraper was used to detach microbial cells from surface of microcosm or from glass beads/wool.
+Two ml of phosphate buffer was added and the final solution thoroughly mixed.
+The final solution was centrifuged at 4,700 g for ten minutes at 4&deg;C and pelet was separated from supernatant.
 Supernatant was anaylzed for protein concentration (see later).
-Pellet was resuspended in 0.5 ml and kept deep frozen (-80$^{\circ}C$) until the assay was conducted.
+Pellet was resuspended in 0.5 ml and kept deep frozen (-80&deg;C) until the assay was conducted.
 Cell pelet was bead beated in ethanol solution for 5 minutes to lyse the cells.
@@
 1)
+
+###Cellular protein quantification
+As a complementary method for microbial biomass quantification, protein concentration was assesed in cell lysate by bicinchoninic acid assay.
+**_(specification of spectrophotometer)_** was used to measure protein concentration.
+Alike DNA, protein concentration was conversed to microbial biomass carbon using several conversion factors published in literature (see Fig.
+1)
+
+###Extracellular protein concentration
+Protein concentration was further assesed in the supernatant (see above).
+Before the analysis, supernatant was filtered through 10 kDa Supor membrane.
+Protein concentration was measured in >10 kDa fraction.
+This fraction is assumed to be dominated be extracellular enzymes produced by microbial cells.
+Extracellular enzymes are assumed to contain 46% of carbon [@Vrede2004].
+
+###Mathematical description
 
@@
 
-#colnames(tab1)<-c(" ", "Log Likelihood", "N", "$R^{2}$", "AIC")
-
-kable(tab1, align='c', digits=c(0, 1, 0, 2, 1), "latex", booktabs=T, col.names = c(" ", "Log Likelihood", "N", "R^2^", "AIC"), caption = "Correspondence between measured resiration rate and respiration rate modeled by five different models (First order decay, Monod growth, two different Microbial Enzyme Models (MEM and MEND) and Metabolic Microbial Enzyme Model (MMEM)).
+kable(tab1, align='c', digits=c(0, 1, 0, 2, 1), "latex", booktabs=T, escape=F, col.names = c(" ", "Log Likelihood", "N", "R^2^", "AIC"), caption = "Correspondence between measured resiration rate and respiration rate modeled by five different models (First order decay, Monod growth, two different Microbial Enzyme Models (MEM and MEND) and Metabolic Microbial Enzyme Model (MMEM)).
 Log Likelihood, coefficient of determination ($R^{2}$), Akaike Information Criterion (AIC) and the number of estimated parameters are reported.
 All models are calibrated against measured respiration rate only.") %>%
+  kable_as_image()
+
+```
+
+```{r Tab. 2, echo=FALSE, fig.width=130, message=FALSE, warning=FALSE, results='asis'}
+# tab2a<-rbind(monod_all_results2$ll_r, mem_all_results2$ll_r,
+#              mend_all_results2$ll_r, mmem_all_results2$ll_r,
+#              monod_all_results4$ll_r, mem_all_results4$ll_r,
+#              mend_all_results4$ll_r, mmem_all_results4$ll_r,
+#              monod_all_results5$ll_r, mem_all_results5$ll_r,
+#              mend_all_results5$ll_r, mmem_all_results5$ll_r,
+#              monod_all_results6$ll_r, mem_all_results6$ll_r,
+#              mend_all_results6$ll_r, mmem_all_results6$ll_r)
+# tab2a[tab2a[,3]<0,"rsq"]<-c(0)
+#  
+# tab2b<-rbind(monod_all_results2$ll_Cmic, mem_all_results2$ll_Cmic,
+#              mend_all_results2$ll_Cmic, mmem_all_results2$ll_Cmic,
+#              monod_all_results4$ll_Cmic, mem_all_results4$ll_Cmic,
+#              mend_all_results4$ll_Cmic, mmem_all_results4$ll_Cmic,
+#              monod_all_results5$ll_Cmic, mem_all_results5$ll_Cmic,
+#              mend_all_results5$ll_Cmic, mmem_all_results5$ll_Cmic,
+#              monod_all_results6$ll_Cmic, mem_all_results6$ll_Cmic,
+#              mend_all_results6$ll_Cmic, mmem_all_results6$ll_Cmic)
+# tab2b[tab2b[,3]<0,"rsq"]<-c(0)
+#  
+# tab2<-cbind(tab2a, tab2b)
+#  
+# write.csv(tab2, "./tab2.csv")
+
+tab2<-read.csv("tab2.csv")
+
+#colnames(tab1)<-c("_Log Likelihood_", "_N_", "_R^{2}_", "_AIC_")
+# rownames(tab2)<-c("Monod growth", "MEM", "MEND", "MMEM","Monod growth", "MEM", "MEND", "MMEM","Monod growth", "MEM", "MEND", "MMEM","Monod growth", "MEM", "MEND", "MMEM")
+#write.csv(tab1, "./tab1.csv")
+
+kable(tab2, align='c', digits=c(0, 1, 0, 2, 1, 1, 0, 2, 1), "latex", escape=F, booktabs=T, col.names = c(" ", "Log Likelihood", "N", "R^2^", "AIC","Log Likelihood", "N", "R^2^", "AIC"), caption = "Correspondence between measured resiration rate and respiration rate modeled by five different models (First order decay, Monod growth, two different Microbial Enzyme Models (MEM and MEND) and Metabolic Microbial Enzyme Model (MMEM)).
+Log Likelihood, coefficient of determination ($R^{2}$), Akaike Information Criterion (AIC) and the number of estimated parameters are reported.
+All models are calibrated against measured respiration rate only.") %>%
+  add_header_above(c(" " = 1 ,"Respiration rate" = 4, "Microbial biomass" = 4))  %>%
+  group_rows("Beck et al., 2018 (DNA ~ 0.9% $C_{MB}$)", 1 , 4) %>%
+  group_rows("Median of DNA abundance across studies (DNA ~ 2.39% $C_{MB}$)", 5 , 8) %>%
+  group_rows("Beck et al., 2018 (Protein ~ 60%% $C_{MB}$)", 9 , 12) %>%
+  group_rows("Hanegraaf and Muller, 2001 (Protein ~ 182%% $C_{MB}$)", 13 , 16) %>%
+  kable_styling(bootstrap_options = "striped", full_width = F, font_size = 7) %>%
   kable_as_image()
@@
 ```{r Fig.
-1, echo=FALSE, resultsw = 'asis', message=FALSE, warning=FALSE, fig.height=14, fig.width=9, fig.cap="_Distribution of microbial biomass carbon data as directly measured by the DNA or cellular protein concentration and calculated by different conversion factors published in the literature [@Baart2008; @Beck2018; @VanDuuren2013; @Marstorp2000, @Christensen1993; Christensen1993; @Yokoyama2017; @VanPutten1995; @Henriksen1996; @Hanegraaf2001].
+1, echo=FALSE, resultsw = 'asis', message=FALSE, warning=FALSE, fig.height=10, fig.width=7, fig.cap="_Distribution of microbial biomass carbon data as directly measured by the DNA or cellular protein concentration and calculated by different conversion factors published in the literature [@Baart2008; @Beck2018; @VanDuuren2013; @Marstorp2000, @Christensen1993; Christensen1993; @Yokoyama2017; @VanPutten1995; @Henriksen1996; @Hanegraaf2001].
 Middle line corresonds to the median, lower and upper hinges to first and third quartile respectively, lower and upper whisker extends from lower and upper hinge to 1.5 time the interquartile range.

### 2018-07-24 manuscript_draft.Rmd
mathematical description

@@
 ###Mathematical description
+Five models of different complexity were used to predict patterns of respiration rate, microbial biomass carbon ($C_{MB}$) or extracellular enzyme concentration ($C_{Enz}$) during the experiment.
+The emphasis is given to respiration rate prediction.
+
+####First order decay (Fig. 1a)
+In this simplest model, organic carbon ($C_{S}$) decays into $CO_{2}$ at the constant rate (k):
+
+[1] $~~~~~~~~~~\frac{dC_{S}}{dt}~=~k~\times~C_{S}$,
+
+organic substrate (glucose, cellobiose or their mixture) and the microbial biomass are considered as $C_{S}$ in this model.
+$C_{MB}$ and $C_{Enz}$ cannot be predicted by the model since they are not explicitely assumed to affect the respiration rate.
+
+####Monod growth (Fig. 1b)
+In this model, $C_{S}$ is consumed by microbial biomass and transformed to $CO_{2}$ or $C_{MB}$. Microbial biomass is dying at constant rate returning used organic carbon to $C_{S}$ pool:
+
+[2] $~~~~~~~~~~\frac{dC_{MB}}{dt}~=~uptake~\times~CUE~-~k_{MB}~\times~C_{MB}$,
+
+[3] $~~~~~~~~~~\frac{dC_{S}}{dt}~=~-uptake~+~k_{MB}~\times~C_{MB}$.
+
+In eqs. 2 and 3, CUE is carbon use efficiency and $k_{MB}$ is death rate constant of microbial biomass decay process. Microbial carbon uptake is defined as hyperbolic function of $C_{S}$ and $C_{MB}$:
+
+[4] $~~~~~~~~~~uptake~=\frac{V_{MAX}~\times~C_{S}~\times~C_{MB}}{K_{M}~+~C_{S}}$,
+
+in which $V_{MAX}$ is maximum velocity constant and $K_{M}$ is affinity constant.
+Respiration rate is then given by $uptake~\times~(1~-~CUE)$.
+Monod growth model assumes that the production of extracellular enzymes is negligible.
+
+####Microbial enzyme models (Fig. 1c)
+In microbial enzyme models, microbial biomass produce extracellular enzymes. The way the carbon is partitioned between $C_{MB}$, $CO_{2}$ and $C_{Enz}$ can follow different mechanisms.
 
@@
 
-kable(tab1, align='c', digits=c(0, 1, 0, 2, 1), "latex", booktabs=T, escape=F, col.names = c(" ", "Log Likelihood", "N", "R^2^", "AIC"), caption = "Correspondence between measured resiration rate and respiration rate modeled by five different models (First order decay, Monod growth, two different Microbial Enzyme Models (MEM and MEND) and Metabolic Microbial Enzyme Model (MMEM)).
+kable(tab1, align='c', digits=c(0, 1, 0, 2, 1), "latex", booktabs=T, escape=F, col.names = c(" ", "Log Likelihood", "N", "R^2^", "AIC"), caption = "**Tab.
+1:** Correspondence between measured resiration rate and respiration rate modeled by five different models (First order decay, Monod growth, two different Microbial Enzyme Models (MEM and MEND) and Metabolic Microbial Enzyme Model (MMEM)).
 Log Likelihood, coefficient of determination ($R^{2}$), Akaike Information Criterion (AIC) and the number of estimated parameters are reported.
@@
 #   
-# tab2<-cbind(tab2a, tab2b)
+# tab2<-as.data.frame(cbind(tab2a, tab2b))
+# tab2$Model<-c("Monod growth", "MEM", "MEND", "MMEM","Monod growth", "MEM", "MEND", "MMEM","Monod growth", "MEM", "MEND", "MMEM","Monod growth", "MEM", "MEND", "MMEM")
+# tab2<-tab2[,c(9, 1:8)]   
 # 
@@
 
-tab2<-read.csv("tab2.csv")
+tab2<-read.csv("tab2.csv")[,-1]
 
 #colnames(tab1)<-c("_Log Likelihood_", "_N_", "_R^{2}_", "_AIC_")
-# rownames(tab2)<-c("Monod growth", "MEM", "MEND", "MMEM","Monod growth", "MEM", "MEND", "MMEM","Monod growth", "MEM", "MEND", "MMEM","Monod growth", "MEM", "MEND", "MMEM")
 #write.csv(tab1, "./tab1.csv")
 
-kable(tab2, align='c', digits=c(0, 1, 0, 2, 1, 1, 0, 2, 1), "latex", escape=F, booktabs=T, col.names = c(" ", "Log Likelihood", "N", "R^2^", "AIC","Log Likelihood", "N", "R^2^", "AIC"), caption = "Correspondence between measured resiration rate and respiration rate modeled by five different models (First order decay, Monod growth, two different Microbial Enzyme Models (MEM and MEND) and Metabolic Microbial Enzyme Model (MMEM)).
+kable(tab2, align='c', digits=c(0, 1, 0, 2, 1, 1, 0, 2, 1), "latex", escape=F, booktabs=T, col.names = c(" ", "Log Likelihood", "N", "R^2^", "AIC","Log Likelihood", "N", "R^2^", "AIC"), caption = "**Tab.
+2:** Correspondence between measured and modelled resiration rate and microbial biomass carbon ($C_{MB}$).
+Four different models (Monod growth, two different Microbial Enzyme Models (MEM and MEND) and Metabolic Microbial Enzyme Model (MMEM)) were tested.
 Log Likelihood, coefficient of determination ($R^{2}$), Akaike Information Criterion (AIC) and the number of estimated parameters are reported.
-All models are calibrated against measured respiration rate only.") %>%
+All models are calibrated against measured respiration rate and $C_{MB}$, which was calculated from measured DNA or cellular protein content using four different conversion factors reported in literature.") %>%
   add_header_above(c(" " = 1 ,"Respiration rate" = 4, "Microbial biomass" = 4))  %>%
-  group_rows("Beck et al., 2018 (DNA ~ 0.9% $C_{MB}$)", 1 , 4) %>%
-  group_rows("Median of DNA abundance across studies (DNA ~ 2.39% $C_{MB}$)", 5 , 8) %>%
-  group_rows("Beck et al., 2018 (Protein ~ 60%% $C_{MB}$)", 9 , 12) %>%
-  group_rows("Hanegraaf and Muller, 2001 (Protein ~ 182%% $C_{MB}$)", 13 , 16) %>%
+  group_rows("Beck et al., 2018 (DNA ~ 0.9% Cmb)", 1 , 4) %>%
+  group_rows("Median of DNA abundance across studies (DNA ~ 2.39% Cmb)", 5 , 8) %>%
+  group_rows("Beck et al., 2018 (Protein ~ 60% Cmb)", 9 , 12) %>%
+  group_rows("Hanegraaf and Muller, 2001 (Protein ~ 182% Cmb)", 13 , 16) %>%
   kable_styling(bootstrap_options = "striped", full_width = F, font_size = 7) %>%
+  kable_as_image()
+
+```
+
+```{r Tab. 3, echo=FALSE, fig.width=130, message=FALSE, warning=FALSE, results='asis'}
+# tab3a<-rbind(t(rbind(mem_all_results7$ll_r, mend_all_results7$ll_r, mmem_all_results7$ll_r)),
+#              t(rbind(mem_all_results8$ll_r, mend_all_results8$ll_r, mmem_all_results8$ll_r)),
+#              t(rbind(mem_all_results9$ll_r,mend_all_results9$ll_r, mmem_all_results9$ll_r)),
+#              t(rbind(mem_all_results10$ll_r, mend_all_results10$ll_r, mmem_all_results10$ll_r)))
+# colnames(tab3a)<-c("MEM", "MEND", "MMEM")
+#  
+# tab3b<-rbind(t(rbind(mem_all_results7$ll_Cmic, mend_all_results7$ll_Cmic, mmem_all_results7$ll_Cmic)),
+#              t(rbind(mem_all_results8$ll_Cmic, mend_all_results8$ll_Cmic, mmem_all_results8$ll_Cmic)),
+#              t(rbind(mem_all_results9$ll_Cmic,mend_all_results9$ll_Cmic, mmem_all_results9$ll_Cmic)),
+#              t(rbind(mem_all_results10$ll_Cmic, mend_all_results10$ll_Cmic, mmem_all_results10$ll_Cmic)))
+# colnames(tab3b)<-c("MEM", "MEND", "MMEM")
+# tab3b[c(7,11,15),]<-c(0)
+# 
+# 
+# tab3c<-rbind(t(rbind(mem_all_results7$ll_E, mend_all_results7$ll_E, mmem_all_results7$ll_E)),
+#              t(rbind(mem_all_results8$ll_E, mend_all_results8$ll_E, mmem_all_results8$ll_E)),
+#              t(rbind(mem_all_results9$ll_E,mend_all_results9$ll_E, mmem_all_results9$ll_E)),
+#              t(rbind(mem_all_results10$ll_E, mend_all_results10$ll_E, mmem_all_results10$ll_E)))
+# colnames(tab3c)<-c("MEM", "MEND", "MMEM")
+# tab3c[c(3, 7,11,15),]<-c(0)
+# 
+# tab3<-as.data.frame(cbind(tab3a, tab3b, tab3c))
+#    
+# tab3$par<-rep(c("Log Likelihood", "N", "R^2^", "AIC"), times=4)
+# tab3<-tab3[,c(10, 1:9)]   
+#  
+# write.csv(tab3, "./tab3.csv")
+
+tab3<-read.csv("tab3.csv")[,-1]
+
+#colnames(tab1)<-c("_Log Likelihood_", "_N_", "_R^{2}_", "_AIC_")
+#write.csv(tab1, "./tab1.csv")
+
+kable(tab3, align='c', "latex", escape=F, booktabs=T, digits=2, col.names = c(" ", "MEM", "MEND", "MMEM","MEM", "MEND", "MMEM","MEM", "MEND", "MMEM"), caption = "**Tab.
+3:** Correspondence between measured and modelled resiration rate, microbial biomass carbon ($C_{MB}$) and extracellular enzymes.
+Three different models (two different Microbial Enzyme Models (MEM and MEND) and Metabolic Microbial Enzyme Model (MMEM)) were tested.
+Log Likelihood, coefficient of determination ($R^{2}$), Akaike Information Criterion (AIC) and the number of estimated parameters are reported.
+All models are calibrated against measured respiration rate, extracellular enzyme concentration and $C_{MB}$, which was calculated from measured DNA or cellular protein content using four different conversion factors reported in literature.") %>%
+  add_header_above(c(" " = 1 ,"Respiration rate" = 3, "Microbial biomass" = 3, "Enzymes" = 3))  %>%
+  group_rows("Beck et al., 2018 (DNA ~ 0.9% Cmb)", 1 , 4) %>%
+  group_rows("Median of DNA abundance across studies (DNA ~ 2.39% Cmb)", 5 , 8) %>%
+  group_rows("Beck et al., 2018 (Protein ~ 60% Cmb)", 9 , 12) %>%
+  group_rows("Hanegraaf and Muller, 2001 (Protein ~ 182% Cmb)", 13 , 16) %>%
+  kable_styling(bootstrap_options = "condensed", full_width = F, font_size = 4.5) %>%
   kable_as_image()
@@
 
-```{r Scheme 1, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width = 20, fig.cap="_Schematic representation of three mathematical models, which are commonly used to predict the respiration rate.
+```{r Scheme 1, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width = 20, fig.cap="**Fig.
+1:** _Schematic representation of three mathematical models, which are commonly used to predict the respiration rate.
 $C_{S}$ - carbon substrate, $C_{MB}$ - microbial biomass carbon, $C_{Enz}$ - enzymes carbon._"}
@@
 ```{r Fig.
-1, echo=FALSE, resultsw = 'asis', message=FALSE, warning=FALSE, fig.height=10, fig.width=7, fig.cap="_Distribution of microbial biomass carbon data as directly measured by the DNA or cellular protein concentration and calculated by different conversion factors published in the literature [@Baart2008; @Beck2018; @VanDuuren2013; @Marstorp2000, @Christensen1993; Christensen1993; @Yokoyama2017; @VanPutten1995; @Henriksen1996; @Hanegraaf2001].
+1, echo=FALSE, resultsw = 'asis', message=FALSE, warning=FALSE, fig.height=8, fig.width=7, fig.cap="**Fig.
+2:** _Distribution of microbial biomass carbon data as directly measured by the DNA or cellular protein concentration and calculated by different conversion factors published in the literature [@Baart2008; @Beck2018; @VanDuuren2013; @Marstorp2000, @Christensen1993; Christensen1993; @Yokoyama2017; @VanPutten1995; @Henriksen1996; @Hanegraaf2001].
 Middle line corresonds to the median, lower and upper hinges to first and third quartile respectively, lower and upper whisker extends from lower and upper hinge to 1.5 time the interquartile range.

### 2018-07-24 data_analysis.R

@@ -399,10 +399,12 @@ mend_unique_results$ll_r
 ###############################################################################################
 source("../mmem_function.R")
 
+
 #across all treatments
 mmem_all_results<-mmem_function(data=dat, SUB = TRUE, FACT = 4, Vars = c("Substrate", "r", "Time"),
                                 ColM = c("time", "r", "r1", "r2"),Cmic=mean(dat$Cmic.dna.init, na.rm = T),
                                 VarsCmic = c("Substrate", "Cmic.dna", "Time"), Niter = 10000)
+
 
 #no_cors<-detectCores()-1
 #cl<-makeCluster(no_cors)
@@ -454,6 +456,8 @@ mmem_structures_results$ll_r
 mmem_substrates_results$ll_r
 mmem_unique_results$ll_r
 
+mmem_all_results_m$ll_r
+
 plot(obs_E~mod_E, mmem_unique_results$OvP_E)
 ####################################################################################################
 ####################################################################################################
@@ -473,22 +477,22 @@ Cmic.dna.init_H<-mean(m0$DNA.init/0.004*0.45/12.01/3*0.25, na.rm=T)
 #monod growth
 monod_all_results2<-monod_growth_function(data=dat, SUB = TRUE, FACT = 4, Vars = c("Substrate", "r", "Cmic.dna_H","Time"),
                                          ColM = c("time", "r", "Cmic","r1", "Cmic1","r2", "Cmic2"),Cmic=Cmic.dna.init_H,
-                                         VarsCmic = c("Substrate", "Cmic.dna_H", "Time"), Niter = 10000)
+                                         VarsCmic = c("Substrate", "Cmic.dna_H", "Time"), Niter = 100000)
 
 #mem
 mem_all_results2<-mem_function(data=dat, SUB = TRUE, FACT = 4, Vars = c("Substrate", "r", "Cmic.dna_H","Time"),
                               ColM = c("time", "r", "Cmic","r1", "Cmic1","r2", "Cmic2"),Cmic=Cmic.dna.init_H,
-                              VarsCmic = c("Substrate", "Cmic.dna_H", "Time"), Niter = 10000)
+                              VarsCmic = c("Substrate", "Cmic.dna_H", "Time"), Niter = 100000)
 
[…]

### 2018-08-15 manuscript_draft.Rmd
for different conversion factors

@@
 ####Microbial enzyme models (Fig. 1c)
-In microbial enzyme models, microbial biomass produce extracellular enzymes. The way the carbon is partitioned between $C_{MB}$, $CO_{2}$ and $C_{Enz}$ can follow different mechanisms.
+In microbial enzyme models, microbial biomass produce extracellular enzymes. The way the carbon is partitioned between $C_{MB}$, $CO_{2}$ and $C_{Enz}$ can follow different rules.
+
+* Classical microbial enzyme model (MEM)
+
+Microbial enzyme model presented by Allison et al. [-@Allison2010] assumes that extracellular enzymes are produced constitutively by the rate ($r_{Enz}$), which is proportional to biomass:
+
+[5] $~~~~~~~~~~enzyme~production~=~r_{Enz}~\times~C_{MB}$.
+
+When extracellular enzymes are produced, microbial biomass is loosing carbon. Therefore, the mass balance equation for microbial biomass is:
+
+[6] $~~~~~~~~~~\frac{dC_{MB}}{dt}~=~uptake~\times~CUE~-~death~-~enzyme~production$.
 

### 2018-08-15 data_analysis.R

@@ -1,14 +1,14 @@
 #################################################MinT#################################################
 ##libraries
 library(deSolve)
-library(dplyr)
+#library(dplyr)
 library(FME)
 library(reshape)
-library(reshape2)
+#library(reshape2)
 library(ggplot2)
 library(foreach)
 library(doParallel)
-library(data.table)
+#library(data.table)
 library(openxlsx)
 library(plyr)
 
@@ -192,23 +192,60 @@ ggplot(conversions, aes(Reference, Cmic))+geom_boxplot(cex=0.8, aes(colour=Organ
   facet_grid(.~Proxy)+coord_flip()+theme_min+theme(axis.title.y = element_blank())+
   ylab(expression(paste("Microbial biomass carbon (", mu, "mol ", ml^{-1},")")))
 
-######################################################################################################
-#carefull here, there is slightly different recalculation
-m0$Cmic.dna.init<-m0$DNA.init/0.0494*0.44/12.01/3*0.25
-######################################################################################################
-
-
 #extracellular protein to extracellular protein carbon (46% of carbon in protein - Vrede et al., 2004)
 m0$E<-m0$Prot.out*0.46/12.01/4
 
 
 #extracting only the columns of interest
-mint<-m0[,c("Structure", "Substrate", "r", "Time", "DOCinit", "Cmic.dna", "Cmic.dna.init", "E", "Cmic.prot.filled")]
+mint<-m0[,c("Structure", "Substrate", "r", "Time", "DOCinit", "DNA", "Prot.in", "E")]
 mint<-mint[!is.na(mint$Substrate), ]
 
 
 dat<-subset(mint, Substrate!="Free" & Substrate!="Celluloze")
 summary(dat)
+
+#Four datasets of microbial biomass are defined 
+
+#1. Highest DNA
+#measured microbial biomass
+dat$Cmic.dna_H<-m0[(m0$Substrate!="Free" & m0$Substrate!="Celluloze" & !is.na(m0$Substrate)), "DNA"]/0.004*0.45/12.01/4
+#measured initial microbial biomass
+Cmic.dna.init_H<-mean(m0$DNA.init/0.004*0.45/12.01/3*0.25, na.rm=T)
+
+#2. Median DNA - this is 2.39% of biomass - the best choice of a modeller
+#measured microbial biomass
[…]

### 2018-08-15 mem_function.R

@@ -1,4 +1,4 @@
-mem_function<-function(data, SUB, FACT, Vars, ColM, ColV, VarsCmic, Cmic, Niter){
+mem_function<-function(data, SUB, FACT, Vars, ColM, Mean, Cmic, Niter){
   
   #SUB - needs to be calibrated in matrix
   #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
@@ -72,22 +72,54 @@ mem_function<-function(data, SUB, FACT, Vars, ColM, ColV, VarsCmic, Cmic, Niter)
     #create cost function
     estim<-function(data){
       
-      Obs_decay<-select(data, Vars)
+      ###Glucose
+      gl<-select(filter(data, Substrate=="Glucose"), Vars)
+      colnames(gl)<-c("time", "r", "E" ,"Cmic")
+      gl2<-select(filter(data, Substrate!="Glucose"), Vars[1])
+      colnames(gl2)<-"time"
+      gl2$r<-NA
+      gl2$E<-NA
+      gl2$Cmic<-NA
+      m1<-rbind(gl, gl2)
+      m1<-m1[order(m1$time),]
       
-      m1<-merge(select(filter(Obs_decay, Substrate=="Glucose"), Vars[-1]),
-                select(filter(Obs_decay, Substrate=="Cellobiose"), Vars[-1]),all = T, by="Time")
+      ###Cellobiose
+      cel<-select(filter(data, Substrate=="Cellobiose"), Vars)
+      colnames(cel)<-c("time", "r1", "E1" ,"Cmic1")
+      cel2<-select(filter(data, Substrate!="Cellobiose"), Vars[1])
+      colnames(cel2)<-"time"
+      cel2$r1<-NA
+      cel2$E1<-NA
+      cel2$Cmic1<-NA
+      m2<-rbind(cel, cel2)
+      m2<-m2[order(m2$time),]
       
-      m2<-merge(m1, 
-                select(filter(Obs_decay, Substrate=="Mix"), Vars[-1]),all = T, by="Time")
+      ###Mix
+      mix<-select(filter(data, Substrate=="Mix"), Vars)
+      colnames(mix)<-c("time", "r2", "E2" ,"Cmic2")
+      mix2<-select(filter(data, Substrate!="Mix"), Vars[1])
+      colnames(mix2)<-"time"
+      mix2$r2<-NA
+      mix2$E2<-NA
+      mix2$Cmic2<-NA
[…]

### 2018-08-15 mend_function.R

@@ -1,4 +1,4 @@
-mend_function<-function(data, SUB, FACT, Vars, ColM, ColV, VarsCmic, Cmic, Niter){
+mend_function<-function(data, SUB, FACT, Vars, ColM, Mean, Cmic, Niter){
   
   #SUB - needs to be calibrated in matrix
   #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
@@ -87,22 +87,75 @@ mend_function<-function(data, SUB, FACT, Vars, ColM, ColV, VarsCmic, Cmic, Niter
     #create cost function
     estim<-function(data){
       
-      Obs_decay<-select(data, Vars)
+      ###Glucose
+      gl<-select(filter(data, Substrate=="Glucose" ), Vars)
+      colnames(gl)<-c("time", "r", "E" ,"Cmic")
+      #round the time to one digit
+      gl$time<-round(gl$time, 1)
       
-      m1<-merge(select(filter(Obs_decay, Substrate=="Glucose"), Vars[-1]),
-                select(filter(Obs_decay, Substrate=="Cellobiose"), Vars[-1]),all = T, by="Time")
+      gl2<-select(filter(data, Substrate!="Glucose" ), Vars[1])
+      colnames(gl2)<-"time"
+      #round the time to one digit
+      gl2$time<-round(gl2$time, 1)
       
-      m2<-merge(m1, 
-                select(filter(Obs_decay, Substrate=="Mix"), Vars[-1]),all = T, by="Time")
+      gl2$r<-NA
+      gl2$E<-NA
+      gl2$Cmic<-NA
+      m1<-rbind(gl, gl2)
+      m1<-m1[order(m1$time),]
       
-      colnames(m2)<-ColM
+      ###Cellobiose
+      cel<-select(filter(data, Substrate=="Cellobiose" ), Vars)
+      colnames(cel)<-c("time", "r1", "E1" ,"Cmic1")
+      #round the time to one digit
+      cel$time<-round(cel$time, 1)
+      
+      cel2<-select(filter(data, Substrate!="Cellobiose" ), Vars[1])
+      colnames(cel2)<-"time"
+      #round the time to one digit
+      cel2$time<-round(cel2$time, 1)
+      
+      
+      cel2$r1<-NA
+      cel2$E1<-NA
+      cel2$Cmic1<-NA
+      m2<-rbind(cel, cel2)
+      m2<-m2[order(m2$time),]
+      
+      ###Mix
[…]

### 2018-08-15 mmem_function.R

@@ -1,4 +1,4 @@
-mmem_function<-function(data, SUB, FACT, Vars, ColM, ColV, VarsCmic, Cmic, Niter){
+mmem_function<-function(data, SUB, FACT, Vars, ColM, Mean, Cmic, Niter){
   
   #SUB - needs to be calibrated in matrix
   #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
@@ -53,15 +53,15 @@ mmem_function<-function(data, SUB, FACT, Vars, ColM, ColV, VarsCmic, Cmic, Niter
           #defining stochiometric coefficients
           #PO ratio
           #PO=1.5*(1-((Cu/Cmic)/((Cu/Cmic)+Sover)))
-          y=2*1.5*Yatp
+          y=2*PO*Yatp
           
           #x
           #x=(Cu/Cmic)/((Cu/Cmic)+Sprod)
           
           #coefficients
-          psi.r=1/((y*x)+(y*0.64*(1-x))+1)
-          psi.g=y*x/((y*x)+(y*0.64*(1-x))+1)
-          psi.e=0.64*y*(1-x)/((y*x)+(y*0.64*(1-x))+1)
+          psi.r=1/((y*x)+(y*w*(1-x))+1)
+          psi.g=y*x/((y*x)+(y*w*(1-x))+1)
+          psi.e=w*y*(1-x)/((y*x)+(y*w*(1-x))+1)
           
           
           dCmic<--kmic*Cmic+Cu*psi.g
@@ -90,30 +90,62 @@ mmem_function<-function(data, SUB, FACT, Vars, ColM, ColV, VarsCmic, Cmic, Niter
     #create cost function
     estim<-function(data){
       
-      Obs_decay<-select(data, Vars)
+      ###Glucose
+      gl<-select(filter(data, Substrate=="Glucose" ), Vars)
+      colnames(gl)<-c("time", "r", "E" ,"Cmic")
+      gl2<-select(filter(data, Substrate!="Glucose" ), Vars[1])
+      colnames(gl2)<-"time"
+      gl2$r<-NA
+      gl2$E<-NA
+      gl2$Cmic<-NA
+      m1<-rbind(gl, gl2)
+      m1<-m1[order(m1$time),]
       
-      m1<-merge(select(filter(Obs_decay, Substrate=="Glucose"), Vars[-1]),
[…]

### 2018-08-15 monod_growth_function.R

@@ -1,5 +1,5 @@
 ###monod growth function
-monod_growth_function<-function(data, SUB, FACT, Vars, ColM, ColV, VarsCmic, Cmic, Niter){
+monod_growth_function<-function(data, SUB, FACT, Vars, ColM, Mean, Cmic, Niter){
   
   #SUB - needs to be calibrated in matrix
   #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
@@ -72,22 +72,64 @@ monod_growth_function<-function(data, SUB, FACT, Vars, ColM, ColV, VarsCmic, Cmi
     #create cost function
     estim<-function(data){
       
-      Obs_decay<-select(data, Vars)
+      # Obs_decay<-select(data, Vars)
+      # 
+      # m1<-merge(select(filter(Obs_decay, Substrate=="Glucose"), Vars[-1]),
+      #           select(filter(Obs_decay, Substrate=="Cellobiose"), Vars[-1]),all = T, by="Time")
+      # 
+      # m2<-merge(m1, 
+      #           select(filter(Obs_decay, Substrate=="Mix"), Vars[-1]),all = T, by="Time")
+      # 
+      # colnames(m2)<-ColM
       
-      m1<-merge(select(filter(Obs_decay, Substrate=="Glucose"), Vars[-1]),
-                select(filter(Obs_decay, Substrate=="Cellobiose"), Vars[-1]),all = T, by="Time")
+      #more accurate
+      ###Glucose
+      gl<-select(filter(data, Substrate=="Glucose"), Vars)
+      colnames(gl)<-c("time", "r", "E" ,"Cmic")
+      gl2<-select(filter(data, Substrate!="Glucose" ), Vars[1])
+      colnames(gl2)<-"time"
+      gl2$r<-NA
+      gl2$E<-NA
+      gl2$Cmic<-NA
+      m1<-rbind(gl, gl2)
+      m1<-m1[order(m1$time),]
       
-      m2<-merge(m1, 
-                select(filter(Obs_decay, Substrate=="Mix"), Vars[-1]),all = T, by="Time")
+      ###Cellobiose
+      cel<-select(filter(data, Substrate=="Cellobiose"), Vars)
+      colnames(cel)<-c("time", "r1", "E1" ,"Cmic1")
+      cel2<-select(filter(data, Substrate!="Cellobiose"), Vars[1])
[…]

### 2018-08-15 data_analysis.R
initial state variable estimation functions

@@ -247,6 +247,9 @@ Cmic.prot.init_H<-mean((m0$DNA.init*fill_coefs[2]+fill_coefs[1])/0.82*0.45/12.01
 dat$DNAc<-dat$DNA*0.51/12.01/4
 dat$Protc<-dat$Prot.in*0.46/12.01/4
 
+#initial DNA concentration
+DNAci<-mean(m0$DNA.init, na.rm = T)*0.46/12.01/3*0.25
+
 #######################################################################################
 ##############################First order decay########################################
 #######################################################################################
@@ -2624,6 +2627,134 @@ an_each_allb$ll
 an_each_allc$ll
 an_each_alld$ll
 
+#############################################################################################
+##########################################Initial condition estimation#######################
+#############################################################################################
+#becuase conversion factors give extremely different results, all models are calibrated against
+#DNA or protein content. The percentage of DNA/Protein content in microbial biomass and thus,
+#initial biomass as well, is estimated as two additional parameters
+
+###########################################################################################
+#########################################MEm##############################################
+###########################################################################################
+source("../mem_init.R")
+
+#across all treatments
+mem_alli<-mem_init(data=dat, SUB = TRUE, FACT = 4, Mean=TRUE, DNAci=DNAci, Niter = 10000)
+
+
+no_cors<-detectCores()-1
+cl<-makeCluster(no_cors)
+registerDoParallel(cl)
+
+#for different structures
+mem_structuresi<-mem_init(data=dat, SUB = TRUE, FACT = 2, Mean=TRUE, DNAci=DNAci, Niter = 10000)
+
+#for different substrates
+mem_substratesi<-mem_init(data=dat, SUB = FALSE, FACT = 1, Mean=TRUE, DNAci=DNAci, Niter = 10000)
+
+#for each separately 
+mem_eachi<-mem_init(data=dat, SUB = FALSE, FACT = 3, Mean=TRUE, DNAci=DNAci, Niter = 10000)
+
+stopImplicitCluster()
+
+
+mem_alli$ll
+mem_structuresi$ll
+mem_substratesi$ll
+mend_eachi$ll
+
+
+###########################################################################################
[…]

### 2018-08-15 andrew_function.R
state variables estimation functions

@@ -1,0 +1,839 @@
+andrew_function<-function(data, SUB, FACT, Vars, ColM, Mean, Cmic, Niter){
+  
+  #SUB - needs to be calibrated in matrix
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  #Vars = names of variables to calibrated the model against
+  #ColM = colnames for these varaibles in matrix
+  #ColV = colnames for these varaibles in normal setting
+  #VarsCmic = which biomass to use for likelihood calculation (it should follow the same notation as Vars)
+  #Cmic = initial microbial biomass C
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-setDT(data)[, id := .GRP, by = .(Substrate)]
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-setDT(data)[, id := .GRP, by = .(Structure)]
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-setDT(data)[, id := .GRP, by = .(Substrate,Structure)]
+      }else{
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  ############################################################################################
+  #if parameters are estimated across different substrates, different initial C conc
+  #has to be acknowledged
+  
+  if(SUB==TRUE){
+    
+    #this is the mmem model fitted across different substrates 
+    #with different initial carbon concentrations
+    mmem<-function(X, pars, t){
+      
+      #mmem model
+      deriv<-function(time, state, pars){
+        
+        with(as.list(c(state, pars)),{
+          
+          #specific growth rate
+          
+          u=umax*C/(Km+C+Ki*C^2)
+          
+          #respiration rate
+          r=m*Cmic+(1/Ymic-1)*Cmic*u+(1/Ye-1)*(Cmic*u*alfa)
+          
+          
+          dCmic<-Cmic*u-kmic*Cmic
+          dC<--(1/Ymic)*Cmic*u-(1/Ye)*(Cmic*u*alfa)-m*Cmic+kmic*Cmic+ke*E
+          dE<-Cmic*u*alfa-ke*E
+          
+          return(list(c(dCmic, dC, dE), r=r))
+          
+        })
+      }
+      
+      #function to apply acros 3 different initial substrate concentrations
[…]

### 2018-08-15 andrew_init.R

@@ -1,0 +1,932 @@
+andrew_init<-function(data, SUB, FACT, Mean, DNAci, Niter){
+  
+  #SUB - needs to be calibrated in matrix
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  #Vars = names of variables to calibrated the model against
+  #ColM = colnames for these varaibles in matrix
+  #ColV = colnames for these varaibles in normal setting
+  #VarsCmic = which biomass to use for likelihood calculation (it should follow the same notation as Vars)
+  #Cmic = initial microbial biomass C
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-setDT(data)[, id := .GRP, by = .(Substrate)]
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-setDT(data)[, id := .GRP, by = .(Structure)]
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-setDT(data)[, id := .GRP, by = .(Substrate,Structure)]
+      }else{
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  ############################################################################################
+  #if parameters are estimated across different substrates, different initial C conc
+  #has to be acknowledged
+  
+  if(SUB==TRUE){
+    
+    #this is the mmem model fitted across different substrates 
+    #with different initial carbon concentrations
+    mmem<-function(X, pars, t){
+      
+      #mmem model
+      deriv<-function(time, state, pars){
+        
+        with(as.list(c(state, pars)),{
+          
+          #specific growth rate
+          
+          u=umax*C/(Km+C+Ki*C^2)
+          
+          #respiration rate
+          r=m*Cmic+(1/Ymic-1)*Cmic*u+(1/Ye-1)*(Cmic*u*alfa)
+          
+          
+          dCmic<-Cmic*u-kmic*Cmic
+          dC<--(1/Ymic)*Cmic*u-(1/Ye)*(Cmic*u*alfa)-m*Cmic+kmic*Cmic+ke*E
+          dE<-Cmic*u*alfa-ke*E
+          
+          #DNA and Protein content in biomass
+          DNAc=fd*Cmic
+          Protc=fp*Cmic
+          
+          return(list(c(dCmic, dC, dE), r=r, DNAc=DNAc, Protc=Protc))
+          
[…]

### 2018-08-15 deb_function.R

@@ -1,0 +1,942 @@
+deb_function<-function(data, SUB, FACT, Mean, Niter){
+  
+  #SUB - needs to be calibrated in matrix
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  #Vars = names of variables to calibrated the model against
+  #ColM = colnames for these varaibles in matrix
+  #ColV = colnames for these varaibles in normal setting
+  #VarsCmic = which biomass to use for likelihood calculation (it should follow the same notation as Vars)
+  #Cmic = initial microbial biomass C
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-setDT(data)[, id := .GRP, by = .(Substrate)]
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-setDT(data)[, id := .GRP, by = .(Structure)]
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-setDT(data)[, id := .GRP, by = .(Substrate,Structure)]
+      }else{
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  ############################################################################################
+  #if parameters are estimated across different substrates, different initial C conc
+  #has to be acknowledged
+  
+  if(SUB==TRUE){
+    
+    #this is the mmem model fitted across different substrates 
+    #with different initial carbon concentrations
+    mmem<-function(X, pars, t){
+      
+      #mmem model
+      deriv<-function(time, state, pars){
+        
+        with(as.list(c(state, pars)),{
+          
+          #Carbon uptake
+          Cu=Vmax*C*S/(Km+C)
+          
+          #Assimilation
+          #A=(Amax/Vmax)*Cu
+          
+          #maintnance
+          m=S*m0
+          
+          #reserve mobilization rate for growth and enzyme production
+          an=f*R-S*m0
+          
+          #growth
+          g=x*an
+          
+          #enzyme production
+          p=(1-x)*an
+          
+          #calibrated variables
+          #DNA and protein abundance in reserves and structures are mean values reported 
[…]

### 2018-08-15 mem_init.R

@@ -1,0 +1,890 @@
+mem_init<-function(data, SUB, FACT, Mean, Niter, DNAci){
+  
+  #SUB - needs to be calibrated in matrix
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  #Vars = names of variables to calibrated the model against
+  #ColM = colnames for these varaibles in matrix
+  #ColV = colnames for these varaibles in normal setting
+  #VarsCmic = which biomass to use for likelihood calculation (it should follow the same notation as Vars)
+  #Cmic = initial microbial biomass C
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-setDT(data)[, id := .GRP, by = .(Substrate)]
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-setDT(data)[, id := .GRP, by = .(Structure)]
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-setDT(data)[, id := .GRP, by = .(Substrate,Structure)]
+      }else{
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  ############################################################################################
+  #if parameters are estimated across different substrates, different initial C conc
+  #has to be acknowledged
+  
+  if(SUB==TRUE){
+    
+    #this is the mem model fitted across different substrates 
+    #with different initial carbon concentrations
+    mem<-function(X, pars, t){
+      
+      #mem model
+      deriv<-function(time, state, pars){
+        
+        with(as.list(c(state, pars)),{
+          
+          dCmic<--kmic*Cmic+CUE*Vmax*Cmic*C/(Km+C)-pe*Cmic
+          dC<-kmic*Cmic-Vmax*Cmic*C/(Km+C)+ke*E
+          dE<-pe*Cmic-ke*E
+          
+          #DNA and Protein calculation
+          DNAc=fd*Cmic
+          Protc=fp*Cmic
+          
+          return(list(c(dCmic, dC, dE),  DNAc=DNAc, Protc=Protc, r=(1-CUE)*Vmax*Cmic*C/(Km+C)))
+          
+        })
+      }
+      
+      #function to apply acros 3 different initial substrate concentrations
[…]

### 2018-08-15 mend_init.R

@@ -1,0 +1,923 @@
+mend_init<-function(data, SUB, FACT, Mean, Niter, DNAci){
+  
+  #SUB - needs to be calibrated in matrix
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  #Vars = names of variables to calibrated the model against
+  #ColM = colnames for these varaibles in matrix
+  #ColV = colnames for these varaibles in normal setting
+  #VarsCmic = which biomass to use for likelihood calculation (it should follow the same notation as Vars)
+  #Cmic = initial microbial biomass C
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-setDT(data)[, id := .GRP, by = .(Substrate)]
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-setDT(data)[, id := .GRP, by = .(Structure)]
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-setDT(data)[, id := .GRP, by = .(Substrate,Structure)]
+      }else{
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  ############################################################################################
+  #if parameters are estimated across different substrates, different initial C conc
+  #has to be acknowledged
+  
+  if(SUB==TRUE){
+    
+    #this is the mend model fitted across different substrates 
+    #with different initial carbon concentrations
+    mend<-function(X, pars, t){
+      
+      #mend model
+      deriv<-function(time, state, pars){
+        
+        with(as.list(c(state, pars)),{
+          
+          #equations
+          #C uptake
+          F1=1/CUE*(Vmax+mr)*Cmic*C/(Km+C)
+          #growth respiration
+          F4=(1/CUE-1)*Vmax*Cmic*C/(Km+C)
+          #maintnance respiration
+          F5=(1/CUE-1)*mr*Cmic*C/(Km+C)
+          #microbial mortality
+          F8=(1-pe)*mr*Cmic
+          #enzyme production
+          F9=pe*mr*Cmic
+          #enzyme decay
+          F10=ke*E
+          
+          #states
+          dCmic<-F1-(F4+F5)-F8-F9
[…]

### 2018-08-15 mmem_init.R

@@ -1,0 +1,940 @@
+mmem_init<-function(data, SUB, FACT, Mean, DNAci, Niter){
+  
+  #SUB - needs to be calibrated in matrix
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  #Vars = names of variables to calibrated the model against
+  #ColM = colnames for these varaibles in matrix
+  #ColV = colnames for these varaibles in normal setting
+  #VarsCmic = which biomass to use for likelihood calculation (it should follow the same notation as Vars)
+  #Cmic = initial microbial biomass C
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-setDT(data)[, id := .GRP, by = .(Substrate)]
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-setDT(data)[, id := .GRP, by = .(Structure)]
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-setDT(data)[, id := .GRP, by = .(Substrate,Structure)]
+      }else{
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  ############################################################################################
+  #if parameters are estimated across different substrates, different initial C conc
+  #has to be acknowledged
+  
+  if(SUB==TRUE){
+    
+    #this is the mmem model fitted across different substrates 
+    #with different initial carbon concentrations
+    mmem<-function(X, pars, t){
+      
+      #mmem model
+      deriv<-function(time, state, pars){
+        
+        with(as.list(c(state, pars)),{
+          
+          #Carbon uptake
+          
+          Cu=Vmax*C*Cmic/(Km+C)
+          
+          #defining stochiometric coefficients
+          #PO ratio
+          #PO=1.5*(1-((Cu/Cmic)/((Cu/Cmic)+Sover)))
+          y=2*PO*Yatp
+          
+          #x
+          #x=(Cu/Cmic)/((Cu/Cmic)+Sprod)
+          
+          #coefficients
+          psi.r=1/((y*x)+(y*w*(1-x))+1)
+          psi.g=y*x/((y*x)+(y*w*(1-x))+1)
+          psi.e=w*y*(1-x)/((y*x)+(y*w*(1-x))+1)
+          
+          
[…]

### 2018-08-16 andrew_init.R
initial condition estimation functions

@@ -801,7 +801,7 @@ andrew_init<-function(data, SUB, FACT, Mean, DNAci, Niter){
       mtrue<-select(mtrue, c("time", "DNAc"))
       
       out<-as.data.frame(ode(y=c(Cmic=pars[["Cmic_0"]], C=cinit, E=0), parms=pars, times=seq(0,130), func=deriv))
-      cost<-modCost(model = out, obs = Obs_dat)
+      cost<-modCost(model = out, obs = mtrue)
       
       return(cost)
       
@@ -843,7 +843,7 @@ andrew_init<-function(data, SUB, FACT, Mean, DNAci, Niter){
       mtrue<-select(mtrue, c("time", "Protc"))
       
       out<-as.data.frame(ode(y=c(Cmic=pars[["Cmic_0"]], C=cinit, E=0), parms=pars, times=seq(0,130), func=deriv))
-      cost<-modCost(model = out, obs = Obs_dat)
+      cost<-modCost(model = out, obs = mtrue)
       
       return(cost)
       

### 2018-08-16 data_analysis.R

@@ -1,14 +1,14 @@
 #################################################MinT#################################################
 ##libraries
 library(deSolve)
-#library(dplyr)
+library(dplyr)
 library(FME)
 library(reshape)
 #library(reshape2)
 library(ggplot2)
 library(foreach)
 library(doParallel)
-#library(data.table)
+library(data.table)
 library(openxlsx)
 library(plyr)
 
@@ -2633,10 +2633,68 @@ an_each_alld$ll
 #becuase conversion factors give extremely different results, all models are calibrated against
 #DNA or protein content. The percentage of DNA/Protein content in microbial biomass and thus,
 #initial biomass as well, is estimated as two additional parameters
+#I will try to calibrate against DNA only, protein only and DNA and Protein together
 
 ###########################################################################################
-#########################################MEm##############################################
+#########################################MEM##############################################
 ###########################################################################################
+#Using DNA
+source("../mem_init_a.R")
+
+#across all treatments
+mem_alli_a<-mem_init_a(data=dat, SUB = TRUE, FACT = 4, Mean=TRUE, DNAci=DNAci, Niter = 10000)
+
+
+no_cors<-detectCores()-1
+cl<-makeCluster(no_cors)
+registerDoParallel(cl)
+
+#for different structures
+mem_structuresi_a<-mem_init_a(data=dat, SUB = TRUE, FACT = 2, Mean=TRUE, DNAci=DNAci, Niter = 10000)
+
+#for different substrates
+mem_substratesi_a<-mem_init_a(data=dat, SUB = FALSE, FACT = 1, Mean=TRUE, DNAci=DNAci, Niter = 10000)
+
+#for each separately 
+mem_eachi_a<-mem_init_a(data=dat, SUB = FALSE, FACT = 3, Mean=TRUE, DNAci=DNAci, Niter = 10000)
+
+stopImplicitCluster()
+
+
+mem_alli_a$ll
+mem_structuresi_a$ll
+mem_substratesi_a$ll
[…]

### 2018-08-16 mem_init.R

@@ -759,7 +759,7 @@ mem_init<-function(data, SUB, FACT, Mean, Niter, DNAci){
       mtrue<-select(mtrue, c("time", "DNAc"))
       
       out<-as.data.frame(ode(y=c(Cmic=pars[["Cmic_0"]], C=cinit, E=0), parms=pars, times=seq(0,130), func=deriv))
-      cost<-modCost(model = out, obs = Obs_dat)
+      cost<-modCost(model = out, obs = mtrue)
       
       return(cost)
       
@@ -801,7 +801,7 @@ mem_init<-function(data, SUB, FACT, Mean, Niter, DNAci){
       mtrue<-select(mtrue, c("time", "Protc"))
       
       out<-as.data.frame(ode(y=c(Cmic=pars[["Cmic_0"]], C=cinit, E=0), parms=pars, times=seq(0,130), func=deriv))
-      cost<-modCost(model = out, obs = Obs_dat)
+      cost<-modCost(model = out, obs = mtrue)
       
       return(cost)
       

### 2018-08-16 mend_init.R

@@ -792,7 +792,7 @@ mend_init<-function(data, SUB, FACT, Mean, Niter, DNAci){
       mtrue<-select(mtrue, c("time", "DNAc"))
       
       out<-as.data.frame(ode(y=c(Cmic=pars[["Cmic_0"]], C=cinit, E=0), parms=pars, times=seq(0,130), func=deriv))
-      cost<-modCost(model = out, obs = Obs_dat)
+      cost<-modCost(model = out, obs = mtrue)
       
       return(cost)
       
@@ -834,7 +834,7 @@ mend_init<-function(data, SUB, FACT, Mean, Niter, DNAci){
       mtrue<-select(mtrue, c("time", "Protc"))
       
       out<-as.data.frame(ode(y=c(Cmic=pars[["Cmic_0"]], C=cinit, E=0), parms=pars, times=seq(0,130), func=deriv))
-      cost<-modCost(model = out, obs = Obs_dat)
+      cost<-modCost(model = out, obs = mtrue)
       
       return(cost)
       

### 2018-08-16 mmem_init.R

@@ -809,7 +809,7 @@ mmem_init<-function(data, SUB, FACT, Mean, DNAci, Niter){
       mtrue<-select(mtrue, c("time", "DNAc"))
       
       out<-as.data.frame(ode(y=c(Cmic=pars[["Cmic_0"]], C=cinit, E=0), parms=pars, times=seq(0,130), func=deriv))
-      cost<-modCost(model = out, obs = Obs_dat)
+      cost<-modCost(model = out, obs = mtrue)
       
       return(cost)
       
@@ -851,7 +851,7 @@ mmem_init<-function(data, SUB, FACT, Mean, DNAci, Niter){
       mtrue<-select(mtrue, c("time", "Protc"))
       
       out<-as.data.frame(ode(y=c(Cmic=pars[["Cmic_0"]], C=cinit, E=0), parms=pars, times=seq(0,130), func=deriv))
-      cost<-modCost(model = out, obs = Obs_dat)
+      cost<-modCost(model = out, obs = mtrue)
       
       return(cost)
       

### 2018-08-16 andrew_init_a.R
specific initial condition estimation function - for DNA or protein content

@@ -1,0 +1,932 @@
+andrew_init_a<-function(data, SUB, FACT, Mean, DNAci, Niter){
+  
+  #SUB - needs to be calibrated in matrix
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  #Vars = names of variables to calibrated the model against
+  #ColM = colnames for these varaibles in matrix
+  #ColV = colnames for these varaibles in normal setting
+  #VarsCmic = which biomass to use for likelihood calculation (it should follow the same notation as Vars)
+  #Cmic = initial microbial biomass C
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-setDT(data)[, id := .GRP, by = .(Substrate)]
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-setDT(data)[, id := .GRP, by = .(Structure)]
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-setDT(data)[, id := .GRP, by = .(Substrate,Structure)]
+      }else{
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  ############################################################################################
+  #if parameters are estimated across different substrates, different initial C conc
+  #has to be acknowledged
+  
+  if(SUB==TRUE){
+    
+    #this is the mmem model fitted across different substrates 
+    #with different initial carbon concentrations
+    mmem<-function(X, pars, t){
+      
+      #mmem model
+      deriv<-function(time, state, pars){
+        
+        with(as.list(c(state, pars)),{
+          
+          #specific growth rate
+          
+          u=umax*C/(Km+C+Ki*C^2)
+          
+          #respiration rate
+          r=m*Cmic+(1/Ymic-1)*Cmic*u+(1/Ye-1)*(Cmic*u*alfa)
+          
+          
+          dCmic<-Cmic*u-kmic*Cmic
+          dC<--(1/Ymic)*Cmic*u-(1/Ye)*(Cmic*u*alfa)-m*Cmic+kmic*Cmic+ke*E
+          dE<-Cmic*u*alfa-ke*E
+          
+          #DNA and Protein content in biomass
+          DNAc=fd*Cmic
+          #Protc=fp*Cmic
+          
+          return(list(c(dCmic, dC, dE), r=r, DNAc=DNAc))
+          
+        })
[…]

### 2018-08-16 andrew_init_b.R

@@ -1,0 +1,924 @@
+andrew_init_b<-function(data, SUB, FACT, Mean, DNAci, Niter){
+  
+  #SUB - needs to be calibrated in matrix
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  #Vars = names of variables to calibrated the model against
+  #ColM = colnames for these varaibles in matrix
+  #ColV = colnames for these varaibles in normal setting
+  #VarsCmic = which biomass to use for likelihood calculation (it should follow the same notation as Vars)
+  #Cmic = initial microbial biomass C
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-setDT(data)[, id := .GRP, by = .(Substrate)]
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-setDT(data)[, id := .GRP, by = .(Structure)]
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-setDT(data)[, id := .GRP, by = .(Substrate,Structure)]
+      }else{
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  ############################################################################################
+  #if parameters are estimated across different substrates, different initial C conc
+  #has to be acknowledged
+  
+  if(SUB==TRUE){
+    
+    #this is the mmem model fitted across different substrates 
+    #with different initial carbon concentrations
+    mmem<-function(X, pars, t){
+      
+      #mmem model
+      deriv<-function(time, state, pars){
+        
+        with(as.list(c(state, pars)),{
+          
+          #specific growth rate
+          
+          u=umax*C/(Km+C+Ki*C^2)
+          
+          #respiration rate
+          r=m*Cmic+(1/Ymic-1)*Cmic*u+(1/Ye-1)*(Cmic*u*alfa)
+          
+          
+          dCmic<-Cmic*u-kmic*Cmic
+          dC<--(1/Ymic)*Cmic*u-(1/Ye)*(Cmic*u*alfa)-m*Cmic+kmic*Cmic+ke*E
+          dE<-Cmic*u*alfa-ke*E
+          
+          #DNA and Protein content in biomass
+          #DNAc=fd*Cmic
+          Protc=fp*Cmic
+          
+          return(list(c(dCmic, dC, dE), r=r, Protc=Protc))
+          
+        })
+      }
[…]

### 2018-08-16 mem_init_a.R

@@ -1,0 +1,890 @@
+mem_init_a<-function(data, SUB, FACT, Mean, Niter, DNAci){
+  
+  #SUB - needs to be calibrated in matrix
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  #Vars = names of variables to calibrated the model against
+  #ColM = colnames for these varaibles in matrix
+  #ColV = colnames for these varaibles in normal setting
+  #VarsCmic = which biomass to use for likelihood calculation (it should follow the same notation as Vars)
+  #Cmic = initial microbial biomass C
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-setDT(data)[, id := .GRP, by = .(Substrate)]
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-setDT(data)[, id := .GRP, by = .(Structure)]
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-setDT(data)[, id := .GRP, by = .(Substrate,Structure)]
+      }else{
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  ############################################################################################
+  #if parameters are estimated across different substrates, different initial C conc
+  #has to be acknowledged
+  
+  if(SUB==TRUE){
+    
+    #this is the mem model fitted across different substrates 
+    #with different initial carbon concentrations
+    mem<-function(X, pars, t){
+      
+      #mem model
+      deriv<-function(time, state, pars){
+        
+        with(as.list(c(state, pars)),{
+          
+          dCmic<--kmic*Cmic+CUE*Vmax*Cmic*C/(Km+C)-pe*Cmic
+          dC<-kmic*Cmic-Vmax*Cmic*C/(Km+C)+ke*E
+          dE<-pe*Cmic-ke*E
+          
+          #DNA and Protein calculation
+          DNAc=fd*Cmic
+          #Protc=fp*Cmic
+          
+          return(list(c(dCmic, dC, dE),  DNAc=DNAc, r=(1-CUE)*Vmax*Cmic*C/(Km+C)))
+          
+        })
+      }
+      
+      #function to apply acros 3 different initial substrate concentrations
[…]

### 2018-08-16 mem_init_b.R

@@ -1,0 +1,876 @@
+mem_init_b<-function(data, SUB, FACT, Mean, Niter, DNAci){
+  
+  #SUB - needs to be calibrated in matrix
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  #Vars = names of variables to calibrated the model against
+  #ColM = colnames for these varaibles in matrix
+  #ColV = colnames for these varaibles in normal setting
+  #VarsCmic = which biomass to use for likelihood calculation (it should follow the same notation as Vars)
+  #Cmic = initial microbial biomass C
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-setDT(data)[, id := .GRP, by = .(Substrate)]
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-setDT(data)[, id := .GRP, by = .(Structure)]
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-setDT(data)[, id := .GRP, by = .(Substrate,Structure)]
+      }else{
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  ############################################################################################
+  #if parameters are estimated across different substrates, different initial C conc
+  #has to be acknowledged
+  
+  if(SUB==TRUE){
+    
+    #this is the mem model fitted across different substrates 
+    #with different initial carbon concentrations
+    mem<-function(X, pars, t){
+      
+      #mem model
+      deriv<-function(time, state, pars){
+        
+        with(as.list(c(state, pars)),{
+          
+          dCmic<--kmic*Cmic+CUE*Vmax*Cmic*C/(Km+C)-pe*Cmic
+          dC<-kmic*Cmic-Vmax*Cmic*C/(Km+C)+ke*E
+          dE<-pe*Cmic-ke*E
+          
+          #DNA and Protein calculation
+          #DNAc=fd*Cmic
+          Protc=fp*Cmic
+          
+          return(list(c(dCmic, dC, dE),  Protc=Protc, r=(1-CUE)*Vmax*Cmic*C/(Km+C)))
+          
+        })
+      }
+      
+      #function to apply acros 3 different initial substrate concentrations
[…]

### 2018-08-16 mend_init_a.R

@@ -1,0 +1,923 @@
+mend_init_a<-function(data, SUB, FACT, Mean, Niter, DNAci){
+  
+  #SUB - needs to be calibrated in matrix
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  #Vars = names of variables to calibrated the model against
+  #ColM = colnames for these varaibles in matrix
+  #ColV = colnames for these varaibles in normal setting
+  #VarsCmic = which biomass to use for likelihood calculation (it should follow the same notation as Vars)
+  #Cmic = initial microbial biomass C
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-setDT(data)[, id := .GRP, by = .(Substrate)]
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-setDT(data)[, id := .GRP, by = .(Structure)]
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-setDT(data)[, id := .GRP, by = .(Substrate,Structure)]
+      }else{
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  ############################################################################################
+  #if parameters are estimated across different substrates, different initial C conc
+  #has to be acknowledged
+  
+  if(SUB==TRUE){
+    
+    #this is the mend model fitted across different substrates 
+    #with different initial carbon concentrations
+    mend<-function(X, pars, t){
+      
+      #mend model
+      deriv<-function(time, state, pars){
+        
+        with(as.list(c(state, pars)),{
+          
+          #equations
+          #C uptake
+          F1=1/CUE*(Vmax+mr)*Cmic*C/(Km+C)
+          #growth respiration
+          F4=(1/CUE-1)*Vmax*Cmic*C/(Km+C)
+          #maintnance respiration
+          F5=(1/CUE-1)*mr*Cmic*C/(Km+C)
+          #microbial mortality
+          F8=(1-pe)*mr*Cmic
+          #enzyme production
+          F9=pe*mr*Cmic
+          #enzyme decay
+          F10=ke*E
+          
+          #states
+          dCmic<-F1-(F4+F5)-F8-F9
[…]

### 2018-08-16 mend_init_b.R

@@ -1,0 +1,914 @@
+mend_init_b<-function(data, SUB, FACT, Mean, Niter, DNAci){
+  
+  #SUB - needs to be calibrated in matrix
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  #Vars = names of variables to calibrated the model against
+  #ColM = colnames for these varaibles in matrix
+  #ColV = colnames for these varaibles in normal setting
+  #VarsCmic = which biomass to use for likelihood calculation (it should follow the same notation as Vars)
+  #Cmic = initial microbial biomass C
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-setDT(data)[, id := .GRP, by = .(Substrate)]
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-setDT(data)[, id := .GRP, by = .(Structure)]
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-setDT(data)[, id := .GRP, by = .(Substrate,Structure)]
+      }else{
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  ############################################################################################
+  #if parameters are estimated across different substrates, different initial C conc
+  #has to be acknowledged
+  
+  if(SUB==TRUE){
+    
+    #this is the mend model fitted across different substrates 
+    #with different initial carbon concentrations
+    mend<-function(X, pars, t){
+      
+      #mend model
+      deriv<-function(time, state, pars){
+        
+        with(as.list(c(state, pars)),{
+          
+          #equations
+          #C uptake
+          F1=1/CUE*(Vmax+mr)*Cmic*C/(Km+C)
+          #growth respiration
+          F4=(1/CUE-1)*Vmax*Cmic*C/(Km+C)
+          #maintnance respiration
+          F5=(1/CUE-1)*mr*Cmic*C/(Km+C)
+          #microbial mortality
+          F8=(1-pe)*mr*Cmic
+          #enzyme production
+          F9=pe*mr*Cmic
+          #enzyme decay
+          F10=ke*E
+          
+          #states
+          dCmic<-F1-(F4+F5)-F8-F9
[…]

### 2018-08-16 mmem_init_a.R

@@ -1,0 +1,939 @@
+mmem_init_a<-function(data, SUB, FACT, Mean, DNAci, Niter){
+  
+  #SUB - needs to be calibrated in matrix
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  #Vars = names of variables to calibrated the model against
+  #ColM = colnames for these varaibles in matrix
+  #ColV = colnames for these varaibles in normal setting
+  #VarsCmic = which biomass to use for likelihood calculation (it should follow the same notation as Vars)
+  #Cmic = initial microbial biomass C
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-setDT(data)[, id := .GRP, by = .(Substrate)]
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-setDT(data)[, id := .GRP, by = .(Structure)]
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-setDT(data)[, id := .GRP, by = .(Substrate,Structure)]
+      }else{
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  ############################################################################################
+  #if parameters are estimated across different substrates, different initial C conc
+  #has to be acknowledged
+  
+  if(SUB==TRUE){
+    
+    #this is the mmem model fitted across different substrates 
+    #with different initial carbon concentrations
+    mmem<-function(X, pars, t){
+      
+      #mmem model
+      deriv<-function(time, state, pars){
+        
+        with(as.list(c(state, pars)),{
+          
+          #Carbon uptake
+          
+          Cu=Vmax*C*Cmic/(Km+C)
+          
+          #defining stochiometric coefficients
+          #PO ratio
+          #PO=1.5*(1-((Cu/Cmic)/((Cu/Cmic)+Sover)))
+          y=2*PO*Yatp
+          
+          #x
+          #x=(Cu/Cmic)/((Cu/Cmic)+Sprod)
+          
+          #coefficients
+          psi.r=1/((y*x)+(y*w*(1-x))+1)
+          psi.g=y*x/((y*x)+(y*w*(1-x))+1)
+          psi.e=w*y*(1-x)/((y*x)+(y*w*(1-x))+1)
+          
+          
[…]

### 2018-08-16 mmem_init_b.R

@@ -1,0 +1,930 @@
+mmem_init_b<-function(data, SUB, FACT, Mean, DNAci, Niter){
+  
+  #SUB - needs to be calibrated in matrix
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  #Vars = names of variables to calibrated the model against
+  #ColM = colnames for these varaibles in matrix
+  #ColV = colnames for these varaibles in normal setting
+  #VarsCmic = which biomass to use for likelihood calculation (it should follow the same notation as Vars)
+  #Cmic = initial microbial biomass C
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-setDT(data)[, id := .GRP, by = .(Substrate)]
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-setDT(data)[, id := .GRP, by = .(Structure)]
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-setDT(data)[, id := .GRP, by = .(Substrate,Structure)]
+      }else{
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  ############################################################################################
+  #if parameters are estimated across different substrates, different initial C conc
+  #has to be acknowledged
+  
+  if(SUB==TRUE){
+    
+    #this is the mmem model fitted across different substrates 
+    #with different initial carbon concentrations
+    mmem<-function(X, pars, t){
+      
+      #mmem model
+      deriv<-function(time, state, pars){
+        
+        with(as.list(c(state, pars)),{
+          
+          #Carbon uptake
+          
+          Cu=Vmax*C*Cmic/(Km+C)
+          
+          #defining stochiometric coefficients
+          #PO ratio
+          #PO=1.5*(1-((Cu/Cmic)/((Cu/Cmic)+Sover)))
+          y=2*PO*Yatp
+          
+          #x
+          #x=(Cu/Cmic)/((Cu/Cmic)+Sprod)
+          
+          #coefficients
+          psi.r=1/((y*x)+(y*w*(1-x))+1)
+          psi.g=y*x/((y*x)+(y*w*(1-x))+1)
+          psi.e=w*y*(1-x)/((y*x)+(y*w*(1-x))+1)
+          
+          
[…]

### 2018-08-17 andrew_init_a.R
data specific functions for different parameters estimation

@@ -216,7 +216,7 @@ andrew_init_a<-function(data, SUB, FACT, Mean, DNAci, Niter){
     for(i in unique(dat$id)){
       for(n in 1:((ncol(parameters)-1)/3)){
         
-        parameters[i, n]<-summary(res[[i]])[1,n]
+        parameters[i, n]<-res[[i]]$bestpar[n]
       }
     }
     #lower quartile
@@ -303,8 +303,8 @@ andrew_init_a<-function(data, SUB, FACT, Mean, DNAci, Niter){
     
     for(i in unique(dat$id)){
       
-      obs_r<-append(obs_r, cost_r(pars=summary(res[[i]])[1,], data=dat[dat$id==i, ])$residuals$obs)
-      mod_r<-append(mod_r, cost_r(pars=summary(res[[i]])[1,], data=dat[dat$id==i, ])$residuals$mod)
+      obs_r<-append(obs_r, cost_r(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$obs)
+      mod_r<-append(mod_r, cost_r(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$mod)
       
     }
     
@@ -385,8 +385,8 @@ andrew_init_a<-function(data, SUB, FACT, Mean, DNAci, Niter){
     }
     
     for(i in unique(dat$id)){
-      obs_DNAc<-append(obs_DNAc, cost_DNAc(pars=summary(res[[i]])[1,], data=dat[dat$id==i, ])$residuals$obs)
-      mod_DNAc<-append(mod_DNAc, cost_DNAc(pars=summary(res[[i]])[1,], data=dat[dat$id==i, ])$residuals$mod)
+      obs_DNAc<-append(obs_DNAc, cost_DNAc(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$obs)
+      mod_DNAc<-append(mod_DNAc, cost_DNAc(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$mod)
     }
     
     OvP_DNAc<-data.frame(obs_DNAc, mod_DNAc)
@@ -543,15 +543,15 @@ andrew_init_a<-function(data, SUB, FACT, Mean, DNAci, Niter){
       mtrue<-select(mtrue, c("time", "E", "E1", "E2"))
       
       out<-mmem(X=c(25, 25, 16.5), pars = pars, t=seq(0,130))
[…]

### 2018-08-17 andrew_init_b.R

@@ -120,8 +120,8 @@ andrew_init_b<-function(data, SUB, FACT, Mean, DNAci, Niter){
       m3<-rbind(mix, mix2)
       m3<-m3[order(m3$time),]
       
-      m1[,c(6:9)]<-m2[,c(2:5)]
-      m1[,c(10:13)]<-m3[,c(2:5)]
+      m1[,c(5:7)]<-m2[,c(2:4)]
+      m1[,c(8:10)]<-m3[,c(2:4)]
       
       mtrue<-m1
       
@@ -214,7 +214,7 @@ andrew_init_b<-function(data, SUB, FACT, Mean, DNAci, Niter){
     for(i in unique(dat$id)){
       for(n in 1:((ncol(parameters)-1)/3)){
         
-        parameters[i, n]<-summary(res[[i]])[1,n]
+        parameters[i, n]<-res[[i]]$bestpar[n]
       }
     }
     #lower quartile
@@ -283,8 +283,8 @@ andrew_init_b<-function(data, SUB, FACT, Mean, DNAci, Niter){
       m3<-rbind(mix, mix2)
       m3<-m3[order(m3$time),]
       
-      m1[,c(6:9)]<-m2[,c(2:5)]
-      m1[,c(10:13)]<-m3[,c(2:5)]
+      m1[,c(5:7)]<-m2[,c(2:4)]
+      m1[,c(8:10)]<-m3[,c(2:4)]
       
       mtrue<-m1
       
@@ -299,8 +299,8 @@ andrew_init_b<-function(data, SUB, FACT, Mean, DNAci, Niter){
     
     for(i in unique(dat$id)){
       
-      obs_r<-append(obs_r, cost_r(pars=summary(res[[i]])[1,], data=dat[dat$id==i, ])$residuals$obs)
-      mod_r<-append(mod_r, cost_r(pars=summary(res[[i]])[1,], data=dat[dat$id==i, ])$residuals$mod)
+      obs_r<-append(obs_r, cost_r(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$obs)
+      mod_r<-append(mod_r, cost_r(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$mod)
       
     }
     
@@ -445,8 +445,8 @@ andrew_init_b<-function(data, SUB, FACT, Mean, DNAci, Niter){
       m3<-rbind(mix, mix2)
       m3<-m3[order(m3$time),]
       
[…]

### 2018-08-17 mem_function.R

@@ -176,7 +176,7 @@ mem_function<-function(data, SUB, FACT, Vars, ColM, Mean, Cmic, Niter){
     for(i in unique(dat$id)){
       for(n in 1:6){
         
-        parameters[i, n]<-summary(res[[i]])[6,n]
+        parameters[i, n]<-res[[i]]$bestpar[n]
       }
     }
     #lower quartile
@@ -256,8 +256,8 @@ mem_function<-function(data, SUB, FACT, Vars, ColM, Mean, Cmic, Niter){
     
     for(i in unique(dat$id)){
       
-      obs_r<-append(obs_r, cost_r(pars=summary(res[[i]])[6,], data=dat[dat$id==i, ])$residuals$obs)
-      mod_r<-append(mod_r, cost_r(pars=summary(res[[i]])[6,], data=dat[dat$id==i, ])$residuals$mod)
+      obs_r<-append(obs_r, cost_r(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$obs)
+      mod_r<-append(mod_r, cost_r(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$mod)
       
     }
     
@@ -330,8 +330,8 @@ mem_function<-function(data, SUB, FACT, Vars, ColM, Mean, Cmic, Niter){
     }
     
     for(i in unique(dat$id)){
-      obs_Cmic<-append(obs_Cmic, cost_Cmic(pars=summary(res[[i]])[6,], data=dat[dat$id==i, ])$residuals$obs)
-      mod_Cmic<-append(mod_Cmic, cost_Cmic(pars=summary(res[[i]])[6,], data=dat[dat$id==i, ])$residuals$mod)
+      obs_Cmic<-append(obs_Cmic, cost_Cmic(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$obs)
+      mod_Cmic<-append(mod_Cmic, cost_Cmic(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$mod)
     }
     
     OvP_Cmic<-data.frame(obs_Cmic, mod_Cmic)
@@ -396,19 +396,15 @@ mem_function<-function(data, SUB, FACT, Vars, ColM, Mean, Cmic, Niter){
       mtrue<-select(m1, c("time", "E", "E1", "E2"))
       
       out<-mem(X=c(25, 25, 16.5), pars = pars, t=seq(0,130))
[…]

### 2018-08-17 mend_init.R

@@ -210,7 +210,7 @@ mend_init<-function(data, SUB, FACT, Mean, Niter, DNAci){
     for(i in unique(dat$id)){
       for(n in 1:((ncol(parameters)-1)/3)){
         
-        parameters[i, n]<-summary(res[[i]])[1,n]
+        parameters[i, n]<-res[[i]]$bestpar[n]
       }
     }
     #lower quartile
@@ -297,8 +297,8 @@ mend_init<-function(data, SUB, FACT, Mean, Niter, DNAci){
     
     for(i in unique(dat$id)){
       
-      obs_r<-append(obs_r, cost_r(pars=summary(res[[i]])[1,], data=dat[dat$id==i, ])$residuals$obs)
-      mod_r<-append(mod_r, cost_r(pars=summary(res[[i]])[1,], data=dat[dat$id==i, ])$residuals$mod)
+      obs_r<-append(obs_r, cost_r(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$obs)
+      mod_r<-append(mod_r, cost_r(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$mod)
       
     }
     
@@ -379,8 +379,8 @@ mend_init<-function(data, SUB, FACT, Mean, Niter, DNAci){
     }
     
     for(i in unique(dat$id)){
-      obs_DNAc<-append(obs_DNAc, cost_DNAc(pars=summary(res[[i]])[1,], data=dat[dat$id==i, ])$residuals$obs)
-      mod_DNAc<-append(mod_DNAc, cost_DNAc(pars=summary(res[[i]])[1,], data=dat[dat$id==i, ])$residuals$mod)
+      obs_DNAc<-append(obs_DNAc, cost_DNAc(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$obs)
+      mod_DNAc<-append(mod_DNAc, cost_DNAc(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$mod)
     }
     
     OvP_DNAc<-data.frame(obs_DNAc, mod_DNAc)
@@ -461,8 +461,8 @@ mend_init<-function(data, SUB, FACT, Mean, Niter, DNAci){
     }
     
     for(i in unique(dat$id)){
[…]

### 2018-08-17 mend_init_a.R

@@ -210,7 +210,7 @@ mend_init_a<-function(data, SUB, FACT, Mean, Niter, DNAci){
     for(i in unique(dat$id)){
       for(n in 1:((ncol(parameters)-1)/3)){
         
-        parameters[i, n]<-summary(res[[i]])[1,n]
+        parameters[i, n]<-res[[i]]$bestpar[n]
       }
     }
     #lower quartile
@@ -297,8 +297,8 @@ mend_init_a<-function(data, SUB, FACT, Mean, Niter, DNAci){
     
     for(i in unique(dat$id)){
       
-      obs_r<-append(obs_r, cost_r(pars=summary(res[[i]])[1,], data=dat[dat$id==i, ])$residuals$obs)
-      mod_r<-append(mod_r, cost_r(pars=summary(res[[i]])[1,], data=dat[dat$id==i, ])$residuals$mod)
+      obs_r<-append(obs_r, cost_r(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$obs)
+      mod_r<-append(mod_r, cost_r(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$mod)
       
     }
     
@@ -379,8 +379,8 @@ mend_init_a<-function(data, SUB, FACT, Mean, Niter, DNAci){
     }
     
     for(i in unique(dat$id)){
-      obs_DNAc<-append(obs_DNAc, cost_DNAc(pars=summary(res[[i]])[1,], data=dat[dat$id==i, ])$residuals$obs)
-      mod_DNAc<-append(mod_DNAc, cost_DNAc(pars=summary(res[[i]])[1,], data=dat[dat$id==i, ])$residuals$mod)
+      obs_DNAc<-append(obs_DNAc, cost_DNAc(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$obs)
+      mod_DNAc<-append(mod_DNAc, cost_DNAc(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$mod)
     }
     
     OvP_DNAc<-data.frame(obs_DNAc, mod_DNAc)
@@ -537,15 +537,15 @@ mend_init_a<-function(data, SUB, FACT, Mean, Niter, DNAci){
       mtrue<-select(mtrue, c("time", "E", "E1", "E2"))
       
       out<-mend(X=c(25, 25, 16.5), pars = pars, t=seq(0,130))
[…]

### 2018-08-17 mmem_function.R

@@ -200,7 +200,7 @@ mmem_function<-function(data, SUB, FACT, Vars, ColM, Mean, Cmic, Niter){
     for(i in unique(dat$id)){
       for(n in 1:8){
         
-        parameters[i, n]<-summary(res[[i]])[6,n]
+        parameters[i, n]<-res[[i]]$bestpar[n]
       }
     }
     #lower quartile
@@ -280,8 +280,8 @@ mmem_function<-function(data, SUB, FACT, Vars, ColM, Mean, Cmic, Niter){
     
     for(i in unique(dat$id)){
       
-      obs_r<-append(obs_r, cost_r(pars=summary(res[[i]])[6,], data=dat[dat$id==i, ])$residuals$obs)
-      mod_r<-append(mod_r, cost_r(pars=summary(res[[i]])[6,], data=dat[dat$id==i, ])$residuals$mod)
+      obs_r<-append(obs_r, cost_r(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$obs)
+      mod_r<-append(mod_r, cost_r(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$mod)
       
     }
     
@@ -354,8 +354,8 @@ mmem_function<-function(data, SUB, FACT, Vars, ColM, Mean, Cmic, Niter){
     }
     
     for(i in unique(dat$id)){
-      obs_Cmic<-append(obs_Cmic, cost_Cmic(pars=summary(res[[i]])[6,], data=dat[dat$id==i, ])$residuals$obs)
-      mod_Cmic<-append(mod_Cmic, cost_Cmic(pars=summary(res[[i]])[6,], data=dat[dat$id==i, ])$residuals$mod)
+      obs_Cmic<-append(obs_Cmic, cost_Cmic(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$obs)
+      mod_Cmic<-append(mod_Cmic, cost_Cmic(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$mod)
     }
     
     OvP_Cmic<-data.frame(obs_Cmic, mod_Cmic)
@@ -420,15 +420,15 @@ mmem_function<-function(data, SUB, FACT, Vars, ColM, Mean, Cmic, Niter){
       mtrue<-select(m1, c("time", "E", "E1", "E2"))
       
[…]

### 2018-08-17 mmem_init.R

@@ -219,7 +219,7 @@ mmem_init<-function(data, SUB, FACT, Mean, DNAci, Niter){
     for(i in unique(dat$id)){
       for(n in 1:((ncol(parameters)-1)/3)){
         
-        parameters[i, n]<-summary(res[[i]])[1,n]
+        parameters[i, n]<-res[[i]]$bestpar[n]
       }
     }
     #lower quartile
@@ -306,8 +306,8 @@ mmem_init<-function(data, SUB, FACT, Mean, DNAci, Niter){
     
     for(i in unique(dat$id)){
       
-      obs_r<-append(obs_r, cost_r(pars=summary(res[[i]])[1,], data=dat[dat$id==i, ])$residuals$obs)
-      mod_r<-append(mod_r, cost_r(pars=summary(res[[i]])[1,], data=dat[dat$id==i, ])$residuals$mod)
+      obs_r<-append(obs_r, cost_r(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$obs)
+      mod_r<-append(mod_r, cost_r(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$mod)
       
     }
     
@@ -388,8 +388,8 @@ mmem_init<-function(data, SUB, FACT, Mean, DNAci, Niter){
     }
     
     for(i in unique(dat$id)){
-      obs_DNAc<-append(obs_DNAc, cost_DNAc(pars=summary(res[[i]])[1,], data=dat[dat$id==i, ])$residuals$obs)
-      mod_DNAc<-append(mod_DNAc, cost_DNAc(pars=summary(res[[i]])[1,], data=dat[dat$id==i, ])$residuals$mod)
+      obs_DNAc<-append(obs_DNAc, cost_DNAc(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$obs)
+      mod_DNAc<-append(mod_DNAc, cost_DNAc(pars=res[[i]]$bestpar, data=dat[dat$id==i, ])$residuals$mod)
     }
     
     OvP_DNAc<-data.frame(obs_DNAc, mod_DNAc)
@@ -470,8 +470,8 @@ mmem_init<-function(data, SUB, FACT, Mean, DNAci, Niter){
     }
     
     for(i in unique(dat$id)){
[…]

### 2018-08-17 mmem_init_b.R

@@ -120,7 +120,7 @@ mmem_init_b<-function(data, SUB, FACT, Mean, DNAci, Niter){
       
       ###Mix
       mix<-select(filter(data, Substrate=="Mix" ), c("Time", "r", "E" , "Protc"))
-      colnames(mix)<-c("time", "r2", "E2" ,"DNAc2", "Protc2")
+      colnames(mix)<-c("time", "r2", "E2" ,"Protc2")
       mix2<-select(filter(data, Substrate!="Mix" ), "Time")
       colnames(mix2)<-"time"
       mix2$r2<-NA
@@ -217,7 +217,7 @@ mmem_init_b<-function(data, SUB, FACT, Mean, DNAci, Niter){
     for(i in unique(dat$id)){
       for(n in 1:((ncol(parameters)-1)/3)){
         
-        parameters[i, n]<-summary(res[[i]])[1,n]
+        parameters[i, n]<-res[[i]]$bestpar[n]
       }
     }
     #lower quartile
@@ -276,7 +276,7 @@ mmem_init_b<-function(data, SUB, FACT, Mean, DNAci, Niter){
       
       ###Mix
       mix<-select(filter(data, Substrate=="Mix" ), c("Time", "r", "E" , "Protc"))
-      colnames(mix)<-c("time", "r2", "E2" ,"DNAc2", "Protc2")
+      colnames(mix)<-c("time", "r2", "E2" ,"Protc2")
       mix2<-select(filter(data, Substrate!="Mix" ), "Time")
       colnames(mix2)<-"time"
       mix2$r2<-NA
@@ -290,6 +290,7 @@ mmem_init_b<-function(data, SUB, FACT, Mean, DNAci, Niter){
       m1[,c(8:10)]<-m3[,c(2:4)]
       
       mtrue<-m1
+      
       mtrue<-select(mtrue, c("time", "r", "r1", "r2"))
       
       out<-mmem(X=c(25, 25, 16.5), pars = pars, t=seq(0,130))
@@ -301,8 +302,8 @@ mmem_init_b<-function(data, SUB, FACT, Mean, DNAci, Niter){
     
     for(i in unique(dat$id)){
       
-      obs_r<-append(obs_r, cost_r(pars=summary(res[[i]])[1,], data=dat[dat$id==i, ])$residuals$obs)
[…]

### 2018-11-27 deb.R
old scripts removed from master file

@@ -1,255 +1,0 @@
-deb<-function(data, FACT){
-
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-
-
-  #if factor needs to be included, its levels are defined
-  if(FACT==1){
-
-    dat<-data
-    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
-
-  }else{
-
-    if(FACT==2){
-
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-
-    }else{
-
-      if(FACT==3){
-
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-
-
-      }else{
-
-        dat<-data
-        dat$id<-c(1)
-
-      }
-    }
-  }
-
-    #deb model
-    deriv<-function(time, state, pars){
-
-      with(as.list(c(state, pars)),{
-
-        #Carbon uptake
-        Cu=Vmax*C*S/(Km+C)
-
-        #maintnance
-        m=S*m0
-
-        #reserve mobilization rate for growth and enzyme production
-        an=f*R-m
-
-        #calibrated variables
-        #protein abundance in reserves and structures is mean value reported
-        #by Hanegraaf and Muller, 2001
-        r=m+pmax(an*(1-Yu), 0)
-
-        Protc=0.7095*S+0.6085*R
-        #Protc=fps*S+fpr*R
-
-        dR<-Cu-f*R
-        dS<-pmax(an*Yu,0)+pmin(0, an)
-        dC<--Cu
-
-
-        return(list(c(dR, dS, dC), r=r, Protc=Protc))
-
-      })
-    }
-    #define names of parameters
-    parnames<-c("Vmax", "Km", "m0", "f", "Yu", "R_0", "S_0")
-
-    #parameters estimation function
-    estim<-function(odeset){
-
-      #defining cost function
-      cost<-function(x){
-
-        par<-x[1:length(parnames)]
-
-        names(par)<-parnames
-
-        #first, pars dependent output from ode is matched with measured values
[…]

### 2018-11-27 deb_i.R

@@ -1,255 +1,0 @@
-deb_i<-function(data, FACT){
-  
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-  
-  
-  #if factor needs to be included, its levels are defined
-  if(FACT==1){
-    
-    dat<-data
-    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
-    
-  }else{
-    
-    if(FACT==2){
-      
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-      
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-      
-    }else{
-      
-      if(FACT==3){
-        
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-        
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-        
-        
-      }else{
-        
-        dat<-data
-        dat$id<-c(1)
-        
-      }
-    }
-  }
-  
-    #deb model
-    deriv<-function(time, state, pars){
-      
-      with(as.list(c(state, pars)),{
-        
-        #Carbon uptake
-        Cu=Vmax*C*S/(Km+C)
-        
-        #maintnance
-        m=S*m0
-        
-        #reserve mobilization rate for growth and enzyme production
-        an=f*R-m
-        
-        #calibrated variables
-        #protein abundance in reserves and structures is mean value reported 
-        #by Hanegraaf and Muller, 2001
-        r=m+pmax(an*(1-Yu), 0)
-        
-        #Protc=0.7095*S+0.6085*R
-        Protc=fpr*R+fps*S
-        
-        dR<-Cu-f*R
-        dS<-pmax(an*Yu,0)+pmin(0, an)
-        dC<--Cu
-        
-        
-        return(list(c(dR, dS, dC), r=r, Protc=Protc))
-        
-      })
-    }
-    #define names of parameters
-    parnames<-c("Vmax", "Km", "m0", "f", "Yu", "fpr", "fps", "R_0", "S_0")
-    
-    #parameters estimation function
-    estim<-function(odeset){
-      
-      #defining cost function
-      cost<-function(x){
-        
-        par<-x[1:length(parnames)]
-        
-        names(par)<-parnames
-        
[…]

### 2018-11-27 mend.R

@@ -1,245 +1,0 @@
-mend<-function(data, FACT){
-  
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-  
-  if(FACT==1){
-    
-    dat<-data
-    dat$id<-ifelse(d$Substrate=="Glucose", 1, 2)
-    
-  }else{
-    
-    if(FACT==2){
-      
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-      
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-      
-    }else{
-      
-      if(FACT==3){
-        
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-        
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-        
-        
-      }else{
-        
-        dat<-data
-        dat$id<-c(1)
-        
-      }
-    }
-  }
-  
-    #mend model
-    deriv<-function(time, state, pars){
-      
-      with(as.list(c(state, pars)),{
-        
-        #equations
-        #C uptake
-        F1=1/CUE*(Vmax+mr)*Cmic*C/(Km+C)
-        #growth respiration
-        F4=(1/CUE-1)*Vmax*Cmic*C/(Km+C)
-        #maintnance respiration
-        F5=(1/CUE-1)*mr*Cmic*C/(Km+C)
-        #microbial mortality
-        F8=mr*Cmic
-        
-        #states
-        dCmic<-F1-(F4+F5)-F8
-        dC<--F1+F8
-        
-        return(list(c(dCmic, dC), r=F4+F5, Protc=0.548*Cmic))
-        
-      })
-    }
-    #define names of parameters
-    parnames<-c("Vmax", "Km", "CUE", "mr", "Cmic_0")
-    
-    #parameters estimation function
-    estim<-function(odeset){
-      
-      #defining cost function
-      cost<-function(x){
-        
-        par<-x[1:length(parnames)]
-        
-        names(par)<-parnames
-        
-        #first, pars dependent output from ode is matched with measured values
-        yhat_all<-as.data.frame(ode(y=c(Cmic=par[["Cmic_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))
-        
[…]

### 2018-11-27 monod.R

@@ -1,238 +1,0 @@
-###monod growth function
-monod<-function(data, FACT){
-  
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-  
-  if(FACT==1){
-    
-    dat<-data
-    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
-    
-  }else{
-    
-    if(FACT==2){
-      
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-      
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-      
-    }else{
-      
-      if(FACT==3){
-        
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-        
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-        
-        
-      }else{
-        
-        dat<-data
-        dat$id<-c(1)
-        
-      }
-    }
-  }
-  
-  
-    #monod growth function
-    deriv<-function(time, state, pars){
-      
-      with(as.list(c(state, pars)),{
-        
-        dCmic<--k*Cmic+CUE*Vmax*Cmic*C/(Km+C)
-        dC<-k*Cmic-Vmax*Cmic*C/(Km+C)
-        
-        return(list(c(dCmic, dC), r=(1-CUE)*Vmax*Cmic*C/(Km+C), Protc=0.548*Cmic))
-        
-      })
-    }
-    
-    #define names of parameters
-    parnames<-c("Vmax", "Km", "CUE", "k", "Cmic_0")
-    
-    #parameters estimation function
-    estim<-function(odeset){
-      
-      #defining cost function
-      cost<-function(x){
-        
-        par<-x[1:length(parnames)]
-        
-        names(par)<-parnames
-        
-        #first, pars dependent output from ode is matched with measured values
-        yhat_all<-as.data.frame(ode(y=c(Cmic=par[["Cmic_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))
-        
-        #select time and the measured variables 
-        yhat<-select(yhat_all, c("time", "r", "Protc"))
-        
-        #reformat to long format data frame
-        Yhat<-melt(yhat, id.vars = "time")
-        
-        #add the measured data to a data frame
[…]

### 2018-11-27 SI_code.Rmd
uploading SI

---
title: Variable Composition of Microbial Biomass Affects the Accuracy of Microbial
  Explicit Models Simulation
author: "Petr Capek, Katherine Todd-Brown, Natalie Sadler, Nancy Hess, Kirsten Hofmockel"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Supplementary Information
## Step-by-step analysis and code

The analysis requires several libraries

```{r libraries loading, message=FALSE, warning=FALSE}

##libraries
library(deSolve)
library(dplyr)
library(FME)
library(reshape)
library(ggplot2)
library(foreach)
library(doParallel)
library(openxlsx)
library(DEoptim)
library(gridExtra)
library(kableExtra)
```

Data are available at https://github.com/petacapek/MinT.

```{r data loading, echo=FALSE}

#respiration rates in umol/ml/h
resp<-read.csv(file=c("data_checked_respiration_raw.csv"))

#arrange the data
resp.ordered<-resp[order(resp$Sample, resp$Day), ]

#biological data
biology<-read.csv(file=c("data_checked_biology_raw.csv"))

#arrange the data in the same way
biology.ordered<-biology[order(biology$Sample, biology$Day), -c(2,3)]
biology.ordered$Sample<-paste0("CSub3_", biology.ordered$Sample)

#merging both data sets
m0<-merge(resp.ordered, biology.ordered, by.x=c("Sample", "Day"), by.y = c("Sample", "Day"), all=T)

```

The conducted experiment is large. For this analysis, only subset of data is used. We use treatments with Glucose or Celluloze as a single source of organic carbon and all three levels of incubation system complexity (i.e. broth, glass wool and mix of glass beads).

```{r data filtering}

d<-subset(m0, Substrate=="Glucose" | Substrate=="Cellobiose")
summary(d)

#removing outliers
#respiration rate
ggplot(d, aes(Time, r))+geom_point(cex=6)+facet_wrap(Structure~Substrate, scales="free")

d[(d$Substrate=="Glucose" & d$Structure=="Glass wool" & d$Time>20 & d$Time<30 & d$r<0.4 & !is.na(d$r)), "r"]<-NA

#Proteins
ggplot(d, aes(Time, Prot.in))+geom_point(cex=6)+facet_wrap(Structure~Substrate, scales="free")

d[(d$Substrate=="Cellobiose" & d$Structure=="Broth" &  d$Time>75), "Prot.in"]<-NA
d[(d$Substrate=="Glucose" & d$Structure=="Broth" & d$Time>75), "Prot.in"]<-NA
d[(d$Substrate=="Cellobiose" & d$Structure=="Mixed glass" & d$Time>75), "Prot.in"]<-NA
d[(d$Substrate=="Cellobiose" & d$Structure=="Glass wool" & d$Time>75 & d$Prot.in>110 & !is.na(d$Prot.in)), "Prot.in"]<-NA

```

DNA and protein concentrations are in $mu$grams per incubation vial. Both values are recalculated to unit of carbon per one ml of media in the vial (4 ml) using the conversion factors reported in Vrede et al. (2004).

```{r conversions}

d$Protc<-d$Prot.in*0.46/12.01/4
d$DNAc<-d$DNA*0.51/12.01/4

```

###All models are first calibrated against respiration rate
Calibration of each model is done in four steps. First, the model is calibrated against data across all treatments. Second, the model is calibrated for each level of incubation system complexity separately. Third, the model is calibrated for Glucose and Cellobiose treatments separately. Fourth, the model is calibrated for each experimental treatment separately. 
Goodness of correspondence between measurements and model simulations are calculated across all treatments. By doing so, the effect of experimental treatment is evaluated. If the model simulation fits the data significantly better when model parameters are estimated for different treatments separately, effect of this treatment is significant.

The calibration of the models is not run within this document because it takes couple of hours to finish. Results of the previous iteration is printed instead. 
To run the code and reproduce the results, the option "eval" need to be changed to TRUE. Please note that the parallel computing is used. Number of cores is set to total number of cores of the computer minus 1. This can be changed if necessary.

####Monod model
```{r monod model respiration, eval=F, echo=T}
monod_r<-function(data, FACT){

  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No

  if(FACT==1){

    dat<-data
    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)

  }else{

    if(FACT==2){

      dat<-data
      dat<-data
      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
      ids$id<-seq(1:nrow(ids))

      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))

    }else{

      if(FACT==3){

        dat<-data
        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
        ids$id<-seq(1:nrow(ids))

        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))

      }else{

        dat<-data
        dat$id<-c(1)

      }
    }
  }

    #monod growth function
    deriv<-function(time, state, pars){

      with(as.list(c(state, pars)),{

        dCmic<--k*Cmic+CUE*Vmax*Cmic*C/(Km+C)
        dC<-k*Cmic-Vmax*Cmic*C/(Km+C)

        return(list(c(dCmic, dC), r=(1-CUE)*Vmax*Cmic*C/(Km+C), Protc=0.548*Cmic))

      })
    }

    #define names of parameters
    parnames<-c("Vmax", "Km", "CUE", "k", "Cmic_0")

    #parameters estimation function
    estim<-function(odeset){

      #defining cost function
      cost<-function(x){

        par<-x[1:length(parnames)]

        names(par)<-parnames

        #first, pars dependent output from ode is matched with measured values
        yhat_all<-as.data.frame(ode(y=c(Cmic=par[["Cmic_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))

        #select time and the measured variables 
        yhat<-select(yhat_all, c("time", "r", "Protc"))

        #reformat to long format data frame
        Yhat<-melt(yhat, id.vars = "time")

        #add the measured data to a data frame
        Yhat$obs<-c(odeset[order(odeset$Time), c("r")], odeset[order(odeset$Time), "Protc"])

        #add the weighting factor
        #I want to have the weighting factor to be proportional to mean of the given variable 
        yweights<-Yhat %>% group_by(variable) %>% summarise(weights=mean(value, na.rm = T))

        #match with the Yhat data frame
        Yhat<-merge(Yhat, yweights, by.x=c("variable"), by.y=c("variable"))

        Yhat<-Yhat[Yhat$variable!="Protc", ]

        #now, the root mean square error is calculated
        NRMSE<-as.numeric(Yhat %>% summarise(NRMSE=sum((((value-obs)/mean(weights))^2), na.rm = T)))

        return(NRMSE)

      }

      #defining goodness of fit function 
      rsq_ode<-function(x){

        par<-x[1:length(parnames)]

        names(par)<-parnames

        #first, pars dependent output from ode is matched with measured values
        yhat_all<-as.data.frame(ode(y=c(Cmic=par[["Cmic_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))

        #select time and the measured variables 
        yhat<-select(yhat_all, c("time", "r", "Protc"))

        #reformat to long format data frame
        Yhat<-melt(yhat, id.vars = "time")

        #add the measured data to a data frame
        Yhat$obs<-c(odeset[order(odeset$Time), c("r")], odeset[order(odeset$Time), "Protc"])
        Yhat$Substrate<-rep(odeset[order(odeset$Time), "Substrate"], times=2)
        Yhat$Structure<-rep(odeset[order(odeset$Time), "Structure"], times=2)

        #rsquared calculation for each variable
        Gfit<-Yhat %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
                                                        SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                        ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        Gfit$R2<-with(Gfit, 1-SSres/SStot)
        Gfit$N<-c(5)
        Gfit$AIC<-with(Gfit, 2*N-2*ll)

        rsq_out<-list(Yhat=Yhat, Gfit=Gfit)

        return(rsq_out)

      }

      #approximate parameter estimation is done by MCMC method
      par_mcmc<-modMCMC(f=cost, p=c(Vmax=0.1, Km=3, CUE=0.5, k=1e-3, Cmic_0=0.01), 
                        lower=c(Vmax=1e-6, Km=1e-3, CUE=0, k=1e-6, Cmic_0=0.0001),
                        upper=c(Vmax=10, Km=100, CUE=1, k=10, Cmic_0=10), niter=10000)

      #lower and upper limits for parameters are extracted
      pl<-summary(par_mcmc)["min",]
      pu<-summary(par_mcmc)["max",]

      #these limits are used to find global optimum by DEoptim
      opt_par<-DEoptim(fn=cost, lower=pl, upper=pu, 
                       control = c(itermax = 10000, steptol = 50, reltol = 1e-8, 
                                   trace=FALSE, strategy=3, NP=250))

      #global optimum parameters are further used in MCMC to find parameters distribution
      par_prof<-modMCMC(f=cost, p=opt_par$optim$bestmem, 
                        lower=pl,
                        upper=pu, niter=5000)

      #goodness of fit
      fit<-rsq_ode(opt_par$optim$bestmem)

      #best parameters
      p<-opt_par$optim$bestmem
      names(p)<-parnames

      #return list with opt_par and par_prof
      estim_out<-list(pars=p, par_prof=par_prof, fit=fit)

      return(estim_out)

    }

    #parameter estimation
    if(FACT==4){

      res<-vector("list", length = 1)
      res[[1]]<-estim(odeset=dat)

    }else{

      res<-foreach(i=unique(dat$id), .combine=list, .multicombine = TRUE,
                   .packages=c("FME", "dplyr", "DEoptim", "reshape")) %dopar% {

                     estim(odeset=dat[dat$id==i,])

                   }
    }

    #calculation of overall goodness of fit from individual results
    if(FACT==4){

      res$goodness<-res[[1]]$fit$Gfit
      res$OvP<-res[[1]]$fit$Yhat

    }else{

      if(FACT==1){

        res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat)

        #rsquared calculation for each variable
        res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
                                                                   SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                                   ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
        res$goodness$N<-rep(length(parnames)*2, times=2)
        res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

      }else{

        if(FACT==2){

          res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat, res[[3]]$fit$Yhat)

          #rsquared calculation for each variable
          res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
                                                                     SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                                     ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
          res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
          res$goodness$N<-rep(length(parnames)*3, times=2)
          res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

        }else{

          res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat, res[[3]]$fit$Yhat, res[[4]]$fit$Yhat, res[[5]]$fit$Yhat, res[[6]]$fit$Yhat)

          #rsquared calculation for each variable
          res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
                                                                     SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                                     ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
          res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
          res$goodness$N<-rep(length(parnames)*6, times=2)
          res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

        }

      }

    }

    return(res)

}

```

```{r monod respiration calibration, eval=F, echo=T}
#across all treatments
monod_r1<-monod_r(data=d, FACT = 4)

no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)

#for different structures
monod_r2<-monod_r(data=d, FACT = 2)

#for different substrates
monod_r3<-monod_r(data=d, FACT = 1)

#for each separately 
monod_r4<-monod_r(data=d, FACT = 3)

stopImplicitCluster()

monod_r1$goodness
monod_r2$goodness
monod_r3$goodness
monod_r4$goodness
```

```{r monod respiration calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

#results of the previous iteration (not run)
# monod_r_results_par<-rbind(monod_r1[[1]]$pars,
#                        monod_r2[[1]]$pars, monod_r2[[2]]$pars, monod_r2[[3]]$pars,
#                        monod_r3[[1]]$pars, monod_r3[[2]]$pars,
#                        monod_r4[[1]]$pars, monod_r4[[2]]$pars, monod_r4[[3]]$pars,
#                        monod_r4[[4]]$pars, monod_r4[[5]]$pars, monod_r4[[6]]$pars)
# monod_r_results_good<-rbind(monod_r1$goodness[1,c(4:7)],
#                        monod_r2$goodness[1,c(4:7)],
#                        monod_r3$goodness[1,c(4:7)],
#                        monod_r4$goodness[1,c(4:7)])
# write.csv(monod_r_results_par, file = c("../monod_r_results_par.csv"), row.names = F)
# write.csv(monod_r_results_good, file = c("../monod_r_results_good.csv"), row.names = F)

monod_r_results_par<-read.csv("monod_r_results_par.csv")
monod_r_results_good<-read.csv("monod_r_results_good.csv")

t.s1.1<-data.frame(Vmax=monod_r_results_par[,1],
                 Km=monod_r_results_par[,2],
                 CUE=monod_r_results_par[,3],
                 k=monod_r_results_par[,4])
row.names(t.s1.1)<-c("All", "Broth", "Glass", "Wool",
                    "Celluloze", "Glucose",
                    "Broth - Celluloze", "Broth - Glucose",
                    "Glass - Celluloze", "Glass - Glucose",
                    "Wool - Celluloze", "Wool - Glucose")

options(knitr.kable.NA = '')

kable(t.s1.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$'), row.names=T, caption = "_**Table S1:** Monod model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monod_r_results_good, digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S2:** Goodness of fit of monod model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 2) %>%
  group_rows("Substrate", 3, 4) %>%
  group_rows("Each", 4, 4)

```

####MEND model
```{r MEND model respiration, eval=F, echo=T}
mend_r<-function(data, FACT){

  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No

  if(FACT==1){

    dat<-data
    dat$id<-ifelse(d$Substrate=="Glucose", 1, 2)

  }else{

    if(FACT==2){

      dat<-data
      dat<-data
      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
      ids$id<-seq(1:nrow(ids))

      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))

    }else{

      if(FACT==3){

        dat<-data
        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
        ids$id<-seq(1:nrow(ids))

        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))

      }else{

        dat<-data
        dat$id<-c(1)

      }
    }
  }

    #mend model
    deriv<-function(time, state, pars){

      with(as.list(c(state, pars)),{

        #equations
        #C uptake
        F1=1/CUE*(Vmax+mr)*Cmic*C/(Km+C)
        #growth respiration
        F4=(1/CUE-1)*Vmax*Cmic*C/(Km+C)
        #maintnance respiration
        F5=(1/CUE-1)*mr*Cmic*C/(Km+C)
        #microbial mortality
        F8=mr*Cmic

        #states
        dCmic<-F1-(F4+F5)-F8
        dC<--F1+F8

        return(list(c(dCmic, dC), r=F4+F5, Protc=0.548*Cmic))

      })
    }
    #define names of parameters
    parnames<-c("Vmax", "Km", "CUE", "mr", "Cmic_0")

    #parameters estimation function
    estim<-function(odeset){

      #defining cost function
      cost<-function(x){

        par<-x[1:length(parnames)]

        names(par)<-parnames

        #first, pars dependent output from ode is matched with measured values
        yhat_all<-as.data.frame(ode(y=c(Cmic=par[["Cmic_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))

        #select time and the measured variables 
        yhat<-select(yhat_all, c("time", "Protc", "r"))

        #reformat to long format data frame
        Yhat<-melt(yhat, id.vars = "time")

        #add the measured data to a data frame
        Yhat$obs<-c(odeset[order(odeset$Time), "Protc"], odeset[order(odeset$Time), c("r")])

        #add the weighting factor
        #I want to have the weighting factor to be proportional to mean of the given variable 
        yweights<-Yhat %>% group_by(variable) %>% summarise(weights=mean(value, na.rm = T))

        #match with the Yhat data frame
        Yhat<-merge(Yhat, yweights, by.x=c("variable"), by.y=c("variable"))

        Yhat<-Yhat[Yhat$variable!="Protc", ]

        #now, the root mean square error is calculated
        NRMSE<-as.numeric(Yhat %>% summarise(NRMSE=sum((((value-obs)/mean(weights))^2), na.rm = T)))

        return(NRMSE)

      }

      #defining goodness of fit function 
      rsq_ode<-function(x){

        par<-x[1:length(parnames)]

        names(par)<-parnames

        #first, pars dependent output from ode is matched with measured values
        yhat_all<-as.data.frame(ode(y=c(Cmic=par[["Cmic_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))

        #select time and the measured variables 
        yhat<-select(yhat_all, c("time", "Protc", "r"))

        #reformat to long format data frame
        Yhat<-melt(yhat, id.vars = "time")

        #add the measured data to a data frame
        Yhat$obs<-c(odeset[order(odeset$Time), "Protc"], odeset[order(odeset$Time), c("r")])
        Yhat$Substrate<-rep(odeset[order(odeset$Time), "Substrate"], times=2)
        Yhat$Structure<-rep(odeset[order(odeset$Time), "Structure"], times=2)

        #rsquared calculation for each variable
        Gfit<-Yhat %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
                                                        SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                        ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        Gfit$R2<-with(Gfit, 1-SSres/SStot)
        Gfit$N<-c(5)
        Gfit$AIC<-with(Gfit, 2*N-2*ll)

        rsq_out<-list(Yhat=Yhat, Gfit=Gfit)

        return(rsq_out)

      }

      #approximate parameter estimation is done by MCMC method
      par_mcmc<-modMCMC(f=cost, p=c(Vmax=0.1, Km=3, CUE=0.5, mr=0.01, Cmic_0=0.01), 
                          lower=c(Vmax=1e-3, Km=1e-3, CUE=0, mr=1e-5,  Cmic_0=1e-5),
                          upper=c(Vmax=10, Km=100, CUE=1, mr=10, Cmic_0=5), niter=10000)

      #lower and upper limits for parameters are extracted
      pl<-summary(par_mcmc)["min",]
      pu<-summary(par_mcmc)["max",]

      #these limits are used to find global optimum by DEoptim
      opt_par<-DEoptim(fn=cost, lower=pl, upper=pu, 
                       control = c(itermax = 10000, steptol = 50, reltol = 1e-8, 
                                   trace=FALSE, strategy=3, NP=250))

      #global optimum parameters are further used in MCMC to find parameters distribution
      par_prof<-modMCMC(f=cost, p=opt_par$optim$bestmem, 
                        lower=pl,
                        upper=pu, niter=5000)

      #goodness of fit
      fit<-rsq_ode(opt_par$optim$bestmem)

      #best parameters
      p<-opt_par$optim$bestmem
      names(p)<-parnames

      #return list with opt_par and par_prof
      estim_out<-list(pars=p, par_prof=par_prof, fit=fit)

      return(estim_out)

    }

    #parameter estimation
    if(FACT==4){

      res<-vector("list", length = 1)
      res[[1]]<-estim(odeset=dat)

    }else{

      res<-foreach(i=unique(dat$id), .combine=list, .multicombine = TRUE,
                   .packages=c("FME", "dplyr", "DEoptim", "reshape")) %dopar% {

                     estim(odeset=dat[dat$id==i,])

                   }
    }

    #calculation of overall goodness of fit from individual results
    if(FACT==4){

      res$goodness<-res[[1]]$fit$Gfit
      res$OvP<-res[[1]]$fit$Yhat

    }else{

      if(FACT==1){

        res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat)

        #rsquared calculation for each variable
        res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
                                                               SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                               ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
        res$goodness$N<-rep(length(parnames)*2, times=2)
        res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

      }else{

        if(FACT==2){

          res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat, res[[3]]$fit$Yhat)

          #rsquared calculation for each variable
          res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
                                                                 SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                                 ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
          res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
          res$goodness$N<-rep(length(parnames)*3, times=2)
          res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

        }else{

          res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat, res[[3]]$fit$Yhat, res[[4]]$fit$Yhat, res[[5]]$fit$Yhat, res[[6]]$fit$Yhat)

          #rsquared calculation for each variable
          res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
                                                                 SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                                 ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
          res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
          res$goodness$N<-rep(length(parnames)*6, times=2)
          res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

        }

      }

    }

  return(res)

}

```

```{r MEND respiration calibration, eval=F, echo=T}
#across all treatments
mend_r1<-mend_r(data=d, FACT = 4)

no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)

#for different structures
mend_r2<-mend_r(data=d, FACT = 2)

#for different substrates
mend_r3<-mend_r(data=d, FACT = 1)

#for each separately 
mend_r4<-mend_r(data=d, FACT = 3)

stopImplicitCluster()

mend_r1$goodness
mend_r2$goodness
mend_r3$goodness
mend_r4$goodness

```

```{r MEND respiration calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

#results of the previous iteration (not run)
# mend_r_results_par<-rbind(mend_r1[[1]]$pars,
#                       mend_r2[[1]]$pars, mend_r2[[2]]$pars, mend_r2[[3]]$pars,
#                       mend_r3[[1]]$pars, mend_r3[[2]]$pars,
#                       mend_r4[[1]]$pars, mend_r4[[2]]$pars, mend_r4[[3]]$pars,
#                       mend_r4[[4]]$pars, mend_r4[[5]]$pars, mend_r4[[6]]$pars)
# mend_r_results_good<-rbind(mend_r1$goodness[2,c(4:7)],
#                      mend_r2$goodness[2,c(4:7)],
#                      mend_r3$goodness[2,c(4:7)],
#                      mend_r4$goodness[2,c(4:7)])
# write.csv(mend_r_results_par, file = c("../mend_r_results_par.csv"), row.names = F)
# write.csv(mend_r_results_good, file = c("../mend_r_results_good.csv"), row.names = F)

mend_r_results_par<-read.csv("mend_r_results_par.csv")
mend_r_results_good<-read.csv("mend_r_results_good.csv")

t.s2.1<-data.frame(Vmax=mend_r_results_par[,1],
                 Km=mend_r_results_par[,2],
                 CUE=mend_r_results_par[,3],
                 mr=mend_r_results_par[,4])
row.names(t.s2.1)<-c("All", "Broth", "Glass", "Wool",
                    "Celluloze", "Glucose",
                    "Broth - Celluloze", "Broth - Glucose",
                    "Glass - Celluloze", "Glass - Glucose",
                    "Wool - Celluloze", "Wool - Glucose")

options(knitr.kable.NA = '')

kable(t.s2.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$'), row.names=T, caption = "_**Table S3:** MEND model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(mend_r_results_good, digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S4:** Goodness of fit of MEND model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 2) %>%
  group_rows("Substrate", 3, 4) %>%
  group_rows("Each", 4, 4)

```

####DEB model
```{r DEB model respiration, eval=F, echo=T}
deb_r<-function(data, FACT){

  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No

  #if factor needs to be included, its levels are defined
  if(FACT==1){

    dat<-data
    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)

  }else{

    if(FACT==2){

      dat<-data
      dat<-data
      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
      ids$id<-seq(1:nrow(ids))

      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))

    }else{

      if(FACT==3){

        dat<-data
        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
        ids$id<-seq(1:nrow(ids))

        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))

      }else{

        dat<-data
        dat$id<-c(1)

      }
    }
  }

    #deb model
    deriv<-function(time, state, pars){

      with(as.list(c(state, pars)),{

        #Carbon uptake
        Cu=Vmax*C*S/(Km+C)

        #maintnance
        m=S*m0

        #reserve mobilization rate for growth
        an=f*R-m

        #calibrated variables
        #protein abundance in reserves and structures is mean value reported
        #by Hanegraaf and Muller, 2001
        r=m+pmax(an*(1-Yu), 0)

        Protc=0.7095*S+0.6085*R
        #Protc=fps*S+fpr*R

        dR<-Cu-f*R
        dS<-pmax(an*Yu,0)+pmin(0, an)
        dC<--Cu

        return(list(c(dR, dS, dC), r=r, Protc=Protc))

      })
    }
    #define names of parameters
    parnames<-c("Vmax", "Km", "m0", "f", "Yu", "R_0", "S_0")

    #parameters estimation function
    estim<-function(odeset){

      #defining cost function
      cost<-function(x){

        par<-x[1:length(parnames)]

        names(par)<-parnames

        #first, pars dependent output from ode is matched with measured values
        yhat_all<-as.data.frame(ode(y=c(R=par[["R_0"]], S=par[["S_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))

        #select time and the measured variables
        yhat<-select(yhat_all, c("time", "r", "Protc"))

        #reformat to long format data frame
        Yhat<-melt(yhat, id.vars = "time")

        #add the measured data to a data frame
        Yhat$obs<-c(odeset[order(odeset$Time), c("r")], odeset[order(odeset$Time), "Protc"])

        #add the weighting factor
        #I want to have the weighting factor to be proportional to mean of the given variable
        yweights<-Yhat %>% group_by(variable) %>% summarise(weights=mean(value, na.rm = T))

        #match with the Yhat data frame
        Yhat<-merge(Yhat, yweights, by.x=c("variable"), by.y=c("variable"))

        Yhat<-Yhat[Yhat$variable!="Protc", ]

        #now, the root mean square error is calculated
        NRMSE<-as.numeric(Yhat %>% summarise(NRMSE=sum((((value-obs)/mean(weights))^2), na.rm = T)))

        return(NRMSE)

      }

      #defining goodness of fit function
      rsq_ode<-function(x){

        par<-x[1:length(parnames)]

        names(par)<-parnames

        #first, pars dependent output from ode is matched with measured values
        yhat_all<-as.data.frame(ode(y=c(R=par[["R_0"]], S=par[["S_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))

        #select time and the measured variables
        yhat<-select(yhat_all, c("time", "r", "Protc"))

        #reformat to long format data frame
        Yhat<-melt(yhat, id.vars = "time")

        #add the measured data to a data frame
        Yhat$obs<-c(odeset[order(odeset$Time), c("r")], odeset[order(odeset$Time), "Protc"])
        Yhat$Substrate<-rep(odeset[order(odeset$Time), "Substrate"], times=2)
        Yhat$Structure<-rep(odeset[order(odeset$Time), "Structure"], times=2)

        #rsquared calculation for each variable
        Gfit<-Yhat %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T),
                                                        SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                        ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        Gfit$R2<-with(Gfit, 1-SSres/SStot)
        Gfit$N<-c(7)
        Gfit$AIC<-with(Gfit, 2*N-2*ll)

        rsq_out<-list(Yhat=Yhat, Gfit=Gfit)

        return(rsq_out)

      }

      #approximate parameter estimation is done by MCMC method
      par_mcmc<-modMCMC(f=cost, p=c(Vmax=0.1, Km=3, m0=0.01, f=4, Yu=0.8,R_0=0.03483481, S_0=0.07661063),
                          lower=c(Vmax=1e-5, Km=1e-2, m0=1e-5, f=1e-3, Yu=0.01, R_0=1e-4, S_0=1e-4),
                          upper=c(Vmax=1, Km=20, m0=1, f=100, Yu=0.8, R_0=2, S_0=2), niter=10000)

      #lower and upper limits for parameters are extracted
      pl<-summary(par_mcmc)["min",]
      pu<-summary(par_mcmc)["max",]

      #these limits are used to find global optimum by DEoptim
      opt_par<-DEoptim(fn=cost, lower=pl, upper=pu,
                       control = c(itermax = 10000, steptol = 50, reltol = 1e-8,
                                   trace=FALSE, strategy=3, NP=250))

      #global optimum parameters are further used in MCMC to find parameters distribution
      par_prof<-modMCMC(f=cost, p=opt_par$optim$bestmem,
                        lower=pl,
                        upper=pu, niter=5000)

      #goodness of fit
      fit<-rsq_ode(opt_par$optim$bestmem)

      #best parameters
      p<-opt_par$optim$bestmem
      names(p)<-parnames

      #return list with opt_par and par_prof
      estim_out<-list(pars=p, par_prof=par_prof, fit=fit)

      return(estim_out)

    }

    #parameter estimation
    if(FACT==4){

      res<-vector("list", length = 1)
      res[[1]]<-estim(odeset=dat)

    }else{

      res<-foreach(i=unique(dat$id), .combine=list, .multicombine = TRUE,
                   .packages=c("FME", "dplyr", "DEoptim", "reshape")) %dopar% {

                     estim(odeset=dat[dat$id==i,])

                   }
    }

    #calculation of overall goodness of fit from individual results
    if(FACT==4){

      res$goodness<-res[[1]]$fit$Gfit
      res$OvP<-res[[1]]$fit$Yhat

    }else{

      if(FACT==1){

        res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat)

        #rsquared calculation for each variable
        res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T),
                                                               SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                               ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
        res$goodness$N<-rep(length(parnames)*2, times=2)
        res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

      }else{

        if(FACT==2){

          res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat, res[[3]]$fit$Yhat)

          #rsquared calculation for each variable
          res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T),
                                                                 SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                                 ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
          res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
          res$goodness$N<-rep(length(parnames)*3, times=2)
          res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

        }else{

          res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat, res[[3]]$fit$Yhat, res[[4]]$fit$Yhat, res[[5]]$fit$Yhat, res[[6]]$fit$Yhat)

          #rsquared calculation for each variable
          res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T),
                                                                 SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                                 ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
          res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
          res$goodness$N<-rep(length(parnames)*6, times=2)
          res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

        }

      }

    }

  return(res)

}

```

```{r DEB respiration calibration, eval=F, echo=T}

#across all treatments
deb_r1<-deb_r(data=d, FACT = 4)

no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)

#for different structures
deb_r2<-deb_r(data=d, FACT = 2)

#for different substrates
deb_r3<-deb_r(data=d, FACT = 1)

#for each separately 
deb_r4<-deb_r(data=d, FACT = 3)

stopImplicitCluster()

deb_r1$goodness
deb_r2$goodness
deb_r3$goodness
deb_r4$goodness

```

```{r DEB respiration calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

#results of the previous iteration (not run)
# deb_r_results_par<-rbind(deb_r1[[1]]$pars,
#                      deb_r2[[1]]$pars, deb_r2[[2]]$pars, deb_r2[[3]]$pars,
#                      deb_r3[[1]]$pars, deb_r3[[2]]$pars,
#                      deb_r4[[1]]$pars, deb_r4[[2]]$pars, deb_r4[[3]]$pars,
#                      deb_r4[[4]]$pars, deb_r4[[5]]$pars, deb_r4[[6]]$pars)
# deb_r_results_good<-rbind(deb_r1$goodness[1,c(4:7)],
#                    deb_r2$goodness[1,c(4:7)],
#                    deb_r3$goodness[1,c(4:7)],
#                    deb_r4$goodness[1,c(4:7)])
# write.csv(deb_r_results_par, file = c("../deb_r_results_par.csv"), row.names = F)
# write.csv(deb_r_results_good, file = c("../deb_r_results_good.csv"), row.names = F)

deb_r_results_par<-read.csv("deb_r_results_par.csv")
deb_r_results_good<-read.csv("deb_r_results_good.csv")

t.s3.1<-data.frame(Vmax=deb_r_results_par[,1],
                 Km=deb_r_results_par[,2],
                 mo=deb_r_results_par[,3],
                 f=deb_r_results_par[,4],
                 Yu=deb_r_results_par[,5])
row.names(t.s3.1)<-c("All", "Broth", "Glass", "Wool",
                    "Celluloze", "Glucose",
                    "Broth - Celluloze", "Broth - Glucose",
                    "Glass - Celluloze", "Glass - Glucose",
                    "Wool - Celluloze", "Wool - Glucose")

options(knitr.kable.NA = '')

kable(t.s3.1, digits = c(2, 2, 3, 2, 2), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', '$\\m_{R}$', '$\\f_{0}$', '$\\Y_{S}$'), row.names=T, caption = "_**Table S5:** DEB model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(deb_r_results_good, digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S6:** Goodness of fit of DEB model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 2) %>%
  group_rows("Substrate", 3, 4) %>%
  group_rows("Each", 4, 4)

```

###All models are calibrated against respiration rate, proteins and DNA concentration
The calibration is done the same way as before. In addition to model parameters, the conversion factors between $C_{MB}$ (Monod and MEND model), $R$, $S$ (DEB model) and protein/DNA concetration are estimated. This is done for two reasons. First, conversion factors reported in literature vary widely. Second, using the mean or median of all reported conversion factors do not simulate $C_{MB}$, $R$ and $S$ dynamic accuratelly (data not shown).
There is one important difference between Monod/MEND model and DEB model. Temporal dynamic of Proteins and DNA concentration is different (see Fig. 2 of the main text) suggesting the variable composition of microbial biomass. Monod and MEND models do not account for this variability and thus, conversion factor have to be a fixed value. However, model parameters can theoretically compansate for the variability in microbial biomass composition but only if the model is calibrated against protein or DNA separately. Becuase the temporal dynamic of protein and DNA is different, the compensation have to occur differently. In contrast, DEB model account for the biomass composition variability explicitely and therefore, it can be calibrated against protein and DNA concentration at the same time. The following code reflects this difference.
Note that the calibration against respiration rate and protein/DNA concentration at the same time lead to a decrease of the the correspondence between Monod/MEND model respiration rate simulation and observations in several instances.

####Monod model
```{r monod model, eval=F, echo=T}
#######################################Proteins#############################################
monod_i<-function(data, FACT){

  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No

  if(FACT==1){

    dat<-data
    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)

  }else{

    if(FACT==2){

      dat<-data
      dat<-data
      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
      ids$id<-seq(1:nrow(ids))

      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))

    }else{

      if(FACT==3){

        dat<-data
        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
        ids$id<-seq(1:nrow(ids))

        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))

      }else{

        dat<-data
        dat$id<-c(1)

      }
    }
  }

    #monod growth function
    deriv<-function(time, state, pars){

      with(as.list(c(state, pars)),{

        Protc=fp*Cmic

        dCmic<--k*Cmic+CUE*Vmax*Cmic*C/(Km+C)
        dC<-k*Cmic-Vmax*Cmic*C/(Km+C)

        return(list(c(dCmic, dC), r=(1-CUE)*Vmax*Cmic*C/(Km+C), Protc=Protc))

      })
    }

    #define names of parameters
    parnames<-c("Vmax", "Km", "CUE", "k", "fp", "Cmic_0")

    #parameters estimation function
    estim<-function(odeset){

      #defining cost function
      cost<-function(x){

        par<-x[1:length(parnames)]

        names(par)<-parnames

        #first, pars dependent output from ode is matched with measured values
        yhat_all<-as.data.frame(ode(y=c(Cmic=par[["Cmic_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))

        #select time and the measured variables 
        yhat<-select(yhat_all, c("time", "r", "Protc"))

        #reformat to long format data frame
        Yhat<-melt(yhat, id.vars = "time")

        #add the measured data to a data frame
        Yhat$obs<-c(odeset[order(odeset$Time), c("r")], odeset[order(odeset$Time), "Protc"])

        #add the weighting factor
        #I want to have the weighting factor to be proportional to mean of the given variable 
        yweights<-Yhat %>% group_by(variable) %>% summarise(weights=mean(value, na.rm = T))

        #match with the Yhat data frame
        Yhat<-merge(Yhat, yweights, by.x=c("variable"), by.y=c("variable"))

        #now, the root mean square error is calculated
        NRMSE<-as.numeric(Yhat %>% group_by(variable) %>% summarise(NRMSE=sum((((value-obs)/mean(weights))^2), na.rm = T)) %>% 
                            summarise(NRMSE=sum(NRMSE)))

        return(NRMSE)

      }

      #defining goodness of fit function 
      rsq_ode<-function(x){

        par<-x[1:length(parnames)]

        names(par)<-parnames

        #first, pars dependent output from ode is matched with measured values
        yhat_all<-as.data.frame(ode(y=c(Cmic=par[["Cmic_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))

        #select time and the measured variables 
        yhat<-select(yhat_all, c("time", "r", "Protc"))

        #reformat to long format data frame
        Yhat<-melt(yhat, id.vars = "time")

        #add the measured data to a data frame
        Yhat$obs<-c(odeset[order(odeset$Time), c("r")], odeset[order(odeset$Time), "Protc"])
        Yhat$Substrate<-rep(odeset[order(odeset$Time), "Substrate"], times=2)
        Yhat$Structure<-rep(odeset[order(odeset$Time), "Structure"], times=2)

        #rsquared calculation for each variable
        Gfit<-Yhat %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
                                                        SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                        ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        Gfit$R2<-with(Gfit, 1-SSres/SStot)
        Gfit$N<-c(6)
        Gfit$AIC<-with(Gfit, 2*N-2*ll)

        rsq_out<-list(Yhat=Yhat, Gfit=Gfit)

        return(rsq_out)

      }

      #approximate parameter estimation is done by MCMC method
      par_mcmc<-modMCMC(f=cost, p=c(Vmax=0.1, Km=3, CUE=0.5, k=1e-3, fp=0.5, Cmic_0=0.01), 
                        lower=c(Vmax=1e-6, Km=1e-3, CUE=0, k=1e-6, fp=0, Cmic_0=0.0001),
                        upper=c(Vmax=10, Km=100, CUE=1, k=10, fp=1, Cmic_0=10), niter=10000)

      #lower and upper limits for parameters are extracted
      pl<-summary(par_mcmc)["min",]
      pu<-summary(par_mcmc)["max",]

      #these limits are used to find global optimum by DEoptim
      opt_par<-DEoptim(fn=cost, lower=pl, upper=pu, 
                       control = c(itermax = 10000, steptol = 50, reltol = 1e-8, 
                                   trace=FALSE, strategy=3, NP=250))

      #global optimum parameters are further used in MCMC to find parameters distribution
      par_prof<-modMCMC(f=cost, p=opt_par$optim$bestmem, 
                        lower=pl,
                        upper=pu, niter=5000)

      #goodness of fit
      fit<-rsq_ode(opt_par$optim$bestmem)

      #best parameters
      p<-opt_par$optim$bestmem
      names(p)<-parnames

      #return list with opt_par and par_prof
      estim_out<-list(pars=p, par_prof=par_prof, fit=fit)

      return(estim_out)

    }

    #parameter estimation
    if(FACT==4){

      res<-vector("list", length = 1)
      res[[1]]<-estim(odeset=dat)

    }else{

      res<-foreach(i=unique(dat$id), .combine=list, .multicombine = TRUE,
                   .packages=c("FME", "dplyr", "DEoptim", "reshape")) %dopar% {

                     estim(odeset=dat[dat$id==i,])

                   }
    }

    #calculation of overall goodness of fit from individual results
    if(FACT==4){

      res$goodness<-res[[1]]$fit$Gfit
      res$OvP<-res[[1]]$fit$Yhat

    }else{

      if(FACT==1){

        res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat)

        #rsquared calculation for each variable
        res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
                                                                   SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                                   ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
        res$goodness$N<-rep(length(parnames)*2, times=2)
        res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

      }else{

        if(FACT==2){

          res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat, res[[3]]$fit$Yhat)

          #rsquared calculation for each variable
          res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
                                                                     SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                                     ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
          res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
          res$goodness$N<-rep(length(parnames)*3, times=2)
          res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

        }else{

          res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat, res[[3]]$fit$Yhat, res[[4]]$fit$Yhat, res[[5]]$fit$Yhat, res[[6]]$fit$Yhat)

          #rsquared calculation for each variable
          res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
                                                                     SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                                     ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
          res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
          res$goodness$N<-rep(length(parnames)*6, times=2)
          res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

        }

      }

    }

    return(res)
}

##########################################DNA###########################################
###monod growth function
monod_i_DNA<-function(data, FACT){

  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No

  if(FACT==1){

    dat<-data
    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)

  }else{

    if(FACT==2){

      dat<-data
      dat<-data
      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
      ids$id<-seq(1:nrow(ids))

      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))

    }else{

      if(FACT==3){

        dat<-data
        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
        ids$id<-seq(1:nrow(ids))

        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))

      }else{

        dat<-data
        dat$id<-c(1)

      }
    }
  }

    #monod growth function
    deriv<-function(time, state, pars){

      with(as.list(c(state, pars)),{

        DNAc=fd*Cmic

        dCmic<--k*Cmic+CUE*Vmax*Cmic*C/(Km+C)
        dC<-k*Cmic-Vmax*Cmic*C/(Km+C)

        return(list(c(dCmic, dC), r=(1-CUE)*Vmax*Cmic*C/(Km+C), DNAc=DNAc))

      })
    }

    #define names of parameters
    parnames<-c("Vmax", "Km", "CUE", "k", "fd", "Cmic_0")

    #parameters estimation function
    estim<-function(odeset){

      #defining cost function
      cost<-function(x){

        par<-x[1:length(parnames)]

        names(par)<-parnames

        #first, pars dependent output from ode is matched with measured values
        yhat_all<-as.data.frame(ode(y=c(Cmic=par[["Cmic_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))

        #select time and the measured variables
        yhat<-select(yhat_all, c("time", "r", "DNAc"))

        #reformat to long format data frame
        Yhat<-melt(yhat, id.vars = "time")

        #add the measured data to a data frame
        Yhat$obs<-c(odeset[order(odeset$Time), c("r")], odeset[order(odeset$Time), "DNAc"])

        #add the weighting factor
        #I want to have the weighting factor to be proportional to mean of the given variable
        yweights<-Yhat %>% group_by(variable) %>% summarise(weights=mean(value, na.rm = T))

        #match with the Yhat data frame
        Yhat<-merge(Yhat, yweights, by.x=c("variable"), by.y=c("variable"))

        #now, the root mean square error is calculated
        NRMSE<-as.numeric(Yhat %>% group_by(variable) %>% summarise(NRMSE=sum((((value-obs)/mean(weights))^2), na.rm = T)) %>%
                            summarise(NRMSE=sum(NRMSE)))

        return(NRMSE)

      }

      #defining goodness of fit function
      rsq_ode<-function(x){

        par<-x[1:length(parnames)]

        names(par)<-parnames

        #first, pars dependent output from ode is matched with measured values
        yhat_all<-as.data.frame(ode(y=c(Cmic=par[["Cmic_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))

        #select time and the measured variables
        yhat<-select(yhat_all, c("time", "r", "DNAc"))

        #reformat to long format data frame
        Yhat<-melt(yhat, id.vars = "time")

        #add the measured data to a data frame
        Yhat$obs<-c(odeset[order(odeset$Time), c("r")], odeset[order(odeset$Time), "DNAc"])
        Yhat$Substrate<-rep(odeset[order(odeset$Time), "Substrate"], times=2)
        Yhat$Structure<-rep(odeset[order(odeset$Time), "Structure"], times=2)

        #rsquared calculation for each variable
        Gfit<-Yhat %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T),
                                                        SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                        ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        Gfit$R2<-with(Gfit, 1-SSres/SStot)
        Gfit$N<-c(6)
        Gfit$AIC<-with(Gfit, 2*N-2*ll)

        rsq_out<-list(Yhat=Yhat, Gfit=Gfit)

        return(rsq_out)

      }

      #approximate parameter estimation is done by MCMC method
      par_mcmc<-modMCMC(f=cost, p=c(Vmax=0.1, Km=3, CUE=0.5, k=1e-3, fp=0.5, Cmic_0=0.01),
                        lower=c(Vmax=1e-6, Km=1e-3, CUE=0, k=1e-6, fp=0, Cmic_0=0.0001),
                        upper=c(Vmax=10, Km=100, CUE=1, k=10, fp=1, Cmic_0=10), niter=10000)

      #lower and upper limits for parameters are extracted
      pl<-summary(par_mcmc)["min",]
      pu<-summary(par_mcmc)["max",]

      #these limits are used to find global optimum by DEoptim
      opt_par<-DEoptim(fn=cost, lower=pl, upper=pu,
                       control = c(itermax = 10000, steptol = 50, reltol = 1e-8,
                                   trace=FALSE, strategy=3, NP=250))

      #global optimum parameters are further used in MCMC to find parameters distribution
      par_prof<-modMCMC(f=cost, p=opt_par$optim$bestmem,
                        lower=pl,
                        upper=pu, niter=5000)

      #goodness of fit
      fit<-rsq_ode(opt_par$optim$bestmem)

      #best parameters
      p<-opt_par$optim$bestmem
      names(p)<-parnames

      #return list with opt_par and par_prof
      estim_out<-list(pars=p, par_prof=par_prof, fit=fit)

      return(estim_out)

    }

    #parameter estimation
    if(FACT==4){

      res<-vector("list", length = 1)
      res[[1]]<-estim(odeset=dat)

    }else{

      res<-foreach(i=unique(dat$id), .combine=list, .multicombine = TRUE,
                   .packages=c("FME", "dplyr", "DEoptim", "reshape")) %dopar% {

                     estim(odeset=dat[dat$id==i,])

                   }
    }

    #calculation of overall goodness of fit from individual results
    if(FACT==4){

      res$goodness<-res[[1]]$fit$Gfit
      res$OvP<-res[[1]]$fit$Yhat

    }else{

      if(FACT==1){

        res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat)

        #rsquared calculation for each variable
        res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T),
                                                                   SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                                   ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
        res$goodness$N<-rep(length(parnames)*2, times=2)
        res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

      }else{

        if(FACT==2){

          res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat, res[[3]]$fit$Yhat)

          #rsquared calculation for each variable
          res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T),
                                                                     SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                                     ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
          res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
          res$goodness$N<-rep(length(parnames)*3, times=2)
          res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

        }else{

          res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat, res[[3]]$fit$Yhat, res[[4]]$fit$Yhat, res[[5]]$fit$Yhat, res[[6]]$fit$Yhat)

          #rsquared calculation for each variable
          res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T),
                                                                     SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                                     ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
          res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
          res$goodness$N<-rep(length(parnames)*6, times=2)
          res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

        }

      }

    }

    return(res)
}

```

```{r, eval=F, echo=T}
#Proteins
#across all treatments
monod_i1<-monod_i(data=d, FACT = 4)

no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)

#for different structures
monod_i2<-monod_i(data=d, FACT = 2)

#for different substrates
monod_i3<-monod_i(data=d, FACT = 1)

#for each separately 
monod_i4<-monod_i(data=d, FACT = 3)

stopImplicitCluster()

monod_i1$goodness
monod_i2$goodness
monod_i3$goodness
monod_i4$goodness

#DNA
#across all treatments
monod_i1_DNA<-monod_i(data=d, FACT = 4)

no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)

#for different structures
monod_i2_DNA<-monod_i_DNA(data=d, FACT = 2)

#the initial Cmb concentration is estimated for each treatment separately
#the following code is run with the fixed initial Cmb concentration 
#across treatments (observed mean)
#this is done only with instances in which the model simulation can explain more than 0% 
#of the variability in protein and/or DNA dynamic  
monod_i2_DNA_fixed<-monod_i_DNA_fixed(data=d, FACT = 2)

#for different substrates
monod_i3_DNA<-monod_i_DNA(data=d, FACT = 1)

#for each separately 
monod_i4_DNA<-monod_i_DNA(data=d, FACT = 3)

stopImplicitCluster()

monod_i1_DNA$goodness
monod_i2_DNA$goodness
monod_i3_DNA$goodness
monod_i4_DNA$goodness

```

```{r Monod overall calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

#results of the previous iteration (not run)
# monod_all_results_par<-rbind(monod_i1[[1]]$pars, monod_i1_DNA[[1]]$pars,
#                     monod_i2[[1]]$pars, monod_i2[[2]]$pars, monod_i2[[3]]$pars,
#                     monod_i2_DNA_fixed[[1]]$pars, monod_i2_DNA_fixed[[2]]$pars, monod_i2_DNA_fixed[[3]]$pars,
#                     monod_i3[[1]]$pars, monod_i3[[2]]$pars,
#                     monod_i3_DNA[[1]]$pars, monod_i3_DNA[[2]]$pars,
#                     monod_i4[[1]]$pars, monod_i4[[2]]$pars, monod_i4[[3]]$pars,
#                     monod_i4[[4]]$pars, monod_i4[[5]]$pars, monod_i4[[6]]$pars,
#                     monod_i4_DNA[[1]]$pars, monod_i4_DNA[[2]]$pars, monod_i4_DNA[[3]]$pars,
#                     monod_i4_DNA[[4]]$pars, monod_i4_DNA[[5]]$pars, monod_i4_DNA[[6]]$pars)
# 
# monod_all_results_good<-rbind(monod_i1$goodness[,c(4:7)],
#                   monod_i2$goodness[,c(4:7)],
#                   monod_i3$goodness[,c(4:7)],
#                   monod_i4$goodness[,c(4:7)],
#                   monod_i1_DNA$goodness[,c(4:7)],
#                   monod_i2_DNA_fixed$goodness[,c(4:7)],
#                   monod_i3_DNA$goodness[,c(4:7)],
#                   monod_i4_DNA$goodness[,c(4:7)])
# 
# monod_all_results_good[monod_all_results_good$R2<0,"R2"]<-c(0)
# monod_all_results_good$Name<-c(rep(c("Respiration rate", "Protein"), times=4),
#                                rep(c("Respiration rate", "DNA"), times=4))
# monod_all_results_good<-monod_all_results_good[,c(5,1,2,3,4)]
# 
# 
# write.csv(monod_all_results_par, file = c("../monod_all_results_par.csv"), row.names = F)
# write.csv(monod_all_results_good, file = c("../monod_all_results_good.csv"), row.names = F)

monod_all_results_par<-read.csv("monod_all_results_par.csv")
monod_all_results_good<-read.csv("monod_all_results_good.csv")

t.s4.1<-data.frame(Name=rep(c("All", "Broth", "Glass", "Wool",
                    "Celluloze", "Glucose",
                    "Broth - Celluloze", "Broth - Glucose",
                    "Glass - Celluloze", "Glass - Glucose",
                    "Wool - Celluloze", "Wool - Glucose"), each =2),
                   Parameter=rep(c("Protein", "DNA"), times=12),
                   Vmax=monod_all_results_par[,1],
                   Km=monod_all_results_par[,2],
                   CUE=monod_all_results_par[,3],
                   k=monod_all_results_par[,4],
                   conv=monod_all_results_par[,5])

options(knitr.kable.NA = '')

kable(t.s4.1, digits = c(0, 0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" "," ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$', 'Conversion factor'), row.names=F, caption = "_**Table S7:** Monod model parameters calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 2) %>%
  group_rows("Structure", 3, 8) %>%
  group_rows("Substrate", 9, 12) %>%
  group_rows("Each", 13, 24)

kable(monod_all_results_good, digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S8:** Goodness of fit of Monod model calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("Protein", 1, 8, label_row_css = "background-color: #666; color: #fff;") %>%
  group_rows("All", 1, 2) %>%
  group_rows("Structure", 3, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 8) %>%
  group_rows("DNA", 9, 16, label_row_css = "background-color: #666; color: #fff;") %>%
  group_rows("All", 9, 10) %>%
  group_rows("Structure",11, 12) %>%
  group_rows("Substrate", 13, 14) %>%
  group_rows("Each", 15, 16)

```

####Mend model
```{r MEND model, eval=F, echo=T}
############################################Protein############################################
mend_i<-function(data, FACT){

  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No

  if(FACT==1){

    dat<-data
    dat$id<-ifelse(d$Substrate=="Glucose", 1, 2)

  }else{

    if(FACT==2){

      dat<-data
      dat<-data
      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
      ids$id<-seq(1:nrow(ids))

      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))

    }else{

      if(FACT==3){

        dat<-data
        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
        ids$id<-seq(1:nrow(ids))

        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))

      }else{

        dat<-data
        dat$id<-c(1)

      }
    }
  }

    #mend model
    deriv<-function(time, state, pars){

      with(as.list(c(state, pars)),{

        #equations
        #C uptake
        F1=1/CUE*(Vmax+mr)*Cmic*C/(Km+C)
        #growth respiration
        F4=(1/CUE-1)*Vmax*Cmic*C/(Km+C)
        #maintnance respiration
        F5=(1/CUE-1)*mr*Cmic*C/(Km+C)
        #microbial mortality
        F8=mr*Cmic

        Protc=fp*Cmic

        #states
        dCmic<-F1-(F4+F5)-F8
        dC<--F1+F8

        return(list(c(dCmic, dC), r=F4+F5, Protc=Protc))

      })
    }
    #define names of parameters
    parnames<-c("Vmax", "Km", "CUE", "mr", "fp", "Cmic_0")

    #parameters estimation function
    estim<-function(odeset){

      #defining cost function
      cost<-function(x){

        par<-x[1:length(parnames)]

        names(par)<-parnames

        #first, pars dependent output from ode is matched with measured values
        yhat_all<-as.data.frame(ode(y=c(Cmic=par[["Cmic_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))

        #select time and the measured variables 
        yhat<-select(yhat_all, c("time", "Protc", "r"))

        #reformat to long format data frame
        Yhat<-melt(yhat, id.vars = "time")

        #add the measured data to a data frame
        Yhat$obs<-c(odeset[order(odeset$Time), "Protc"], odeset[order(odeset$Time), c("r")])

        #add the weighting factor
        #I want to have the weighting factor to be proportional to mean of the given variable 
        yweights<-Yhat %>% group_by(variable) %>% summarise(weights=mean(value, na.rm = T))

        #match with the Yhat data frame
        Yhat<-merge(Yhat, yweights, by.x=c("variable"), by.y=c("variable"))

        #now, the root mean square error is calculated
        NRMSE<-as.numeric(Yhat %>% group_by(variable) %>% summarise(NRMSE=sum((((value-obs)/mean(weights))^2), na.rm = T)) %>% 
                            summarise(NRMSE=sum(NRMSE)))

        return(NRMSE)

      }

      #defining goodness of fit function 
      rsq_ode<-function(x){

        par<-x[1:length(parnames)]

        names(par)<-parnames

        #first, pars dependent output from ode is matched with measured values
        yhat_all<-as.data.frame(ode(y=c(Cmic=par[["Cmic_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))

        #select time and the measured variables 
        yhat<-select(yhat_all, c("time", "Protc", "r"))

        #reformat to long format data frame
        Yhat<-melt(yhat, id.vars = "time")

        #add the measured data to a data frame
        Yhat$obs<-c(odeset[order(odeset$Time), "Protc"], odeset[order(odeset$Time), c("r")])
        Yhat$Substrate<-rep(odeset[order(odeset$Time), "Substrate"], times=2)
        Yhat$Structure<-rep(odeset[order(odeset$Time), "Structure"], times=2)

        #rsquared calculation for each variable
        Gfit<-Yhat %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
                                                        SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                        ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        Gfit$R2<-with(Gfit, 1-SSres/SStot)
        Gfit$N<-c(6)
        Gfit$AIC<-with(Gfit, 2*N-2*ll)

        rsq_out<-list(Yhat=Yhat, Gfit=Gfit)

        return(rsq_out)

      }

      #approximate parameter estimation is done by MCMC method
      par_mcmc<-modMCMC(f=cost, p=c(Vmax=0.1, Km=3, CUE=0.5, mr=0.01, fp=0.5, Cmic_0=0.01), 
                          lower=c(Vmax=1e-3, Km=1e-3, CUE=0, mr=1e-5, fp=0, Cmic_0=1e-5),
                          upper=c(Vmax=10, Km=100, CUE=1, mr=10, fp=1, Cmic_0=5), niter=10000)

      #lower and upper limits for parameters are extracted
      pl<-summary(par_mcmc)["min",]
      pu<-summary(par_mcmc)["max",]

      #these limits are used to find global optimum by DEoptim
      opt_par<-DEoptim(fn=cost, lower=pl, upper=pu, 
                       control = c(itermax = 10000, steptol = 50, reltol = 1e-8, 
                                   trace=FALSE, strategy=3, NP=250))

      #global optimum parameters are further used in MCMC to find parameters distribution
      par_prof<-modMCMC(f=cost, p=opt_par$optim$bestmem, 
                        lower=pl,
                        upper=pu, niter=5000)

      #goodness of fit
      fit<-rsq_ode(opt_par$optim$bestmem)

      #best parameters
      p<-opt_par$optim$bestmem
      names(p)<-parnames

      #return list with opt_par and par_prof
      estim_out<-list(pars=p, par_prof=par_prof, fit=fit)

      return(estim_out)

    }

    #parameter estimation
    if(FACT==4){

      res<-vector("list", length = 1)
      res[[1]]<-estim(odeset=dat)

    }else{

      res<-foreach(i=unique(dat$id), .combine=list, .multicombine = TRUE,
                   .packages=c("FME", "dplyr", "DEoptim", "reshape")) %dopar% {

                     estim(odeset=dat[dat$id==i,])

                   }
    }

    #calculation of overall goodness of fit from individual results
    if(FACT==4){

      res$goodness<-res[[1]]$fit$Gfit
      res$OvP<-res[[1]]$fit$Yhat

    }else{

      if(FACT==1){

        res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat)

        #rsquared calculation for each variable
        res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
                                                               SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                               ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
        res$goodness$N<-rep(length(parnames)*2, times=2)
        res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

      }else{

        if(FACT==2){

          res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat, res[[3]]$fit$Yhat)

          #rsquared calculation for each variable
          res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
                                                                 SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                                 ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
          res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
          res$goodness$N<-rep(length(parnames)*3, times=2)
          res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

        }else{

          res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat, res[[3]]$fit$Yhat, res[[4]]$fit$Yhat, res[[5]]$fit$Yhat, res[[6]]$fit$Yhat)

          #rsquared calculation for each variable
          res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
                                                                 SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                                 ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
          res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
          res$goodness$N<-rep(length(parnames)*6, times=2)
          res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

        }

      }

    }

  return(res)

}
##############################################DNA##############################################
mend_i_DNA<-function(data, FACT){

  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No

  if(FACT==1){

    dat<-data
    dat$id<-ifelse(d$Substrate=="Glucose", 1, 2)

  }else{

    if(FACT==2){

      dat<-data
      dat<-data
      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
      ids$id<-seq(1:nrow(ids))

      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))

    }else{

      if(FACT==3){

        dat<-data
        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
        ids$id<-seq(1:nrow(ids))

        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))

      }else{

        dat<-data
        dat$id<-c(1)

      }
    }
  }

    #mend model
    deriv<-function(time, state, pars){

      with(as.list(c(state, pars)),{

        #equations
        #C uptake
        F1=1/CUE*(Vmax+mr)*Cmic*C/(Km+C)
        #growth respiration
        F4=(1/CUE-1)*Vmax*Cmic*C/(Km+C)
        #maintnance respiration
        F5=(1/CUE-1)*mr*Cmic*C/(Km+C)
        #microbial mortality
        F8=mr*Cmic

        DNAc=fd*Cmic

        #states
        dCmic<-F1-(F4+F5)-F8
        dC<--F1+F8

        return(list(c(dCmic, dC), r=F4+F5, DNAc=DNAc))

      })
    }
    #define names of parameters
    parnames<-c("Vmax", "Km", "CUE", "mr", "fd", "Cmic_0")

    #parameters estimation function
    estim<-function(odeset){

      #defining cost function
      cost<-function(x){

        par<-x[1:length(parnames)]

        names(par)<-parnames

        #first, pars dependent output from ode is matched with measured values
        yhat_all<-as.data.frame(ode(y=c(Cmic=par[["Cmic_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))

        #select time and the measured variables 
        yhat<-select(yhat_all, c("time", "DNAc", "r"))

        #reformat to long format data frame
        Yhat<-melt(yhat, id.vars = "time")

        #add the measured data to a data frame
        Yhat$obs<-c(odeset[order(odeset$Time), "DNAc"], odeset[order(odeset$Time), c("r")])

        #add the weighting factor
        #I want to have the weighting factor to be proportional to mean of the given variable 
        yweights<-Yhat %>% group_by(variable) %>% summarise(weights=mean(value, na.rm = T))

        #match with the Yhat data frame
        Yhat<-merge(Yhat, yweights, by.x=c("variable"), by.y=c("variable"))

        #now, the root mean square error is calculated
        NRMSE<-as.numeric(Yhat %>% group_by(variable) %>% summarise(NRMSE=sum((((value-obs)/mean(weights))^2), na.rm = T)) %>% 
                            summarise(NRMSE=sum(NRMSE)))

        return(NRMSE)

      }

      #defining goodness of fit function 
      rsq_ode<-function(x){

        par<-x[1:length(parnames)]

        names(par)<-parnames

        #first, pars dependent output from ode is matched with measured values
        yhat_all<-as.data.frame(ode(y=c(Cmic=par[["Cmic_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))

        #select time and the measured variables 
        yhat<-select(yhat_all, c("time", "DNAc", "r"))

        #reformat to long format data frame
        Yhat<-melt(yhat, id.vars = "time")

        #add the measured data to a data frame
        Yhat$obs<-c(odeset[order(odeset$Time), "DNAc"], odeset[order(odeset$Time), c("r")])
        Yhat$Substrate<-rep(odeset[order(odeset$Time), "Substrate"], times=2)
        Yhat$Structure<-rep(odeset[order(odeset$Time), "Structure"], times=2)

        #rsquared calculation for each variable
        Gfit<-Yhat %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
                                                        SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                        ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        Gfit$R2<-with(Gfit, 1-SSres/SStot)
        Gfit$N<-c(6)
        Gfit$AIC<-with(Gfit, 2*N-2*ll)

        rsq_out<-list(Yhat=Yhat, Gfit=Gfit)

        return(rsq_out)

      }

      #approximate parameter estimation is done by MCMC method
      par_mcmc<-modMCMC(f=cost, p=c(Vmax=0.1, Km=3, CUE=0.5, mr=0.01, fp=0.5, Cmic_0=0.01), 
                          lower=c(Vmax=1e-3, Km=1e-3, CUE=0, mr=1e-5, fp=0, Cmic_0=1e-5),
                          upper=c(Vmax=10, Km=100, CUE=1, mr=10, fp=1, Cmic_0=5), niter=10000)

      #lower and upper limits for parameters are extracted
      pl<-summary(par_mcmc)["min",]
      pu<-summary(par_mcmc)["max",]

      #these limits are used to find global optimum by DEoptim
      opt_par<-DEoptim(fn=cost, lower=pl, upper=pu, 
                       control = c(itermax = 10000, steptol = 50, reltol = 1e-8, 
                                   trace=FALSE, strategy=3, NP=250))

      #global optimum parameters are further used in MCMC to find parameters distribution
      par_prof<-modMCMC(f=cost, p=opt_par$optim$bestmem, 
                        lower=pl,
                        upper=pu, niter=5000)

      #goodness of fit
      fit<-rsq_ode(opt_par$optim$bestmem)

      #best parameters
      p<-opt_par$optim$bestmem
      names(p)<-parnames

      #return list with opt_par and par_prof
      estim_out<-list(pars=p, par_prof=par_prof, fit=fit)

      return(estim_out)

    }

    #parameter estimation
    if(FACT==4){

      res<-vector("list", length = 1)
      res[[1]]<-estim(odeset=dat)

    }else{

      res<-foreach(i=unique(dat$id), .combine=list, .multicombine = TRUE,
                   .packages=c("FME", "dplyr", "DEoptim", "reshape")) %dopar% {

                     estim(odeset=dat[dat$id==i,])

                   }
    }

    #calculation of overall goodness of fit from individual results
    if(FACT==4){

      res$goodness<-res[[1]]$fit$Gfit
      res$OvP<-res[[1]]$fit$Yhat

    }else{

      if(FACT==1){

        res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat)

        #rsquared calculation for each variable
        res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
                                                               SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                               ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
        res$goodness$N<-rep(length(parnames)*2, times=2)
        res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

      }else{

        if(FACT==2){

          res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat, res[[3]]$fit$Yhat)

          #rsquared calculation for each variable
          res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
                                                                 SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                                 ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
          res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
          res$goodness$N<-rep(length(parnames)*3, times=2)
          res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

        }else{

          res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat, res[[3]]$fit$Yhat, res[[4]]$fit$Yhat, res[[5]]$fit$Yhat, res[[6]]$fit$Yhat)

          #rsquared calculation for each variable
          res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
                                                                 SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                                 ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
          res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
          res$goodness$N<-rep(length(parnames)*6, times=2)
          res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

        }

      }

    }

  return(res)

}
```

```{r eval=F, echo=T}
#Proteins
#across all treatments
mend_i1<-mend_i(data=d, FACT = 4)

no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)

#for different structures
mend_i2<-mend_i(data=d, FACT = 2)

#for different substrates
mend_i3<-mend_i(data=d, FACT = 1)

#for each separately 
mend_i4<-mend_i(data=d, FACT = 3)

stopImplicitCluster()

mend_i1$goodness
mend_i2$goodness
mend_i3$goodness
mend_i4$goodness

#DNA
#across all treatments
mend_i1_DNA<-mend_i_DNA(data=d, FACT = 4)

no_cors<-detectCores()
cl<-makeCluster(no_cors)
registerDoParallel(cl)

#for different structures
mend_i2_DNA<-mend_i_DNA(data=d, FACT = 2)

#for different substrates
mend_i3_DNA<-mend_i_DNA(data=d, FACT = 1)

#for each separately 
mend_i4_DNA<-mend_i_DNA(data=d, FACT = 3)

stopImplicitCluster()

mend_i1_DNA$goodness
mend_i2_DNA$goodness
mend_i3_DNA$goodness
mend_i4_DNA$goodness

```

```{r MEND overall calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

#results of the previous iteration (not run)
# mend_all_results_par<-rbind(mend_i1[[1]]$pars, mend_i1_DNA[[1]]$pars,
#                    mend_i2[[1]]$pars, mend_i2[[2]]$pars, mend_i2[[3]]$pars,
#                    mend_i2_DNA[[1]]$pars, mend_i2_DNA[[2]]$pars, mend_i2_DNA[[3]]$pars,
#                    mend_i3[[1]]$pars, mend_i3[[2]]$pars,
#                    mend_i3_DNA[[1]]$pars, mend_i3_DNA[[2]]$pars,
#                    mend_i4[[1]]$pars, mend_i4[[2]]$pars, mend_i4[[3]]$pars,
#                    mend_i4[[4]]$pars, mend_i4[[5]]$pars, mend_i4[[6]]$pars,
#                    mend_i4_DNA[[1]]$pars, mend_i4_DNA[[2]]$pars, mend_i4_DNA[[3]]$pars,
#                    mend_i4_DNA[[4]]$pars, mend_i4_DNA[[5]]$pars, mend_i4_DNA[[6]]$pars)
# 
# mend_all_results_good<-rbind(mend_i1$goodness[c(2,1),c(4:7)],
#                  mend_i2$goodness[c(2,1),c(4:7)],
#                  mend_i3$goodness[c(2,1),c(4:7)],
#                  mend_i4$goodness[c(2,1),c(4:7)],
#                  mend_i1_DNA$goodness[c(2,1),c(4:7)],
#                  mend_i2_DNA$goodness[c(2,1),c(4:7)],
#                  mend_i3_DNA$goodness[c(2,1),c(4:7)],
#                  mend_i4_DNA$goodness[c(2,1),c(4:7)])
# 
# mend_all_results_good[mend_all_results_good$R2<0,"R2"]<-c(0)
# mend_all_results_good$Name<-c(rep(c("Respiration rate", "Protein"), times=4),
#                               rep(c("Respiration rate", "DNA"), times=4))
# mend_all_results_good<-mend_all_results_good[,c(5,1,2,3,4)]
#  
# 
# write.csv(mend_all_results_par, file = c("../mend_all_results_par.csv"), row.names = F)
# write.csv(mend_all_results_good, file = c("../mend_all_results_good.csv"), row.names = F)

mend_all_results_par<-read.csv("mend_all_results_par.csv")
mend_all_results_good<-read.csv("mend_all_results_good.csv")

t.s5.1<-data.frame(Name=rep(c("All", "Broth", "Glass", "Wool",
                    "Celluloze", "Glucose",
                    "Broth - Celluloze", "Broth - Glucose",
                    "Glass - Celluloze", "Glass - Glucose",
                    "Wool - Celluloze", "Wool - Glucose"), each =2),
                   Parameter=rep(c("Protein", "DNA"), times=12),
                   Vmax=mend_all_results_par[,1],
                   Km=mend_all_results_par[,2],
                   CUE=mend_all_results_par[,3],
                   mr=mend_all_results_par[,4],
                   conv=mend_all_results_par[,5])

options(knitr.kable.NA = '')

kable(t.s5.1, digits = c(0, 0, 2, 2, 2, 3, 2), align = 'c', "html", booktabs=T, col.names = c(" "," ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$', 'Conversion factor'), row.names=F, caption = "_**Table S9:** MEND model parameters calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 2) %>%
  group_rows("Structure", 3, 8) %>%
  group_rows("Substrate", 9, 12) %>%
  group_rows("Each", 13, 24)

kable(mend_all_results_good, digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S10:** Goodness of fit of MEND model calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("Protein", 1, 8, label_row_css = "background-color: #666; color: #fff;") %>%
  group_rows("All", 1, 2) %>%
  group_rows("Structure", 3, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 8) %>%
  group_rows("DNA", 9, 16, label_row_css = "background-color: #666; color: #fff;") %>%
  group_rows("All", 9, 10) %>%
  group_rows("Structure",11, 12) %>%
  group_rows("Substrate", 13, 14) %>%
  group_rows("Each", 15, 16)

```

####DEB model
```{r DEB model, eval=F, echo=T}
deb_i_all<-function(data, FACT){

  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No

  #if factor needs to be included, its levels are defined
  if(FACT==1){

    dat<-data
    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)

  }else{

    if(FACT==2){

      dat<-data
      dat<-data
      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
      ids$id<-seq(1:nrow(ids))

      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))

    }else{

      if(FACT==3){

        dat<-data
        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
        ids$id<-seq(1:nrow(ids))

        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))

      }else{

        dat<-data
        dat$id<-c(1)

      }
    }
  }

    #deb model
    deriv<-function(time, state, pars){

      with(as.list(c(state, pars)),{

        #Carbon uptake
        Cu=Vmax*C*S/(Km+C)

        #maintnance
        m=S*m0

        #reserve mobilization rate for growth and enzyme production
        an=f*R-m

        #calibrated variables
        #protein abundance in reserves and structures is mean value reported
        #by Hanegraaf and Muller, 2001
        r=m+pmax(an*(1-Yu), 0)

        #Protc=0.7095*S+0.6085*R
        Protc=fpr*R+fps*S
        DNAc=fds*S

        dR<-Cu-f*R
        dS<-pmax(an*Yu,0)+pmin(0, an)
        dC<--Cu

        return(list(c(dR, dS, dC), r=r, Protc=Protc, DNAc=DNAc))

      })
    }
    #define names of parameters
    parnames<-c("Vmax", "Km", "m0", "f", "Yu", "fpr", "fps", "fds", "R_0", "S_0")

    #parameters estimation function
    estim<-function(odeset){

      #defining cost function
      cost<-function(x){

        par<-x[1:length(parnames)]

        names(par)<-parnames

        #first, pars dependent output from ode is matched with measured values
        yhat_all<-as.data.frame(ode(y=c(R=par[["R_0"]], S=par[["S_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))

        #select time and the measured variables
        yhat<-select(yhat_all, c("time", "r", "Protc", "DNAc"))

        #reformat to long format data frame
        Yhat<-melt(yhat, id.vars = "time")

        #add the measured data to a data frame
        Yhat$obs<-c(odeset[order(odeset$Time), c("r")], odeset[order(odeset$Time), "Protc"], odeset[order(odeset$Time), "DNAc"])

        #add the weighting factor
        #I want to have the weighting factor to be proportional to mean of the given variable
        yweights<-Yhat %>% group_by(variable) %>% summarise(weights=mean(value, na.rm = T))

        #match with the Yhat data frame
        Yhat<-merge(Yhat, yweights, by.x=c("variable"), by.y=c("variable"))

        #now, the root mean square error is calculated
        NRMSE<-as.numeric(Yhat %>% group_by(variable) %>% summarise(NRMSE=sum((((value-obs)/mean(weights))^2), na.rm = T)) %>%
                            summarise(NRMSE=sum(NRMSE)))

        return(NRMSE)

      }

      #defining goodness of fit function
      rsq_ode<-function(x){

        par<-x[1:length(parnames)]

        names(par)<-parnames

        #first, pars dependent output from ode is matched with measured values
        yhat_all<-as.data.frame(ode(y=c(R=par[["R_0"]], S=par[["S_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))

        #select time and the measured variables
        yhat<-select(yhat_all, c("time", "r", "Protc", "DNAc"))

        #reformat to long format data frame
        Yhat<-melt(yhat, id.vars = "time")

        #add the measured data to a data frame
        Yhat$obs<-c(odeset[order(odeset$Time), c("r")], odeset[order(odeset$Time), "Protc"], odeset[order(odeset$Time), "DNAc"])
        Yhat$Substrate<-rep(odeset[order(odeset$Time), "Substrate"], times=3)
        Yhat$Structure<-rep(odeset[order(odeset$Time), "Structure"], times=3)

        #rsquared calculation for each variable
        Gfit<-Yhat %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T),
                                                        SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                        ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        Gfit$R2<-with(Gfit, 1-SSres/SStot)
        Gfit$N<-c(10)
        Gfit$AIC<-with(Gfit, 2*N-2*ll)

        rsq_out<-list(Yhat=Yhat, Gfit=Gfit)

        return(rsq_out)

      }

      #approximate parameter estimation is done by MCMC method
      par_mcmc<-modMCMC(f=cost, p=c(Vmax=0.1, Km=3, m0=0.01, f=4, Yu=0.8, fpr=0.5, fps=0.5, fds=0.03, R_0=0.03483481, S_0=0.07661063),
                          lower=c(Vmax=1e-5, Km=1e-2, m0=1e-5, f=1e-3, Yu=0.01, fpr=0, fps=0, fds=0, R_0=1e-4, S_0=1e-4),
                          upper=c(Vmax=1, Km=20, m0=1, f=100, Yu=0.8, fpr=1, fps=1, fds=1, R_0=2, S_0=2), niter=10000)

      #lower and upper limits for parameters are extracted
      pl<-summary(par_mcmc)["min",]
      pu<-summary(par_mcmc)["max",]

      #these limits are used to find global optimum by DEoptim
      opt_par<-DEoptim(fn=cost, lower=pl, upper=pu,
                       control = c(itermax = 10000, steptol = 50, reltol = 1e-8,
                                   trace=FALSE, strategy=3, NP=250))

      #global optimum parameters are further used in MCMC to find parameters distribution
      par_prof<-modMCMC(f=cost, p=opt_par$optim$bestmem,
                        lower=pl,
                        upper=pu, niter=5000)

      #goodness of fit
      fit<-rsq_ode(opt_par$optim$bestmem)

      #best parameters
      p<-opt_par$optim$bestmem
      names(p)<-parnames

      #return list with opt_par and par_prof
      estim_out<-list(pars=p, par_prof=par_prof, fit=fit)

      return(estim_out)

    }

    #parameter estimation
    if(FACT==4){

      res<-vector("list", length = 1)
      res[[1]]<-estim(odeset=dat)

    }else{

      res<-foreach(i=unique(dat$id), .combine=list, .multicombine = TRUE,
                   .packages=c("FME", "dplyr", "DEoptim", "reshape")) %dopar% {

                     estim(odeset=dat[dat$id==i,])

                   }
    }

    #calculation of overall goodness of fit from individual results
    if(FACT==4){

      res$goodness<-res[[1]]$fit$Gfit
      res$OvP<-res[[1]]$fit$Yhat

    }else{

      if(FACT==1){

        res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat)

        #rsquared calculation for each variable
        res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T),
                                                               SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                               ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
        res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
        res$goodness$N<-rep(length(parnames)*2, times=3)
        res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

      }else{

        if(FACT==2){

          res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat, res[[3]]$fit$Yhat)

          #rsquared calculation for each variable
          res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T),
                                                                 SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                                 ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
          res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
          res$goodness$N<-rep(length(parnames)*3, times=3)
          res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

        }else{

          res$OvP<-rbind(res[[1]]$fit$Yhat, res[[2]]$fit$Yhat, res[[3]]$fit$Yhat, res[[4]]$fit$Yhat, res[[5]]$fit$Yhat, res[[6]]$fit$Yhat)

          #rsquared calculation for each variable
          res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T),
                                                                 SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
                                                                 ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
          res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
          res$goodness$N<-rep(length(parnames)*6, times=3)
          res$goodness$AIC<-with(res$goodness, 2*N-2*ll)

        }

      }

    }

  return(res)

}
```

```{r, eval=F, echo=T}

#across all treatments
deb_i1_all<-deb_i_all(data=d, FACT = 4)

no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)

#for different structures
deb_i2_all<-deb_i_all(data=d, FACT = 2)

#the initial R and S concentrations are estimated for each treatment separately
#the following code is run with the fixed initial R and S concentrations
#across treatments (observed mean)
#this is done only with instances in which the model simulation can explain more than 0% 
#of the variability in protein and/or DNA dynamic  
deb_i2_all_fix<-deb_i_all_fix(data=d, FACT = 2)

#in the following code, the Ys parameter is fixed across treatments
#becuase its variability across treatments is minimal (data not shown)
deb_i2_all_fix_Yu<-deb_i_all_fix_Yu(data=d, FACT = 2)

#for different substrates
deb_i3_all<-deb_i_all(data=d, FACT = 1)

#for each separately 
deb_i4_all<-deb_i_all(data=d, FACT = 3)

deb_i4_all_fix<-deb_i_all_fix(data=d, FACT = 3)

stopImplicitCluster()

deb_i1_all$goodness
deb_i2_all$goodness
deb_i3_all$goodness
deb_i4_all$goodness

```

```{r DEB overall calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

#results of the previous iteration (not run)
# deb_all_results_par<-rbind(deb_i1_all[[1]]$pars[1:8], 
#                    c(deb_i2_all_fix_Yu[[1]]$pars[1:4], Yu=0.6, deb_i2_all_fix_Yu[[1]]$pars[5:7]), c(deb_i2_all_fix_Yu[[2]]$pars[1:4], Yu=0.6, deb_i2_all_fix_Yu[[2]]$pars[5:7]), c(deb_i2_all_fix_Yu[[3]]$pars[1:4], Yu=0.6, deb_i2_all_fix_Yu[[3]]$pars[5:7]),
#                    deb_i3_all[[1]]$pars[1:8], deb_i3_all[[2]]$pars[1:8],
#                    deb_i4_all[[1]]$pars[1:8], deb_i4_all[[2]]$pars[1:8], deb_i4_all[[3]]$pars[1:8],
#                    deb_i4_all[[4]]$pars[1:8], deb_i4_all[[5]]$pars[1:8], deb_i4_all[[6]]$pars[1:8])
# 
# deb_all_results_good<-rbind(deb_i1_all$goodness[,c(4:7)],
#                  deb_i2_all_fix_Yu$goodness[,c(4:7)],
#                  deb_i3_all$goodness[,c(4:7)],
#                  deb_i4_all$goodness[,c(4:7)])
# 
# deb_all_results_good[deb_all_results_good$R2<0,"R2"]<-c(0)
# deb_all_results_good$Name<-rep(c("Respiration rate", "Protein", "DNA"), times=4)
# deb_all_results_good<-deb_all_results_good[,c(5,1,2,3,4)]
#  
# 
# write.csv(deb_all_results_par, file = c("../deb_all_results_par.csv"), row.names = F)
# write.csv(deb_all_results_good, file = c("../deb_all_results_good.csv"), row.names = F)

deb_all_results_par<-read.csv("deb_all_results_par")
deb_all_results_good<-read.csv("deb_all_results_good")

t.s6.1<-data.frame(Name=c("All", "Broth", "Glass", "Wool",
                    "Celluloze", "Glucose",
                    "Broth - Celluloze", "Broth - Glucose",
                    "Glass - Celluloze", "Glass - Glucose",
                    "Wool - Celluloze", "Wool - Glucose"),
                   Vmax=deb_all_results_par[,1],
                   Km=deb_all_results_par[,2],
                   m0=deb_all_results_par[,3],
                   f=deb_all_results_par[,4],
                   Yu=deb_all_results_par[,5],
                   fpr=deb_all_results_par[,6],
                   fps=deb_all_results_par[,7],
                   fds=deb_all_results_par[,8])

options(knitr.kable.NA = '')

kable(t.s6.1, digits = c(0, 2, 2, 3, 2, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" ",'$\\V_{MAX}$', '$\\K_{M}$', '$\\m_{R}$', '$\\f_{0}$', '$\\Y_{S}$', '$\\R_{Proteins}$', '$\\R_{DNA}$', '$\\"S"_{DNA}$'), row.names=F, caption = "_**Table S11:** DEB model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(deb_all_results_good, digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S12:** Goodness of fit of DEB model calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 3) %>%
  group_rows("Structure", 4, 6) %>%
  group_rows("Substrate", 7, 9) %>%
  group_rows("Each", 10, 12)

```

### 2019-02-28 SI_code.Rmd
cleaning old scripts

@@
 ```
+##Statistical analysis
+###Respiration
+Cumulative respiration is calculated from the measured respiration rate using the local polynomial regression.
 
@@
 row.names(t.s1.1)<-c("All", "Broth", "Glass", "Wool",
-                    "Celluloze", "Glucose",
-                    "Broth - Celluloze", "Broth - Glucose",
-                    "Glass - Celluloze", "Glass - Glucose",
-                    "Wool - Celluloze", "Wool - Glucose")
+                    "Cellobiose", "Glucose",
+                    "Broth - Cellobiose", "Broth - Glucose",
+                    "Glass - Cellobiose", "Glass - Glucose",
+                    "Wool - Cellobiose", "Wool - Glucose")
 
@@
                     "Celluloze", "Glucose",
-                    "Broth - Celluloze", "Broth - Glucose",
-                    "Glass - Celluloze", "Glass - Glucose",
-                    "Wool - Celluloze", "Wool - Glucose")
+                    "Broth - Cellobiose", "Broth - Glucose",
+                    "Glass - Cellobiose", "Glass - Glucose",
+                    "Wool - Cellobiose", "Wool - Glucose")
 
@@
 row.names(t.s3.1)<-c("All", "Broth", "Glass", "Wool",
-                    "Celluloze", "Glucose",
-                    "Broth - Celluloze", "Broth - Glucose",
-                    "Glass - Celluloze", "Glass - Glucose",
-                    "Wool - Celluloze", "Wool - Glucose")
+                    "Cellobiose", "Glucose",
+                    "Broth - Cellobiose", "Broth - Glucose",
+                    "Glass - Cellobiose", "Glass - Glucose",
+                    "Wool - Cellobiose", "Wool - Glucose")
 
@@
 t.s4.1<-data.frame(Name=rep(c("All", "Broth", "Glass", "Wool",
-                    "Celluloze", "Glucose",
-                    "Broth - Celluloze", "Broth - Glucose",
-                    "Glass - Celluloze", "Glass - Glucose",
-                    "Wool - Celluloze", "Wool - Glucose"), each =2),
+                    "Cellobiose", "Glucose",
+                    "Broth - Cellobiose", "Broth - Glucose",
+                    "Glass - Cellobiose", "Glass - Glucose",
+                    "Wool - Cellobiose", "Wool - Glucose"), each =2),
                    Parameter=rep(c("Protein", "DNA"), times=12),
@@
 t.s5.1<-data.frame(Name=rep(c("All", "Broth", "Glass", "Wool",
-                    "Celluloze", "Glucose",
-                    "Broth - Celluloze", "Broth - Glucose",
-                    "Glass - Celluloze", "Glass - Glucose",
-                    "Wool - Celluloze", "Wool - Glucose"), each =2),
+                    "Cellobiose", "Glucose",
+                    "Broth - Cellobiose", "Broth - Glucose",
+                    "Glass - Cellobiose", "Glass - Glucose",
+                    "Wool - Cellobiose", "Wool - Glucose"), each =2),
                    Parameter=rep(c("Protein", "DNA"), times=12),
@@
 t.s6.1<-data.frame(Name=c("All", "Broth", "Glass", "Wool",
-                    "Celluloze", "Glucose",
-                    "Broth - Celluloze", "Broth - Glucose",
-                    "Glass - Celluloze", "Glass - Glucose",
-                    "Wool - Celluloze", "Wool - Glucose"),
+                    "Cellobiose", "Glucose",
+                    "Broth - Cellobiose", "Broth - Glucose",
+                    "Glass - Cellobiose", "Glass - Glucose",
+                    "Wool - Cellobiose", "Wool - Glucose"),
                    Vmax=deb_all_results_par[,1],

### 2019-02-28 SI_code.Rmd
cleaning

@@
 library(kableExtra)
+library(lmerTest)
+library(psycho)
+library(ggeffects)
 ```
@@
 ##Statistical analysis
-###Respiration
-Cumulative respiration is calculated from the measured respiration rate using the local polynomial regression.
+Linear Mixed-Effect models are used to evaluate the effect of substrate and the type of physical barrier on measured respiration rate, DNA and protein concentration.
+Becuase all three variables display significant variability over time, the time of measurement is treated as a random effect.
+Type of the barrier is nested within the substrate.
+
+```{r lme analysis}
+
+```
 

### 2019-02-28 data_analysis.R

@@ -1,475 +1,0 @@
-#################################################MinT#################################################
-##libraries
-library(deSolve)
-library(dplyr)
-library(FME)
-library(reshape)
-#library(reshape2)
-library(ggplot2)
-library(foreach)
-library(doParallel)
-library(openxlsx)
-library(DEoptim)
-library(gridExtra)
-
-##ggplot theme
-theme_min<-theme(axis.text.x=element_text(vjust=0.2, size=18, colour="black"),
-                 axis.text.y=element_text(hjust=0.2, size=18, colour="black"),
-                 axis.title=element_text(size=18, colour="black"),
-                 axis.line=element_line(size=0.5, colour="black"),
-                 strip.text=element_text(size=18, face="bold"),
-                 axis.ticks=element_line(size=1, colour="black"),
-                 axis.ticks.length=unit(-0.05, "cm"),
-                 panel.background=element_rect(colour="black", fill="white"),
-                 panel.grid=element_line(linetype=0),
-                 legend.text=element_text(size=14, colour="black"),
-                 legend.title=element_text(size=14, colour="black"),
-                 legend.position=c("right"),
-                 legend.key.size=unit(1, "cm"),
-                 strip.background=element_rect(fill="grey98", colour="black"),
-                 legend.key=element_rect(fill="white", size=1.2),
-                 legend.spacing=unit(0.5, "cm"),
-                 plot.title=element_text(size=18, face="bold", hjust=-0.05))
-
-#######################################################################################################
-
-##data 
-#respiration rates in umol/ml/h
-resp<-read.xlsx(xlsxFile = c("C:/Users/cape159/Documents/pracovni/data_statistika/minT/data_checked_respiration.xlsx"),
-                sheet = 1)
-
-#arrange the data
-resp.ordered<-resp[order(resp$Sample, resp$Day), ]
-
-#biological data
-biology<-read.xlsx(xlsxFile = c("C:/Users/cape159/Documents/pracovni/data_statistika/minT/data_checked_biology.xlsx"),
-                   sheet = 1)
-
-#arrange the data in the same way
-biology.ordered<-biology[order(biology$Sample, biology$Day), -c(2,3)]
-biology.ordered$Sample<-paste0("CSub3_", biology.ordered$Sample)
-
-#merging both data sets
[…]

### 2019-02-28 deb_i_all.R

@@ -1,256 +1,0 @@
-deb_i_all<-function(data, FACT){
-
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-
-
-  #if factor needs to be included, its levels are defined
-  if(FACT==1){
-
-    dat<-data
-    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
-
-  }else{
-
-    if(FACT==2){
-
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-
-    }else{
-
-      if(FACT==3){
-
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-
-
-      }else{
-
-        dat<-data
-        dat$id<-c(1)
-
-      }
-    }
-  }
-
-    #deb model
-    deriv<-function(time, state, pars){
-
-      with(as.list(c(state, pars)),{
-
-        #Carbon uptake
-        Cu=Vmax*C*S/(Km+C)
-
-        #maintnance
-        m=S*m0
-
-        #reserve mobilization rate for growth and enzyme production
-        an=f*R-m
-
-        #calibrated variables
-        #protein abundance in reserves and structures is mean value reported
-        #by Hanegraaf and Muller, 2001
-        r=m+pmax(an*(1-Yu), 0)
-
-        #Protc=0.7095*S+0.6085*R
-        Protc=fpr*R+fps*S
-        DNAc=fds*S
-
-        dR<-Cu-f*R
-        dS<-pmax(an*Yu,0)+pmin(0, an)
-        dC<--Cu
-
-
-        return(list(c(dR, dS, dC), r=r, Protc=Protc, DNAc=DNAc))
-
-      })
-    }
-    #define names of parameters
-    parnames<-c("Vmax", "Km", "m0", "f", "Yu", "fpr", "fps", "fds", "R_0", "S_0")
-
-    #parameters estimation function
-    estim<-function(odeset){
-
-      #defining cost function
-      cost<-function(x){
-
-        par<-x[1:length(parnames)]
-
-        names(par)<-parnames
-
[…]

### 2019-02-28 deb_i_all_fix_Yu.R

@@ -1,256 +1,0 @@
-deb_i_all_fix_Yu<-function(data, FACT){
-
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-
-
-  #if factor needs to be included, its levels are defined
-  if(FACT==1){
-
-    dat<-data
-    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
-
-  }else{
-
-    if(FACT==2){
-
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-
-    }else{
-
-      if(FACT==3){
-
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-
-
-      }else{
-
-        dat<-data
-        dat$id<-c(1)
-
-      }
-    }
-  }
-
-    #deb model
-    deriv<-function(time, state, pars){
-
-      with(as.list(c(state, pars)),{
-
-        #Carbon uptake
-        Cu=Vmax*C*S/(Km+C)
-
-        #maintnance
-        m=S*m0
-
-        #reserve mobilization rate for growth and enzyme production
-        an=f*R-m
-
-        #calibrated variables
-        #protein abundance in reserves and structures is mean value reported
-        #by Hanegraaf and Muller, 2001
-        r=m+pmax(an*(1-0.6944324), 0)
-
-        #Protc=0.7095*S+0.6085*R
-        Protc=fpr*R+fps*S
-        DNAc=fds*S
-
-        dR<-Cu-f*R
-        dS<-pmax(an*0.6944324,0)+pmin(0, an)
-        dC<--Cu
-
-
-        return(list(c(dR, dS, dC), r=r, Protc=Protc, DNAc=DNAc))
-
-      })
-    }
-    #define names of parameters
-    parnames<-c("Vmax", "Km", "m0", "f", "fpr", "fps", "fds")
-
-    #parameters estimation function
-    estim<-function(odeset){
-
-      #defining cost function
-      cost<-function(x){
-
-        par<-x[1:length(parnames)]
-
-        names(par)<-parnames
-
-        #first, pars dependent output from ode is matched with measured values
[…]

### 2019-02-28 deb_r.R

@@ -1,256 +1,0 @@
-deb_r<-function(data, FACT){
-
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-
-
-  #if factor needs to be included, its levels are defined
-  if(FACT==1){
-
-    dat<-data
-    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
-
-  }else{
-
-    if(FACT==2){
-
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-
-    }else{
-
-      if(FACT==3){
-
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-
-
-      }else{
-
-        dat<-data
-        dat$id<-c(1)
-
-      }
-    }
-  }
-
-    #deb model
-    deriv<-function(time, state, pars){
-
-      with(as.list(c(state, pars)),{
-
-        #Carbon uptake
-        Cu=Vmax*C*S/(Km+C)
-
-        #maintnance
-        m=S*m0
-
-        #reserve mobilization rate for growth and enzyme production
-        an=f*R-m
-
-        #calibrated variables
-        #protein abundance in reserves and structures is mean value reported
-        #by Hanegraaf and Muller, 2001
-        r=m+pmax(an*(1-Yu), 0)
-
-        Protc=0.7095*S+0.6085*R
-        #Protc=fps*S+fpr*R
-
-        dR<-Cu-f*R
-        dS<-pmax(an*Yu,0)+pmin(0, an)
-        dC<--Cu
-
-
-        return(list(c(dR, dS, dC), r=r, Protc=Protc))
-
-      })
-    }
-    #define names of parameters
-    parnames<-c("Vmax", "Km", "m0", "f", "Yu", "R_0", "S_0")
-
-    #parameters estimation function
-    estim<-function(odeset){
-
-      #defining cost function
-      cost<-function(x){
-
-        par<-x[1:length(parnames)]
-
-        names(par)<-parnames
-
-        #first, pars dependent output from ode is matched with measured values
[…]

### 2019-02-28 mend_i.R

@@ -1,247 +1,0 @@
-mend_i<-function(data, FACT){
-  
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-  
-  if(FACT==1){
-    
-    dat<-data
-    dat$id<-ifelse(d$Substrate=="Glucose", 1, 2)
-    
-  }else{
-    
-    if(FACT==2){
-      
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-      
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-      
-    }else{
-      
-      if(FACT==3){
-        
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-        
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-        
-        
-      }else{
-        
-        dat<-data
-        dat$id<-c(1)
-        
-      }
-    }
-  }
-  
-    #mend model
-    deriv<-function(time, state, pars){
-      
-      with(as.list(c(state, pars)),{
-        
-        #equations
-        #C uptake
-        F1=1/CUE*(Vmax+mr)*Cmic*C/(Km+C)
-        #growth respiration
-        F4=(1/CUE-1)*Vmax*Cmic*C/(Km+C)
-        #maintnance respiration
-        F5=(1/CUE-1)*mr*Cmic*C/(Km+C)
-        #microbial mortality
-        F8=mr*Cmic
-        
-        Protc=fp*Cmic
-        
-        #states
-        dCmic<-F1-(F4+F5)-F8
-        dC<--F1+F8
-        
-        return(list(c(dCmic, dC), r=F4+F5, Protc=Protc))
-        
-      })
-    }
-    #define names of parameters
-    parnames<-c("Vmax", "Km", "CUE", "mr", "fp", "Cmic_0")
-    
-    #parameters estimation function
-    estim<-function(odeset){
-      
-      #defining cost function
-      cost<-function(x){
-        
-        par<-x[1:length(parnames)]
-        
-        names(par)<-parnames
-        
-        #first, pars dependent output from ode is matched with measured values
[…]

### 2019-02-28 mend_i_DNA.R

@@ -1,247 +1,0 @@
-mend_i_DNA<-function(data, FACT){
-  
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-  
-  if(FACT==1){
-    
-    dat<-data
-    dat$id<-ifelse(d$Substrate=="Glucose", 1, 2)
-    
-  }else{
-    
-    if(FACT==2){
-      
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-      
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-      
-    }else{
-      
-      if(FACT==3){
-        
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-        
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-        
-        
-      }else{
-        
-        dat<-data
-        dat$id<-c(1)
-        
-      }
-    }
-  }
-  
-    #mend model
-    deriv<-function(time, state, pars){
-      
-      with(as.list(c(state, pars)),{
-        
-        #equations
-        #C uptake
-        F1=1/CUE*(Vmax+mr)*Cmic*C/(Km+C)
-        #growth respiration
-        F4=(1/CUE-1)*Vmax*Cmic*C/(Km+C)
-        #maintnance respiration
-        F5=(1/CUE-1)*mr*Cmic*C/(Km+C)
-        #microbial mortality
-        F8=mr*Cmic
-        
-        DNAc=fd*Cmic
-        
-        #states
-        dCmic<-F1-(F4+F5)-F8
-        dC<--F1+F8
-        
-        return(list(c(dCmic, dC), r=F4+F5, DNAc=DNAc))
-        
-      })
-    }
-    #define names of parameters
-    parnames<-c("Vmax", "Km", "CUE", "mr", "fd", "Cmic_0")
-    
-    #parameters estimation function
-    estim<-function(odeset){
-      
-      #defining cost function
-      cost<-function(x){
-        
-        par<-x[1:length(parnames)]
-        
-        names(par)<-parnames
-        
-        #first, pars dependent output from ode is matched with measured values
[…]

### 2019-02-28 mend_r.R

@@ -1,246 +1,0 @@
-mend_r<-function(data, FACT){
-  
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-  
-  if(FACT==1){
-    
-    dat<-data
-    dat$id<-ifelse(d$Substrate=="Glucose", 1, 2)
-    
-  }else{
-    
-    if(FACT==2){
-      
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-      
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-      
-    }else{
-      
-      if(FACT==3){
-        
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-        
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-        
-        
-      }else{
-        
-        dat<-data
-        dat$id<-c(1)
-        
-      }
-    }
-  }
-  
-    #mend model
-    deriv<-function(time, state, pars){
-      
-      with(as.list(c(state, pars)),{
-        
-        #equations
-        #C uptake
-        F1=1/CUE*(Vmax+mr)*Cmic*C/(Km+C)
-        #growth respiration
-        F4=(1/CUE-1)*Vmax*Cmic*C/(Km+C)
-        #maintnance respiration
-        F5=(1/CUE-1)*mr*Cmic*C/(Km+C)
-        #microbial mortality
-        F8=mr*Cmic
-        
-        #states
-        dCmic<-F1-(F4+F5)-F8
-        dC<--F1+F8
-        
-        return(list(c(dCmic, dC), r=F4+F5, Protc=0.548*Cmic))
-        
-      })
-    }
-    #define names of parameters
-    parnames<-c("Vmax", "Km", "CUE", "mr", "Cmic_0")
-    
-    #parameters estimation function
-    estim<-function(odeset){
-      
-      #defining cost function
-      cost<-function(x){
-        
-        par<-x[1:length(parnames)]
-        
-        names(par)<-parnames
-        
-        #first, pars dependent output from ode is matched with measured values
-        yhat_all<-as.data.frame(ode(y=c(Cmic=par[["Cmic_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))
[…]

### 2019-02-28 monod_i.R

@@ -1,240 +1,0 @@
-###monod growth function
-monod_i<-function(data, FACT){
-  
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-  
-  if(FACT==1){
-    
-    dat<-data
-    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
-    
-  }else{
-    
-    if(FACT==2){
-      
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-      
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-      
-    }else{
-      
-      if(FACT==3){
-        
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-        
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-        
-        
-      }else{
-        
-        dat<-data
-        dat$id<-c(1)
-        
-      }
-    }
-  }
-  
-  
-    #monod growth function
-    deriv<-function(time, state, pars){
-      
-      with(as.list(c(state, pars)),{
-        
-        Protc=fp*Cmic
-        
-        dCmic<--k*Cmic+CUE*Vmax*Cmic*C/(Km+C)
-        dC<-k*Cmic-Vmax*Cmic*C/(Km+C)
-        
-        return(list(c(dCmic, dC), r=(1-CUE)*Vmax*Cmic*C/(Km+C), Protc=Protc))
-        
-      })
-    }
-    
-    #define names of parameters
-    parnames<-c("Vmax", "Km", "CUE", "k", "fp", "Cmic_0")
-    
-    #parameters estimation function
-    estim<-function(odeset){
-      
-      #defining cost function
-      cost<-function(x){
-        
-        par<-x[1:length(parnames)]
-        
-        names(par)<-parnames
-        
-        #first, pars dependent output from ode is matched with measured values
-        yhat_all<-as.data.frame(ode(y=c(Cmic=par[["Cmic_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))
-        
-        #select time and the measured variables 
-        yhat<-select(yhat_all, c("time", "r", "Protc"))
-        
-        #reformat to long format data frame
[…]

### 2019-02-28 monod_i_DNA.R

@@ -1,240 +1,0 @@
-###monod growth function
-monod_i_DNA<-function(data, FACT){
-
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-
-  if(FACT==1){
-
-    dat<-data
-    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
-
-  }else{
-
-    if(FACT==2){
-
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-
-    }else{
-
-      if(FACT==3){
-
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-
-
-      }else{
-
-        dat<-data
-        dat$id<-c(1)
-
-      }
-    }
-  }
-
-
-    #monod growth function
-    deriv<-function(time, state, pars){
-
-      with(as.list(c(state, pars)),{
-
-        DNAc=fd*Cmic
-
-        dCmic<--k*Cmic+CUE*Vmax*Cmic*C/(Km+C)
-        dC<-k*Cmic-Vmax*Cmic*C/(Km+C)
-
-        return(list(c(dCmic, dC), r=(1-CUE)*Vmax*Cmic*C/(Km+C), DNAc=DNAc))
-
-      })
-    }
-
-    #define names of parameters
-    parnames<-c("Vmax", "Km", "CUE", "k", "fd", "Cmic_0")
-
-    #parameters estimation function
-    estim<-function(odeset){
-
-      #defining cost function
-      cost<-function(x){
-
-        par<-x[1:length(parnames)]
-
-        names(par)<-parnames
-
-        #first, pars dependent output from ode is matched with measured values
-        yhat_all<-as.data.frame(ode(y=c(Cmic=par[["Cmic_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))
-
-        #select time and the measured variables
-        yhat<-select(yhat_all, c("time", "r", "DNAc"))
-
-        #reformat to long format data frame
-        Yhat<-melt(yhat, id.vars = "time")
-
[…]

### 2019-02-28 monod_i_DNA_fixed.R

@@ -1,240 +1,0 @@
-###monod growth function
-monod_i_DNA_fixed<-function(data, FACT){
-
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-
-  if(FACT==1){
-
-    dat<-data
-    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
-
-  }else{
-
-    if(FACT==2){
-
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-
-    }else{
-
-      if(FACT==3){
-
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-
-
-      }else{
-
-        dat<-data
-        dat$id<-c(1)
-
-      }
-    }
-  }
-
-
-    #monod growth function
-    deriv<-function(time, state, pars){
-
-      with(as.list(c(state, pars)),{
-
-        DNAc=fd*Cmic
-
-        dCmic<--k*Cmic+CUE*Vmax*Cmic*C/(Km+C)
-        dC<-k*Cmic-Vmax*Cmic*C/(Km+C)
-
-        return(list(c(dCmic, dC), r=(1-CUE)*Vmax*Cmic*C/(Km+C), DNAc=DNAc))
-
-      })
-    }
-
-    #define names of parameters
-    parnames<-c("Vmax", "Km", "CUE", "k", "fd")
-
-    #parameters estimation function
-    estim<-function(odeset){
-
-      #defining cost function
-      cost<-function(x){
-
-        par<-x[1:length(parnames)]
-
-        names(par)<-parnames
-
-        #first, pars dependent output from ode is matched with measured values
-        yhat_all<-as.data.frame(ode(y=c(Cmic=0.05943, C=25), parms=par, deriv, times=sort(odeset$Time)))
-
-        #select time and the measured variables
-        yhat<-select(yhat_all, c("time", "r", "DNAc"))
-
-        #reformat to long format data frame
-        Yhat<-melt(yhat, id.vars = "time")
-
-        #add the measured data to a data frame
[…]

### 2019-02-28 monod_r.R

@@ -1,239 +1,0 @@
-###monod growth function
-monod_r<-function(data, FACT){
-  
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-  
-  if(FACT==1){
-    
-    dat<-data
-    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
-    
-  }else{
-    
-    if(FACT==2){
-      
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-      
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-      
-    }else{
-      
-      if(FACT==3){
-        
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-        
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-        
-        
-      }else{
-        
-        dat<-data
-        dat$id<-c(1)
-        
-      }
-    }
-  }
-  
-  
-    #monod growth function
-    deriv<-function(time, state, pars){
-      
-      with(as.list(c(state, pars)),{
-        
-        dCmic<--k*Cmic+CUE*Vmax*Cmic*C/(Km+C)
-        dC<-k*Cmic-Vmax*Cmic*C/(Km+C)
-        
-        return(list(c(dCmic, dC), r=(1-CUE)*Vmax*Cmic*C/(Km+C), Protc=0.548*Cmic))
-        
-      })
-    }
-    
-    #define names of parameters
-    parnames<-c("Vmax", "Km", "CUE", "k", "Cmic_0")
-    
-    #parameters estimation function
-    estim<-function(odeset){
-      
-      #defining cost function
-      cost<-function(x){
-        
-        par<-x[1:length(parnames)]
-        
-        names(par)<-parnames
-        
-        #first, pars dependent output from ode is matched with measured values
-        yhat_all<-as.data.frame(ode(y=c(Cmic=par[["Cmic_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))
-        
-        #select time and the measured variables 
-        yhat<-select(yhat_all, c("time", "r", "Protc"))
-        
-        #reformat to long format data frame
-        Yhat<-melt(yhat, id.vars = "time")
-        
-        #add the measured data to a data frame
[…]

### 2019-03-01 SI_code.Rmd
statistical analysis added

@@
 library(ggeffects)
+library(rcompanion)
+
+##ggplot theme
+theme_min<-theme(axis.text.x=element_text(vjust=0.2, size=18, colour="black"),
+                 axis.text.y=element_text(hjust=0.2, size=18, colour="black"),
+                 axis.title=element_text(size=18, colour="black"),
+                 axis.line=element_line(size=0.5, colour="black"),
+                 strip.text=element_text(size=18, face="bold"),
+                 axis.ticks=element_line(size=1, colour="black"),
+                 axis.ticks.length=unit(-0.05, "cm"),
+                 panel.background=element_rect(colour="black", fill="white"),
+                 panel.grid=element_line(linetype=0),
+                 legend.text=element_text(size=14, colour="black"),
+                 legend.title=element_text(size=14, colour="black"),
+                 legend.position=c("right"),
+                 legend.key.size=unit(1, "cm"),
+                 strip.background=element_rect(fill="grey98", colour="black"),
+                 legend.key=element_rect(fill="white", size=1.2),
+                 legend.spacing=unit(0.5, "cm"),
+                 plot.title=element_text(size=18, face="bold", hjust=-0.05))
 ```
@@
 
-DNA and protein concentrations are in $mu$grams per incubation vial.
+DNA and protein concentrations are in $\mu$grams per incubation vial.
 Both values are recalculated to unit of carbon per one ml of media in the vial (4 ml) using the conversion factors reported in Vrede et al.
@@
 
-```{r lme analysis}
+```{r lme analysis, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+#respiration rate
+lme1<-lmer(r~Substrate+Structure/Substrate + (1|Day), d)
+#Proteins
+lme2<-lmer(Protc~Substrate+Structure/Substrate + (1|Day), d)
+#DNA
+lme3<-lmer(DNAc~Substrate+Structure/Substrate + (1|Day), d)
+
+t.s0<-as.data.frame(cbind(anova(lme1)[c(5,3,6)], anova(lme3)[c(5,3,6)], anova(lme3)[c(5,3,6)]))
+
+row.names(t.s0)<-c("Substrate", "Structure", "Substrate*Structure")
+
+kable(t.s0, digits = c(2, 1, 3, 2, 1, 3, 2, 1, 3), align = 'c', "html", booktabs=T, col.names = c("F", "df", "p", "F", "df", "p", "F", "df", "p"), row.names=T, caption = "_**Table S1:** Results of linear mixed-effects model analysis.
+Effect of substrate, the type physical barrier (Structure) and their interaction was evaluated.
+Statistical analysis was conducted for measured respiration rate, cellular protien concentration and DNA.
+F values, degrees of freedom (df) and the probabality (p) are reported._") %>%
+  kable_styling() %>% 
+  add_header_above(c(" " = 1, "Respiration rate" = 3, "Cellular Proteins" = 3, "DNA" = 3),
+                   italic=T)
 
 ```
-
+$~$
+$~$
+$~$
+$~$
+$~$
+$~$
+
+```{r lme effects, echo=FALSE, fig.height=5, fig.width=15, message=FALSE, paged.print=FALSE, fig.cap="Fig.
+S1: Marginal effects of the substrate (Glucose or Cellobiose) and the type of physical barrier (BROTH, GLASS, WOOL) on the respiration rate (a), cellular proteins (b) and DNA concentration (c) estimated by the linear mixed-effect model."}
+#respiration
+lme1_contr<-get_contrasts(lme1)
+lme1_effects<-ggpredict(lme1, c("Substrate", "Structure"), type="fe")
+lme1_effects$x<-c(rep("Cellobiose", 3), rep("Glucose", 3))
+lme1_effects$group<-c(rep(c("BROTH", "WOOL", "GLASS"), 2))
+lme1_effects<-lme1_effects[order(lme1_effects$group), ]
+lme1_effects$letters<-c(as.character(cldList(p ~ Contrast, data      = lme1_contr, threshold = 0.05)[[2]])[c(1:2, 5:6)], as.character(cldList(p ~ Contrast, data      = lme1_contr, threshold = 0.05)[[2]])[c(3:4)])
+#Proteins
+lme2_contr<-get_contrasts(lme2)
+lme2_effects<-ggpredict(lme2, c("Substrate", "Structure"), type="fe")
+lme2_effects$x<-c(rep("Cellobiose", 3), rep("Glucose", 3))
+lme2_effects$group<-c(rep(c("BROTH", "WOOL", "GLASS"), 2))
+lme2_effects<-lme2_effects[order(lme2_effects$group), ]
+lme2_effects$letters<-c(as.character(cldList(p ~ Contrast, data      = lme2_contr, threshold = 0.05)[[2]])[c(1:2, 5:6)], as.character(cldList(p ~ Contrast, data      = lme2_contr, threshold = 0.05)[[2]])[c(3:4)])
+#DNA
+lme3_contr<-get_contrasts(lme3)
+lme3_effects<-ggpredict(lme3, c("Substrate", "Structure"), type="fe")
+lme3_effects$x<-c(rep("Cellobiose", 3), rep("Glucose", 3))
+lme3_effects$group<-c(rep(c("BROTH", "WOOL", "GLASS"), 2))
+lme3_effects<-lme3_effects[order(lme3_effects$group), ]
+lme3_effects$letters<-c(as.character(cldList(p ~ Contrast, data      = lme3_contr, threshold = 0.05)[[2]])[c(1:2, 5:6)], as.character(cldList(p ~ Contrast, data      = lme3_contr, threshold = 0.05)[[2]])[c(3:4)])
+
+p1<-ggplot(lme1_effects, aes(x, predicted))+
+    geom_point(position = position_dodge(width = 0.5), aes(fill=group), shape=21, cex=6, show.legend = F)+
+    scale_fill_manual(values = c("black", "grey", "white"))+
+    scale_colour_manual(values = c("black", "black", "black"))+
+    geom_errorbar(aes(ymin=predicted-std.error, ymax=predicted+std.error, colour=group),
+                  position=position_dodge(width = 0.5), width=0.1, show.legend = F)+
+    ylab(expression(paste("Respiration rate (", mu, "mol",(C-CO[2])~ml^{-1}~h^{-1}, ")")))+
+    ggtitle("a)")+
+    geom_text(aes(label=letters, colour=group, vjust=-10, fontface="bold.italic"),
+              cex=5, position=position_dodge(width = 0.5), show.legend = F)+
+    ylim(-0.01,0.3)+
+    theme_min+theme(axis.title.x = element_blank(),
+                    legend.title = element_blank())
+p2<-ggplot(lme2_effects, aes(x, predicted))+
+    geom_point(position = position_dodge(width = 0.5), aes(fill=group), shape=21, cex=6, show.legend = F)+
+    scale_fill_manual(values = c("black", "grey", "white"))+
+    scale_colour_manual(values = c("black", "black", "black"))+
+    geom_errorbar(aes(ymin=predicted-std.error, ymax=predicted+std.error, colour=group),
+                  position=position_dodge(width = 0.5), width=0.1, show.legend = F)+
+    ylab(expression(paste("Cellular Proteins (", mu, "mol",(C[Proteins])~ml^{-1}, ")")))+
+    ggtitle("b)")+
+    geom_text(aes(label=letters, colour=group, vjust=-3, fontface="bold.italic"),
+              cex=5, position=position_dodge(width = 0.5), show.legend = F)+
+    ylim(0,1.2)+
+    theme_min+theme(axis.title.x = element_blank(),
+                    legend.title = element_blank())
+p3<-ggplot(lme3_effects, aes(x, predicted))+
+    geom_point(position = position_dodge(width = 0.5), aes(fill=group), shape=21, cex=6)+
+    scale_fill_manual(values = c("black", "grey", "white"))+
+    scale_colour_manual(values = c("black", "black", "black"))+
+    geom_errorbar(aes(ymin=predicted-std.error, ymax=predicted+std.error, colour=group),
+                  position=position_dodge(width = 0.5), width=0.1)+
+    ylab(expression(paste("DNA (", mu, "mol",(C[DNA])~ml^{-1}, ")")))+
+    ggtitle("c)")+
+    geom_text(aes(label=letters, colour=group, vjust=-3, fontface="bold.italic"),
+              cex=5, position=position_dodge(width = 0.5), show.legend = F)+
+    ylim(0,0.08)+
+    theme_min+theme(axis.title.x = element_blank(),
+                    legend.title = element_blank(),
+                    legend.position = c(0.25,0.2))
+
+grid.arrange(p1, p2, p3, nrow=1)
+
+```
+
+##Microbial-explicit models analysis
 ###All models are first calibrated against respiration rate
@@
 
-kable(t.s1.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$'), row.names=T, caption = "_**Table S1:** Monod model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s1.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$'), row.names=T, caption = "_**Table S2:** Monod model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(monod_r_results_good, digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S2:** Goodness of fit of monod model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(monod_r_results_good, digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S3:** Goodness of fit of monod model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(t.s2.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$'), row.names=T, caption = "_**Table S3:** MEND model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s2.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$'), row.names=T, caption = "_**Table S4:** MEND model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(mend_r_results_good, digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S4:** Goodness of fit of MEND model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(mend_r_results_good, digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S5:** Goodness of fit of MEND model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(t.s3.1, digits = c(2, 2, 3, 2, 2), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', '$\\m_{R}$', '$\\f_{0}$', '$\\Y_{S}$'), row.names=T, caption = "_**Table S5:** DEB model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s3.1, digits = c(2, 2, 3, 2, 2), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', '$\\m_{R}$', '$\\f_{0}$', '$\\Y_{S}$'), row.names=T, caption = "_**Table S6:** DEB model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(deb_r_results_good, digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S6:** Goodness of fit of DEB model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(deb_r_results_good, digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S7:** Goodness of fit of DEB model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(t.s4.1, digits = c(0, 0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" "," ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$', 'Conversion factor'), row.names=F, caption = "_**Table S7:** Monod model parameters calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s4.1, digits = c(0, 0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" "," ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$', 'Conversion factor'), row.names=F, caption = "_**Table S8:** Monod model parameters calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(monod_all_results_good, digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S8:** Goodness of fit of Monod model calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(monod_all_results_good, digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S9:** Goodness of fit of Monod model calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(t.s5.1, digits = c(0, 0, 2, 2, 2, 3, 2), align = 'c', "html", booktabs=T, col.names = c(" "," ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$', 'Conversion factor'), row.names=F, caption = "_**Table S9:** MEND model parameters calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s5.1, digits = c(0, 0, 2, 2, 2, 3, 2), align = 'c', "html", booktabs=T, col.names = c(" "," ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$', 'Conversion factor'), row.names=F, caption = "_**Table S10:** MEND model parameters calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(mend_all_results_good, digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S10:** Goodness of fit of MEND model calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(mend_all_results_good, digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S11:** Goodness of fit of MEND model calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-deb_all_results_par<-read.csv("deb_all_results_par")
-deb_all_results_good<-read.csv("deb_all_results_good")
+deb_all_results_par<-read.csv("deb_all_results_par.csv")
+deb_all_results_good<-read.csv("deb_all_results_good.csv")
 
@@
 
-kable(t.s6.1, digits = c(0, 2, 2, 3, 2, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" ",'$\\V_{MAX}$', '$\\K_{M}$', '$\\m_{R}$', '$\\f_{0}$', '$\\Y_{S}$', '$\\R_{Proteins}$', '$\\R_{DNA}$', '$\\"S"_{DNA}$'), row.names=F, caption = "_**Table S11:** DEB model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s6.1, digits = c(0, 2, 2, 3, 2, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" ",'$\\V_{MAX}$', '$\\K_{M}$', '$\\m_{R}$', '$\\f_{0}$', '$\\Y_{S}$', '$\\R_{Proteins}$', '$\\S_{Proteins}$', '$\\S_{DNA}$'), row.names=F, caption = "_**Table S12:** DEB model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(deb_all_results_good, digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S12:** Goodness of fit of DEB model calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(deb_all_results_good, digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S13:** Goodness of fit of DEB model calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%

### 2019-03-11 SI_code.Rmd
...

@@
 ---
-title: Variable Composition of Microbial Biomass Affects the Accuracy of Microbial
-  Explicit Models Simulation
+title: The effect of restricted substrate diffusion on microbial community function and structure
 author: "Petr Capek, Katherine Todd-Brown, Natalie Sadler, Nancy Hess, Kirsten Hofmockel"
@@
 library(rcompanion)
+library(vegan)
 
@@
 
+#Bacteria
+bac<-as.data.frame(t(read.csv(file=c("bacteria.csv"), skip = 4)[,-1]))
+colnames(bac)<-t(read.csv(file=c("bacteria.csv"), skip = 4)[,1])
+
+bac_env<-as.data.frame(t(read.csv(file=c("bacteria.csv"), nrows = 4, header = F)[,-1]))
+colnames(bac_env)<-t(read.csv(file=c("bacteria.csv"), nrows = 4, header = F)[,1])
+
+bac_env$Structure<-recode(bac_env$Structure, 'Broth'="BROTH", 'Glass wool'="WOOL",
+                          'Mixed glass'="GLASS")
+
+#Fungi
+fungi<-as.data.frame(t(read.csv(file=c("fungi.csv"), skip = 4)[,-1]))
+colnames(fungi)<-t(read.csv(file=c("fungi.csv"), skip = 4)[,1])
+
+fungi_env<-as.data.frame(t(read.csv(file=c("fungi.csv"), nrows = 4, header = F)[,-1]))
+colnames(fungi_env)<-t(read.csv(file=c("fungi.csv"), nrows = 4, header = F)[,1])
+
+fungi_env$Structure<-recode(fungi_env$Structure, 'Broth'="BROTH", 'Glass wool'="WOOL",
+                          'Mixed glass'="GLASS")
+
 ```
@@
 
-t.s0<-as.data.frame(cbind(anova(lme1)[c(5,3,6)], anova(lme3)[c(5,3,6)], anova(lme3)[c(5,3,6)]))
+t.s0<-as.data.frame(cbind(anova(lme1)[c(5,3,6)], anova(lme2)[c(5,3,6)], anova(lme3)[c(5,3,6)]))
 
@@
 kable(t.s0, digits = c(2, 1, 3, 2, 1, 3, 2, 1, 3), align = 'c', "html", booktabs=T, col.names = c("F", "df", "p", "F", "df", "p", "F", "df", "p"), row.names=T, caption = "_**Table S1:** Results of linear mixed-effects model analysis.
-Effect of substrate, the type physical barrier (Structure) and their interaction was evaluated.
-Statistical analysis was conducted for measured respiration rate, cellular protien concentration and DNA.
-F values, degrees of freedom (df) and the probabality (p) are reported._") %>%
+Effect of substrate, the type of physical barrier (Structure) and their interaction was evaluated.
+Statistical analysis was conducted for measured respiration rate, cellular protein concentration and DNA.
+F values, degrees of freedom (df) and the probability (p) are reported._") %>%
   kable_styling() %>% 
@@
 ```
+$~$
+$~$
+$~$
+Next, the effect of the substrate and the presence of a physical barrier on microbial community structure and diversity is evaluated.
+First, the jaccard distance indices are calculated on a normalized data.
+Then, Permutational Multivariate Analysis of Variance is used to estimate the treatment effects.
+Analysis permutations are constrained to an individual times of incubation to avoid the potential confounding effect of time.
+
+```{r vegan analysis, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+#Bacteria
+bac.norm<-decostand(bac, method=c("normalize"))
+bac.dist<-vegdist(bac.norm, method = "jaccard")
+bac.ad<-adonis(bac.dist~Substrate+Structure/Substrate, bac_env, strata = bac_env$Day)
+
+#Fungi
+fungi.norm<-decostand(fungi, method=c("normalize"))
+fungi.dist<-vegdist(fungi.norm, method = "jaccard")
+fungi.ad<-adonis(fungi.dist~Substrate+Structure/Substrate, fungi_env, strata = fungi_env$Day)
+
+t.s01<-cbind(as.data.frame(bac.ad$aov[c(4, 1, 6)])[c(1:3), ], as.data.frame(fungi.ad$aov[c(4, 1, 6)])[c(1:3), ])
+
+row.names(t.s01)<-c("Substrate", "Structure", "Substrate*Structure")
+
+kable(t.s01, digits = c(2, 1, 3, 2, 1, 3), align = 'c', "html", booktabs=T, col.names = c("F", "df", "p", "F", "df", "p"), row.names=T, caption = "_**Table S2:** Results of permutational multivariate analysis of variance.
+Effect of substrate, the type of physical barrier (Structure) and their interaction was evaluated.
+Statistical analysis was conducted for bacterial and fungal part of the microbial community separately.
+F values, degrees of freedom (df) and the probability (p) are reported._") %>%
+  kable_styling() %>% 
+  add_header_above(c(" " = 1, "Bacteria" = 3, "Fungi" = 3),
+                   italic=T)
+
+```
+$~$
+$~$
+$~$
+$~$
+$~$
+$~$
+```{r community figure, echo=FALSE, fig.height=5, fig.width=10, fig.cap="Fig.
+S2: Multivariate distances between six experimental groups that combine two different substrate treatments (Glucose or Cellobiose) and three different types of physical barrier (BROTH, WOOL, GLASS - black, grey and empty symbols respectively) in respect to structure of bacterial ((a); 16S) and fungal ((b); ITS) part of microbial community."}
+#Bacteria
+bacm<-as.matrix(bac.dist)
+#Intercept - Glucose - BROTH
+mGB<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
+       intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Glucose"))]
+mGB[mGB==0]<-NA
+#Glucose - GLASS
+mGG<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
+       intersect(which(bac_env$Structure=="GLASS"),which(bac_env$Substrate=="Glucose"))]
+mGG[mGG==0]<-NA
+#Glucose - WOOL
+mGW<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
+       intersect(which(bac_env$Structure=="WOOL"),which(bac_env$Substrate=="Glucose"))]
+mGW[mGW==0]<-NA
+#Cellobiose - BROTH
+mCB<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
+       intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose"))]
+mCB[mCB==0]<-NA
+#Cellobiose - GLASS
+mCG<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
+       intersect(which(bac_env$Structure=="GLASS"),which(bac_env$Substrate=="Cellobiose"))]
+mCG[mCG==0]<-NA
+#Cellobiose - WOOL
+mCW<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
+       intersect(which(bac_env$Structure=="WOOL"),which(bac_env$Substrate=="Cellobiose"))]
+mCW[mCW==0]<-NA
+
+bacdiff<-data.frame(Dist=c(mean(mGB, na.rm=T), mean(mGG, na.rm=T), mean(mGW, na.rm=T), 0, mean(mCG, na.rm=T), mean(mCW, na.rm=T)),
+                    Dist.sd=c(sd(mGB, na.rm=T), sd(mGG, na.rm=T), sd(mGW, na.rm=T), sd(mCB, na.rm=T), sd(mCG, na.rm=T), sd(mCW, na.rm=T)),
+                    Substrate=c(rep("Glucose", 3), rep("Cellobiose", 3)),
+                    Structure=rep(c("BROTH", "GLASS", "WOOL"), 2))
+
+#Fungi
+fm<-as.matrix(fungi.dist)
+#Intercept - Glucose - BROTH
+fGB<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
+          intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Glucose"))]
+fGB[fGB==0]<-NA
+#Glucose - GLASS
+fGG<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
+          intersect(which(fungi_env$Structure=="GLASS"),which(fungi_env$Substrate=="Glucose"))]
+fGG[fGG==0]<-NA
+#Glucose - WOOL
+fGW<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
+          intersect(which(fungi_env$Structure=="WOOL"),which(fungi_env$Substrate=="Glucose"))]
+fGW[fGW==0]<-NA
+#Cellobiose - BROTH
+fCB<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
+          intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose"))]
+fCB[fCB==0]<-NA
+#Cellobiose - GLASS
+fCG<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
+          intersect(which(fungi_env$Structure=="GLASS"),which(fungi_env$Substrate=="Cellobiose"))]
+fCG[fCG==0]<-NA
+#Cellobiose - WOOL
+fCW<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
+          intersect(which(fungi_env$Structure=="WOOL"),which(fungi_env$Substrate=="Cellobiose"))]
+fCW[fCW==0]<-NA
+
+fdiff<-data.frame(Dist=c(mean(fGB, na.rm=T), mean(fGG, na.rm=T), mean(fGW, na.rm=T), 0, mean(fCG, na.rm=T), mean(fCW, na.rm=T)),
+                    Dist.sd=c(sd(fGB, na.rm=T), sd(fGG, na.rm=T), sd(fGW, na.rm=T), sd(fCB, na.rm=T), sd(fCG, na.rm=T), sd(fCW, na.rm=T)),
+                    Substrate=c(rep("Glucose", 3), rep("Cellobiose", 3)),
+                    Structure=rep(c("BROTH", "GLASS", "WOOL"), 2))
+
+bacplot<-ggplot(bacdiff, aes(Substrate, Dist))+
+  geom_point(position = position_dodge(width = 0.5), aes(fill=Structure), shape=21, cex=6)+
+  scale_fill_manual(values = c("black", "grey", "white"))+
+  scale_colour_manual(values = c("black", "black", "black"))+
+  geom_errorbar(aes(ymin=Dist-Dist.sd, ymax=Dist+Dist.sd, colour=Structure),
+                position=position_dodge(width = 0.5), width=0.1)+
+  ylab("16S Distance")+
+  ggtitle("a)")+
+  theme_min+theme(axis.title.x = element_blank(),
+                  legend.title = element_blank(),
+                  legend.position = c(0.8, 0.2))
+fplot<-ggplot(fdiff, aes(Substrate, Dist))+
+  geom_point(position = position_dodge(width = 0.5), aes(fill=Structure), shape=21, cex=6, show.legend = F)+
+  scale_fill_manual(values = c("black", "grey", "white"))+
+  scale_colour_manual(values = c("black", "black", "black"))+
+  geom_errorbar(aes(ymin=Dist-Dist.sd, ymax=Dist+Dist.sd, colour=Structure),
+                position=position_dodge(width = 0.5), width=0.1, show.legend = F)+
+  ylab("ITS Distance")+
+  ggtitle("b)")+
+  theme_min+theme(axis.title.x = element_blank(),
+                  legend.title = element_blank())
+
+grid.arrange(bacplot, fplot, nrow=1)
+
+```
+
+$~$
+$~$
+$~$
 
@@
 
-kable(t.s1.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$'), row.names=T, caption = "_**Table S2:** Monod model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s1.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$'), row.names=T, caption = "_**Table S3:** Monod model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(monod_r_results_good, digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S3:** Goodness of fit of monod model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(monod_r_results_good, digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S4:** Goodness of fit of monod model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(t.s2.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$'), row.names=T, caption = "_**Table S4:** MEND model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s2.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$'), row.names=T, caption = "_**Table S5:** MEND model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(mend_r_results_good, digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S5:** Goodness of fit of MEND model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(mend_r_results_good, digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S6:** Goodness of fit of MEND model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(t.s3.1, digits = c(2, 2, 3, 2, 2), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', '$\\m_{R}$', '$\\f_{0}$', '$\\Y_{S}$'), row.names=T, caption = "_**Table S6:** DEB model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s3.1, digits = c(2, 2, 3, 2, 2), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', '$\\m_{R}$', '$\\f_{0}$', '$\\Y_{S}$'), row.names=T, caption = "_**Table S7:** DEB model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(deb_r_results_good, digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S7:** Goodness of fit of DEB model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(deb_r_results_good, digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S8:** Goodness of fit of DEB model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(t.s4.1, digits = c(0, 0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" "," ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$', 'Conversion factor'), row.names=F, caption = "_**Table S8:** Monod model parameters calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s4.1, digits = c(0, 0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" "," ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$', 'Conversion factor'), row.names=F, caption = "_**Table S9:** Monod model parameters calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(monod_all_results_good, digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S9:** Goodness of fit of Monod model calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(monod_all_results_good, digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S10:** Goodness of fit of Monod model calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(t.s5.1, digits = c(0, 0, 2, 2, 2, 3, 2), align = 'c', "html", booktabs=T, col.names = c(" "," ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$', 'Conversion factor'), row.names=F, caption = "_**Table S10:** MEND model parameters calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s5.1, digits = c(0, 0, 2, 2, 2, 3, 2), align = 'c', "html", booktabs=T, col.names = c(" "," ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$', 'Conversion factor'), row.names=F, caption = "_**Table S11:** MEND model parameters calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(mend_all_results_good, digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S11:** Goodness of fit of MEND model calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(mend_all_results_good, digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S12:** Goodness of fit of MEND model calibrated against respiration rate and protein/DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(t.s6.1, digits = c(0, 2, 2, 3, 2, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" ",'$\\V_{MAX}$', '$\\K_{M}$', '$\\m_{R}$', '$\\f_{0}$', '$\\Y_{S}$', '$\\R_{Proteins}$', '$\\S_{Proteins}$', '$\\S_{DNA}$'), row.names=F, caption = "_**Table S12:** DEB model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s6.1, digits = c(0, 2, 2, 3, 2, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" ",'$\\V_{MAX}$', '$\\K_{M}$', '$\\m_{R}$', '$\\f_{0}$', '$\\Y_{S}$', '$\\R_{Proteins}$', '$\\S_{Proteins}$', '$\\S_{DNA}$'), row.names=F, caption = "_**Table S13:** DEB model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(deb_all_results_good, digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S13:** Goodness of fit of DEB model calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(deb_all_results_good, digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S14:** Goodness of fit of DEB model calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 ```
+
+$~$
+$~$
+$~$
+$~$
+$~$
+$~$
+```{r correlation figure, echo=FALSE, fig.height=5, fig.width=10, fig.cap="Fig.
+S3: Effect of structure of bacterial and fungal part of microbial community (expressed as multivariate mean distance between groups; see Matrial and Methods section) on overall microbial community maintnance respiration rate ($m_{R}$ - (a)) and protein concentration in Reserves pool ($R_{Protein}$ - (b)).
+The value of both parameters are expressed as a mean difference between groups in order to allow unbiased correlation with microbial community structure metric (see Matrial and Methods section)."}
+
+dpars<-t.s6.1[c(7:12), -1]
+dpars$Substrate<-rep(c("Cellobiose", "Glucose"), 3)
+dpars$Structure<-c("BROTH", "BROTH", "GLASS", "GLASS", "WOOL", "WOOL")
+
+dpars2<-merge(dpars, bacdiff, by.y=c("Substrate", "Structure"), by.x=c("Substrate", "Structure"))
+
+dpars3<-merge(dpars, fdiff, by.y=c("Substrate", "Structure"), by.x=c("Substrate", "Structure"))
+
+for(i in 3:10){
+  dpars2[, i]<-sqrt((dpars2[, i]-as.numeric(dpars2[1, i]))^2)
+}
+
+for(i in 3:10){
+  dpars3[, i]<-sqrt((dpars3[, i]-as.numeric(dpars3[1, i]))^2)
+}
+
+dpars2$group<-c("Bacteria")
+dpars3$group<-c("Fungi")
+dpars4<-rbind(dpars2, dpars3)
+
+corplot1.p<-coefficients(nls(m0~Dist^x, dpars4, start = list(x=2)))
+
+corplot1<-ggplot(dpars4[dpars4$group=="Bacteria", ], aes(Dist, m0*1e4))+
+  geom_point(cex=6, aes(fill=Structure), show.legend=T, shape=21)+
+  geom_point(data=dpars4[dpars4$group=="Fungi", ], cex=6, 
+             aes(x=Dist, y=m0*1e4, fill=Structure), show.legend=F, shape=22)+
+  scale_fill_manual(values = c("black", "grey", "white"))+
+  stat_function(fun = function(x){(x^corplot1.p)*1e4}, lwd=1, colour="grey30")+
+  xlab("16S and ITS Distance")+
+  ggtitle("a)")+
+  ylab(expression(paste(m[R]~difference)))+
+  theme_min+theme(legend.title = element_blank(),
+                  legend.position = c(0.2,0.8))
+
+corplot2.p<-coefficients(nls(fpr~Dist^x, dpars4, start = list(x=2)))
+
+corplot2<-ggplot(dpars4[dpars4$group=="Bacteria", ], aes(Dist, fpr))+
+  geom_point(cex=6, aes(fill=Structure), show.legend=F, shape=21)+
+  geom_point(data=dpars4[dpars4$group=="Fungi", ], cex=6, 
+             aes(x=Dist, y=fpr, fill=Structure), show.legend=F, shape=22)+
+  scale_fill_manual(values = c("black", "grey", "white"))+
+  stat_function(fun = function(x){(x^corplot2.p)}, lwd=1, colour="grey30")+
+  xlab("16S and ITS Distance")+
+  ggtitle("b)")+
+  ylab(expression(paste(R[Proteins]~difference)))+
+  theme_min
+
+grid.arrange(corplot1, corplot2, nrow=1)
+
+```

### 2019-03-25 SI_code.Rmd
MEND restructured

@@
         #C uptake
-        F1=1/CUE*(Vmax+mr)*Cmic*C/(Km+C)
-        #growth respiration
-        F4=(1/CUE-1)*Vmax*Cmic*C/(Km+C)
-        #maintnance respiration
-        F5=(1/CUE-1)*mr*Cmic*C/(Km+C)
-        #microbial mortality
-        F8=mr*Cmic
+        Cu=Vmax*Cmic*C/(Km+C)
+        #respiration
+        r=mr*Cmic+(1-CUE)*Cu
+
+        #Protc=Cmic*fp
+        #DNAc=Cmic*fd
 
         #states
-        dCmic<-F1-(F4+F5)-F8
-        dC<--F1+F8
-
-        return(list(c(dCmic, dC), r=F4+F5, Protc=0.548*Cmic))
+        dCmic<-CUE*Cu-mr*Cmic
+        dC<--Cu
+
+        return(list(c(dCmic, dC), r=r))
 
@@
         #select time and the measured variables 
-        yhat<-select(yhat_all, c("time", "Protc", "r"))
+        yhat<-select(yhat_all, c("time", "r"))
+        colnames(yhat)<-c("time", "value")
 
         #reformat to long format data frame
-        Yhat<-melt(yhat, id.vars = "time")
+        #Yhat<-melt(yhat, id.vars = "time")
 
         #add the measured data to a data frame
-        Yhat$obs<-c(odeset[order(odeset$Time), "Protc"], odeset[order(odeset$Time), c("r")])
+        yhat$obs<-c(odeset[order(odeset$Time), c("r")])
 
@@
         #I want to have the weighting factor to be proportional to mean of the given variable 
-        yweights<-Yhat %>% group_by(variable) %>% summarise(weights=mean(value, na.rm = T))
+        yhat$weights<-rep(mean(yhat$value, na.rm = T), times=nrow(yhat))
 
         #match with the Yhat data frame
-        Yhat<-merge(Yhat, yweights, by.x=c("variable"), by.y=c("variable"))
-
-        Yhat<-Yhat[Yhat$variable!="Protc", ]
+        #Yhat<-merge(Yhat, yweights, by.x=c("variable"), by.y=c("variable"))
+
+        #Yhat<-Yhat[Yhat$variable!="Protc", ]
 
         #now, the root mean square error is calculated
-        NRMSE<-as.numeric(Yhat %>% summarise(NRMSE=sum((((value-obs)/mean(weights))^2), na.rm = T)))
+        NRMSE<-with(yhat, sum((((value-obs)/mean(weights))^2), na.rm = T))
 
@@
         #select time and the measured variables 
-        yhat<-select(yhat_all, c("time", "Protc", "r"))
+        yhat<-select(yhat_all, c("time", "r"))
+        colnames(yhat)<-c("time", "value")
 
         #reformat to long format data frame
-        Yhat<-melt(yhat, id.vars = "time")
+        #Yhat<-melt(yhat, id.vars = "time")
 
         #add the measured data to a data frame
-        Yhat$obs<-c(odeset[order(odeset$Time), "Protc"], odeset[order(odeset$Time), c("r")])
-        Yhat$Substrate<-rep(odeset[order(odeset$Time), "Substrate"], times=2)
-        Yhat$Structure<-rep(odeset[order(odeset$Time), "Structure"], times=2)
+        yhat$obs<-c(odeset[order(odeset$Time), c("r")])
+
+        #add the weighting factor
+        #I want to have the weighting factor to be proportional to mean of the given variable 
+        #yhat$weights<-rep(mean(yhat$value, na.rm = T), times=nrow(yhat))
+
+        #match with the Yhat data frame
+        #Yhat<-merge(Yhat, yweights, by.x=c("variable"), by.y=c("variable"))
+
+        #Yhat<-Yhat[Yhat$variable!="Protc", ]
 
         #rsquared calculation for each variable
-        Gfit<-Yhat %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
-                                                        SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
-                                                        ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
-        Gfit$R2<-with(Gfit, 1-SSres/SStot)
-        Gfit$N<-c(5)
-        Gfit$AIC<-with(Gfit, 2*N-2*ll)
-
-        rsq_out<-list(Yhat=Yhat, Gfit=Gfit)
+        SSres=with(yhat, sum(((obs-value)^2), na.rm = T))
+        SStot=with(yhat, sum(((obs-mean(obs, na.rm = T))^2), na.rm = T))
+        ll=with(yhat, -sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
+        R2<-1-SSres/SStot
+        N<-length(x)
+        AIC<-2*N-2*ll
+
+        Gfit<-c(R2=R2, N=N, AIC=AIC, ll=ll)
+
+        rsq_out<-list(Yhat=yhat, Gfit=Gfit)
 
@@
         #rsquared calculation for each variable
-        res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
-                                                               SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
-                                                               ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
-        res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
-        res$goodness$N<-rep(length(parnames)*2, times=2)
-        res$goodness$AIC<-with(res$goodness, 2*N-2*ll)
+        SSres=with(res$OvP, sum(((obs-value)^2), na.rm = T))
+        SStot=with(res$OvP, sum(((obs-mean(obs, na.rm = T))^2), na.rm = T))
+        ll=with(res$OvP, -sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
+        R2<-1-SSres/SStot
+        N<-length(parnames)*2
+        AIC<-2*N-2*ll
+
+        res$goodness<-c(R2=R2, N=N, AIC=AIC, ll=ll)
 
@@
         #rsquared calculation for each variable
-          res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
-                                                                 SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
-                                                                 ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
-          res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
-          res$goodness$N<-rep(length(parnames)*3, times=2)
-          res$goodness$AIC<-with(res$goodness, 2*N-2*ll)
+        SSres=with(res$OvP, sum(((obs-value)^2), na.rm = T))
+        SStot=with(res$OvP, sum(((obs-mean(obs, na.rm = T))^2), na.rm = T))
+        ll=with(res$OvP, -sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
+        R2<-1-SSres/SStot
+        N<-length(parnames)*3
+        AIC<-2*N-2*ll
+
+        res$goodness<-c(R2=R2, N=N, AIC=AIC, ll=ll)
 
@@
         #rsquared calculation for each variable
-          res$goodness<-res$OvP %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T), 
-                                                                 SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
-                                                                 ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
-          res$goodness$R2<-with(res$goodness, 1-SSres/SStot)
-          res$goodness$N<-rep(length(parnames)*6, times=2)
-          res$goodness$AIC<-with(res$goodness, 2*N-2*ll)
-
+        SSres=with(res$OvP, sum(((obs-value)^2), na.rm = T))
+        SStot=with(res$OvP, sum(((obs-mean(obs, na.rm = T))^2), na.rm = T))
+        ll=with(res$OvP, -sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
+        R2<-1-SSres/SStot
+        N<-length(parnames)*6
+        AIC<-2*N-2*ll
+
+        res$goodness<-c(R2=R2, N=N, AIC=AIC, ll=ll)
         }
@@
 
-#results of the previous iteration (not run)
+# #results of the previous iteration (not run)
 # mend_r_results_par<-rbind(mend_r1[[1]]$pars,
@@
 #                      mend_r4[[4]]$pars, mend_r4[[5]]$pars, mend_r4[[6]]$pars)
-# mend_r_results_good<-rbind(mend_r1$goodness[2,c(4:7)],
-#                      mend_r2$goodness[2,c(4:7)],
-#                      mend_r3$goodness[2,c(4:7)],
-#                      mend_r4$goodness[2,c(4:7)])
+# mend_r_results_good<-as.data.frame(rbind(mend_r1$goodness,
+#                     mend_r2$goodness,
+#                     mend_r3$goodness,
+#                     mend_r4$goodness))
 # write.csv(mend_r_results_par, file = c("../mend_r_results_par.csv"), row.names = F)
@@
         Gfit$R2<-with(Gfit, 1-SSres/SStot)
-        Gfit$N<-c(6)
+        Gfit$N<-length(x)
         Gfit$AIC<-with(Gfit, 2*N-2*ll)
@@
         #C uptake
-        F1=1/CUE*(Vmax+mr)*Cmic*C/(Km+C)
-        #growth respiration
-        F4=(1/CUE-1)*Vmax*Cmic*C/(Km+C)
-        #maintnance respiration
-        F5=(1/CUE-1)*mr*Cmic*C/(Km+C)
-        #microbial mortality
-        F8=mr*Cmic
-
-        Protc=fp*Cmic
+        Cu=Vmax*Cmic*C/(Km+C)
+        #respiration
+        r=mr*Cmic+(1-CUE)*Cu
+
+        Protc=Cmic*fp
+        #DNAc=Cmic*fd
 
         #states
-        dCmic<-F1-(F4+F5)-F8
-        dC<--F1+F8
-
-        return(list(c(dCmic, dC), r=F4+F5, Protc=Protc))
+        dCmic<-CUE*Cu-mr*Cmic
+        dC<--Cu
+
+        return(list(c(dCmic, dC), r=r, Protc=Protc))
 
@@
         #C uptake
-        F1=1/CUE*(Vmax+mr)*Cmic*C/(Km+C)
-        #growth respiration
-        F4=(1/CUE-1)*Vmax*Cmic*C/(Km+C)
-        #maintnance respiration
-        F5=(1/CUE-1)*mr*Cmic*C/(Km+C)
-        #microbial mortality
-        F8=mr*Cmic
-
-        DNAc=fd*Cmic
+        Cu=Vmax*Cmic*C/(Km+C)
+        #respiration
+        r=mr*Cmic+(1-CUE)*Cu
+
+        #Protc=Cmic*fp
+        DNAc=Cmic*fd
 
         #states
-        dCmic<-F1-(F4+F5)-F8
-        dC<--F1+F8
-
-        return(list(c(dCmic, dC), r=F4+F5, DNAc=DNAc))
+        dCmic<-CUE*Cu-mr*Cmic
+        dC<--Cu
+
+        return(list(c(dCmic, dC), r=r, DNAc=DNAc))
 
@@
       #approximate parameter estimation is done by MCMC method
-      par_mcmc<-modMCMC(f=cost, p=c(Vmax=0.1, Km=3, CUE=0.5, mr=0.01, fp=0.5, Cmic_0=0.01), 
-                          lower=c(Vmax=1e-3, Km=1e-3, CUE=0, mr=1e-5, fp=0, Cmic_0=1e-5),
-                          upper=c(Vmax=10, Km=100, CUE=1, mr=10, fp=1, Cmic_0=5), niter=10000)
+      par_mcmc<-modMCMC(f=cost, p=c(Vmax=0.1, Km=3, CUE=0.5, mr=0.01, fd=0.5, Cmic_0=0.01), 
+                          lower=c(Vmax=1e-3, Km=1e-3, CUE=0, mr=1e-5, fd=0, Cmic_0=1e-5),
+                          upper=c(Vmax=10, Km=100, CUE=1, mr=10, fd=1, Cmic_0=5), niter=10000)
 
@@
 
-#results of the previous iteration (not run)
+# #results of the previous iteration (not run)
 # mend_all_results_par<-rbind(mend_i1[[1]]$pars, mend_i1_DNA[[1]]$pars,
@@
 ```
+
+```{r picina example, eval=FALSE, include=TRUE}
+
+#######################################Monod model##################################################
+monod_ex<-function(data){
+
+   #monod growth function
+    deriv<-function(time, state, pars){
+
+      with(as.list(c(state, pars)),{
+
+        Cb<-matrix(state[1:5], ncol = 1)
+        C<-matrix(state[6:10], ncol = 1)
+
+        dCb<--k*Cb+CUE*Vmax*Cb*C/(Km+C)
+        dC<-k*Cb-Vmax*Cb*C/(Km+C)
+
+        return(list(c(dCb, dC), r=(1-CUE)*Vmax*Cb*C/(Km+C), Cmic=kec*Cb))
+
+      })
+    }
+
+    #define names of parameters
+    parnames<-c("Vmax", "Km", "CUE", "k", "kec")
+
+    #parameters estimation function
+    estim<-function(odeset){
+
+      #defining cost function
+      cost<-function(x){
+
+        par<-x[1:length(parnames)]
+
+        names(par)<-parnames
+
+        #initial conditions
+        Cmicinit=93.6/par[["kec"]]
+        Cinit=315
+
+        #first, pars dependent output from ode is matched with measured values
+        yhat_all<-as.data.frame(ode(y=c(Cb1=Cmicinit, Cb2=Cmicinit, Cb3=Cmicinit, Cb4=Cmicinit, Cb5=Cmicinit,
+                                        C1=315, C2=315, C3=315, C4=315, C5=315), parms=par, deriv, times=c(0, 8, 16, 24, 48, 72)))
+
+        #select time and the measured variables
+        yhat<-select(yhat_all, c("time", "Cmic1", "Cmic2", "Cmic3","Cmic4", "Cmic5",
+                               "C1", "C2", "C3", "C4", "C5",
+                               "r1", "r2", "r3", "r4", "r5"))
+
+        #reformat to long format data frame
+        Yhat<-melt(yhat, id.vars = "time")
+
+        #add the measured data to a data frame
+        Yhat$obs<-c(as.numeric(odeset[c(1,6,11,16, 21, 26), c("Cmic")]),
+                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("Cmic")]),
+                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("Cmic")]),
+                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("Cmic")]),
+                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("Cmic")]),
+                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("DOC")]),
+                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("DOC")]),
+                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("DOC")]),
+                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("DOC")]),
+                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("DOC")]),
+                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("r")]),
+                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("r")]),
+                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("r")]),
+                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("r")]),
+                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("r")]))
+
+        #add the weighting factor
+        #I want to have the weighting factor to be proportional to mean of the given variable
+        yweights<-Yhat %>% group_by(variable) %>% summarise(weights=mean(value, na.rm = T))
+
+        #match with the Yhat data frame
+        Yhat<-merge(Yhat, yweights, by.x=c("variable"), by.y=c("variable"))
+
+        #now, the root mean square error is calculated
+        NRMSE<-as.numeric(Yhat %>% group_by(variable) %>% summarise(NRMSE=sum((((value-obs)/mean(weights))^2), na.rm = T)) %>%
+                            summarise(NRMSE=sum(NRMSE)))
+
+        return(NRMSE)
+
+      }
+
+      #defining goodness of fit function
+      rsq_ode<-function(x){
+
+        par<-x[1:length(parnames)]
+
+        names(par)<-parnames
+
+        #initial conditions
+        Cmicinit=93.6/par[["kec"]]
+        Cinit=315
+
+        #first, pars dependent output from ode is matched with measured values
+        yhat_all<-as.data.frame(ode(y=c(Cb1=Cmicinit, Cb2=Cmicinit, Cb3=Cmicinit, Cb4=Cmicinit, Cb5=Cmicinit,
+                                        C1=315, C2=315, C3=315, C4=315, C5=315), parms=par, deriv, times=c(0, 8, 16, 24, 48, 72)))
+
+        #select time and the measured variables
+        yhat<-select(yhat_all, c("time", "Cmic1", "Cmic2", "Cmic3","Cmic4", "Cmic5",
+                               "C1", "C2", "C3", "C4", "C5",
+                               "r1", "r2", "r3", "r4", "r5"))
+
+        #reformat to long format data frame
+        Yhat<-melt(yhat, id.vars = "time")
+
+        #add the measured data to a data frame
+        Yhat$obs<-c(as.numeric(odeset[c(1,6,11,16, 21, 26), c("Cmic")]),
+                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("Cmic")]),
+                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("Cmic")]),
+                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("Cmic")]),
+                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("Cmic")]),
+                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("DOC")]),
+                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("DOC")]),
+                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("DOC")]),
+                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("DOC")]),
+                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("DOC")]),
+                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("r")]),
+                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("r")]),
+                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("r")]),
+                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("r")]),
+                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("r")]))
+
+        Yhat$variable<-c(rep("Cmic", 30), rep("C", 30), rep("r", 30))
+
+        #rsquared calculation for each variable
+        Gfit<-Yhat %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T),
+                                                        SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
+                                                        ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
+        Gfit$R2<-with(Gfit, 1-SSres/SStot)
+        Gfit$N<-length(x)
+        Gfit$AIC<-with(Gfit, 2*N-2*ll)
+
+        rsq_out<-list(Yhat=Yhat, Gfit=Gfit)
+
+        return(rsq_out)
+
+      }
+
+      #approximate parameter estimation is done by MCMC method
+      par_mcmc<-modMCMC(f=cost, p=c(Vmax=0.1, Km=3, CUE=0.5, k=0.01, kec=0.38),
+                          lower=c(Vmax=1e-5, Km=1e-2, CUE=0, k=1e-5, kec=0),
+                          upper=c(Vmax=1, Km=20, CUE=1, k=10, kec=1), niter=10000)
+
+      #lower and upper limits for parameters are extracted
+      pl<-summary(par_mcmc)["min",]
+      pu<-summary(par_mcmc)["max",]
+
+      #these limits are used to find global optimum by DEoptim
+      opt_par<-DEoptim(fn=cost, lower=pl, upper=pu,
+                       control = c(itermax = 10000, steptol = 50, reltol = 1e-8,
+                                   trace=FALSE, strategy=3, NP=250))
+
+      #global optimum parameters are further used in MCMC to find parameters distribution
+      par_prof<-modMCMC(f=cost, p=opt_par$optim$bestmem,
+                        lower=pl,
+                        upper=pu, niter=5000)
+
+      #goodness of fit
+      fit<-rsq_ode(opt_par$optim$bestmem)
+
+      #best parameters
+      p<-opt_par$optim$bestmem
+      names(p)<-parnames
+
+      #return list with opt_par and par_prof
+      estim_out<-list(pars=p, par_prof=par_prof, fit=fit)
+
+      return(estim_out)
+
+    }
+
+    res<-estim(odeset=data)
+
+  return(res)
+
+}
+
+#################################################MEND#######################################################
+mend_ex<-function(data){
+
+   #mend growth function
+    deriv<-function(time, state, pars){
+
+      with(as.list(c(state, pars)),{
+
+        Cb<-matrix(state[1:5], ncol = 1)
+        C<-matrix(state[6:10], ncol = 1)
+
+        dCb<-CUE*Vmax*Cb*C/(Km+C)-mr*Cb
+        dC<--Vmax*Cb*C/(Km+C)
+
+        return(list(c(dCb, dC), r=(1-CUE)*Vmax*Cb*C/(Km+C)+mr*Cb, Cmic=kec*Cb))
+
+      })
+    }
+
+    #define names of parameters
+    parnames<-c("Vmax", "Km", "CUE", "mr", "kec")
+
+    #parameters estimation function
+    estim<-function(odeset){
+
+      #defining cost function
+      cost<-function(x){
+
+        par<-x[1:length(parnames)]
+
+        names(par)<-parnames
+
+        #initial conditions
+        Cmicinit=93.6/par[["kec"]]
+        Cinit=315
+
+        #first, pars dependent output from ode is matched with measured values
+        yhat_all<-as.data.frame(ode(y=c(Cb1=Cmicinit, Cb2=Cmicinit, Cb3=Cmicinit, Cb4=Cmicinit, Cb5=Cmicinit,
+                                        C1=315, C2=315, C3=315, C4=315, C5=315), parms=par, deriv, times=c(0, 8, 16, 24, 48, 72)))
+
+        #select time and the measured variables
+        yhat<-select(yhat_all, c("time", "Cmic1", "Cmic2", "Cmic3","Cmic4", "Cmic5",
+                               "C1", "C2", "C3", "C4", "C5",
+                               "r1", "r2", "r3", "r4", "r5"))
+
+        #reformat to long format data frame
+        Yhat<-melt(yhat, id.vars = "time")
+
+        #add the measured data to a data frame
+        Yhat$obs<-c(as.numeric(odeset[c(1,6,11,16, 21, 26), c("Cmic")]),
+                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("Cmic")]),
+                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("Cmic")]),
+                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("Cmic")]),
+                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("Cmic")]),
+                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("DOC")]),
+                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("DOC")]),
+                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("DOC")]),
+                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("DOC")]),
+                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("DOC")]),
+                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("r")]),
+                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("r")]),
+                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("r")]),
+                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("r")]),
+                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("r")]))
+
+        #add the weighting factor
+        #I want to have the weighting factor to be proportional to mean of the given variable
+        yweights<-Yhat %>% group_by(variable) %>% summarise(weights=mean(value, na.rm = T))
+
+        #match with the Yhat data frame
+        Yhat<-merge(Yhat, yweights, by.x=c("variable"), by.y=c("variable"))
+
+        #now, the root mean square error is calculated
+        NRMSE<-as.numeric(Yhat %>% group_by(variable) %>% summarise(NRMSE=sum((((value-obs)/mean(weights))^2), na.rm = T)) %>%
+                            summarise(NRMSE=sum(NRMSE)))
+
+        return(NRMSE)
+
+      }
+
+      #defining goodness of fit function
+      rsq_ode<-function(x){
+
+        par<-x[1:length(parnames)]
+
+        names(par)<-parnames
+
+        #initial conditions
+        Cmicinit=93.6/par[["kec"]]
+        Cinit=315
+
+        #first, pars dependent output from ode is matched with measured values
+        yhat_all<-as.data.frame(ode(y=c(Cb1=Cmicinit, Cb2=Cmicinit, Cb3=Cmicinit, Cb4=Cmicinit, Cb5=Cmicinit,
+                                        C1=315, C2=315, C3=315, C4=315, C5=315), parms=par, deriv, times=c(0, 8, 16, 24, 48, 72)))
+
+        #select time and the measured variables
+        yhat<-select(yhat_all, c("time", "Cmic1", "Cmic2", "Cmic3","Cmic4", "Cmic5",
+                               "C1", "C2", "C3", "C4", "C5",
+                               "r1", "r2", "r3", "r4", "r5"))
+
+        #reformat to long format data frame
+        Yhat<-melt(yhat, id.vars = "time")
+
+        #add the measured data to a data frame
+        Yhat$obs<-c(as.numeric(odeset[c(1,6,11,16, 21, 26), c("Cmic")]),
+                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("Cmic")]),
+                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("Cmic")]),
+                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("Cmic")]),
+                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("Cmic")]),
+                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("DOC")]),
+                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("DOC")]),
+                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("DOC")]),
+                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("DOC")]),
+                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("DOC")]),
+                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("r")]),
+                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("r")]),
+                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("r")]),
+                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("r")]),
+                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("r")]))
+
+        Yhat$variable<-c(rep("Cmic", 30), rep("C", 30), rep("r", 30))
+
+        #rsquared calculation for each variable
+        Gfit<-Yhat %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T),
+                                                        SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
+                                                        ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
+        Gfit$R2<-with(Gfit, 1-SSres/SStot)
+        Gfit$N<-length(x)
+        Gfit$AIC<-with(Gfit, 2*N-2*ll)
+
+        rsq_out<-list(Yhat=Yhat, Gfit=Gfit)
+
+        return(rsq_out)
+
+      }
+
+      #approximate parameter estimation is done by MCMC method
+      par_mcmc<-modMCMC(f=cost, p=c(Vmax=0.1, Km=3, CUE=0.5, mr=0.01,kec=0.38),
+                          lower=c(Vmax=1e-5, Km=1e-2, CUE=0, mr=1e-5, kec=0),
+                          upper=c(Vmax=1, Km=20, CUE=1, mr=10, kec=1), niter=10000)
+
+      #lower and upper limits for parameters are extracted
+      pl<-summary(par_mcmc)["min",]
+      pu<-summary(par_mcmc)["max",]
+
+      #these limits are used to find global optimum by DEoptim
+      opt_par<-DEoptim(fn=cost, lower=pl, upper=pu,
+                       control = c(itermax = 10000, steptol = 50, reltol = 1e-8,
+                                   trace=FALSE, strategy=3, NP=250))
+
+      #global optimum parameters are further used in MCMC to find parameters distribution
+      par_prof<-modMCMC(f=cost, p=opt_par$optim$bestmem,
+                        lower=pl,
+                        upper=pu, niter=5000)
+
+      #goodness of fit
+      fit<-rsq_ode(opt_par$optim$bestmem)
+
+      #best parameters
+      p<-opt_par$optim$bestmem
+      names(p)<-parnames
+
+      #return list with opt_par and par_prof
+      estim_out<-list(pars=p, par_prof=par_prof, fit=fit)
+
+      return(estim_out)
+
+    }
+
+    res<-estim(odeset=data)
+
+  return(res)
+
+}
+
+#################################################DEB#######################################################
+deb_ex<-function(data){
+
+   #deb function
+    deriv<-function(time, state, pars){
+
+      with(as.list(c(state, pars)),{
+
+      R<-matrix(state[1:5], ncol = 1)
+      S<-matrix(state[6:10], ncol = 1)
+      C<-matrix(state[11:15], ncol = 1)
+
+      #C uptake
+      Cu=Vmax*S*C/(Km+C)
+
+      #udrzovani
+      m=S*m0
+
+      #mobilizace rezerv
+      an=f*R-m
+
+      #respirace
+      r=m+pmax(an*(1-Yu), 0)
+
+      #microbialni uhlik je neco z rezerv a neco ze struktur
+      Cmic=fr*R+fs*S
+
+      dR<-Cu-f*R
+      dS<-pmax(an*Yu,0)+pmin(0, an)
+      dC<--Cu
+
+      return(list(c(dR, dS, dC), Cmic=Cmic, r=r))
+
+      })
+    }
+
+    #define names of parameters
+    parnames<-c("Vmax", "Km", "m0", "f", "Yu", "fr", "fs", "Rinit")
+
+    #parameters estimation function
+    estim<-function(odeset){
+
+      #defining cost function
+      cost<-function(x){
+
+        par<-x
+
+        names(par)<-parnames
+
+        Cmicinit=93.6
+        Cinit=315
+
+        #initial R
+        Rinit<-par[["Rinit"]]
+
+        #initial S is then
+        Sinit<-(Cmicinit-Rinit*par[["fr"]])/par[["fs"]]
+
+        #first, pars dependent output from ode is matched with measured values
+        yhat_all<-as.data.frame(ode(y=c(R1=Rinit, R2=Rinit, R3=Rinit, R4=Rinit, R5=Rinit,
+                                      S1=Sinit, S2=Sinit, S3=Sinit, S4=Sinit, S5=Sinit,
+                                      C1=315, C2=315, C3=315, C4=315, C5=315), 
+                                  parms=par[1:7], 
+                                  deriv, 
+                                  times=c(0, 8, 16, 24, 48, 72)))
+         #select time and the measured variables 
+         yhat<-select(yhat_all, c("time", "Cmic1", "Cmic2", "Cmic3","Cmic4", "Cmic5",
+                               "C1", "C2", "C3", "C4", "C5",
+                               "r1", "r2", "r3", "r4", "r5"))
+
+        #reformat to long format data frame
+        Yhat<-melt(yhat, id.vars = "time")
+
+        #add the measured data to a data frame
+        Yhat$obs<-c(as.numeric(odeset[c(1,6,11,16, 21, 26), c("Cmic")]),
+                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("Cmic")]),
+                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("Cmic")]),
+                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("Cmic")]),
+                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("Cmic")]),
+                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("DOC")]),
+                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("DOC")]),
+                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("DOC")]),
+                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("DOC")]),
+                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("DOC")]),
+                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("r")]),
+                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("r")]),
+                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("r")]),
+                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("r")]),
+                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("r")]))
+
+        #add the weighting factor
+        #I want to have the weighting factor to be proportional to mean of the given variable
+        yweights<-Yhat %>% group_by(variable) %>% summarise(weights=mean(value, na.rm = T))
+
+        #match with the Yhat data frame
+        Yhat<-merge(Yhat, yweights, by.x=c("variable"), by.y=c("variable"))
+
+        #now, the root mean square error is calculated
+        NRMSE<-as.numeric(Yhat %>% group_by(variable) %>% summarise(NRMSE=sum((((value-obs)/mean(weights))^2), na.rm = T)) %>%
+                            summarise(NRMSE=sum(NRMSE)))
+
+        return(NRMSE)
+
+      }
+
+      #defining goodness of fit function
+      rsq_ode<-function(x){
+
+        par<-x
+
+        names(par)<-parnames
+
+        #initial conditions
+        Cmicinit=93.6
+        Cinit=315
+
+        #initial R
+        Rinit<-par[["Rinit"]]
+
+        #initial S is then
+        Sinit<-(Cmicinit-Rinit*par[["fr"]])/par[["fs"]]
+
+        #first, pars dependent output from ode is matched with measured values
+        yhat_all<-as.data.frame(ode(y=c(R1=Rinit, R2=Rinit, R3=Rinit, R4=Rinit, R5=Rinit,
+                                      S1=Sinit, S2=Sinit, S3=Sinit, S4=Sinit, S5=Sinit,
+                                      C1=315, C2=315, C3=315, C4=315, C5=315), 
+                                  parms=par[1:7], 
+                                  deriv, 
+                                  times=c(0, 8, 16, 24, 48, 72)))
+         #select time and the measured variables 
+         yhat<-select(yhat_all, c("time", "Cmic1", "Cmic2", "Cmic3","Cmic4", "Cmic5",
+                               "C1", "C2", "C3", "C4", "C5",
+                               "r1", "r2", "r3", "r4", "r5"))
+
+        #reformat to long format data frame
+        Yhat<-melt(yhat, id.vars = "time")
+
+        #add the measured data to a data frame
+        Yhat$obs<-c(as.numeric(odeset[c(1,6,11,16, 21, 26), c("Cmic")]),
+                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("Cmic")]),
+                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("Cmic")]),
+                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("Cmic")]),
+                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("Cmic")]),
+                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("DOC")]),
+                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("DOC")]),
+                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("DOC")]),
+                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("DOC")]),
+                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("DOC")]),
+                  as.numeric(odeset[c(1,6,11,16, 21, 26), c("r")]),
+                  as.numeric(odeset[c(2,7,12,17, 22, 27), c("r")]),
+                  as.numeric(odeset[c(3,8,13,18, 23, 28), c("r")]),
+                  as.numeric(odeset[c(4,9,14,19, 24, 29), c("r")]),
+                  as.numeric(odeset[c(5,10,15,20, 25, 30), c("r")]))
+
+        Yhat$variable<-c(rep("Cmic", 30), rep("C", 30), rep("r", 30))
+
+        #rsquared calculation for each variable
+        Gfit<-Yhat %>% group_by(variable) %>% summarise(SSres=sum(((obs-value)^2), na.rm = T),
+                                                        SStot=sum(((obs-mean(obs, na.rm = T))^2), na.rm = T),
+                                                        ll=-sum(((obs-value)^2), na.rm = T)/2/(sd(obs, na.rm = T)^2))
+        Gfit$R2<-with(Gfit, 1-SSres/SStot)
+        Gfit$N<-length(x)
+        Gfit$AIC<-with(Gfit, 2*N-2*ll)
+
+        rsq_out<-list(Yhat=Yhat, Gfit=Gfit)
+
+        return(rsq_out)
+
+      }
+
+      #approximate parameter estimation is done by MCMC method
+      par_mcmc<-modMCMC(f=cost, p=c(Vmax=0.1, Km=3, m0=1e-3, f=1e-2, Yu=0.5, fr=0.8, fs=0.38, Rinit=40),
+                          lower=c(Vmax=1e-5, Km=1e-2, m0=1e-5, f=1e-5, Yu=0, fr=0, fs=0, Rinit=93*0.01),
+                          upper=c(Vmax=1, Km=20, m0=10, f=10, Yu=1, fr=1, fs=1, Rinit=93*0.9), niter=10000)
+
+      #lower and upper limits for parameters are extracted
+      pl<-summary(par_mcmc)["min",]
+      pu<-summary(par_mcmc)["max",]
+
+      #these limits are used to find global optimum by DEoptim
+      opt_par<-DEoptim(fn=cost, lower=pl, upper=pu,
+                       control = c(itermax = 10000, steptol = 50, reltol = 1e-8,
+                                   trace=FALSE, strategy=3, NP=250))
+
+      #global optimum parameters are further used in MCMC to find parameters distribution
+      par_prof<-modMCMC(f=cost, p=opt_par$optim$bestmem,
+                        lower=pl,
+                        upper=pu, niter=5000)
+
+      #goodness of fit
+      fit<-rsq_ode(opt_par$optim$bestmem)
+
+      #best parameters
+      p<-opt_par$optim$bestmem
+      names(p)<-parnames
+
+      #return list with opt_par and par_prof
+      estim_out<-list(pars=p, par_prof=par_prof, fit=fit)
+
+      return(estim_out)
+
+    }
+
+    res<-estim(odeset=data)
+
+  return(res)
+
+}
+```

### 2019-04-17 SI_code.Rmd
MEND model changed

@@
 ```
+
+```{r}
+#read the data
+picina<-read.xlsx(xlsxFile = c("C:/Users/cape159/Documents/pracovni/data_statistika/kopackuv_grant/michaci_pokus/DB_theory/Hasan_picek/hasan_picek12.xlsx"))
+
+picina_monod<-monod_ex(picina)
+picina_mend<-mend_ex(picina)
+picina_deb<-deb_ex(picina)
+
+picina_monod$fit$Gfit
+picina_mend$fit$Gfit
+picina_deb$fit$Gfit
+```
+
+```{r}
+
+#Define the time
+t_ex<-seq(0, 72)
+
+#Define monod model
+monodf<-function(time, state, pars){
+      with(as.list(c(state, pars)),{
+        dCb<--k*Cb+CUE*Vmax*Cb*C/(Km+C)
+        dC<-k*Cb-Vmax*Cb*C/(Km+C)
+        return(list(c(dCb, dC), r=(1-CUE)*Vmax*Cb*C/(Km+C), Cmic=kec*Cb))
+        })
+}
+
+#Define initial conditions
+Cmicinit=93.6
+Cinit=315
+
+#Use estimated parameters
+monodf_pars<-picina_monod$pars
+
+#Do the simulation
+monod_sim<-as.data.frame(ode(y=c(Cb=Cmicinit/monodf_pars[["kec"]], C=Cinit), 
+                                  parms=monodf_pars, monodf, times=t_ex))
+
+######################################################################################
+#Define MEND model
+mendf<-function(time, state, pars){
+      with(as.list(c(state, pars)),{
+        dCb<--mr*Cb+CUE*Vmax*Cb*C/(Km+C)
+        dC<--Vmax*Cb*C/(Km+C)
+        return(list(c(dCb, dC), r=(1-CUE)*Vmax*Cb*C/(Km+C)+mr*Cb, Cmic=kec*Cb))
+        })
+}
+
+#Use estimated parameters
+mendf_pars<-picina_mend$pars
+
+#Do the simulation
+mend_sim<-as.data.frame(ode(y=c(Cb=Cmicinit/mendf_pars[["kec"]], C=Cinit), 
+                                  parms=mendf_pars, mendf, times=t_ex))
+
+#########################################################################################
+#Define DEB model
+debf<-function(time, state, pars){
+      with(as.list(c(state, pars)),{
+      #C uptake
+      Cu=Vmax*S*C/(Km+C)
+      #udrzovani
+      m=S*m0
+      #mobilizace rezerv
+      an=f*R-m
+      #respirace
+      r=m+pmax(an*(1-Yu), 0)
+      #microbialni uhlik je neco z rezerv a neco ze struktur
+      Cmic=fr*R+fs*S
+      dR<-Cu-f*R
+      dS<-pmax(an*Yu,0)+pmin(0, an)
+      dC<--Cu
+      return(list(c(dR, dS, dC), Cmic=Cmic, r=r))
+      })
+}
+
+#Use estimated parameters
+debf_pars<-picina_deb$pars
+
+#Do the simulation
+deb_sim<-as.data.frame(ode(y=c(R=debf_pars[["Rinit"]], 
+                               S=(Cmicinit-debf_pars[["Rinit"]]*debf_pars[["fr"]])/debf_pars[["fs"]], 
+                               C=Cinit),
+                           parms=debf_pars, debf, times=t_ex))
+
+###################################################################
+#Combine data
+sim_all<-rbind(monod_sim[,c("time", "Cmic", "r")],
+               mend_sim[,c("time", "Cmic", "r")],
+               deb_sim[,c("time", "Cmic", "r")])
+
+sim_all$Model<-c(rep("Monod", times=73), rep("MEND", times=73), rep("DEB", times=73))
+sim_all$Conversion<-numeric(length = 219)
+sim_all[sim_all$Model=="Monod", "Conversion"]<-picina_monod$pars[["kec"]]
+sim_all[sim_all$Model=="MEND", "Conversion"]<-picina_mend$pars[["kec"]]
+sim_all[sim_all$Model=="DEB", "Conversion"]<-with(deb_sim, Cmic/(R+S))
+
+#Cmic
+lbcmic_monod<-paste("R^2 == ", round(as.numeric(picina_monod$fit$Gfit[2,5]), 2))
+lbcmic_mend<-paste("R^2 == ", round(as.numeric(picina_mend$fit$Gfit[2,5]), 2))
+lbcmic_deb<-paste("R^2 == ", round(as.numeric(picina_deb$fit$Gfit[2,5]), 2))
+
+#r
+lbr_monod<-paste("R^2 == ", round(as.numeric(picina_monod$fit$Gfit[3,5]), 2))
+lbr_mend<-paste("R^2 == ", round(as.numeric(picina_mend$fit$Gfit[3,5]), 2))
+lbr_deb<-paste("R^2 == ", round(as.numeric(picina_deb$fit$Gfit[3,5]), 2))
+
+sim1<-ggplot(picina, aes(time, Cmic))+geom_point(cex=6, pch=21, fill="grey")+
+  geom_line(data=sim_all, aes(time, Cmic, linetype=Model), lwd=1.5)+theme_min+
+  xlab("Time (h)")+
+  ylab(expression(paste("Chloroform labile C (", mu, "g ", g^{-1}, ")")))+
+  theme(legend.title = element_blank(),
+        legend.position = c(0.65,0.2))+
+  annotate("text", x=66, y=103.5, label=lbcmic_deb, parse=T, size=5.5)+
+  annotate("text", x=66, y=99, label=lbcmic_mend, parse=T, size=5.5)+
+  annotate("text", x=66, y=94.5, label=lbcmic_monod, parse=T, size=5.5)+
+  ggtitle("a)")
+
+sim2<-ggplot(picina, aes(time, r))+geom_point(cex=6, pch=21, fill="grey")+
+  geom_line(data=sim_all, aes(time, r, linetype=Model), lwd=1.5)+theme_min+
+  xlab("Time (h)")+
+  ylab(expression(paste("Respiration rate (", mu, "g ", g^{-1}~h^{-1}, ")")))+
+  theme(legend.title = element_blank(),
+        legend.position = c(0.65,0.7))+
+  annotate("text", x=66, y=3.35, label=lbr_deb, parse=T, size=5.5)+
+  annotate("text", x=66, y=3.1, label=lbr_mend, parse=T, size=5.5)+
+  annotate("text", x=66, y=2.8, label=lbr_monod, parse=T, size=5.5)+
+  ggtitle("b)")
+
+sim3<-ggplot(sim_all, aes(time, Conversion))+geom_line(aes(linetype=Model), lwd=1.5, show.legend = F)+
+  theme_min+ylim(0.18, 0.25)+
+  xlab("Time (h)")+
+  ylab("Conversion factor")+
+  theme(legend.title = element_blank(),
+        legend.position = c(0.8,0.55))+
+  ggtitle("c)")
+
+grid.arrange(sim1, sim2, sim3, nrow=1)
+
+```

### 2019-04-19 SI_code.Rmd
Supplementary figure data and code added to markdown document

--- 2019-03-25 SI_code.Rmd
@@
 title: The effect of restricted substrate diffusion on microbial community function and structure
-author: "Petr Capek, Katherine Todd-Brown, Natalie Sadler, Nancy Hess, Kirsten Hofmockel"
+author: "Petr Capek, Katherine Todd-Brown, Natalie Sadler, Jianqiu Zheng, Nancy Hess, Kirsten Hofmockel"
 date: "`r Sys.Date()`"
@@
 ```
+##Supplementary figure
+```{r Monod and MEND model parameters, echo=FALSE, fig.height=24, fig.width=20, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Fig.
+S1: Monod (a) and MMR (b) model parameters ($V_{MAX}$ - maximum velocity constant, $K_{M}$ - affinity constant, CUE - carbon use efficiency, $k_{MB}$ - microbial biomass decay rate constant, $m_{R}$ - specific maintenance rate constant) estimated separately for three different types of physical barriers presented in incubation bottle (BROTH - black symbols, GLASS - grey symbols and WOOL - empty symbols).
+Symbols represent best parameters values estimated by the Differential Evolution Algorithm and error bars represent standard deviation of posterior parameters distribution estimated by Constrained Markov Chain Monte Carlo simulation.
+Parameters estimation was constrained by either observed respiration rates, observed respiration rates and cellular proteins concentration or observed respiration rates and DNA concentration."}
+
+# # Monod model calibrated against the respiration rate (not run)
+# # uncertainty of model parameters
+# monod_r_sd<-as.data.frame(rbind(apply(monod_r2[[1]]$par_prof$pars, 2, sd),
+#                                 apply(monod_r2[[2]]$par_prof$pars, 2, sd),
+#                                 apply(monod_r2[[3]]$par_prof$pars, 2, sd)))[,-5]
+# colnames(monod_r_sd)<-c("Vmax.sd", "Km.sd", "CUE.sd", "Km.sd")
+# #Combine with the estimated parameters
+# monod_rs1<-cbind(monod_r_results_par[c(2:4), c(1:4)], monod_r_sd)
+# monod_rs1$Legend<-rep("Respiration", times=3)
+# 
+# # Monod model calibrated against the respiration rate and cellular proteins (not run)
+# # uncertainty of model parameters
+# monod_all_sd<-as.data.frame(rbind(apply(monod_i2[[1]]$par_prof$pars, 2, sd),
+#                                   apply(monod_i2[[2]]$par_prof$pars, 2, sd),
+#                                   apply(monod_i2[[3]]$par_prof$pars, 2, sd),
+#                                   apply(monod_i2_DNA[[1]]$par_prof$pars, 2, sd),
+#                                   apply(monod_i2_DNA[[2]]$par_prof$pars, 2, sd),
+#                                   apply(monod_i2_DNA[[3]]$par_prof$pars, 2, sd)))[,-c(5,6)]
+# colnames(monod_all_sd)<-c("Vmax.sd", "Km.sd", "CUE.sd", "Km.sd")
+# #Combine with the estimated parameters
+# monod_alls<-cbind(monod_all_results_par[c(2:4, 14:16), c(1:4)], monod_all_sd)
+# monod_alls$Legend<-c(rep("Respiration & Cellular Proteins", times=3),
+#                      rep("Respiration & DNA", times=3))
+# 
+# #Combine monod_alls and monod_rs1
+# monod_s<-rbind(monod_rs1, monod_alls)
+# #Add labels
+# monod_s$Structure<-rep(c("BROTH", "GLASS", "WOOL"), times=3)
+# 
+# # MEND model calibrated against the respiration rate (not run)
+# # uncertainty of model parameters
+# mend_r_sd<-as.data.frame(rbind(apply(mend_r2[[1]]$par_prof$pars, 2, sd),
+#                                 apply(mend_r2[[2]]$par_prof$pars, 2, sd),
+#                                 apply(mend_r2[[3]]$par_prof$pars, 2, sd)))[,-5]
+# colnames(mend_r_sd)<-c("Vmax.sd", "Km.sd", "CUE.sd", "Km.sd")
+# #Combine with the estimated parameters
+# mend_rs1<-cbind(mend_r_results_par[c(2:4), c(1:4)], mend_r_sd)
+# mend_rs1$Legend<-rep("Respiration", times=3)
+# 
+# # MEND model calibrated against the respiration rate and cellular proteins (not run)
+# # uncertainty of model parameters
+# mend_all_sd<-as.data.frame(rbind(apply(mend_i2[[1]]$par_prof$pars, 2, sd),
+#                                   apply(mend_i2[[2]]$par_prof$pars, 2, sd),
+#                                   apply(mend_i2[[3]]$par_prof$pars, 2, sd),
+#                                   apply(mend_i2_DNA[[1]]$par_prof$pars, 2, sd),
+#                                   apply(mend_i2_DNA[[2]]$par_prof$pars, 2, sd),
+#                                   apply(mend_i2_DNA[[3]]$par_prof$pars, 2, sd)))[,-c(5,6)]
+# colnames(mend_all_sd)<-c("Vmax.sd", "Km.sd", "CUE.sd", "Km.sd")
+# #Combine with the estimated parameters
+# mend_alls<-cbind(mend_all_results_par[c(2:4, 14:16), c(1:4)], mend_all_sd)
+# mend_alls$Legend<-c(rep("Respiration & Cellular Proteins", times=3),
+#                      rep("Respiration & DNA", times=3))
+# 
+# #Combine mend_alls and mend_rs1
+# mend_s<-rbind(mend_rs1, mend_alls)
+# #Add labels
+# mend_s$Structure<-rep(c("BROTH", "GLASS", "WOOL"), times=3)
+# 
+# write.csv(monod_s, file = c("../monod_FigS1.csv"), row.names = F)
+# write.csv(mend_s, file = c("../mend_FigS1.csv"), row.names = F)
+
+monod_s<-read.csv("monod_FigS1.csv")
+mend_s<-read.csv("mend_FigS1.csv")
+
+Monod_s<-melt(monod_s[, c(1:4, 9, 10)], id.vars=c("Legend", "Structure"))
+Monod_s$sd<-melt(monod_s[, c(5:8, 9, 10)], id.vars=c("Legend", "Structure"))[, "value"]
+
+#add proper parameter labels
+mlab<-c('Vmax' = 'V[MAX]',
+        'Km' = 'K[M]',
+        'CUE' = 'CUE',
+        'k' = 'k[MB]')
+
+figS1a<-ggplot(Monod_s, aes(Legend, value))+
+  geom_point(shape=21, cex=6, aes(fill=Structure), position = position_dodge(width = 0.6))+
+  geom_errorbar(aes(ymin=value-sd, ymax=value+sd, colour=Structure), 
+                position = position_dodge(width = 0.6))+
+  scale_fill_manual(values = c("black", "grey", "white"))+
+  scale_colour_manual(values = c("black", "black", "black"))+theme_min+
+  theme(axis.title.x = element_blank(),
+        legend.title = element_blank(),
+        axis.text.x=element_text(vjust=0.2, size=18, colour="black", angle = 90),
+        legend.position = c("top"),
+        plot.title=element_text(size=18, face="bold", hjust=-0.01))+
+  facet_wrap(~variable, scale="free", nrow=1, 
+             labeller = labeller(variable = as_labeller(mlab, label_parsed)))+
+  ylab("Parameter value")+ggtitle("a)")
+
+Mend_s<-melt(mend_s[, c(1:4, 9, 10)], id.vars=c("Legend", "Structure"))
+Mend_s$sd<-melt(mend_s[, c(5:8, 9, 10)], id.vars=c("Legend", "Structure"))[, "value"]
+
+#add proper parameter labels
+mlab2<-c('Vmax' = 'V[MAX]',
+        'Km' = 'K[M]',
+        'CUE' = 'CUE',
+        'mr' = 'm[R]')
+
+figS1b<-ggplot(Mend_s, aes(Legend, value))+
+  geom_point(shape=21, cex=6, aes(fill=Structure), position = position_dodge(width = 0.6),
+             show.legend = F)+
+  geom_errorbar(aes(ymin=value-sd, ymax=value+sd, colour=Structure), 
+                position = position_dodge(width = 0.6), show.legend = F)+
+  scale_fill_manual(values = c("black", "grey", "white"))+
+  scale_colour_manual(values = c("black", "black", "black"))+theme_min+
+  theme(axis.title.x = element_blank(),
+        legend.title = element_blank(),
+        axis.text.x=element_text(vjust=0.2, size=18, colour="black", angle = 90),
+        plot.title=element_text(size=18, face="bold", hjust=-0.01))+
+  facet_wrap(~variable, scale="free", nrow=1, 
+             labeller = labeller(variable = as_labeller(mlab2, label_parsed)))+
+  ylab("Parameter value")+ggtitle("b)")
+
+grid.arrange(figS1a, figS1b, nrow=2)
+
+```

### 2019-06-11 SI_code.Rmd
Figs. 2 and 3 changed

@@
 bac.norm<-decostand(bac, method=c("normalize"))
-bac.dist<-vegdist(bac.norm, method = "jaccard")
+bac.dist<-vegdist(bac.norm, method = "bray")
 bac.ad<-adonis(bac.dist~Substrate+Structure/Substrate, bac_env, strata = bac_env$Day)
@@
 fungi.norm<-decostand(fungi, method=c("normalize"))
-fungi.dist<-vegdist(fungi.norm, method = "jaccard")
+fungi.dist<-vegdist(fungi.norm, method = "bray")
 fungi.ad<-adonis(fungi.dist~Substrate+Structure/Substrate, fungi_env, strata = fungi_env$Day)

### 2019-10-22 SI_code_revised.Rmd
All code revised

---
title: The effect of habitat structure on microbial community function and composition
author: "Petr Capek, Katherine Todd-Brown, Natalie Sadler, Jianqiu Zheng, Nancy Hess, Kirsten Hofmockel"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Supplementary Information
## Step-by-step analysis and code

The analysis requires several libraries

```{r libraries loading, message=FALSE, warning=FALSE}

##libraries
library(deSolve)
library(dplyr)
library(FME)
library(reshape)
library(ggplot2)
library(foreach)
library(doParallel)
library(openxlsx)
library(DEoptim)
library(gridExtra)
library(kableExtra)
library(lmerTest)
library(psycho)
library(ggeffects)
library(rcompanion)
library(vegan)

##ggplot theme
theme_min<-theme(axis.text.x=element_text(vjust=0.2, size=18, colour="black"),
                 axis.text.y=element_text(hjust=0.2, size=18, colour="black"),
                 axis.title=element_text(size=18, colour="black"),
                 axis.line=element_line(size=0.5, colour="black"),
                 strip.text=element_text(size=18, face="bold"),
                 axis.ticks=element_line(size=1, colour="black"),
                 axis.ticks.length=unit(-0.05, "cm"),
                 panel.background=element_rect(colour="black", fill="white"),
                 panel.grid=element_line(linetype=0),
                 legend.text=element_text(size=14, colour="black"),
                 legend.title=element_text(size=14, colour="black"),
                 legend.position=c("right"),
                 legend.key.size=unit(1, "cm"),
                 strip.background=element_rect(fill="grey98", colour="black"),
                 legend.key=element_rect(fill="white", size=1.2),
                 legend.spacing=unit(0.5, "cm"),
                 plot.title=element_text(size=18, face="bold", hjust=-0.05))
```

Data are available at https://github.com/petacapek/MinT.

```{r data loading, echo=FALSE}
#respiration rates in umol/ml/h
resp<-read.csv(file=c("data_checked_respiration_raw.csv"))

#arrange the data
resp.ordered<-resp[order(resp$Sample, resp$Day), ]

#biological data
biology<-read.csv(file=c("data_checked_biology_raw.csv"))

#arrange the data in the same way
biology.ordered<-biology[order(biology$Sample, biology$Day), -c(2,3)]
biology.ordered$Sample<-paste0("CSub3_", biology.ordered$Sample)

#merging both data sets
m0<-merge(resp.ordered, biology.ordered, by.x=c("Sample", "Day"), by.y = c("Sample", "Day"), all=T)

#Bacteria
bac<-as.data.frame(t(read.csv(file=c("bacteria.csv"), skip = 4)[,-1]))
colnames(bac)<-t(read.csv(file=c("bacteria.csv"), skip = 4)[,1])

bac_env<-as.data.frame(t(read.csv(file=c("bacteria.csv"), nrows = 4, header = F)[,-1]))
colnames(bac_env)<-t(read.csv(file=c("bacteria.csv"), nrows = 4, header = F)[,1])

bac_env$Structure<-recode(bac_env$Structure, 'Broth'="BROTH", 'Glass wool'="WOOL",
                          'Mixed glass'="GLASS")

#Fungi
fungi<-as.data.frame(t(read.csv(file=c("fungi.csv"), skip = 4)[,-1]))
colnames(fungi)<-t(read.csv(file=c("fungi.csv"), skip = 4)[,1])

fungi_env<-as.data.frame(t(read.csv(file=c("fungi.csv"), nrows = 4, header = F)[,-1]))
colnames(fungi_env)<-t(read.csv(file=c("fungi.csv"), nrows = 4, header = F)[,1])

fungi_env$Structure<-recode(fungi_env$Structure, 'Broth'="BROTH", 'Glass wool'="WOOL",
                          'Mixed glass'="GLASS")

```

The conducted experiment is large. For this analysis, only subset of data is used. We use treatments with Glucose or Celluloze as a single source of organic carbon and all three levels of incubation system complexity (i.e. broth, glass wool and mix of glass beads).

```{r data filtering}

d<-subset(m0, Substrate=="Glucose" | Substrate=="Cellobiose")
summary(d)

#removing outliers
#respiration rate
ggplot(d, aes(Time, r))+geom_point(cex=6)+facet_wrap(Structure~Substrate, scales="free")

d[(d$Substrate=="Glucose" & d$Structure=="Glass wool" & d$Time>20 & d$Time<30 & d$r<0.4 & !is.na(d$r)), "r"]<-NA

#Proteins
ggplot(d, aes(Time, Prot.in))+geom_point(cex=6)+facet_wrap(Structure~Substrate, scales="free")

d[(d$Substrate=="Cellobiose" & d$Structure=="Broth" &  d$Time>75), "Prot.in"]<-NA
d[(d$Substrate=="Glucose" & d$Structure=="Broth" & d$Time>75), "Prot.in"]<-NA
d[(d$Substrate=="Cellobiose" & d$Structure=="Mixed glass" & d$Time>75), "Prot.in"]<-NA
d[(d$Substrate=="Cellobiose" & d$Structure=="Glass wool" & d$Time>75 & d$Prot.in>110 & !is.na(d$Prot.in)), "Prot.in"]<-NA

```

DNA and protein concentrations are in $\mu$grams per incubation vial. Both values are recalculated to unit of carbon per one ml of media in the vial (4 ml) using the conversion factors reported in Vrede et al. (2004).

```{r conversions}

d$Protc<-d$Prot.in*0.46/12.01/4
d$DNAc<-d$DNA*0.51/12.01/4

```
##Statistical analysis
Linear Mixed-Effect models are used to evaluate the effect of substrate and the type of physical barrier on measured respiration rate, DNA and protein concentration. Becuase all three variables display significant variability over time, the time of measurement is treated as a random effect. Type of the barrier is nested within the substrate.

```{r lme analysis, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#respiration rate
lme1<-lmer(r~Substrate+Structure/Substrate + (1|Day), d)
#Proteins
lme2<-lmer(Protc~Substrate+Structure/Substrate + (1|Day), d)
#DNA
lme3<-lmer(DNAc~Substrate+Structure/Substrate + (1|Day), d)

t.s0<-as.data.frame(cbind(anova(lme1)[c(5,3,6)], anova(lme2)[c(5,3,6)], anova(lme3)[c(5,3,6)]))

row.names(t.s0)<-c("Substrate", "Structure", "Substrate*Structure")

kable(t.s0, digits = c(2, 1, 3, 2, 1, 3, 2, 1, 3), align = 'c', "html", booktabs=T, col.names = c("F", "df", "p", "F", "df", "p", "F", "df", "p"), row.names=T, caption = "_**Table S1:** Results of linear mixed-effects model analysis. Effect of substrate, the type of physical barrier (Structure) and their interaction was evaluated. Statistical analysis was conducted for measured respiration rate, cellular protein concentration and DNA. F values, degrees of freedom (df) and the probability (p) are reported._") %>% 
  kable_styling() %>% 
  add_header_above(c(" " = 1, "Respiration rate" = 3, "Cellular Proteins" = 3, "DNA" = 3),
                   italic=T)

```
$~$
$~$
$~$
$~$
$~$
$~$

```{r lme effects, echo=FALSE, fig.height=5, fig.width=15, message=FALSE, paged.print=FALSE, fig.cap="Fig. S1: Marginal effects of the substrate (Glucose or Cellobiose) and the type of physical barrier (BROTH, GLASS, WOOL) on the respiration rate (a), cellular proteins (b) and DNA concentration (c) estimated by the linear mixed-effect model."}
#respiration
lme1_contr<-get_contrasts(lme1)
lme1_effects<-ggpredict(lme1, c("Substrate", "Structure"), type="fe")
lme1_effects$x<-c(rep("Cellobiose", 3), rep("Glucose", 3))
lme1_effects$group<-c(rep(c("BROTH", "WOOL", "GLASS"), 2))
lme1_effects<-lme1_effects[order(lme1_effects$group), ]
lme1_effects$letters<-c(as.character(cldList(p ~ Contrast, data      = lme1_contr, threshold = 0.05)[[2]])[c(1:2, 5:6)], as.character(cldList(p ~ Contrast, data      = lme1_contr, threshold = 0.05)[[2]])[c(3:4)])
#Proteins
lme2_contr<-get_contrasts(lme2)
lme2_effects<-ggpredict(lme2, c("Substrate", "Structure"), type="fe")
lme2_effects$x<-c(rep("Cellobiose", 3), rep("Glucose", 3))
lme2_effects$group<-c(rep(c("BROTH", "WOOL", "GLASS"), 2))
lme2_effects<-lme2_effects[order(lme2_effects$group), ]
lme2_effects$letters<-c(as.character(cldList(p ~ Contrast, data      = lme2_contr, threshold = 0.05)[[2]])[c(1:2, 5:6)], as.character(cldList(p ~ Contrast, data      = lme2_contr, threshold = 0.05)[[2]])[c(3:4)])
#DNA
lme3_contr<-get_contrasts(lme3)
lme3_effects<-ggpredict(lme3, c("Substrate", "Structure"), type="fe")
lme3_effects$x<-c(rep("Cellobiose", 3), rep("Glucose", 3))
lme3_effects$group<-c(rep(c("BROTH", "WOOL", "GLASS"), 2))
lme3_effects<-lme3_effects[order(lme3_effects$group), ]
lme3_effects$letters<-c(as.character(cldList(p ~ Contrast, data      = lme3_contr, threshold = 0.05)[[2]])[c(1:2, 5:6)], as.character(cldList(p ~ Contrast, data      = lme3_contr, threshold = 0.05)[[2]])[c(3:4)])

p1<-ggplot(lme1_effects, aes(x, predicted))+
    geom_point(position = position_dodge(width = 0.5), aes(fill=group), shape=21, cex=6, show.legend = F)+
    scale_fill_manual(values = c("black", "grey", "white"))+
    scale_colour_manual(values = c("black", "black", "black"))+
    geom_errorbar(aes(ymin=predicted-std.error, ymax=predicted+std.error, colour=group),
                  position=position_dodge(width = 0.5), width=0.1, show.legend = F)+
    ylab(expression(paste("Respiration rate (", mu, "mol",(C-CO[2])~ml^{-1}~h^{-1}, ")")))+
    ggtitle("a)")+
    geom_text(aes(label=letters, colour=group, vjust=-10, fontface="bold.italic"),
              cex=5, position=position_dodge(width = 0.5), show.legend = F)+
    ylim(-0.01,0.3)+
    theme_min+theme(axis.title.x = element_blank(),
                    legend.title = element_blank())
p2<-ggplot(lme2_effects, aes(x, predicted))+
    geom_point(position = position_dodge(width = 0.5), aes(fill=group), shape=21, cex=6, show.legend = F)+
    scale_fill_manual(values = c("black", "grey", "white"))+
    scale_colour_manual(values = c("black", "black", "black"))+
    geom_errorbar(aes(ymin=predicted-std.error, ymax=predicted+std.error, colour=group),
                  position=position_dodge(width = 0.5), width=0.1, show.legend = F)+
    ylab(expression(paste("Cellular Proteins (", mu, "mol",(C[Proteins])~ml^{-1}, ")")))+
    ggtitle("b)")+
    geom_text(aes(label=letters, colour=group, vjust=-3, fontface="bold.italic"),
              cex=5, position=position_dodge(width = 0.5), show.legend = F)+
    ylim(0,1.2)+
    theme_min+theme(axis.title.x = element_blank(),
                    legend.title = element_blank())
p3<-ggplot(lme3_effects, aes(x, predicted))+
    geom_point(position = position_dodge(width = 0.5), aes(fill=group), shape=21, cex=6)+
    scale_fill_manual(values = c("black", "grey", "white"))+
    scale_colour_manual(values = c("black", "black", "black"))+
    geom_errorbar(aes(ymin=predicted-std.error, ymax=predicted+std.error, colour=group),
                  position=position_dodge(width = 0.5), width=0.1)+
    ylab(expression(paste("DNA (", mu, "mol",(C[DNA])~ml^{-1}, ")")))+
    ggtitle("c)")+
    geom_text(aes(label=letters, colour=group, vjust=-3, fontface="bold.italic"),
              cex=5, position=position_dodge(width = 0.5), show.legend = F)+
    ylim(0,0.08)+
    theme_min+theme(axis.title.x = element_blank(),
                    legend.title = element_blank(),
                    legend.position = c(0.25,0.2))

grid.arrange(p1, p2, p3, nrow=1)

```
$~$
$~$
$~$
Next, the effect of the substrate and the presence of a physical barrier on microbial community structure and diversity is evaluated. First, the bray distance indices are calculated on a normalized data. Then, Permutational Multivariate Analysis of Variance is used to estimate the treatment effects. Analysis permutations are constrained to an individual times of incubation to avoid the potential confounding effect of time.

```{r vegan analysis, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#Bacteria
bac.norm<-decostand(bac, method=c("normalize"))
bac.dist<-vegdist(bac.norm, method = "bray")
bac.ad<-adonis(bac.dist~Substrate+Structure/Substrate, bac_env, strata = bac_env$Day)

#Fungi
fungi.norm<-decostand(fungi, method=c("normalize"))
fungi.dist<-vegdist(fungi.norm, method = "bray")
fungi.ad<-adonis(fungi.dist~Substrate+Structure/Substrate, fungi_env, strata = fungi_env$Day)

t.s01<-cbind(as.data.frame(bac.ad$aov[c(4, 1, 6)])[c(1:3), ], as.data.frame(fungi.ad$aov[c(4, 1, 6)])[c(1:3), ])

row.names(t.s01)<-c("Substrate", "Structure", "Substrate*Structure")

kable(t.s01, digits = c(2, 1, 3, 2, 1, 3), align = 'c', "html", booktabs=T, col.names = c("F", "df", "p", "F", "df", "p"), row.names=T, caption = "_**Table S2:** Results of permutational multivariate analysis of variance. Effect of substrate, the type of physical barrier (Structure) and their interaction was evaluated. Statistical analysis was conducted for bacterial and fungal part of the microbial community separately. F values, degrees of freedom (df) and the probability (p) are reported._") %>% 
  kable_styling() %>% 
  add_header_above(c(" " = 1, "Bacteria" = 3, "Fungi" = 3),
                   italic=T)

```
$~$
$~$
$~$
$~$
$~$
$~$
```{r community figure, echo=FALSE, fig.height=5, fig.width=10, fig.cap="Fig. S2: Multivariate distances between six experimental groups that combine two different substrate treatments (Glucose or Cellobiose) and three different types of physical barrier (BROTH, WOOL, GLASS - black, grey and empty symbols respectively) in respect to structure of bacterial ((a); 16S) and fungal ((b); ITS) part of microbial community."}
#Bacteria
bacm<-as.matrix(bac.dist)
#Intercept - Glucose - BROTH
mGB<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
       intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Glucose"))]
mGB[mGB==0]<-NA
#Glucose - GLASS
mGG<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
       intersect(which(bac_env$Structure=="GLASS"),which(bac_env$Substrate=="Glucose"))]
mGG[mGG==0]<-NA
#Glucose - WOOL
mGW<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
       intersect(which(bac_env$Structure=="WOOL"),which(bac_env$Substrate=="Glucose"))]
mGW[mGW==0]<-NA
#Cellobiose - BROTH
mCB<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
       intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose"))]
mCB[mCB==0]<-NA
#Cellobiose - GLASS
mCG<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
       intersect(which(bac_env$Structure=="GLASS"),which(bac_env$Substrate=="Cellobiose"))]
mCG[mCG==0]<-NA
#Cellobiose - WOOL
mCW<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
       intersect(which(bac_env$Structure=="WOOL"),which(bac_env$Substrate=="Cellobiose"))]
mCW[mCW==0]<-NA

bacdiff<-data.frame(Dist=c(mean(mGB, na.rm=T), mean(mGG, na.rm=T), mean(mGW, na.rm=T), 0, mean(mCG, na.rm=T), mean(mCW, na.rm=T)),
                    Dist.sd=c(sd(mGB, na.rm=T), sd(mGG, na.rm=T), sd(mGW, na.rm=T), sd(mCB, na.rm=T), sd(mCG, na.rm=T), sd(mCW, na.rm=T)),
                    Substrate=c(rep("Glucose", 3), rep("Cellobiose", 3)),
                    Structure=rep(c("BROTH", "GLASS", "WOOL"), 2))

#Fungi
fm<-as.matrix(fungi.dist)
#Intercept - Glucose - BROTH
fGB<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
          intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Glucose"))]
fGB[fGB==0]<-NA
#Glucose - GLASS
fGG<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
          intersect(which(fungi_env$Structure=="GLASS"),which(fungi_env$Substrate=="Glucose"))]
fGG[fGG==0]<-NA
#Glucose - WOOL
fGW<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
          intersect(which(fungi_env$Structure=="WOOL"),which(fungi_env$Substrate=="Glucose"))]
fGW[fGW==0]<-NA
#Cellobiose - BROTH
fCB<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
          intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose"))]
fCB[fCB==0]<-NA
#Cellobiose - GLASS
fCG<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
          intersect(which(fungi_env$Structure=="GLASS"),which(fungi_env$Substrate=="Cellobiose"))]
fCG[fCG==0]<-NA
#Cellobiose - WOOL
fCW<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
          intersect(which(fungi_env$Structure=="WOOL"),which(fungi_env$Substrate=="Cellobiose"))]
fCW[fCW==0]<-NA

fdiff<-data.frame(Dist=c(mean(fGB, na.rm=T), mean(fGG, na.rm=T), mean(fGW, na.rm=T), 0, mean(fCG, na.rm=T), mean(fCW, na.rm=T)),
                    Dist.sd=c(sd(fGB, na.rm=T), sd(fGG, na.rm=T), sd(fGW, na.rm=T), sd(fCB, na.rm=T), sd(fCG, na.rm=T), sd(fCW, na.rm=T)),
                    Substrate=c(rep("Glucose", 3), rep("Cellobiose", 3)),
                    Structure=rep(c("BROTH", "GLASS", "WOOL"), 2))

bacplot<-ggplot(bacdiff, aes(Substrate, Dist))+
  geom_point(position = position_dodge(width = 0.5), aes(fill=Structure), shape=21, cex=6)+
  scale_fill_manual(values = c("black", "grey", "white"))+
  scale_colour_manual(values = c("black", "black", "black"))+
  geom_errorbar(aes(ymin=Dist-Dist.sd, ymax=Dist+Dist.sd, colour=Structure),
                position=position_dodge(width = 0.5), width=0.1)+
  ylab("16S Distance")+
  ggtitle("a)")+
  theme_min+theme(axis.title.x = element_blank(),
                  legend.title = element_blank(),
                  legend.position = c(0.8, 0.2))
fplot<-ggplot(fdiff, aes(Substrate, Dist))+
  geom_point(position = position_dodge(width = 0.5), aes(fill=Structure), shape=21, cex=6, show.legend = F)+
  scale_fill_manual(values = c("black", "grey", "white"))+
  scale_colour_manual(values = c("black", "black", "black"))+
  geom_errorbar(aes(ymin=Dist-Dist.sd, ymax=Dist+Dist.sd, colour=Structure),
                position=position_dodge(width = 0.5), width=0.1, show.legend = F)+
  ylab("ITS Distance")+
  ggtitle("b)")+
  theme_min+theme(axis.title.x = element_blank(),
                  legend.title = element_blank())

grid.arrange(bacplot, fplot, nrow=1)

```

$~$
$~$
$~$

##Microbial-explicit models analysis
###All models are first calibrated against respiration rate
Calibration of each model is done in four steps. First, the model is calibrated against data across all treatments. Second, the model is calibrated for each level of incubation system complexity separately. Third, the model is calibrated for Glucose and Cellobiose treatments separately. Fourth, the model is calibrated for each experimental treatment separately. 
Goodness of correspondence between measurements and model simulations are calculated across all treatments. By doing so, the effect of experimental treatment is evaluated. If the model simulation fits the data significantly better when model parameters are estimated for different treatments separately, effect of this treatment is significant.

The calibration of the models is not run within this document because it takes couple of hours to finish. Results of the previous iteration is printed instead. 
To run the code and reproduce the results, the option "eval" need to be changed to TRUE. Please note that the parallel computing is used. Number of cores is set to total number of cores of the computer minus 1. This can be changed if necessary.

When models are calibrated against the respiration rate only, initial concentration of microbial biomass or reserves and structures is estimated as a single parameter. However, when model parameters are estimated for each treatment separately, initial microbial biomass can differ between treatments. This does not correspond to a reality. Therefore, initial concentration of microbial biomass is subsequently fixed across treatments.

####Monod model
```{r monod respiration calibration, eval=F, echo=T}
#Loading the function 
source("../Rscripts/monod_r.R")
source("../Rscripts/monod_r_fix.R")
#across all treatments
monod_r1<-monod_r(data=d, FACT = 4)
#########################################################
no_cors<-detectCores()
cl<-makeCluster(no_cors)
registerDoParallel(cl)
##########################################################
#for different structures
monod_r2<-monod_r(data=d, FACT = 2)
monod_r2_fix<-monod_r_fix(data=d, FACT = 2)
#for different substrates
monod_r3<-monod_r(data=d, FACT = 1)
monod_r3_fix<-monod_r_fix(data=d, FACT = 1)
#for each separately  
monod_r4<-monod_r(data=d, FACT = 3)
monod_r4_fix<-monod_r_fix(data=d, FACT = 3)
###################################################
stopImplicitCluster()
###################################################
# monod_r1$goodness
# monod_r2$goodness
# monod_r3$goodness
# monod_r4$goodness

# mean(c(monod_r4[[1]]$pars[["Cmic_0"]], monod_r4[[2]]$pars[["Cmic_0"]],
#   monod_r4[[3]]$pars[["Cmic_0"]], monod_r4[[4]]$pars[["Cmic_0"]],
#   monod_r4[[5]]$pars[["Cmic_0"]], monod_r4[[6]]$pars[["Cmic_0"]]))

monod_r1$goodness
monod_r2_fix$goodness
monod_r3_fix$goodness
monod_r4_fix$goodness
```

```{r monod respiration calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

#results of the previous iteration (not run)
# monod_r_results_par<-as.data.frame(rbind(monod_r1[[1]]$pars[1:4],
#                       monod_r2_fix[[1]]$pars, monod_r2_fix[[2]]$pars,
#                       monod_r2_fix[[3]]$pars,
#                       monod_r3_fix[[1]]$pars, monod_r3_fix[[2]]$pars,
#                       monod_r4_fix[[1]]$pars, monod_r4_fix[[2]]$pars,
#                       monod_r4_fix[[3]]$pars, monod_r4_fix[[4]]$pars,
#                       monod_r4_fix[[5]]$pars, monod_r4_fix[[6]]$pars))
# monod_r_results$Cmic_init<-c(0.09125118)
# monod_r_results_good<-rbind(monod_r1$goodness,
#                             monod_r2_fix$goodness,
#                             monod_r3_fix$goodness,
#                             monod_r4_fix$goodness)
# write.csv(monod_r_results_par, file = c("../monod_r_results_par.csv"), row.names = F)
# write.csv(monod_r_results_good, file = c("../monod_r_results_good.csv"), row.names = F)

monod_r_results_par<-read.csv("monod_r_results_par.csv")
monod_r_results_good<-read.csv("monod_r_results_good.csv")

t.s1.1<-data.frame(Vmax=monod_r_results_par[,1],
                 Km=monod_r_results_par[,2],
                 CUE=monod_r_results_par[,3],
                 k=monod_r_results_par[,4])
row.names(t.s1.1)<-c("All", "Broth", "Glass", "Wool",
                    "Cellobiose", "Glucose",
                    "Broth - Cellobiose", "Broth - Glucose",
                    "Glass - Cellobiose", "Glass - Glucose",
                    "Wool - Cellobiose", "Wool - Glucose")

options(knitr.kable.NA = '')

kable(t.s1.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$'), row.names=T, caption = "_**Table S3:** Monod model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monod_r_results_good[c(3:6)], digits = c(2, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S4:** Goodness of fit of monod model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 2) %>%
  group_rows("Substrate", 3, 4) %>%
  group_rows("Each", 4, 4)

```

####Monod plus model
```{r Monod plus respiration calibration, eval=F, echo=T}
#Loading the function 
source("../Rscripts/monodplus_r.R")
source("../Rscripts/monodplus_r_fix.R")
#across all treatments
monodplus_r1<-monodplus_r(data=d, FACT = 4)
##################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
##################################################
#for different structures
monodplus_r2<-monodplus_r(data=d, FACT = 2)
monodplus_r2_fix<-monodplus_r_fix(data=d, FACT = 2)
#for different substrates
monodplus_r3<-monodplus_r(data=d, FACT = 1)
monodplus_r3_fix<-monodplus_r_fix(data=d, FACT = 1)
#for each separately 
monodplus_r4<-monodplus_r(data=d, FACT = 3)
monodplus_r4_fix<-monodplus_r_fix(data=d, FACT = 3)
##################################################
stopImplicitCluster()
##################################################
monodplus_r1$goodness
monodplus_r2$goodness
monodplus_r3$goodness
monodplus_r4$goodness

# mean(c(monodplus_r4[[1]]$pars[["Cmic_0"]], monodplus_r4[[2]]$pars[["Cmic_0"]],
#        monodplus_r4[[3]]$pars[["Cmic_0"]], monodplus_r4[[4]]$pars[["Cmic_0"]],
#        monodplus_r4[[5]]$pars[["Cmic_0"]], monodplus_r4[[6]]$pars[["Cmic_0"]]))

monodplus_r1$goodness
monodplus_r2_fix$goodness
monodplus_r3_fix$goodness
monodplus_r4_fix$goodness
```

```{r Monod plus respiration calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

#results of the previous iteration (not run)
# monodplus_r_results_par<-as.data.frame(rbind(monodplus_r1[[1]]$pars[c(1:5)],
#                     monodplus_r2_fix[[1]]$pars, monodplus_r2_fix[[2]]$pars,
#                     monodplus_r2_fix[[3]]$pars,
#                     monodplus_r3_fix[[1]]$pars, monodplus_r3_fix[[2]]$pars,
#                     monodplus_r4_fix[[1]]$pars, monodplus_r4_fix[[2]]$pars,
#                     monodplus_r4_fix[[3]]$pars, monodplus_r4_fix[[4]]$pars,
#                     monodplus_r4_fix[[5]]$pars, monodplus_r4_fix[[6]]$pars))
# monodplus_r_results_good<-as.data.frame(rbind(monodplus_r1$goodness,
#                                               monodplus_r2_fix$goodness,
#                                               monodplus_r3_fix$goodness,
#                                               monodplus_r4_fix$goodness))
# write.csv(monodplus_r_results_par, file = c("../monodplus_r_results_par.csv"), row.names = F)
# write.csv(monodplus_r_results_good, file = c("../monodplus_r_results_good.csv"), row.names = F)

monodplus_r_results_par<-read.csv("monodplus_r_results_par.csv")
monodplus_r_results_good<-read.csv("monodplus_r_results_good.csv")

t.s2.1<-data.frame(Vmax=monodplus_r_results_par[,1],
                 Km=monodplus_r_results_par[,2],
                 CUE=monodplus_r_results_par[,3],
                 mr=monodplus_r_results_par[,4],
                 k=monodplus_r_results_par[,5])
row.names(t.s2.1)<-c("All", "Broth", "Glass", "Wool",
                    "Celluloze", "Glucose",
                    "Broth - Cellobiose", "Broth - Glucose",
                    "Glass - Cellobiose", "Glass - Glucose",
                    "Wool - Cellobiose", "Wool - Glucose")

options(knitr.kable.NA = '')

kable(t.s2.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$', 'k'), row.names=T, caption = "_**Table S5:** Monod plus model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monodplus_r_results_good[,c(4, 1, 2, 3)], digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S6:** Goodness of fit of Monnod plus model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 2) %>%
  group_rows("Substrate", 3, 4) %>%
  group_rows("Each", 4, 4)

```

####Two pool model
```{r Two pool model respiration calibration, eval=F, echo=T}
#Loading the function 
source("../Rscripts/two_pool_r.R")
source("../Rscripts/two_pool_r_fix.R")
#across all treatments
two_pool_r1<-two_pool_r(data=d, FACT = 4)
#################################################
no_cors<-detectCores()
cl<-makeCluster(no_cors)
registerDoParallel(cl)
#################################################
#for different structures
two_pool_r2<-two_pool_r(data=d, FACT = 2)
two_pool_r2_fix<-two_pool_r_fix(data=d, FACT = 2)
#for different substrates
two_pool_r3<-two_pool_r(data=d, FACT = 1)
two_pool_r3_fix<-two_pool_r_fix(data=d, FACT = 1)
#for each separately 
two_pool_r4<-two_pool_r(data=d, FACT = 3)
two_pool_r4_fix<-two_pool_r_fix(data=d, FACT = 3)
#################################################
stopImplicitCluster()
#################################################
# mean(c(two_pool_r4[[1]]$pars[["R_0"]], two_pool_r4[[2]]$pars[["R_0"]],
#       two_pool_r4[[3]]$pars[["R_0"]], two_pool_r4[[4]]$pars[["R_0"]],
#       two_pool_r4[[5]]$pars[["R_0"]], two_pool_r4[[6]]$pars[["R_0"]]))
# mean(c(two_pool_r4[[1]]$pars[["S_0"]], two_pool_r4[[2]]$pars[["S_0"]],
#       two_pool_r4[[3]]$pars[["S_0"]], two_pool_r4[[4]]$pars[["S_0"]],
#       two_pool_r4[[5]]$pars[["S_0"]], two_pool_r4[[6]]$pars[["S_0"]]))

two_pool_r1$goodness
two_pool_r2_fix$goodness
two_pool_r3_fix$goodness
two_pool_r4_fix$goodness
```

```{r Two pool model respiration calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

#results of the previous iteration (not run)
# two_pool_r_results_par<-rbind(two_pool_r1[[1]]$pars[c(1:6)],
#                     two_pool_r2_fix[[1]]$pars, two_pool_r2_fix[[2]]$pars,
#                     two_pool_r2_fix[[3]]$pars,
#                     two_pool_r3_fix[[1]]$pars, two_pool_r3_fix[[2]]$pars,
#                     two_pool_r4_fix[[1]]$pars, two_pool_r4_fix[[2]]$pars,
#                     two_pool_r4_fix[[3]]$pars, two_pool_r4_fix[[4]]$pars,
#                     two_pool_r4_fix[[5]]$pars, two_pool_r4_fix[[6]]$pars)
# two_pool_r_results_good<-rbind(two_pool_r1$goodness,
#                                two_pool_r2$goodness,
#                                two_pool_r3$goodness,
#                                two_pool_r4$goodness)
# write.csv(two_pool_r_results_par, file = c("../two_pool_r_results_par.csv"), row.names = F)
# write.csv(two_pool_r_results_good, file = c("../two_pool_r_results_good.csv"), row.names = F)

two_pool_r_results_par<-read.csv("two_pool_r_results_par.csv")
two_pool_r_results_good<-read.csv("two_pool_r_results_good.csv")

t.s3.1<-data.frame(Vmax=two_pool_r_results_par[,1],
                 Km=two_pool_r_results_par[,2],
                 mo=two_pool_r_results_par[,3],
                 f=two_pool_r_results_par[,4],
                 k=two_pool_r_results_par[,5],
                 Yu=two_pool_r_results_par[,6])
row.names(t.s3.1)<-c("All", "Broth", "Glass", "Wool",
                    "Cellobiose", "Glucose",
                    "Broth - Cellobiose", "Broth - Glucose",
                    "Glass - Cellobiose", "Glass - Glucose",
                    "Wool - Cellobiose", "Wool - Glucose")

options(knitr.kable.NA = '')

kable(t.s3.1, digits = c(2, 2, 3, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', '$\\m_{R}$', '$\\f_{0}$', 'k', '$\\Y_{S}$'), row.names=T, caption = "_**Table S7:** Two pool model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(two_pool_r_results_good[, c(3:6)], digits = c(2, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S8:** Goodness of fit of Two pool model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 2) %>%
  group_rows("Substrate", 3, 4) %>%
  group_rows("Each", 4, 4)

```

###All models are calibrated against respiration rate, proteins and DNA concentration in the cells.
All models are calibrated against all data we measured. In addition to model parameters reported in tables above, conversion factors between proteins/DNA concentration and microbial biomass are estimated. For Monod and Monod plus models, initial concentration of microbial biomass is defined as an initial concentration of the DNA in the incubation vial divided by the conversion factor between the DNA concentration and microbial biomass. For Two pool model, initial concentration of Structures is defined similarly - i.e. initial concentration of the DNA in the incubation vial divided by the conversion factor between the DNA concentration and Structures. However, initial concetration of Reserves cannot be defined because initial concentration of proteins is not know. Therefore, one more parameter is defined. This parameter is denoted as RS_0 in the respective function and define initial concentration of Reserves per unit of Structures. This parameter has to be fixed across treatments to reflect the reality.
In addition, Monod and Monod plus models are calibrated against respiration rate and proteins or DNA concentration separately. When calibrated againnst protein concentration, initial concentration of microbial biomass is estimated and fixed across treatments.

####Monod model
```{r Monod model - biomass components, eval=F, echo=T}
#Proteins
#Loading the function 
source("../Rscripts/monod_prot.R")
source("../Rscripts/monod_prot_fix.R")
#across all treatments
monod_prot1<-monod_prot(data=d, FACT = 4)
###############################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
###############################################################
#for different structures
monod_prot2<-monod_prot(data=d, FACT = 2)
monod_prot2_fix<-monod_prot_fix(data=d, FACT = 2)
#for different substrates
monod_prot3<-monod_prot(data=d, FACT = 1)
monod_prot3_fix<-monod_prot_fix(data=d, FACT = 1)
#for each separately 
monod_prot4<-monod_prot(data=d, FACT = 3)
monod_prot4_fix<-monod_prot_fix(data=d, FACT = 3)
###############################################################
stopImplicitCluster()
###############################################################
monod_prot1$goodness
monod_prot2_fix$goodness
monod_prot3_fix$goodness
monod_prot4_fix$goodness

###############################################################
#DNA
#Loading the function 
source("../Rscripts/monod_i_DNA.R")
#across all treatments
monod_DNA1<-monod_DNA(data=d, FACT = 4)
###############################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
###############################################################
#for different structures
monod_DNA2<-monod_DNA(data=d, FACT = 2)
#for different substrates
monod_DNA3<-monod_DNA(data=d, FACT = 1)
#for each separately 
monod_DNA4<-monod_DNA(data=d, FACT = 3)
###############################################################
stopImplicitCluster()
###############################################################
monod_DNA1$goodness
monod_DNA2$goodness
monod_DNA3$goodness
monod_DNA4$goodness
###############################################################
#Proteins and DNA
#Loading the function 
source("../Rscripts/monod_all.R")
#across all treatments
monod_all1<-monod_all(data=d, FACT = 4)
###############################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
###############################################################
#for different structures
monod_all2<-monod_all(data=d, FACT = 2)
#for different substrates
monod_all3<-monod_all(data=d, FACT = 1)
#for each separately 
monod_all4<-monod_all(data=d, FACT = 3)
###############################################################
stopImplicitCluster()
###############################################################
monod_all1$goodness
monod_all2$goodness
monod_all3$goodness
monod_all4$goodness
```

```{r Monod overall calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# #results of the previous iteration (not run)
# ##All measurements together
# monod_all_results_par<-rbind(monod_all1[[1]]$pars, 
#                              monod_all2[[1]]$pars, monod_all2[[2]]$pars,
#                              monod_all2[[3]]$pars, 
#                              monod_all3[[1]]$pars, monod_all3[[2]]$pars,
#                              monod_all4[[1]]$pars, monod_all4[[2]]$pars,
#                              monod_all4[[3]]$pars, monod_all4[[4]]$pars,
#                              monod_all4[[5]]$pars, monod_all4[[6]]$pars)
# 
# monod_all_results_good<-rbind(monod_all1$goodness,
#                               monod_all2$goodness,
#                               monod_all3$goodness,
#                               monod_all4$goodness)
# 
# write.csv(monod_all_results_par, file = c("../monod_all_results_par.csv"), row.names = F)
# write.csv(monod_all_results_good, file = c("../monod_all_results_good.csv"), row.names = F)
# 
# ##Respiration rate and proteins
# monod_prot_results_par<-rbind(monod_prot1[[1]]$pars[c(1:5)], 
#                              monod_prot2_fix[[1]]$pars, monod_prot2_fix[[2]]$pars,
#                              monod_prot2_fix[[3]]$pars, 
#                              monod_prot3_fix[[1]]$pars, monod_prot3_fix[[2]]$pars,
#                              monod_prot4_fix[[1]]$pars, monod_prot4_fix[[2]]$pars,
#                              monod_prot4_fix[[3]]$pars, monod_prot4_fix[[4]]$pars,
#                              monod_prot4_fix[[5]]$pars, monod_prot4_fix[[6]]$pars)
# 
# monod_prot_results_good<-rbind(monod_prot1$goodness,
#                              monod_prot2_fix$goodness,
#                              monod_prot3_fix$goodness,
#                              monod_prot4_fix$goodness)
# 
# write.csv(monod_prot_results_par, file = c("../monod_prot_results_par.csv"), row.names = F)
# write.csv(monod_prot_results_good, file = c("../monod_prot_results_good.csv"), row.names = F)
# 
# ##Respiration rate and DNA
# monod_DNA_results_par<-rbind(monod_DNA1[[1]]$pars, 
#                              monod_DNA2[[1]]$pars, monod_DNA2[[2]]$pars,
#                              monod_DNA2[[3]]$pars, 
#                              monod_DNA3[[1]]$pars, monod_DNA3[[2]]$pars,
#                              monod_DNA4[[1]]$pars, monod_DNA4[[2]]$pars,
#                              monod_DNA4[[3]]$pars, monod_DNA4[[4]]$pars,
#                              monod_DNA4[[5]]$pars, monod_DNA4[[6]]$pars)
# 
# monod_DNA_results_good<-rbind(monod_DNA1$goodness,
#                               monod_DNA2$goodness,
#                               monod_DNA3$goodness,
#                               monod_DNA4$goodness)
# 
# write.csv(monod_DNA_results_par, file = c("../monod_DNA_results_par.csv"), row.names = F)
# write.csv(monod_DNA_results_good, file = c("../monod_DNA_results_good.csv"), row.names = F)

#All measurements
monod_all_results_par<-read.csv("monod_all_results_par.csv")
monod_all_results_good<-read.csv("monod_all_results_good.csv")
monod_all_results_good$variable<-rep(c("Respiration", "Proteins", "DNA"), 4)
monod_all_results_good[monod_all_results_good$R2<0, "R2"]<-0
#Respiration and proteins
monod_prot_results_par<-read.csv("monod_prot_results_par.csv")
monod_prot_results_good<-read.csv("monod_prot_results_good.csv")
monod_prot_results_good$variable<-rep(c("Respiration", "Proteins"), 4)
monod_prot_results_good[monod_prot_results_good$R2<0, "R2"]<-0
#Respiration and DNA
monod_DNA_results_par<-read.csv("monod_DNA_results_par.csv")
monod_DNA_results_good<-read.csv("monod_DNA_results_good.csv")
monod_DNA_results_good$variable<-rep(c("Respiration", "DNA"), 4)
monod_DNA_results_good[monod_DNA_results_good$R2<0, "R2"]<-0

t.s4.1<-data.frame(Name=c("All", "Broth", "Glass", "Wool",
                    "Cellobiose", "Glucose",
                    "Broth - Cellobiose", "Broth - Glucose",
                    "Glass - Cellobiose", "Glass - Glucose",
                    "Wool - Cellobiose", "Wool - Glucose"),
                   Vmax=monod_all_results_par[,1],
                   Km=monod_all_results_par[,2],
                   CUE=monod_all_results_par[,3],
                   k=monod_all_results_par[,4],
                   fp=monod_all_results_par[,5],
                   fd=monod_all_results_par[,6])

options(knitr.kable.NA = '')

kable(t.s4.1, digits = c(0, 2, 2, 2, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$', '$\\f_{Proteins}$', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S9:** Monod model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monod_all_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S10:** Goodness of fit of Monod model calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 3) %>%
  group_rows("Structure", 4, 6) %>%
  group_rows("Substrate", 7, 9) %>%
  group_rows("Each", 10, 12)

t.s4.2<-data.frame(Name=c("All", "Broth", "Glass", "Wool",
                    "Cellobiose", "Glucose",
                    "Broth - Cellobiose", "Broth - Glucose",
                    "Glass - Cellobiose", "Glass - Glucose",
                    "Wool - Cellobiose", "Wool - Glucose"),
                   Vmax=monod_prot_results_par[,1],
                   Km=monod_prot_results_par[,2],
                   CUE=monod_prot_results_par[,3],
                   k=monod_prot_results_par[,4],
                   fp=monod_prot_results_par[,5])

options(knitr.kable.NA = '')

kable(t.s4.2, digits = c(0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$', '$\\f_{Proteins}$'), row.names=F, caption = "_**Table S11:** Monod model parameters calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monod_prot_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S12:** Goodness of fit of Monod model calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 3, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 8)

t.s4.3<-data.frame(Name=c("All", "Broth", "Glass", "Wool",
                    "Cellobiose", "Glucose",
                    "Broth - Cellobiose", "Broth - Glucose",
                    "Glass - Cellobiose", "Glass - Glucose",
                    "Wool - Cellobiose", "Wool - Glucose"),
                   Vmax=monod_DNA_results_par[,1],
                   Km=monod_DNA_results_par[,2],
                   CUE=monod_DNA_results_par[,3],
                   k=monod_DNA_results_par[,4],
                   fd=monod_DNA_results_par[,5])

options(knitr.kable.NA = '')

kable(t.s4.3, digits = c(0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S13:** Monod model parameters calibrated against respiration rate and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monod_DNA_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S14:** Goodness of fit of Monod model calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 2) %>%
  group_rows("Structure", 3, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 8)

```

####Monod plus model
```{r Monod plus model calibartion, eval=F, echo=T}
#Proteins
#Loading the function 
source("../Rscripts/monodplus_prot.R")
source("../Rscripts/monodplus_prot_fix.R")
#across all treatments
monodplus_prot1<-monodplus_prot(data=d, FACT = 4)
###########################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
###########################################################
#for different structures
monodplus_prot2<-monodplus_prot(data=d, FACT = 2)
monodplus_prot2_fix<-monodplus_prot_fix(data=d, FACT = 2)
#for different substrates
monodplus_prot3<-monodplus_prot(data=d, FACT = 1)
monodplus_prot3_fix<-monodplus_prot_fix(data=d, FACT = 1)
#for each separately 
monodplus_prot4<-monodplus_prot(data=d, FACT = 3)
monodplus_prot4_fix<-monodplus_prot_fix(data=d, FACT = 3)
###########################################################
stopImplicitCluster()
###########################################################
monodplus_prot1$goodness
monodplus_prot2_fix$goodness
monodplus_prot3_fix$goodness
monodplus_prot4_fix$goodness
###########################################################
#DNA
#Loading the function 
source("../Rscripts/monodplus_DNA.R")
#across all treatments
monodplus_DNA1<-monodplus_DNA(data=d, FACT = 4)
###########################################################
no_cors<-detectCores()-2
cl<-makeCluster(no_cors)
registerDoParallel(cl)
###########################################################
#for different structures
monodplus_DNA2<-monodplus_DNA(data=d, FACT = 2)
#for different substrates
monodplus_DNA3<-monodplus_DNA(data=d, FACT = 1)
#for each separately 
monodplus_DNA4<-monodplus_DNA(data=d, FACT = 3)
###########################################################
stopImplicitCluster()
###########################################################
monodplus_DNA1$goodness
monodplus_DNA2$goodness
monodplus_DNA3$goodness
monodplus_DNA4$goodness
###########################################################
#DNA and Proteins
#Loading the function 
source("../Rscripts/monodplus_all.R")
#across all treatments
monodplus_all1<-monodplus_all(data=d, FACT = 4)
###########################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
###########################################################
#for different structures
monodplus_all2<-monodplus_all(data=d, FACT = 2)
#for different substrates
monodplus_all3<-monodplus_all(data=d, FACT = 1)
#for each separately 
monodplus_all4<-monodplus_all(data=d, FACT = 3)
###########################################################
stopImplicitCluster()
###########################################################
monodplus_all1$goodness
monodplus_all2$goodness
monodplus_all3$goodness
monodplus_all4$goodness
```

```{r Monod plus overall calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

# #results of the previous iteration (not run)
# ##All measurements together
# monodplus_all_results_par<-rbind(monodplus_all1[[1]]$pars, 
#                             monodplus_all2[[1]]$pars, monodplus_all2[[2]]$pars,
#                             monodplus_all2[[3]]$pars, 
#                             monodplus_all3[[1]]$pars, monodplus_all3[[2]]$pars,
#                             monodplus_all4[[1]]$pars, monodplus_all4[[2]]$pars,
#                             monodplus_all4[[3]]$pars, monodplus_all4[[4]]$pars,
#                             monodplus_all4[[5]]$pars, monodplus_all4[[6]]$pars)
# 
# monodplus_all_results_good<-rbind(monodplus_all1$goodness[c(2, 3, 1), ],
#                              monodplus_all2$goodness[c(2, 3, 1), ],
#                              monodplus_all3$goodness[c(2, 3, 1), ],
#                              monodplus_all4$goodness[c(2, 3, 1), ])
# 
# write.csv(monodplus_all_results_par, file = c("../monodplus_all_results_par.csv"), row.names = F)
# write.csv(monodplus_all_results_good, file = c("../monodplus_all_results_good.csv"), row.names = F)
# 
# ##Respiration rate and proteins
# monodplus_prot_results_par<-rbind(monodplus_prot1[[1]]$pars[c(1:6)], 
#                             monodplus_prot2_fix[[1]]$pars, monodplus_prot2_fix[[2]]$pars,
#                             monodplus_prot2_fix[[3]]$pars, 
#                             monodplus_prot3_fix[[1]]$pars, monodplus_prot3_fix[[2]]$pars,
#                             monodplus_prot4_fix[[1]]$pars, monodplus_prot4_fix[[2]]$pars,
#                             monodplus_prot4_fix[[3]]$pars, monodplus_prot4_fix[[4]]$pars,
#                             monodplus_prot4_fix[[5]]$pars, monodplus_prot4_fix[[6]]$pars)
# 
# monodplus_prot_results_good<-rbind(monodplus_prot1$goodness[c(2,1), ],
#                              monodplus_prot2_fix$goodness[c(2,1), ],
#                              monodplus_prot3_fix$goodness[c(2,1), ],
#                              monodplus_prot4_fix$goodness[c(2,1), ])
# 
# write.csv(monodplus_prot_results_par, file = c("../monodplus_prot_results_par.csv"), row.names = F)
# write.csv(monodplus_prot_results_good, file = c("../monodplus_prot_results_good.csv"), row.names = F)
# 
# ##Respiration rate and DNA
# monodplus_DNA_results_par<-rbind(monodplus_DNA1[[1]]$pars, 
#                             monodplus_DNA2[[1]]$pars, monodplus_DNA2[[2]]$pars,
#                             monodplus_DNA2[[3]]$pars, 
#                             monodplus_DNA3[[1]]$pars, monodplus_DNA3[[2]]$pars,
#                             monodplus_DNA4[[1]]$pars, monodplus_DNA4[[2]]$pars,
#                             monodplus_DNA4[[3]]$pars, monodplus_DNA4[[4]]$pars,
#                             monodplus_DNA4[[5]]$pars, monodplus_DNA4[[6]]$pars)
#  
# monodplus_DNA_results_good<-rbind(monodplus_DNA1$goodness[c(2, 1), ],
#                              monodplus_DNA2$goodness[c(2, 1), ],
#                              monodplus_DNA3$goodness[c(2, 1), ],
#                              monodplus_DNA4$goodness[c(2, 1), ])
# 
# write.csv(monodplus_DNA_results_par, file = c("../monodplus_DNA_results_par.csv"), row.names = F)
# write.csv(monodplus_DNA_results_good, file = c("../monodplus_DNA_results_good.csv"), row.names = F)

#All measurements
monodplus_all_results_par<-read.csv("monodplus_all_results_par.csv")
monodplus_all_results_good<-read.csv("monodplus_all_results_good.csv")
monodplus_all_results_good$variable<-rep(c("Respiration", "Proteins", "DNA"), 4)
monodplus_all_results_good[monodplus_all_results_good$R2<0, "R2"]<-0
#Respiration and proteins
monodplus_prot_results_par<-read.csv("monodplus_prot_results_par.csv")
monodplus_prot_results_good<-read.csv("monodplus_prot_results_good.csv")
monodplus_prot_results_good$variable<-rep(c("Respiration", "Proteins"), 4)
monodplus_prot_results_good[monodplus_prot_results_good$R2<0, "R2"]<-0
#Respiration and DNA
monodplus_DNA_results_par<-read.csv("monodplus_DNA_results_par.csv")
monodplus_DNA_results_good<-read.csv("monodplus_DNA_results_good.csv")
monodplus_DNA_results_good$variable<-rep(c("Respiration", "DNA"), 4)
monodplus_DNA_results_good[monodplus_DNA_results_good$R2<0, "R2"]<-0

t.s5.1<-data.frame(Name=c("All", "Broth", "Glass", "Wool",
                    "Cellobiose", "Glucose",
                    "Broth - Cellobiose", "Broth - Glucose",
                    "Glass - Cellobiose", "Glass - Glucose",
                    "Wool - Cellobiose", "Wool - Glucose"),
                   Vmax=monodplus_all_results_par[,1],
                   Km=monodplus_all_results_par[,2],
                   CUE=monodplus_all_results_par[,3],
                   mr=monodplus_all_results_par[,4],
                   k=monodplus_all_results_par[,5],
                   fp=monodplus_all_results_par[,6],
                   fd=monodplus_all_results_par[,7])

options(knitr.kable.NA = '')

kable(t.s5.1, digits = c(0, 2, 2, 2, 3, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$', '$\\k_{MB}$', '$\\f_{Proteins}$', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S15:** Monod plus model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monodplus_all_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S16:** Goodness of fit of Monod plus model calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 3) %>%
  group_rows("Structure", 4, 6) %>%
  group_rows("Substrate", 7, 9) %>%
  group_rows("Each", 10, 12)

t.s5.2<-data.frame(Name=c("All", "Broth", "Glass", "Wool",
                    "Cellobiose", "Glucose",
                    "Broth - Cellobiose", "Broth - Glucose",
                    "Glass - Cellobiose", "Glass - Glucose",
                    "Wool - Cellobiose", "Wool - Glucose"),
                   Vmax=monodplus_prot_results_par[,1],
                   Km=monodplus_prot_results_par[,2],
                   CUE=monodplus_prot_results_par[,3],
                   mr=monodplus_prot_results_par[,4],
                   k=monodplus_prot_results_par[,5],
                   fp=monodplus_prot_results_par[,6])

options(knitr.kable.NA = '')

kable(t.s5.2, digits = c(0, 2, 2, 2, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$', '$\\k_{MB}$', '$\\f_{Proteins}$'), row.names=F, caption = "_**Table S17:** Monod plus model parameters calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monodplus_prot_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S18:** Goodness of fit of Monod plus model calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 3, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 8)

t.s5.3<-data.frame(Name=c("All", "Broth", "Glass", "Wool",
                    "Cellobiose", "Glucose",
                    "Broth - Cellobiose", "Broth - Glucose",
                    "Glass - Cellobiose", "Glass - Glucose",
                    "Wool - Cellobiose", "Wool - Glucose"),
                   Vmax=monodplus_DNA_results_par[,1],
                   Km=monodplus_DNA_results_par[,2],
                   CUE=monodplus_DNA_results_par[,3],
                   mr=monodplus_DNA_results_par[,4],
                   k=monodplus_DNA_results_par[,5],
                   fd=monodplus_DNA_results_par[,6])

options(knitr.kable.NA = '')

kable(t.s5.3, digits = c(0, 2, 2, 2, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$', '$\\k_{MB}$', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S19:** Monod plus model parameters calibrated against respiration rate and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(monodplus_DNA_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S20:** Goodness of fit of Monod plus  model calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 2) %>%
  group_rows("Structure", 3, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 8)
```

####Two pool model
Not all model parameters vary significantly acros treatments. Parameters that doesn't show any variability across treatments are fixed.
```{r Two pool model calibration, eval=F, echo=T}
#Loading the function 
source("../Rscripts/two_pool_all.R")
source("../Rscripts/two_pool_all_fix.R")
#across all treatments
two_pool_all1<-two_pool_all(data=d, FACT = 4)
###########################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
###########################################################
#for different structures
two_pool_all2<-two_pool_all(data=d, FACT = 2)
two_pool_all2_fix<-two_pool_all_fix(data=d, FACT = 2)
############################################################################
#in the following code, the Ys and k parameters are fixed across treatments
#becuase its variability across treatments is minimal (data not shown)
source("./Rscripts/two_pool_all_fixYuk.R")
two_pool_all_fixYuk2<-two_pool_all_fixYuk(data=d, FACT = 2)
#Then, parameter f is fixed as well
source("./Rscripts/two_pool_all_fixYukf.R")
two_pool_all_fixYukf2<-two_pool_all_fixYukf(data=d, FACT = 2)
#Then, parameter Vmax is fixed as well
source("./Rscripts/two_pool_all_fixYukfVmax.R")
two_pool_all_fixYukfVmax2<-two_pool_all_fixYukfVmax(data=d, FACT = 2)
#At the end, parameter KM is fixed
source("./Rscripts/two_pool_all_fixYukfVmaxKm.R")
two_pool_all_fixYukfVmaxKm2<-two_pool_all_fixYukfVmaxKm(data=d, FACT = 2)
############################################################################
#for different substrates
two_pool_all3<-two_pool_all(data=d, FACT = 1)
two_pool_all3_fix<-two_pool_all_fix(data=d, FACT = 1)
#for each separately 
two_pool_all4<-two_pool_all(data=d, FACT = 3)
two_pool_all4_fix<-two_pool_all_fix(data=d, FACT = 3)
#in the following code, the Ys, k,f and Vmax parameters are fixed across treatments
two_pool_all_fixYukfVmax4<-two_pool_all_fixYukfVmax(data=d, FACT = 3)
###########################################################
stopImplicitCluster()
###########################################################
two_pool_all1$goodness
two_pool_all_fixYukfVmaxKm2$goodness
two_pool_all3_fix$goodness
two_pool_all_fixYukfVmax4$goodness
```

```{r Two pool model overall calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

#results of the previous iteration (not run)
# two_pool_all_results_par<-rbind(c(two_pool_all1[[1]]$pars[c(1:9)], RS_0 = 0.06185237),
#                           c(Vmax = 0.2761818, 
#                             Km = 6.3136,
#                             two_pool_all_fixYukfVmaxKm2[[1]]$pars[1],
#                             f = 0.2373511,
#                             k = 0.004915818,
#                             Yu = 0.3565256,
#                             two_pool_all_fixYukfVmaxKm2[[1]]$pars[c(2:4)],
#                             RS_0 = 0.06185237),
#                           c(Vmax = 0.2761818, 
#                             Km = 6.3136,
#                             two_pool_all_fixYukfVmaxKm2[[2]]$pars[1],
#                             f = 0.2373511,
#                             k = 0.004915818,
#                             Yu = 0.3565256,
#                             two_pool_all_fixYukfVmaxKm2[[2]]$pars[c(2:4)],
#                             RS_0 = 0.06185237),
#                           c(Vmax = 0.2761818, 
#                             Km = 6.3136,
#                             two_pool_all_fixYukfVmaxKm2[[3]]$pars[1],
#                             f = 0.2373511,
#                             k = 0.004915818,
#                             Yu = 0.3565256,
#                             two_pool_all_fixYukfVmaxKm2[[3]]$pars[c(2:4)],
#                             RS_0 = 0.06185237),
#                           c(two_pool_all3_fix[[1]]$pars[c(1:9)], RS_0 = 0.06185237),
#                           c(two_pool_all3_fix[[2]]$pars[c(1:9)], RS_0 = 0.06185237),
#                           c(Vmax = 0.2761818, 
#                             two_pool_all_fixYukfVmax4[[1]]$pars[c(1:2)],
#                             f = 0.2373511,
#                             k = 0.004915818,
#                             Yu = 0.3565256,
#                             two_pool_all_fixYukfVmax4[[1]]$pars[c(3:5)],
#                             RS_0 = 0.06185237),
#                           c(Vmax = 0.2761818, 
#                             two_pool_all_fixYukfVmax4[[2]]$pars[c(1:2)],
#                             f = 0.2373511,
#                             k = 0.004915818,
#                             Yu = 0.3565256,
#                             two_pool_all_fixYukfVmax4[[2]]$pars[c(3:5)],
#                             RS_0 = 0.06185237),
#                           c(Vmax = 0.2761818, 
#                             two_pool_all_fixYukfVmax4[[3]]$pars[c(1:2)],
#                             f = 0.2373511,
#                             k = 0.004915818,
#                             Yu = 0.3565256,
#                             two_pool_all_fixYukfVmax4[[3]]$pars[c(3:5)],
#                             RS_0 = 0.06185237),
#                           c(Vmax = 0.2761818, 
#                             two_pool_all_fixYukfVmax4[[4]]$pars[c(1:2)],
#                             f = 0.2373511,
#                             k = 0.004915818,
#                             Yu = 0.3565256,
#                             two_pool_all_fixYukfVmax4[[4]]$pars[c(3:5)],
#                             RS_0 = 0.06185237),
#                           c(Vmax = 0.2761818, 
#                             two_pool_all_fixYukfVmax4[[5]]$pars[c(1:2)],
#                             f = 0.2373511,
#                             k = 0.004915818,
#                             Yu = 0.3565256,
#                             two_pool_all_fixYukfVmax4[[5]]$pars[c(3:5)],
#                             RS_0 = 0.06185237),
#                           c(Vmax = 0.2761818, 
#                             two_pool_all_fixYukfVmax4[[6]]$pars[c(1:2)],
#                             f = 0.2373511,
#                             k = 0.004915818,
#                             Yu = 0.3565256,
#                             two_pool_all_fixYukfVmax4[[6]]$pars[c(3:5)],
#                             RS_0 = 0.06185237)
#                           )
# two_pool_all_results_good<-rbind(two_pool_all1$goodness,
#                                  two_pool_all_fixYukfVmaxKm2$goodness,
#                                  two_pool_all3_fix$goodness,
#                                  two_pool_all_fixYukfVmax4$goodness)
# 
# write.csv(two_pool_all_results_par, file = c("../two_pool_all_results_par.csv"), row.names = F)
# write.csv(two_pool_all_results_good, file = c("../two_pool_all_results_good.csv"), row.names = F)

two_pool_all_results_par<-read.csv("two_pool_all_results_par.csv")
two_pool_all_results_good<-read.csv("two_pool_all_results_good.csv")
two_pool_all_results_good$variable<-rep(c("Respiration", "Proteins", "DNA"), 4)
two_pool_all_results_good[two_pool_all_results_good$R2<0, "R2"]<-0

t.s6.1<-data.frame(Name=c("All", "Broth", "Glass", "Wool",
                    "Cellobiose", "Glucose",
                    "Broth - Cellobiose", "Broth - Glucose",
                    "Glass - Cellobiose", "Glass - Glucose",
                    "Wool - Cellobiose", "Wool - Glucose"),
                   Vmax=two_pool_all_results_par[,1],
                   Km=two_pool_all_results_par[,2],
                   m0=two_pool_all_results_par[,3],
                   f=two_pool_all_results_par[,4],
                   k=two_pool_all_results_par[,5],
                   Yu=two_pool_all_results_par[,6],
                   fpr=two_pool_all_results_par[,7],
                   fps=two_pool_all_results_par[,8],
                   fds=two_pool_all_results_par[,9],
                   RS_0=two_pool_all_results_par[,10])

options(knitr.kable.NA = '')

kable(t.s6.1, digits = c(0, 2, 2, 3, 3, 3, 2, 3, 3, 3, 2), align = 'c', "html", booktabs=T, col.names = c(" ","$\\V_{MAX}$", '$\\K_{M}$', '$\\m_{R}$', '$\\f_{0}$', 'k', '$\\Y_{S}$', '$\\R_{Proteins}$', '$\\S_{Proteins}$', '$\\S_{DNA}$', '$\\RS_{0}$'), row.names=F, caption = "_**Table S21:** Two pool model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 1) %>%
  group_rows("Structure", 2, 4) %>%
  group_rows("Substrate", 5, 6) %>%
  group_rows("Each", 7, 12)

kable(two_pool_all_results_good[, c(1, 4:7)], digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S21:** Goodness of fit of Two pool model calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each). See Materials and Methods for symbols explanation._") %>% 
  kable_styling() %>% 
  group_rows("All", 1, 3) %>%
  group_rows("Structure", 4, 6) %>%
  group_rows("Substrate", 7, 9) %>%
  group_rows("Each", 10, 12)

```

$~$
$~$
$~$
$~$
$~$
$~$
```{r correlation figure, echo=FALSE, fig.height=5, fig.width=10, fig.cap="Fig. S3: Effect of structure of bacterial and fungal part of microbial community (expressed as multivariate mean distance between groups; see Matrial and Methods section) on overall microbial community maintnance respiration rate ($m_{R}$ - (a)) and protein concentration in Structures pool ($S_{Proteins}$ - (b)). The value of both parameters are expressed as a mean difference between groups in order to allow unbiased correlation with microbial community structure metric (see Matrial and Methods section)."}

dpars<-t.s6.1[c(2:4), -1]
dpars$Structure<-c("BROTH", "GLASS", "WOOL")

#Bacteria
bacm<-as.matrix(bac.dist)
#Intercept - BROTH
mBB<-bacm[which(bac_env$Structure=="BROTH"), 
       which(bac_env$Structure=="BROTH")]
mBB[mBB==0]<-NA
#BROTH - GLASS
mBG<-bacm[which(bac_env$Structure=="BROTH"), 
       which(bac_env$Structure=="GLASS")]

#BRTOH - WOOL
mBW<-bacm[which(bac_env$Structure=="BROTH"), 
       which(bac_env$Structure=="WOOL")]

bacBGW<-data.frame(Dist=c(0, mean(mBG, na.rm=T), mean(mBW, na.rm=T)),
                    Dist.sd=c(sd(mBB, na.rm=T), sd(mBG, na.rm=T), sd(mBW, na.rm=T)),
                    Structure=c("BROTH", "GLASS", "WOOL"))

#Fungi
#Intercept - BROTH
fBB<-fm[which(fungi_env$Structure=="BROTH"), 
       which(fungi_env$Structure=="BROTH")]
fBB[fBB==0]<-NA
#BROTH - GLASS
fBG<-fm[which(fungi_env$Structure=="BROTH"), 
       which(fungi_env$Structure=="GLASS")]

#BRTOH - WOOL
fBW<-fm[which(fungi_env$Structure=="BROTH"), 
       which(fungi_env$Structure=="WOOL")]

fBGW<-data.frame(Dist=c(0, mean(fBG, na.rm=T), mean(fBW, na.rm=T)),
                    Dist.sd=c(sd(fBB, na.rm=T), sd(fBG, na.rm=T), sd(fBW, na.rm=T)),
                    Structure=c("BROTH", "GLASS", "WOOL"))

dpars2<-merge(dpars, bacBGW, by.y=c("Structure"), by.x=c("Structure"))

dpars3<-merge(dpars, fBGW, by.y=c("Structure"), by.x=c("Structure"))

for(i in 3:13){
  dpars2[, i]<-sqrt((dpars2[, i]-as.numeric(dpars2[1, i]))^2)
}

for(i in 3:13){
  dpars3[, i]<-sqrt((dpars3[, i]-as.numeric(dpars3[1, i]))^2)
}

dpars2$group<-c("Bacteria")
dpars3$group<-c("Fungi")
dpars4<-rbind(dpars2, dpars3)

#corplot1.p<-coef(nls(m0~Dist^x, dpars4, start = list(x=1)))

corplot1<-ggplot(dpars4[dpars4$group=="Bacteria", ], aes(Dist, m0*1e4))+
  geom_point(cex=6, aes(fill=Structure), show.legend=T, shape=21)+
  geom_point(data=dpars4[dpars4$group=="Fungi", ], cex=6, 
             aes(x=Dist, y=m0*1e4, fill=Structure), show.legend=F, shape=22)+
  scale_fill_manual(values = c("black", "grey", "white"))+
  #stat_function(fun = function(x){(x^corplot1.p*1e4)}, lwd=1, colour="grey30")+
  xlab("16S and ITS Distance")+
  ggtitle("a)")+
  ylab(expression(paste(m[R]~difference)))+
  theme_min+theme(legend.title = element_blank(),
                  legend.position = c(0.2,0.8))

#corplot2.p<-coef(nls(fps~Dist^x, dpars4, start = list(x=1)))

corplot2<-ggplot(dpars4[dpars4$group=="Bacteria", ], aes(Dist, fps))+
  geom_point(cex=6, aes(fill=Structure), show.legend=F, shape=21)+
  geom_point(data=dpars4[dpars4$group=="Fungi", ], cex=6, 
             aes(x=Dist, y=fps, fill=Structure), show.legend=F, shape=22)+
  scale_fill_manual(values = c("black", "grey", "white"))+
  #stat_function(fun = function(x){(x^corplot2.p)}, lwd=1, colour="grey30")+
  xlab("16S and ITS Distance")+
  ggtitle("b)")+
  ylab(expression(paste(S[Proteins]~difference)))+
  theme_min

grid.arrange(corplot1, corplot2, nrow=1)

```

```{r Biomass predictions by DEB, eval=FALSE, include=FALSE}
DEB<-function(time, state, pars){
    with(as.list(c(state, pars)),{
      #Carbon uptake
      Cu=0.2761818*C*S/(6.3136+C)
      #Respiration
      r=m0*S+0.2373511*R*(1-0.3565256)

      dR<-Cu-0.2373511*R
      dS<-0.2373511*R*0.3565256 - m0*S - 0.004915818*S
      dC<--Cu + 0.004915818*S

      return(list(c(dR, dS, dC), r=r, Cmic=R+S))
    })
}

#Biomass prediction
##Broth
Bpred<-d[d$Structure=="Broth", c("Structure", "Time", "r")]

Bout <- as.data.frame(ode(y=c(R=0.06185237*(2.01*0.51/12.01/4/deb_k2_all_fixYukfVmaxKm[[1]]$pars[["fds"]]), 
            S=2.01*0.51/12.01/4/deb_k2_all_fixYukfVmaxKm[[1]]$pars[["fds"]], 
            C=25), 
        parms=deb_k2_all_fixYukfVmaxKm[[1]]$pars, func=DEB, times=sort(Bpred$Time))) %>% select(time, Cmic)
###Merge
Bpred<-merge(Bpred, Bout, by.x = c("Time"), by.y = c("time"))

##Mixed glass
Mpred<-d[d$Structure=="Mixed glass", c("Structure", "Time", "r")]

Mout <- as.data.frame(ode(y=c(R=0.06185237*(2.01*0.51/12.01/4/deb_k2_all_fixYukfVmaxKm[[3]]$pars[["fds"]]), 
            S=2.01*0.51/12.01/4/deb_k2_all_fixYukfVmaxKm[[3]]$pars[["fds"]], 
            C=25), 
        parms=deb_k2_all_fixYukfVmaxKm[[3]]$pars, func=DEB, times=sort(Bpred$Time))) %>% select(time, Cmic)
###Merge
Mpred<-merge(Mpred, Mout, by.x = c("Time"), by.y = c("time"))

##Glass wool
Gpred<-d[d$Structure=="Glass wool", c("Structure", "Time", "r")]

Gout <- as.data.frame(ode(y=c(R=0.06185237*(2.01*0.51/12.01/4/deb_k2_all_fixYukfVmaxKm[[2]]$pars[["fds"]]), 
            S=2.01*0.51/12.01/4/deb_k2_all_fixYukfVmaxKm[[2]]$pars[["fds"]], 
            C=25), 
        parms=deb_k2_all_fixYukfVmaxKm[[2]]$pars, func=DEB, times=sort(Bpred$Time))) %>% select(time, Cmic)
###Merge
Gpred<-merge(Gpred, Gout, by.x = c("Time"), by.y = c("time"))

Pd<-rbind(Bpred, Mpred, Gpred)

######################################################
no_cors<-detectCores()-1
cl<-makeCluster(no_cors)
registerDoParallel(cl)
######################################################
#Use monod model to fit the data
source("./Rscripts/monod_test.R")
monod_test_out<-monod_test(data = Pd, FACT = 2)
monod_test_out$goodness

#Use monod model to fit the data
source("../Rscripts/mend_test.R")
mend_test_out<-mend_test(data = Pd, FACT = 2)
mend_test_out$goodness

stopImplicitCluster()
```

##Supplementary figure
```{r Monod and Monod plus model parameters, echo=FALSE, fig.height=24, fig.width=20, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Fig. S1: Monod (a) and Monod plus (b) model parameters ($V_{MAX}$ - maximum velocity constant, $K_{M}$ - affinity constant, CUE - carbon use efficiency, $k_{MB}$ - microbial biomass decay rate constant, $m_{R}$ - specific maintenance rate constant) estimated separately for three different types of physical barriers presented in incubation bottle (BROTH - black symbols, GLASS - grey symbols and WOOL - empty symbols). Symbols represent best parameters values estimated by the Differential Evolution Algorithm and error bars represent standard deviation of posterior parameters distribution estimated by Constrained Markov Chain Monte Carlo simulation. Parameters estimation was constrained by either observed respiration rates, observed respiration rates and cellular proteins concentration, observed respiration rates and DNA concentration or by all measurements at the same time."}

# # Monod model calibrated against the respiration rate (not run)
# #uncertainty of model parameters
# monod_r_sd<-as.data.frame(rbind(apply(monod_r2_fix[[1]]$par_prof$pars, 2, sd),
#                                apply(monod_r2_fix[[2]]$par_prof$pars, 2, sd),
#                                apply(monod_r2_fix[[3]]$par_prof$pars, 2, sd)))[,-5]
# colnames(monod_r_sd)<-c("Vmax.sd", "Km.sd", "CUE.sd", "k.sd")
# 
# #Combine with the estimated parameters
# monod_rs1<-cbind(monod_r_results_par[c(2:4), c(1:4)], monod_r_sd)
# monod_rs1$Legend<-rep("Respiration", times=3)
# 
# #Monod model calibrated against the respiration rate and the rest (not run)
# # uncertainty of model parameters
# monod_rest_sd<-as.data.frame(rbind(apply(monod_prot2_fix[[1]]$par_prof$pars, 2, sd)[c(1:4)],
#                                  apply(monod_prot2_fix[[2]]$par_prof$pars, 2, sd)[c(1:4)],
#                                  apply(monod_prot2_fix[[3]]$par_prof$pars, 2, sd)[c(1:4)],
#                                  apply(monod_DNA2[[1]]$par_prof$pars, 2, sd)[c(1:4)],
#                                  apply(monod_DNA2[[2]]$par_prof$pars, 2, sd)[c(1:4)],
#                                  apply(monod_DNA2[[3]]$par_prof$pars, 2, sd)[c(1:4)],
#                                  apply(monod_all2[[1]]$par_prof$pars, 2, sd)[c(1:4)],
#                                  apply(monod_all2[[2]]$par_prof$pars, 2, sd)[c(1:4)],
#                                  apply(monod_all2[[3]]$par_prof$pars, 2, sd)[c(1:4)]))[,-c(5,6)]
# colnames(monod_rest_sd)<-c("Vmax.sd", "Km.sd", "CUE.sd", "k.sd")
# #Combine with the estimated parameters
# monod_alls<-cbind(rbind(monod_prot_results_par[c(2:4), c(1:4)],
#                         monod_DNA_results_par[c(2:4), c(1:4)],
#                         monod_all_results_par[c(2:4), c(1:4)]), monod_rest_sd)
# 
# 
# monod_alls$Legend<-c(rep("Respiration & Proteins", times=3),
#                     rep("Respiration & DNA", times=3),
#                     rep("All", times=3))
#  
# #Combine monod_alls and monod_rs1
# monod_s<-rbind(monod_rs1, monod_alls)
# #Add labels
# monod_s$Structure<-rep(c("BROTH", "GLASS", "WOOL"), times=4)
# 
# # Monod plus model calibrated against the respiration rate (not run)
# # uncertainty of model parameters
# monodplus_r_sd<-as.data.frame(rbind(apply(monodplus_r2_fix[[1]]$par_prof$pars, 2, sd),
#                                apply(monodplus_r2_fix[[2]]$par_prof$pars, 2, sd),
#                                apply(monodplus_r2_fix[[3]]$par_prof$pars, 2, sd)))
# colnames(monodplus_r_sd)<-c("Vmax.sd", "Km.sd", "CUE.sd", "mr.sd", "k.sd")
# 
# #Combine with the estimated parameters
# monodplus_rs1<-cbind(monodplus_r_results_par[c(2:4), c(1:5)], monodplus_r_sd)
# monodplus_rs1$Legend<-rep("Respiration", times=3)
# 
# #Monod model calibrated against the respiration rate and the rest (not run)
# # uncertainty of model parameters
# monodplus_rest_sd<-as.data.frame(rbind(apply(monodplus_prot2_fix[[1]]$par_prof$pars, 2, sd)[c(1:5)],
#                                  apply(monodplus_prot2_fix[[2]]$par_prof$pars, 2, sd)[c(1:5)],
#                                  apply(monodplus_prot2_fix[[3]]$par_prof$pars, 2, sd)[c(1:5)],
#                                  apply(monodplus_DNA2[[1]]$par_prof$pars, 2, sd)[c(1:5)],
#                                  apply(monodplus_DNA2[[2]]$par_prof$pars, 2, sd)[c(1:5)],
#                                  apply(monodplus_DNA2[[3]]$par_prof$pars, 2, sd)[c(1:5)],
#                                  apply(monodplus_all2[[1]]$par_prof$pars, 2, sd)[c(1:5)],
#                                  apply(monodplus_all2[[2]]$par_prof$pars, 2, sd)[c(1:5)],
#                                  apply(monodplus_all2[[3]]$par_prof$pars, 2, sd)[c(1:5)]))
# colnames(monodplus_rest_sd)<-c("Vmax.sd", "Km.sd", "CUE.sd", "mr.sd", "k.sd")
# #Combine with the estimated parameters
# monodplus_alls<-cbind(rbind(monodplus_prot_results_par[c(2:4), c(1:5)],
#                         monodplus_DNA_results_par[c(2:4), c(1:5)],
#                         monodplus_all_results_par[c(2:4), c(1:5)]), monodplus_rest_sd)
# 
# 
# monodplus_alls$Legend<-c(rep("Respiration & Proteins", times=3),
#                     rep("Respiration & DNA", times=3),
#                     rep("All", times=3))
#  
# #Combine monod_alls and monod_rs1
# monodplus_s<-rbind(monodplus_rs1, monodplus_alls)
# #Add labels
# monodplus_s$Structure<-rep(c("BROTH", "GLASS", "WOOL"), times=4)
# 
# write.csv(monod_s, file = c("../monod_FigS1.csv"), row.names = F)
# write.csv(monodplus_s, file = c("../monodplus_FigS1.csv"), row.names = F)

monod_s<-read.csv("monod_FigS1.csv")
monodplus_s<-read.csv("monodplus_FigS1.csv")

Monod_s<-melt(monod_s[, c(1:4, 9, 10)], id.vars=c("Legend", "Structure"))
Monod_s$sd<-melt(monod_s[, c(5:8, 9, 10)], id.vars=c("Legend", "Structure"))[, "value"]

#add proper parameter labels
mlab<-c('Vmax' = 'V[MAX]',
        'Km' = 'K[M]',
        'CUE' = 'CUE',
        'k' = 'k[MB]')

figS1a<-ggplot(Monod_s, aes(Legend, value))+
  geom_point(shape=21, cex=6, aes(fill=Structure), position = position_dodge(width = 0.6))+
  geom_errorbar(aes(ymin=value-sd, ymax=value+sd, colour=Structure), 
                position = position_dodge(width = 0.6))+
  scale_fill_manual(values = c("black", "grey", "white"))+
  scale_colour_manual(values = c("black", "black", "black"))+theme_min+
  theme(axis.title.x = element_blank(),
        legend.title = element_blank(),
        axis.text.x=element_text(vjust=0.2, size=18, colour="black", angle = 90),
        legend.position = c("top"),
        plot.title=element_text(size=18, face="bold", hjust=-0.01))+
  facet_wrap(~variable, scale="free", nrow=1, 
             labeller = labeller(variable = as_labeller(mlab, label_parsed)))+
  ylab("Parameter value")+ggtitle("a)")

Monodplus_s<-melt(monodplus_s[, c(1:5, 11, 12)], id.vars=c("Legend", "Structure"))
Monodplus_s$sd<-melt(monodplus_s[, c(6:10, 11, 12)], id.vars=c("Legend", "Structure"))[, "value"]

#add proper parameter labels
mlab2<-c('Vmax' = 'V[MAX]',
        'Km' = 'K[M]',
        'CUE' = 'CUE',
        'mr' = 'm[R]',
        'k' = 'k[MB]')

figS1b<-ggplot(Monodplus_s, aes(Legend, value))+
  geom_point(shape=21, cex=6, aes(fill=Structure), position = position_dodge(width = 0.6),
             show.legend = F)+
  geom_errorbar(aes(ymin=value-sd, ymax=value+sd, colour=Structure), 
                position = position_dodge(width = 0.6), show.legend = F)+
  scale_fill_manual(values = c("black", "grey", "white"))+
  scale_colour_manual(values = c("black", "black", "black"))+theme_min+
  theme(axis.title.x = element_blank(),
        legend.title = element_blank(),
        axis.text.x=element_text(vjust=0.2, size=18, colour="black", angle = 90),
        plot.title=element_text(size=18, face="bold", hjust=-0.01))+
  facet_wrap(~variable, scale="free", nrow=1, 
             labeller = labeller(variable = as_labeller(mlab2, label_parsed)))+
  ylab("Parameter value")+ggtitle("b)")

grid.arrange(figS1a, figS1b, nrow=2)

```

### 2019-10-22 Community_analysis/R_Functions/Models_linear/Mr_lin_estMB.R

@@ -85,9 +85,9 @@ Mr_lin_estMB<-function(odeset, par_const){
   #approximate parameter estimation is done by MCMC method
   par_mcmc<-modMCMC(f=cost, p=c(par_const, I = 0.004769861, B=1),
                     lower=c(Ecat=1e-5, chi=0, Yue=0, fpr=0, 
-                            fps=0, fds=0, RSinit=0, I = 0.002871328*1e-2, B=-100),
+                            fps=0, fds=0, RSinit=0, I = 0.002871328*1e-2, B=-5),
                     upper=c(Ecat=10, chi=1, Yue=0.9, fpr=1, 
-                            fps=1, fds=1, RSinit=50, I = 0.002871328*1e2, B=100), niter=10000)
+                            fps=1, fds=1, RSinit=50, I = 0.002871328*1e2, B=5), niter=10000)
   
   #lower and upper limits for parameters are extracted
   pl<-summary(par_mcmc)["min",]

### 2019-10-22 Community_analysis/community_analysis.R

@@ -91,6 +91,21 @@ m0 %>% filter(Substrate=="Celluloze" | Substrate=="Mix") %>%
 ##Bacteria
 bac<-as.data.frame(read.csv(file=c("C:/Users/cape159/Documents/pracovni/data_statistika/minT/MinT/Community_analysis/OTU_bacteria.txt"), header=T,
                             sep="\t"))
+#bac.tax<-as.data.frame(read.csv(file=c("/home/petacapek/Dokumenty/pracovni/data_statistika/minT/MinT/Community_analysis/OTU_bacteria.txt"), header=T,
+#                             sep="\t"))
+bac.tax<-bac.tax[, c(1, 175)]
+bac.tax$X.OTU.ID<-gsub("_","", bac.tax$X.OTU.ID)
+
+#Add taxonomy to bac.norm dataset
+bac.norm_tax<-bac.norm
+taxlab<-character()
+for(i in 1:ncol(bac.norm_tax)){
+  nd<-which(bac.tax$X.OTU.ID==colnames(bac.norm_tax)[i])
+  taxlab<-append(taxlab, as.character(bac.tax$taxonomy[nd]))
+}
+
+bac.norm_tax<-rbind(bac.norm_tax, taxlab)
+
 ###without the taxonomy 
 bacr<-bac[, -175]
 

### 2019-10-22 Rscripts/monod_DNA.R

@@ -1,0 +1,238 @@
+monod_DNA<-function(data, FACT){
+  
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  
+  if(FACT==1){
+    
+    dat<-data
+    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-data
+      dat<-data
+      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
+      ids$id<-seq(1:nrow(ids))
+      
+      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
+      
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-data
+        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
+        ids$id<-seq(1:nrow(ids))
+        
+        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
+        
+        
+      }else{
+        
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  
+  #monod growth function
+  deriv<-function(time, state, pars){
+    
+    with(as.list(c(state, pars)),{
+      
+      DNAc=fd*Cmic
+      
+      dCmic<--k*Cmic+CUE*Vmax*Cmic*C/(Km+C)
+      dC<-k*Cmic-Vmax*Cmic*C/(Km+C)
+      
+      return(list(c(dCmic, dC), r=(1-CUE)*Vmax*Cmic*C/(Km+C), DNAc=DNAc))
+      
+    })
+  }
+  
+  #define names of parameters
+  parnames<-c("Vmax", "Km", "CUE", "k", "fd")
+  
+  #parameters estimation function
+  estim<-function(odeset){
+    
+    #defining cost function
+    cost<-function(x){
+      
+      par<-x[1:length(parnames)]
+      
+      names(par)<-parnames
+      
+      #first, pars dependent output from ode is matched with measured values
+      yhat_all<-as.data.frame(ode(y=c(Cmic=2.01*0.51/12.01/4/par[["fd"]], C=25), parms=par, deriv, times=sort(odeset$Time)))
+      
+      #select time and the measured variables
+      yhat<-select(yhat_all, c("time", "r", "DNAc"))
+      
+      #reformat to long format data frame
[…]

### 2019-10-22 Rscripts/monod_all.R

@@ -1,0 +1,238 @@
+monod_all<-function(data, FACT){
+  
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  
+  if(FACT==1){
+    
+    dat<-data
+    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-data
+      dat<-data
+      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
+      ids$id<-seq(1:nrow(ids))
+      
+      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
+      
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-data
+        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
+        ids$id<-seq(1:nrow(ids))
+        
+        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
+        
+        
+      }else{
+        
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  
+  #monod growth function
+  deriv<-function(time, state, pars){
+    
+    with(as.list(c(state, pars)),{
+      
+      Protc=fp*Cmic
+      DNAc=fd*Cmic
+      
+      dCmic<--k*Cmic+CUE*Vmax*Cmic*C/(Km+C)
+      dC<-k*Cmic-Vmax*Cmic*C/(Km+C)
+      
+      return(list(c(dCmic, dC), r=(1-CUE)*Vmax*Cmic*C/(Km+C), Protc=Protc, DNAc=DNAc))
+      
+    })
+  }
+  
+  #define names of parameters
+  parnames<-c("Vmax", "Km", "CUE", "k", "fp", "fd")
+  
+  #parameters estimation function
+  estim<-function(odeset){
+    
+    #defining cost function
+    cost<-function(x){
+      
+      par<-x[1:length(parnames)]
+      
+      names(par)<-parnames
+      
+      #first, pars dependent output from ode is matched with measured values
+      yhat_all<-as.data.frame(ode(y=c(Cmic=2.01*0.51/12.01/4/par[["fd"]], C=25), parms=par, deriv, times=sort(odeset$Time)))
+      
+      #select time and the measured variables 
+      yhat<-select(yhat_all, c("time", "r", "Protc", "DNAc"))
[…]

### 2019-10-22 Rscripts/monod_prot_fix.R

@@ -1,0 +1,238 @@
+monod_prot_fix<-function(data, FACT){
+  
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  
+  if(FACT==1){
+    
+    dat<-data
+    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-data
+      dat<-data
+      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
+      ids$id<-seq(1:nrow(ids))
+      
+      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
+      
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-data
+        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
+        ids$id<-seq(1:nrow(ids))
+        
+        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
+        
+        
+      }else{
+        
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  
+  #monod growth function
+  deriv<-function(time, state, pars){
+    
+    with(as.list(c(state, pars)),{
+      
+      Protc=fp*Cmic
+      
+      dCmic<--k*Cmic+CUE*Vmax*Cmic*C/(Km+C)
+      dC<-k*Cmic-Vmax*Cmic*C/(Km+C)
+      
+      return(list(c(dCmic, dC), r=(1-CUE)*Vmax*Cmic*C/(Km+C), Protc=Protc))
+      
+    })
+  }
+  
+  #define names of parameters
+  parnames<-c("Vmax", "Km", "CUE", "k", "fp")
+  
+  #parameters estimation function
+  estim<-function(odeset){
+    
+    #defining cost function
+    cost<-function(x){
+      
+      par<-x[1:length(parnames)]
+      
+      names(par)<-parnames
+      
+      #first, pars dependent output from ode is matched with measured values
+      yhat_all<-as.data.frame(ode(y=c(Cmic=0.1001656, C=25), parms=par, deriv, times=sort(odeset$Time)))
+      
+      #select time and the measured variables
+      yhat<-select(yhat_all, c("time", "r", "Protc"))
+      
+      #reformat to long format data frame
+      Yhat<-melt(yhat, id.vars = "time")
+      
+      #add the measured data to a data frame
[…]

### 2019-10-22 Rscripts/monod_r_fix.R

@@ -1,0 +1,220 @@
+###monod growth function
+monod_r_fix<-function(data, FACT){
+  
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  
+  if(FACT==1){
+    
+    dat<-data
+    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-data
+      dat<-data
+      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
+      ids$id<-seq(1:nrow(ids))
+      
+      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
+      
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-data
+        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
+        ids$id<-seq(1:nrow(ids))
+        
+        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
+        
+        
+      }else{
+        
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  
+  #monod growth function
+  deriv<-function(time, state, pars){
+    
+    with(as.list(c(state, pars)),{
+      
+      dCmic<--k*Cmic+CUE*Vmax*Cmic*C/(Km+C)
+      dC<-k*Cmic-Vmax*Cmic*C/(Km+C)
+      
+      return(list(c(dCmic, dC), r=(1-CUE)*Vmax*Cmic*C/(Km+C)))
+      
+    })
+  }
+  
+  #define names of parameters
+  parnames<-c("Vmax", "Km", "CUE", "k")
+  
+  #parameters estimation function
+  estim<-function(odeset){
+    
+    #defining cost function
+    cost<-function(x){
+      
+      par<-x[1:length(parnames)]
+      
+      names(par)<-parnames
+      
+      #first, pars dependent output from ode is matched with measured values
+      yhat_all<-as.data.frame(ode(y=c(Cmic=0.1001656, C=25), parms=par, deriv, times=sort(odeset$Time)))
+      
+      #select time and the measured variables 
+      yhat<-select(yhat_all, c("time", "r"))
+      
+      #add the measured data to a data frame
+      yhat$obs<-c(odeset[order(odeset$Time), c("r")])
+      
+      #now, the root mean square error is calculated
[…]

### 2019-10-22 Rscripts/monod_test.R

@@ -1,0 +1,257 @@
+monod_test<-function(data, FACT){
+  
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  
+  if(FACT==1){
+    
+    dat<-data
+    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-data
+      dat<-data
+      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
+      ids$id<-seq(1:nrow(ids))
+      
+      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
+      
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-data
+        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
+        ids$id<-seq(1:nrow(ids))
+        
+        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
+        
+        
+      }else{
+        
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  
+  #monod growth function
+  deriv<-function(time, state, pars){
+    
+    with(as.list(c(state, pars)),{
+      
+      dCmic<--k*Cmic+CUE*Vmax*Cmic*C/(Km+C)
+      dC<-k*Cmic-Vmax*Cmic*C/(Km+C)
+      
+      return(list(c(dCmic, dC), r=(1-CUE)*Vmax*Cmic*C/(Km+C)))
+      
+    })
+  }
+  
+  #define names of parameters
+  parnames<-c("Vmax", "Km", "CUE", "k")
+  
+  #parameters estimation function
+  estim<-function(odeset){
+    
+    #defining cost function
+    cost<-function(x){
+      
+      par<-x[1:length(parnames)]
+      
+      names(par)<-parnames
+      
+      Cmic0<-numeric()
+      if(odeset$Structure[1]=="Broth"){
+        Cmic0<-0.1287723+2.08193
+      }else{
+        if(odeset$Structure[1]=="Mixed glass"){
+          Cmic0<-0.2931153+4.73895
+        }else{
+          Cmic0<-0.221984+3.588933
+        }
+      }
+      
+      #first, pars dependent output from ode is matched with measured values
[…]

### 2019-10-22 Rscripts/monodplus_DNA.R

@@ -1,0 +1,244 @@
+monodplus_DNA<-function(data, FACT){
+  
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  
+  if(FACT==1){
+    
+    dat<-data
+    dat$id<-ifelse(d$Substrate=="Glucose", 1, 2)
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-data
+      dat<-data
+      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
+      ids$id<-seq(1:nrow(ids))
+      
+      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
+      
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-data
+        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
+        ids$id<-seq(1:nrow(ids))
+        
+        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
+        
+        
+      }else{
+        
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  #mend model
+  deriv<-function(time, state, pars){
+    
+    with(as.list(c(state, pars)),{
+      
+      #equations
+      #C uptake
+      Cu=Vmax*Cmic*C/(Km+C)
+      #respiration
+      r=mr*Cmic+(1-CUE)*Cu
+      
+      #Protc=Cmic*fp
+      DNAc=Cmic*fd
+      
+      #states
+      dCmic<-CUE*Cu-mr*Cmic - k*Cmic
+      dC<--Cu + k*Cmic
+      
+      return(list(c(dCmic, dC), r=r, DNAc=DNAc))
+      
+    })
+  }
+  #define names of parameters
+  parnames<-c("Vmax", "Km", "CUE", "mr", "k", "fd")
+  
+  #parameters estimation function
+  estim<-function(odeset){
+    
+    #defining cost function
+    cost<-function(x){
+      
+      par<-x[1:length(parnames)]
+      
+      names(par)<-parnames
+      
+      #first, pars dependent output from ode is matched with measured values
+      yhat_all<-as.data.frame(ode(y=c(Cmic=2.01*0.51/12.01/4/par[["fd"]], C=25), parms=par, deriv, times=sort(odeset$Time)))
+      
+      #select time and the measured variables 
[…]

### 2019-10-22 Rscripts/monodplus_prot.R

@@ -1,0 +1,244 @@
+monodplus_prot<-function(data, FACT){
+  
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  
+  if(FACT==1){
+    
+    dat<-data
+    dat$id<-ifelse(d$Substrate=="Glucose", 1, 2)
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-data
+      dat<-data
+      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
+      ids$id<-seq(1:nrow(ids))
+      
+      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
+      
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-data
+        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
+        ids$id<-seq(1:nrow(ids))
+        
+        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
+        
+        
+      }else{
+        
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  #mend model
+  deriv<-function(time, state, pars){
+    
+    with(as.list(c(state, pars)),{
+      
+      #equations
+      #C uptake
+      Cu=Vmax*Cmic*C/(Km+C)
+      #respiration
+      r=mr*Cmic+(1-CUE)*Cu
+      
+      Protc=Cmic*fp
+      #DNAc=Cmic*fd
+      
+      #states
+      dCmic<-CUE*Cu-mr*Cmic-k*Cmic
+      dC<--Cu + k*Cmic
+      
+      return(list(c(dCmic, dC), r=r, Protc=Protc))
+      
+      
+    })
+  }
+  #define names of parameters
+  parnames<-c("Vmax", "Km", "CUE", "mr", "k","fp", "Cmic_0")
+  
+  #parameters estimation function
+  estim<-function(odeset){
+    
+    #defining cost function
+    cost<-function(x){
+      
+      par<-x[1:length(parnames)]
+      
+      names(par)<-parnames
+      
+      #first, pars dependent output from ode is matched with measured values
+      yhat_all<-as.data.frame(ode(y=c(Cmic=par[["Cmic_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))
+      
+      #select time and the measured variables 
[…]

### 2019-10-22 Rscripts/monodplus_prot_fix.R

@@ -1,0 +1,244 @@
+monodplus_prot_fix<-function(data, FACT){
+  
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  
+  if(FACT==1){
+    
+    dat<-data
+    dat$id<-ifelse(d$Substrate=="Glucose", 1, 2)
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-data
+      dat<-data
+      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
+      ids$id<-seq(1:nrow(ids))
+      
+      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
+      
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-data
+        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
+        ids$id<-seq(1:nrow(ids))
+        
+        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
+        
+        
+      }else{
+        
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  #mend model
+  deriv<-function(time, state, pars){
+    
+    with(as.list(c(state, pars)),{
+      
+      #equations
+      #C uptake
+      Cu=Vmax*Cmic*C/(Km+C)
+      #respiration
+      r=mr*Cmic+(1-CUE)*Cu
+      
+      Protc=Cmic*fp
+      #DNAc=Cmic*fd
+      
+      #states
+      dCmic<-CUE*Cu-mr*Cmic-k*Cmic
+      dC<--Cu + k*Cmic
+      
+      return(list(c(dCmic, dC), r=r, Protc=Protc))
+      
+      
+    })
+  }
+  #define names of parameters
+  parnames<-c("Vmax", "Km", "CUE", "mr", "k","fp")
+  
+  #parameters estimation function
+  estim<-function(odeset){
+    
+    #defining cost function
+    cost<-function(x){
+      
+      par<-x[1:length(parnames)]
+      
+      names(par)<-parnames
+      
+      #first, pars dependent output from ode is matched with measured values
+      yhat_all<-as.data.frame(ode(y=c(Cmic=0.09125118, C=25), parms=par, deriv, times=sort(odeset$Time)))
+      
+      #select time and the measured variables 
+      yhat<-select(yhat_all, c("time", "Protc", "r"))
[…]

### 2019-10-22 Rscripts/monodplus_r.R

@@ -1,0 +1,236 @@
+monodplus_r<-function(data, FACT){
+  
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  
+  if(FACT==1){
+    
+    dat<-data
+    dat$id<-ifelse(d$Substrate=="Glucose", 1, 2)
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-data
+      dat<-data
+      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
+      ids$id<-seq(1:nrow(ids))
+      
+      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
+      
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-data
+        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
+        ids$id<-seq(1:nrow(ids))
+        
+        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
+        
+        
+      }else{
+        
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  #mend model
+  deriv<-function(time, state, pars){
+    
+    with(as.list(c(state, pars)),{
+      
+      #equations
+      #C uptake
+      Cu=Vmax*Cmic*C/(Km+C)
+      #respiration
+      r=mr*Cmic+(1-CUE)*Cu
+      
+      #Protc=Cmic*fp
+      #DNAc=Cmic*fd
+      
+      #states
+      dCmic<-CUE*Cu-mr*Cmic-k*Cmic
+      dC<--Cu + k*Cmic
+      
+      return(list(c(dCmic, dC), r=r))
+      
+    })
+  }
+  #define names of parameters
+  parnames<-c("Vmax", "Km", "CUE", "mr", "k", "Cmic_0")
+  
+  #parameters estimation function
+  estim<-function(odeset){
+    
+    #defining cost function
+    cost<-function(x){
+      
+      par<-x[1:length(parnames)]
+      
+      names(par)<-parnames
+      
+      #first, pars dependent output from ode is matched with measured values
+      yhat_all<-as.data.frame(ode(y=c(Cmic=par[["Cmic_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))
+      
+      #select time and the measured variables 
[…]

### 2019-10-22 Rscripts/monodplus_r_fix.R

@@ -1,0 +1,236 @@
+monodplus_r_fix<-function(data, FACT){
+  
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  
+  if(FACT==1){
+    
+    dat<-data
+    dat$id<-ifelse(d$Substrate=="Glucose", 1, 2)
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-data
+      dat<-data
+      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
+      ids$id<-seq(1:nrow(ids))
+      
+      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
+      
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-data
+        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
+        ids$id<-seq(1:nrow(ids))
+        
+        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
+        
+        
+      }else{
+        
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  #mend model
+  deriv<-function(time, state, pars){
+    
+    with(as.list(c(state, pars)),{
+      
+      #equations
+      #C uptake
+      Cu=Vmax*Cmic*C/(Km+C)
+      #respiration
+      r=mr*Cmic+(1-CUE)*Cu
+      
+      #Protc=Cmic*fp
+      #DNAc=Cmic*fd
+      
+      #states
+      dCmic<-CUE*Cu-mr*Cmic-k*Cmic
+      dC<--Cu + k*Cmic
+      
+      return(list(c(dCmic, dC), r=r))
+      
+    })
+  }
+  #define names of parameters
+  parnames<-c("Vmax", "Km", "CUE", "mr", "k")
+  
+  #parameters estimation function
+  estim<-function(odeset){
+    
+    #defining cost function
+    cost<-function(x){
+      
+      par<-x[1:length(parnames)]
+      
+      names(par)<-parnames
+      
+      #first, pars dependent output from ode is matched with measured values
+      yhat_all<-as.data.frame(ode(y=c(Cmic=0.09125118, C=25), parms=par, deriv, times=sort(odeset$Time)))
+      
+      #select time and the measured variables 
+      yhat<-select(yhat_all, c("time", "r"))
[…]

### 2019-10-22 Rscripts/monoplus_all.R

@@ -1,0 +1,244 @@
+monodplus_all<-function(data, FACT){
+  
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  
+  if(FACT==1){
+    
+    dat<-data
+    dat$id<-ifelse(d$Substrate=="Glucose", 1, 2)
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-data
+      dat<-data
+      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
+      ids$id<-seq(1:nrow(ids))
+      
+      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
+      
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-data
+        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
+        ids$id<-seq(1:nrow(ids))
+        
+        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
+        
+        
+      }else{
+        
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  #mend model
+  deriv<-function(time, state, pars){
+    
+    with(as.list(c(state, pars)),{
+      
+      #equations
+      #C uptake
+      Cu=Vmax*Cmic*C/(Km+C)
+      #respiration
+      r=mr*Cmic+(1-CUE)*Cu
+      
+      Protc=Cmic*fp
+      DNAc=Cmic*fd
+      
+      #states
+      dCmic<-CUE*Cu-mr*Cmic - k*Cmic
+      dC<--Cu + k*Cmic
+      
+      return(list(c(dCmic, dC), r=r, DNAc=DNAc, Protc=Protc))
+      
+    })
+  }
+  #define names of parameters
+  parnames<-c("Vmax", "Km", "CUE", "mr", "k", "fd", "fp")
+  
+  #parameters estimation function
+  estim<-function(odeset){
+    
+    #defining cost function
+    cost<-function(x){
+      
+      par<-x[1:length(parnames)]
+      
+      names(par)<-parnames
+      
+      #first, pars dependent output from ode is matched with measured values
+      yhat_all<-as.data.frame(ode(y=c(Cmic=2.01*0.51/12.01/4/par[["fd"]], C=25), parms=par, deriv, times=sort(odeset$Time)))
+      
[…]

### 2019-10-22 Rscripts/monoplus_test.R

@@ -1,0 +1,260 @@
+mend_test<-function(data, FACT){
+  
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  
+  if(FACT==1){
+    
+    dat<-data
+    dat$id<-ifelse(d$Substrate=="Glucose", 1, 2)
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-data
+      dat<-data
+      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
+      ids$id<-seq(1:nrow(ids))
+      
+      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
+      
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-data
+        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
+        ids$id<-seq(1:nrow(ids))
+        
+        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
+        
+        
+      }else{
+        
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  #mend model
+  deriv<-function(time, state, pars){
+    
+    with(as.list(c(state, pars)),{
+      
+      #equations
+      #C uptake
+      Cu=Vmax*Cmic*C/(Km+C)
+      #respiration
+      r=mr*Cmic+(1-CUE)*Cu
+      
+      #states
+      dCmic<-CUE*Cu-mr*Cmic-k*Cmic
+      dC<--Cu + k*Cmic
+      
+      return(list(c(dCmic, dC), r=r))
+    })
+  }
+  #define names of parameters
+  parnames<-c("Vmax", "Km", "CUE", "mr", "k")
+  
+  #parameters estimation function
+  estim<-function(odeset){
+    
+    #defining cost function
+    cost<-function(x){
+      
+      par<-x[1:length(parnames)]
+      
+      names(par)<-parnames
+      
+      Cmic0<-numeric()
+      if(odeset$Structure[1]=="Broth"){
+        Cmic0<-0.1287723+2.08193
+      }else{
+        if(odeset$Structure[1]=="Mixed glass"){
+          Cmic0<-0.2931153+4.73895
+        }else{
+          Cmic0<-0.221984+3.588933
+        }
+      }
+      
+      #first, pars dependent output from ode is matched with measured values
[…]

### 2019-10-22 Rscripts/two_pool_all.R

@@ -1,0 +1,252 @@
+two_pool_all<-function(data, FACT){
+  
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-data
+    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-data
+      dat<-data
+      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
+      ids$id<-seq(1:nrow(ids))
+      
+      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
+      
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-data
+        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
+        ids$id<-seq(1:nrow(ids))
+        
+        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
+        
+        
+      }else{
+        
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  #deb model
+  deriv<-function(time, state, pars){
+    
+    with(as.list(c(state, pars)),{
+      
+      #Carbon uptake
+      Cu=Vmax*C*S/(Km+C)
+      #Respiration
+      r=m0*S+f*R*(1-Yu)
+      
+      Protc=fpr*R+fps*S
+      DNAc=fds*S
+      
+      dR<-Cu-f*R
+      dS<-f*R*Yu - m0*S - k*S
+      dC<--Cu + k*S
+      
+      
+      return(list(c(dR, dS, dC), r=r, Protc=Protc, DNAc=DNAc))
+      
+    })
+  }
+  #define names of parameters
+  parnames<-c("Vmax", "Km", "m0", "f", "k", "Yu", "fpr", "fps", "fds", "RS_0")
+  
+  #parameters estimation function
+  estim<-function(odeset){
+    
+    #defining cost function
+    cost<-function(x){
+      
+      par<-x[1:length(parnames)]
+      
+      names(par)<-parnames
+      
+      Sinit=2.01*0.51/12.01/4/par[["fds"]]
+      Rinit=par[["RS_0"]]*Sinit
+      
+      #first, pars dependent output from ode is matched with measured values
[…]

### 2019-10-22 Rscripts/two_pool_all_fix.R

@@ -1,0 +1,252 @@
+two_pool_all_fix<-function(data, FACT){
+  
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-data
+    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-data
+      dat<-data
+      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
+      ids$id<-seq(1:nrow(ids))
+      
+      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
+      
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-data
+        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
+        ids$id<-seq(1:nrow(ids))
+        
+        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
+        
+        
+      }else{
+        
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  #deb model
+  deriv<-function(time, state, pars){
+    
+    with(as.list(c(state, pars)),{
+      
+      #Carbon uptake
+      Cu=Vmax*C*S/(Km+C)
+      #Respiration
+      r=m0*S+f*R*(1-Yu)
+      
+      Protc=fpr*R+fps*S
+      DNAc=fds*S
+      
+      dR<-Cu-f*R
+      dS<-f*R*Yu - m0*S - k*S
+      dC<--Cu + k*S
+      
+      
+      return(list(c(dR, dS, dC), r=r, Protc=Protc, DNAc=DNAc))
+      
+    })
+  }
+  #define names of parameters
+  parnames<-c("Vmax", "Km", "m0", "f", "k", "Yu", "fpr", "fps", "fds")
+  
+  #parameters estimation function
+  estim<-function(odeset){
+    
+    #defining cost function
+    cost<-function(x){
+      
+      par<-x[1:length(parnames)]
+      
+      names(par)<-parnames
+      
+      Sinit=2.01*0.51/12.01/4/par[["fds"]]
+      Rinit=0.06185237*Sinit
+      
+      #first, pars dependent output from ode is matched with measured values
[…]

### 2019-10-22 Rscripts/two_pool_all_fixYuk.R

@@ -1,0 +1,252 @@
+two_pool_all_fixYuk<-function(data, FACT){
+  
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-data
+    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-data
+      dat<-data
+      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
+      ids$id<-seq(1:nrow(ids))
+      
+      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
+      
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-data
+        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
+        ids$id<-seq(1:nrow(ids))
+        
+        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
+        
+        
+      }else{
+        
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  #deb model
+  deriv<-function(time, state, pars){
+    
+    with(as.list(c(state, pars)),{
+      
+      #Carbon uptake
+      Cu=Vmax*C*S/(Km+C)
+      #Respiration
+      r=m0*S+f*R*(1-0.3565256)
+      
+      Protc=fpr*R+fps*S
+      DNAc=fds*S
+      
+      dR<-Cu-f*R
+      dS<-f*R*0.3565256 - m0*S - 0.004915818*S
+      dC<--Cu + 0.004915818*S
+      
+      
+      return(list(c(dR, dS, dC), r=r, Protc=Protc, DNAc=DNAc))
+      
+    })
+  }
+  #define names of parameters
+  parnames<-c("Vmax", "Km", "m0", "f", "fpr", "fps", "fds")
+  
+  #parameters estimation function
+  estim<-function(odeset){
+    
+    #defining cost function
+    cost<-function(x){
+      
+      par<-x[1:length(parnames)]
+      
+      names(par)<-parnames
+      
+      Sinit=2.01*0.51/12.01/4/par[["fds"]]
+      Rinit=0.06185237*Sinit
+      
[…]

### 2019-10-22 Rscripts/two_pool_all_fixYukf.R

@@ -1,0 +1,252 @@
+two_pool_all_fixYukf<-function(data, FACT){
+  
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-data
+    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-data
+      dat<-data
+      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
+      ids$id<-seq(1:nrow(ids))
+      
+      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
+      
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-data
+        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
+        ids$id<-seq(1:nrow(ids))
+        
+        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
+        
+        
+      }else{
+        
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  #deb model
+  deriv<-function(time, state, pars){
+    
+    with(as.list(c(state, pars)),{
+      
+      #Carbon uptake
+      Cu=Vmax*C*S/(Km+C)
+      #Respiration
+      r=m0*S+0.2373511*R*(1-0.3565256)
+      
+      Protc=fpr*R+fps*S
+      DNAc=fds*S
+      
+      dR<-Cu-0.2373511*R
+      dS<-0.2373511*R*0.3565256 - m0*S - 0.004915818*S
+      dC<--Cu + 0.004915818*S
+      
+      
+      return(list(c(dR, dS, dC), r=r, Protc=Protc, DNAc=DNAc))
+      
+    })
+  }
+  #define names of parameters
+  parnames<-c("Vmax", "Km", "m0", "fpr", "fps", "fds")
+  
+  #parameters estimation function
+  estim<-function(odeset){
+    
+    #defining cost function
+    cost<-function(x){
+      
+      par<-x[1:length(parnames)]
+      
+      names(par)<-parnames
+      
+      Sinit=2.01*0.51/12.01/4/par[["fds"]]
+      Rinit=0.06185237*Sinit
+      
+      #first, pars dependent output from ode is matched with measured values
[…]

### 2019-10-22 Rscripts/two_pool_all_fixYukfVmax.R

@@ -1,0 +1,252 @@
+two_pool_all_fixYukfVmax<-function(data, FACT){
+  
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-data
+    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-data
+      dat<-data
+      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
+      ids$id<-seq(1:nrow(ids))
+      
+      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
+      
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-data
+        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
+        ids$id<-seq(1:nrow(ids))
+        
+        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
+        
+        
+      }else{
+        
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  #deb model
+  deriv<-function(time, state, pars){
+    
+    with(as.list(c(state, pars)),{
+      
+      #Carbon uptake
+      Cu=0.2761818*C*S/(Km+C)
+      #Respiration
+      r=m0*S+0.2373511*R*(1-0.3565256)
+      
+      Protc=fpr*R+fps*S
+      DNAc=fds*S
+      
+      dR<-Cu-0.2373511*R
+      dS<-0.2373511*R*0.3565256 - m0*S - 0.004915818*S
+      dC<--Cu + 0.004915818*S
+      
+      
+      return(list(c(dR, dS, dC), r=r, Protc=Protc, DNAc=DNAc))
+      
+    })
+  }
+  #define names of parameters
+  parnames<-c("Km", "m0", "fpr", "fps", "fds")
+  
+  #parameters estimation function
+  estim<-function(odeset){
+    
+    #defining cost function
+    cost<-function(x){
+      
+      par<-x[1:length(parnames)]
+      
+      names(par)<-parnames
+      
+      Sinit=2.01*0.51/12.01/4/par[["fds"]]
+      Rinit=0.06185237*Sinit
+      
[…]

### 2019-10-22 Rscripts/two_pool_all_fixYukfVmaxKm.R

@@ -1,0 +1,252 @@
+two_pool_all_fixYukfVmaxKm<-function(data, FACT){
+  
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-data
+    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-data
+      dat<-data
+      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
+      ids$id<-seq(1:nrow(ids))
+      
+      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
+      
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-data
+        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
+        ids$id<-seq(1:nrow(ids))
+        
+        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
+        
+        
+      }else{
+        
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  #deb model
+  deriv<-function(time, state, pars){
+    
+    with(as.list(c(state, pars)),{
+      
+      #Carbon uptake
+      Cu=0.2761818*C*S/(6.3136+C)
+      #Respiration
+      r=m0*S+0.2373511*R*(1-0.3565256)
+      
+      Protc=fpr*R+fps*S
+      DNAc=fds*S
+      
+      dR<-Cu-0.2373511*R
+      dS<-0.2373511*R*0.3565256 - m0*S - 0.004915818*S
+      dC<--Cu + 0.004915818*S
+      
+      
+      return(list(c(dR, dS, dC), r=r, Protc=Protc, DNAc=DNAc))
+      
+    })
+  }
+  #define names of parameters
+  parnames<-c("m0", "fpr", "fps", "fds")
+  
+  #parameters estimation function
+  estim<-function(odeset){
+    
+    #defining cost function
+    cost<-function(x){
+      
+      par<-x[1:length(parnames)]
+      
+      names(par)<-parnames
+      
+      Sinit=2.01*0.51/12.01/4/par[["fds"]]
+      Rinit=0.06185237*Sinit
+      
[…]

### 2019-10-22 Rscripts/two_pool_r.R

@@ -1,0 +1,228 @@
+two_pool_r<-function(data, FACT){
+  
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-data
+    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-data
+      dat<-data
+      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
+      ids$id<-seq(1:nrow(ids))
+      
+      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
+      
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-data
+        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
+        ids$id<-seq(1:nrow(ids))
+        
+        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
+        
+        
+      }else{
+        
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  #deb model
+  deriv<-function(time, state, pars){
+    
+    with(as.list(c(state, pars)),{
+      
+      #Carbon uptake
+      Cu=Vmax*C*S/(Km+C)
+      #Respiration
+      r=m0*S + f*R*(1-Yu) 
+      
+      dR<-Cu-f*R
+      dS<-f*R*Yu - k*S - m0*S
+      dC<--Cu + k*S
+      
+      return(list(c(dR, dS, dC), r=r))
+      
+    })
+  }
+  #define names of parameters
+  parnames<-c("Vmax", "Km", "m0", "f", "k", "Yu", "R_0", "S_0")
+  
+  #parameters estimation function
+  estim<-function(odeset){
+    
+    #defining cost function
+    cost<-function(x){
+      
+      par<-x[1:length(parnames)]
+      
+      names(par)<-parnames
+      
+      #first, pars dependent output from ode is matched with measured values
+      yhat_all<-as.data.frame(ode(y=c(R=par[["R_0"]], S=par[["S_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))
+      
+      #select time and the measured variables
+      yhat<-select(yhat_all, c("time", "r"))
+      
[…]

### 2019-10-22 Rscripts/two_pool_r_fix.R

@@ -1,0 +1,228 @@
+two_pool_r_fix<-function(data, FACT){
+  
+  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
+  
+  
+  #if factor needs to be included, its levels are defined
+  if(FACT==1){
+    
+    dat<-data
+    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
+    
+  }else{
+    
+    if(FACT==2){
+      
+      dat<-data
+      dat<-data
+      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
+      ids$id<-seq(1:nrow(ids))
+      
+      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
+      
+    }else{
+      
+      if(FACT==3){
+        
+        dat<-data
+        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
+        ids$id<-seq(1:nrow(ids))
+        
+        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
+        
+        
+      }else{
+        
+        dat<-data
+        dat$id<-c(1)
+        
+      }
+    }
+  }
+  
+  #deb model
+  deriv<-function(time, state, pars){
+    
+    with(as.list(c(state, pars)),{
+      
+      #Carbon uptake
+      Cu=Vmax*C*S/(Km+C)
+      #Respiration
+      r=m0*S + f*R*(1-Yu) 
+      
+      dR<-Cu-f*R
+      dS<-f*R*Yu - k*S - m0*S
+      dC<--Cu + k*S
+      
+      return(list(c(dR, dS, dC), r=r))
+      
+    })
+  }
+  #define names of parameters
+  parnames<-c("Vmax", "Km", "m0", "f", "k", "Yu")
+  
+  #parameters estimation function
+  estim<-function(odeset){
+    
+    #defining cost function
+    cost<-function(x){
+      
+      par<-x[1:length(parnames)]
+      
+      names(par)<-parnames
+      
+      #first, pars dependent output from ode is matched with measured values
+      yhat_all<-as.data.frame(ode(y=c(R=0.01908309, S=0.2687341, C=25), parms=par, deriv, times=sort(odeset$Time)))
+      
+      #select time and the measured variables
+      yhat<-select(yhat_all, c("time", "r"))
+      
+      #add the measured data to a data frame
[…]

### 2019-11-07 SI_code_revised.Rmd
Last minor changes

@@
 colnames(bac)<-t(read.csv(file=c("bacteria.csv"), skip = 4)[,1])
+##remove OTUs with all zeros
+bac2<-bac[, colSums(bac!=0)>0]
 
@@
 colnames(fungi)<-t(read.csv(file=c("fungi.csv"), skip = 4)[,1])
+
+##remove OTUs with all zeros
+fungi2<-fungi[, colSums(fungi!=0)>0]
 
@@
 #Bacteria
-bac.norm<-decostand(bac, method=c("normalize"))
+bac.norm<-decostand(bac2, method=c("normalize"))
 bac.dist<-vegdist(bac.norm, method = "bray")
@@
 #Fungi
-fungi.norm<-decostand(fungi, method=c("normalize"))
+fungi.norm<-decostand(fungi2, method=c("normalize"))
 fungi.dist<-vegdist(fungi.norm, method = "bray")
@@
 
-kable(t.s1.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$'), row.names=T, caption = "_**Table S3:** Monod model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s1.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('v', 'k', '$\\Y_{CUE}$', 'm'), row.names=T, caption = "_**Table S3:** Monod model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(t.s2.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$', 'k'), row.names=T, caption = "_**Table S5:** Monod plus model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s2.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('v', 'k', '$\\Y_{CUE}$', 'h', 'm'), row.names=T, caption = "_**Table S5:** Monod plus model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(t.s3.1, digits = c(2, 2, 3, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('$\\V_{MAX}$', '$\\K_{M}$', '$\\m_{R}$', '$\\f_{0}$', 'k', '$\\Y_{S}$'), row.names=T, caption = "_**Table S7:** Two pool model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s3.1, digits = c(2, 2, 3, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('v', 'k', 'h', 'f', 'm', 'Y'), row.names=T, caption = "_**Table S7:** Two pool model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(t.s4.1, digits = c(0, 2, 2, 2, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$', '$\\f_{Proteins}$', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S9:** Monod model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s4.1, digits = c(0, 2, 2, 2, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'm', '$\\f_{Proteins}$', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S9:** Monod model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(t.s4.2, digits = c(0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$', '$\\f_{Proteins}$'), row.names=F, caption = "_**Table S11:** Monod model parameters calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s4.2, digits = c(0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'm', '$\\f_{Proteins}$'), row.names=F, caption = "_**Table S11:** Monod model parameters calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(t.s4.3, digits = c(0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\k_{MB}$', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S13:** Monod model parameters calibrated against respiration rate and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s4.3, digits = c(0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'm', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S13:** Monod model parameters calibrated against respiration rate and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(t.s5.1, digits = c(0, 2, 2, 2, 3, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$', '$\\k_{MB}$', '$\\f_{Proteins}$', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S15:** Monod plus model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s5.1, digits = c(0, 2, 2, 2, 3, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'h', 'm', '$\\f_{Proteins}$', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S15:** Monod plus model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(t.s5.2, digits = c(0, 2, 2, 2, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$', '$\\k_{MB}$', '$\\f_{Proteins}$'), row.names=F, caption = "_**Table S17:** Monod plus model parameters calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s5.2, digits = c(0, 2, 2, 2, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'h', 'm', '$\\f_{Proteins}$'), row.names=F, caption = "_**Table S17:** Monod plus model parameters calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(t.s5.3, digits = c(0, 2, 2, 2, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", '$\\V_{MAX}$', '$\\K_{M}$', 'CUE', '$\\m_{R}$', '$\\k_{MB}$', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S19:** Monod plus model parameters calibrated against respiration rate and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s5.3, digits = c(0, 2, 2, 2, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'h', 'm', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S19:** Monod plus model parameters calibrated against respiration rate and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 ####Two pool model
-Not all model parameters vary significantly acros treatments. Parameters that doesn't show any variability across treatments are fixed.
+Not all model parameters vary significantly across treatments. Parameters that doesn't show any variability across treatments are fixed.
 ```{r Two pool model calibration, eval=F, echo=T}
@@
 two_pool_all_fixYukfVmax2<-two_pool_all_fixYukfVmax(data=d, FACT = 2)
-#At the end, parameter KM is fixed
-source("./Rscripts/two_pool_all_fixYukfVmaxKm.R")
-two_pool_all_fixYukfVmaxKm2<-two_pool_all_fixYukfVmaxKm(data=d, FACT = 2)
 ############################################################################
@@
 two_pool_all1$goodness
-two_pool_all_fixYukfVmaxKm2$goodness
+two_pool_all_fixYukfVmax2$goodness
 two_pool_all3_fix$goodness
@@
 
-#results of the previous iteration (not run)
+# #results of the previous iteration (not run)
 # two_pool_all_results_par<-rbind(c(two_pool_all1[[1]]$pars[c(1:9)], RS_0 = 0.06185237),
 #                          c(Vmax = 0.2761818, 
-#                             Km = 6.3136,
-#                             two_pool_all_fixYukfVmaxKm2[[1]]$pars[1],
+#                            two_pool_all_fixYukfVmax2[[1]]$pars[1],
+#                            two_pool_all_fixYukfVmax2[[1]]$pars[2],
 #                            f = 0.2373511,
@@
 #                            Yu = 0.3565256,
-#                             two_pool_all_fixYukfVmaxKm2[[1]]$pars[c(2:4)],
+#                            two_pool_all_fixYukfVmax2[[1]]$pars[c(3:5)],
 #                            RS_0 = 0.06185237),
 #                          c(Vmax = 0.2761818, 
-#                             Km = 6.3136,
-#                             two_pool_all_fixYukfVmaxKm2[[2]]$pars[1],
+#                            two_pool_all_fixYukfVmax2[[2]]$pars[1],
+#                            two_pool_all_fixYukfVmax2[[2]]$pars[2],
 #                            f = 0.2373511,
@@
 #                            Yu = 0.3565256,
-#                             two_pool_all_fixYukfVmaxKm2[[2]]$pars[c(2:4)],
+#                            two_pool_all_fixYukfVmax2[[2]]$pars[c(3:5)],
 #                            RS_0 = 0.06185237),
 #                          c(Vmax = 0.2761818, 
-#                             Km = 6.3136,
-#                             two_pool_all_fixYukfVmaxKm2[[3]]$pars[1],
+#                            two_pool_all_fixYukfVmax2[[3]]$pars[1],
+#                            two_pool_all_fixYukfVmax2[[3]]$pars[2],
 #                            f = 0.2373511,
@@
 #                            Yu = 0.3565256,
-#                             two_pool_all_fixYukfVmaxKm2[[3]]$pars[c(2:4)],
+#                            two_pool_all_fixYukfVmax2[[3]]$pars[c(3:5)],
 #                            RS_0 = 0.06185237),
@@
 # two_pool_all_results_good<-rbind(two_pool_all1$goodness,
-#                                  two_pool_all_fixYukfVmaxKm2$goodness,
+#                                 two_pool_all_fixYukfVmax2$goodness,
 #                                 two_pool_all3_fix$goodness,
@@
 
-kable(t.s6.1, digits = c(0, 2, 2, 3, 3, 3, 2, 3, 3, 3, 2), align = 'c', "html", booktabs=T, col.names = c(" ","$\\V_{MAX}$", '$\\K_{M}$', '$\\m_{R}$', '$\\f_{0}$', 'k', '$\\Y_{S}$', '$\\R_{Proteins}$', '$\\S_{Proteins}$', '$\\S_{DNA}$', '$\\RS_{0}$'), row.names=F, caption = "_**Table S21:** Two pool model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s6.1, digits = c(0, 2, 2, 3, 3, 3, 2, 3, 3, 3, 2), align = 'c', "html", booktabs=T, col.names = c(" ","v", 'k', 'h', 'f', 'm', 'Y', '$\\B_{R-Proteins}$', '$\\B_{S-Proteins}$', '$\\B_{S-DNA}$', '$\\RS_{0}$'), row.names=F, caption = "_**Table S21:** Two pool model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 $~$
-```{r correlation figure, echo=FALSE, fig.height=5, fig.width=10, fig.cap="Fig.
-S3: Effect of structure of bacterial and fungal part of microbial community (expressed as multivariate mean distance between groups; see Matrial and Methods section) on overall microbial community maintnance respiration rate ($m_{R}$ - (a)) and protein concentration in Structures pool ($S_{Proteins}$ - (b)).
-The value of both parameters are expressed as a mean difference between groups in order to allow unbiased correlation with microbial community structure metric (see Matrial and Methods section)."}
-
+```{r correlation figure, echo=FALSE, fig.height=12, fig.width=16, fig.cap="Fig.
+S3: Relative abundance of bacterial (a) and fungal (b) species in three experimental treatments differing in presence/absence of physical structure.
+Species with relative abundance above 0.5% are plotted.
+Their abundance is visualized by colour scale.
+Effect of bacterial and fungal composition of microbial community (expressed as multivariate mean distance between groups; see Material and Methods section) on overall microbial community affinity to substrate (i.e.
+parameter k - (a)) and abundance of proteins in microbial biomass (i.e.
+sum of protein abundance in biomass reserve and struscture - (b)).
+The values on y-axes are expressed as a mean difference between groups in order to allow unbiased correlation with microbial community structure metric (see Matrial and Methods section)."}
+#Heatmap
+bac.tax<-as.data.frame(read.csv(file=c("./Community_analysis/OTU_bacteria.txt"), header=T,
+                             sep="\t"))
+bac.tax<-bac.tax[, c(1, 175)]
+bac.tax$X.OTU.ID<-gsub("_","", bac.tax$X.OTU.ID)
+
+#Add taxonomy to bac.norm dataset
+bac.norm_tax<-bac
+#Transform the dataset
+bac.norm_taxt<-as.data.frame(t(bac.norm_tax))
+taxlab<-character()
+for(i in 1:nrow(bac.norm_taxt)){
+  nd<-which(bac.tax$X.OTU.ID==rownames(bac.norm_taxt)[i])
+  taxlab<-append(taxlab, as.character(bac.tax$taxonomy[nd]))
+}
+
+bac.norm_taxt$taxonomy<-taxlab
+
+#Extract kingdom
+pattern<-"k__(.+?),"
+regmatches(bac.norm_taxt$taxonomy[1], regexec(pattern, bac.norm_taxt$taxonomy[1]))[[1]][2]
+kingdoms<-character()
+for(i in 1:nrow(bac.norm_taxt)){
+  kingdoms<-append(kingdoms, 
+                   regmatches(bac.norm_taxt$taxonomy[i], regexec(pattern, bac.norm_taxt$taxonomy[i]))[[1]][2])
+}
+kingdoms[which(kingdoms=="?")]<-c("Unknown")
+bac.norm_taxt$Kingdom<-kingdoms
+
+#Extract phylum
+pattern<-"p__(.+?),"
+
+phylum<-character()
+for(i in 1:nrow(bac.norm_taxt)){
+  phylum<-append(phylum, 
+                   regmatches(bac.norm_taxt$taxonomy[i], regexec(pattern, bac.norm_taxt$taxonomy[i]))[[1]][2])
+}
+phylum[which(phylum=="?")]<-c("Unknown")
+bac.norm_taxt$Phylum<-phylum
+
+#Extract class
+pattern<-"c__(.+?),"
+
+cs<-character()
+for(i in 1:nrow(bac.norm_taxt)){
+  cs<-append(cs, 
+                 regmatches(bac.norm_taxt$taxonomy[i], regexec(pattern, bac.norm_taxt$taxonomy[i]))[[1]][2])
+}
+cs[which(cs=="?")]<-c("Unknown")
+bac.norm_taxt$Class<-cs
+
+#Extract order
+pattern<-"o__(.+?),"
+
+os<-character()
+for(i in 1:nrow(bac.norm_taxt)){
+  os<-append(os, 
+             regmatches(bac.norm_taxt$taxonomy[i], regexec(pattern, bac.norm_taxt$taxonomy[i]))[[1]][2])
+}
+os[which(os=="?")]<-c("Unknown")
+bac.norm_taxt$Order<-os
+
+#Extract family
+pattern<-"f__(.+?),"
+
+fs<-character()
+for(i in 1:nrow(bac.norm_taxt)){
+  fs<-append(fs, 
+             regmatches(bac.norm_taxt$taxonomy[i], regexec(pattern, bac.norm_taxt$taxonomy[i]))[[1]][2])
+}
+
+fs[which(fs=="?")]<-c("Unknown")
+bac.norm_taxt$Family<-fs
+
+#Extract genus
+pattern<-"g__(.+?),"
+
+gs<-character()
+for(i in 1:nrow(bac.norm_taxt)){
+  gs<-append(gs, 
+             regmatches(bac.norm_taxt$taxonomy[i], regexec(pattern, bac.norm_taxt$taxonomy[i]))[[1]][2])
+}
+gs[which(gs=="?")]<-c("Unknown")
+bac.norm_taxt$Genus<-gs
+
+#Extract species
+pattern<-"s__(.+?),"
+
+ss<-character()
+for(i in 1:nrow(bac.norm_taxt)){
+  ss<-append(ss, 
+             regmatches(bac.norm_taxt$taxonomy[i], regexec(pattern, bac.norm_taxt$taxonomy[i]))[[1]][2])
+}
+ss[which(is.na(ss))]<-c("Unknown")
+bac.norm_taxt$Specie<-ss
+
+#Plots
+tax_env<-bac_env
+btaxt<-bac.norm_taxt[, 1:71]
+tax_env$IDs<-colnames(btaxt)
+
+##Genus
+btaxtg<-btaxt
+btaxtg$Genus<-bac.norm_taxt$Genus
+
+Btaxtg<-melt(btaxtg, id.vars = c("Genus"))
+colnames(Btaxtg)<-c("Genus", "IDs","Abundance")
+
+Btaxtg<-merge(Btaxtg, tax_env, by.x = "IDs", by.y = "IDs", all.x = TRUE)
+Btaxtg$Total<-numeric(length = nrow(Btaxtg))
+for(i in unique(Btaxtg$IDs)){
+  Btaxtg[Btaxtg$IDs==i, "Total"]<-sum(as.numeric(Btaxtg[Btaxtg$IDs==i, "Abundance"]))
+}
+
+bac_gen_f<-Btaxtg %>% group_by(Genus, Structure, Substrate) %>% 
+  summarize(Abundance=mean(Abundance/Total, na.rm=T))
+
+#Only genus with abundance above 0.05%
+heat1<-ggplot(bac_gen_f[bac_gen_f$Abundance>0.0005, ], aes(Structure, Genus)) +
+  geom_tile(aes(fill = log10(Abundance*100))) +
+  facet_grid(~ Substrate) +
+  scale_fill_gradientn(colours = c("white","darkorange2","grey60"),
+                       breaks = c(-2, -1, 0, 1, 2),
+                       labels=c("0.01", "0.1", "0", "10", "100"),
+                       name = "Relative abundance (%)",
+                       guide = guide_colourbar(direction = "horizontal", title.position = "bottom",
+                                               title.hjust = 0.5, barwidth = 12)) +
+  scale_x_discrete(expand = c(0,0)) + 
+  scale_y_discrete(expand = c(0,0)) +  
+  xlab("Structure") + ylab("")+
+  ggtitle("a)")+
+  theme_min + theme(legend.position = "bottom",
+                    axis.text = element_text(size=14),
+                    strip.text = element_text(size=14))
+
+##Fungi
+f.tax<-as.data.frame(read.csv(file=c("./Community_analysis/OTU_fungi.txt"), header=T,
+                             sep="\t"))
+#Add taxonomy to fungi dataset
+f.norm_tax<-fungi
+#Transform the dataset
+f.norm_taxt<-as.data.frame(t(f.norm_tax))
+ftaxlab<-character()
+for(i in 1:nrow(f.norm_taxt)){
+  nd<-which(f.tax$X.OTU.ID==rownames(f.norm_taxt)[i])
+  ftaxlab<-append(ftaxlab, as.character(f.tax$taxonomy[nd]))
+}
+
+f.norm_taxt$taxonomy<-ftaxlab
+
+#Extract kingdom
+pattern<-"k__(.+?),"
+regmatches(f.norm_taxt$taxonomy[1], regexec(pattern, f.norm_taxt$taxonomy[1]))[[1]][2]
+kingdoms<-character()
+for(i in 1:nrow(f.norm_taxt)){
+  kingdoms<-append(kingdoms, 
+                   regmatches(f.norm_taxt$taxonomy[i], regexec(pattern, f.norm_taxt$taxonomy[i]))[[1]][2])
+}
+kingdoms[which(kingdoms=="?")]<-c("Unknown")
+f.norm_taxt$Kingdom<-kingdoms
+
+#Extract phylum
+pattern<-"p__(.+?),"
+
+phylum<-character()
+for(i in 1:nrow(f.norm_taxt)){
+  phylum<-append(phylum, 
+                 regmatches(f.norm_taxt$taxonomy[i], regexec(pattern, f.norm_taxt$taxonomy[i]))[[1]][2])
+}
+phylum[which(phylum=="?")]<-c("Unknown")
+f.norm_taxt$Phylum<-phylum
+
+#Extract class
+pattern<-"c__(.+?),"
+
+cs<-character()
+for(i in 1:nrow(f.norm_taxt)){
+  cs<-append(cs, 
+             regmatches(f.norm_taxt$taxonomy[i], regexec(pattern, f.norm_taxt$taxonomy[i]))[[1]][2])
+}
+cs[which(cs=="?")]<-c("Unknown")
+f.norm_taxt$Class<-cs
+
+#Extract order
+pattern<-"o__(.+?),"
+
+os<-character()
+for(i in 1:nrow(f.norm_taxt)){
+  os<-append(os, 
+             regmatches(f.norm_taxt$taxonomy[i], regexec(pattern, f.norm_taxt$taxonomy[i]))[[1]][2])
+}
+os[which(os=="?")]<-c("Unknown")
+f.norm_taxt$Order<-os
+
+#Extract family
+pattern<-"f__(.+?),"
+
+fs<-character()
+for(i in 1:nrow(f.norm_taxt)){
+  fs<-append(fs, 
+             regmatches(f.norm_taxt$taxonomy[i], regexec(pattern, f.norm_taxt$taxonomy[i]))[[1]][2])
+}
+
+fs[which(fs=="?")]<-c("Unknown")
+f.norm_taxt$Family<-fs
+
+#Extract genus
+pattern<-"g__(.+?),"
+
+gs<-character()
+for(i in 1:nrow(f.norm_taxt)){
+  gs<-append(gs, 
+             regmatches(f.norm_taxt$taxonomy[i], regexec(pattern, f.norm_taxt$taxonomy[i]))[[1]][2])
+}
+gs[which(gs=="?")]<-c("Unknown")
+f.norm_taxt$Genus<-gs
+
+#Extract species
+pattern<-"s__(.+?),"
+
+ss<-character()
+for(i in 1:nrow(f.norm_taxt)){
+  ss<-append(ss, 
+             regmatches(f.norm_taxt$taxonomy[i], regexec(pattern, f.norm_taxt$taxonomy[i]))[[1]][2])
+}
+ss[which(is.na(ss))]<-c("Unknown")
+f.norm_taxt$Specie<-ss
+
+#Plots
+ftax_env<-fungi_env
+ftaxt<-f.norm_taxt[, 1:64]
+ftax_env$IDs<-colnames(ftaxt)
+
+##Genus
+ftaxtg<-ftaxt
+ftaxtg$Genus<-f.norm_taxt$Genus
+
+Ftaxtg<-melt(ftaxtg, id.vars = c("Genus"))
+colnames(Ftaxtg)<-c("Genus", "IDs","Abundance")
+
+Ftaxtg<-merge(Ftaxtg, ftax_env, by.x = "IDs", by.y = "IDs", all.x = TRUE)
+Ftaxtg$Total<-numeric(length = nrow(Ftaxtg))
+for(i in unique(Ftaxtg$IDs)){
+  Ftaxtg[Ftaxtg$IDs==i, "Total"]<-sum(as.numeric(Ftaxtg[Ftaxtg$IDs==i, "Abundance"]))
+}
+
+f_gen_f<-Ftaxtg %>% group_by(Genus, Structure, Substrate) %>% 
+  summarize(Abundance=mean(Abundance/Total, na.rm=T))
+
+#Genus with abundance above 0.05%
+heat2<-ggplot(f_gen_f[f_gen_f$Abundance>0.0005, ], aes(Structure, Genus)) +
+  geom_tile(aes(fill = log(Abundance*100))) +
+  facet_grid(~ Substrate) +
+  scale_fill_gradientn(colours = c("white","darkorange2","grey60"),
+                       breaks = c(-2, -1, 0, 1, 2),
+                       labels=c("0.01", "0.1", "0", "10", "100"),
+                       name = "Relative abundance (%)",
+                       guide = guide_colourbar(direction = "horizontal", title.position = "bottom",
+                                               title.hjust = 0.5, barwidth = 12)) +
+  scale_x_discrete(expand = c(0,0)) + 
+  scale_y_discrete(expand = c(0,0)) +  
+  xlab("Structure") + ylab("")+
+  ggtitle("b)")+
+  theme_min + theme(legend.position = "bottom",
+                    axis.text = element_text(size=14),
+                    strip.text = element_text(size=14))
+##########################################################################################
 dpars<-t.s6.1[c(2:4), -1]
@@
 
-for(i in 3:13){
+for(i in 2:11){
   dpars2[, i]<-sqrt((dpars2[, i]-as.numeric(dpars2[1, i]))^2)
@@
 
-for(i in 3:13){
+for(i in 2:11){
   dpars3[, i]<-sqrt((dpars3[, i]-as.numeric(dpars3[1, i]))^2)
@@
 
-#corplot1.p<-coef(nls(m0~Dist^x, dpars4, start = list(x=1)))
-
-corplot1<-ggplot(dpars4[dpars4$group=="Bacteria", ], aes(Dist, m0*1e4))+
+corplot1.p<-coefficients(nls(Km~exp(Dist*x)-a, dpars4, start = list(x=10, a=2)))
+
+corplot1<-ggplot(dpars4[dpars4$group=="Bacteria", ], aes(Dist, Km))+
   geom_point(cex=6, aes(fill=Structure), show.legend=T, shape=21)+
   geom_point(data=dpars4[dpars4$group=="Fungi", ], cex=6, 
-             aes(x=Dist, y=m0*1e4, fill=Structure), show.legend=F, shape=22)+
+             aes(x=Dist, y=Km, fill=Structure), show.legend=F, shape=22)+
   scale_fill_manual(values = c("black", "grey", "white"))+
-  #stat_function(fun = function(x){(x^corplot1.p*1e4)}, lwd=1, colour="grey30")+
+  stat_function(fun = function(x){(exp(x*corplot1.p[1])-corplot1.p[2])}, lwd=1, lty=2, colour="grey30")+
   xlab("16S and ITS Distance")+
-  ggtitle("a)")+
-  ylab(expression(paste(m[R]~difference)))+
+  ggtitle("c)")+
+  ylab("Affinity to substrate (relative difference)")+
   theme_min+theme(legend.title = element_blank(),
-                  legend.position = c(0.2,0.8))
-
-#corplot2.p<-coef(nls(fps~Dist^x, dpars4, start = list(x=1)))
-
-corplot2<-ggplot(dpars4[dpars4$group=="Bacteria", ], aes(Dist, fps))+
+                  legend.position = c(0.2,0.7),
+                  plot.margin = unit(c(0.5,1,0.5,2), "in"))
+
+corplot2<-ggplot(dpars4[dpars4$group=="Bacteria", ], aes(Dist, fps+fpr))+
   geom_point(cex=6, aes(fill=Structure), show.legend=F, shape=21)+
   geom_point(data=dpars4[dpars4$group=="Fungi", ], cex=6, 
-             aes(x=Dist, y=fps, fill=Structure), show.legend=F, shape=22)+
+             aes(x=Dist, y=fps+fpr, fill=Structure), show.legend=F, shape=22)+
+  stat_smooth(method = lm, colour="grey30", lty=2, se=F)+
   scale_fill_manual(values = c("black", "grey", "white"))+
-  #stat_function(fun = function(x){(x^corplot2.p)}, lwd=1, colour="grey30")+
-  xlab("16S and ITS Distance")+
-  ggtitle("b)")+
-  ylab(expression(paste(S[Proteins]~difference)))+
-  theme_min
-
-grid.arrange(corplot1, corplot2, nrow=1)
-
-```
-
-```{r Biomass predictions by DEB, eval=FALSE, include=FALSE}
-DEB<-function(time, state, pars){
-    with(as.list(c(state, pars)),{
-      #Carbon uptake
-      Cu=0.2761818*C*S/(6.3136+C)
-      #Respiration
-      r=m0*S+0.2373511*R*(1-0.3565256)
-
-      dR<-Cu-0.2373511*R
-      dS<-0.2373511*R*0.3565256 - m0*S - 0.004915818*S
-      dC<--Cu + 0.004915818*S
-
-      return(list(c(dR, dS, dC), r=r, Cmic=R+S))
-    })
-}
-
-#Biomass prediction
-##Broth
-Bpred<-d[d$Structure=="Broth", c("Structure", "Time", "r")]
-
-Bout <- as.data.frame(ode(y=c(R=0.06185237*(2.01*0.51/12.01/4/deb_k2_all_fixYukfVmaxKm[[1]]$pars[["fds"]]), 
-            S=2.01*0.51/12.01/4/deb_k2_all_fixYukfVmaxKm[[1]]$pars[["fds"]], 
-            C=25), 
-        parms=deb_k2_all_fixYukfVmaxKm[[1]]$pars, func=DEB, times=sort(Bpred$Time))) %>% select(time, Cmic)
-###Merge
-Bpred<-merge(Bpred, Bout, by.x = c("Time"), by.y = c("time"))
-
-##Mixed glass
-Mpred<-d[d$Structure=="Mixed glass", c("Structure", "Time", "r")]
-
-Mout <- as.data.frame(ode(y=c(R=0.06185237*(2.01*0.51/12.01/4/deb_k2_all_fixYukfVmaxKm[[3]]$pars[["fds"]]), 
-            S=2.01*0.51/12.01/4/deb_k2_all_fixYukfVmaxKm[[3]]$pars[["fds"]], 
-            C=25), 
-        parms=deb_k2_all_fixYukfVmaxKm[[3]]$pars, func=DEB, times=sort(Bpred$Time))) %>% select(time, Cmic)
-###Merge
-Mpred<-merge(Mpred, Mout, by.x = c("Time"), by.y = c("time"))
-
-##Glass wool
-Gpred<-d[d$Structure=="Glass wool", c("Structure", "Time", "r")]
-
-Gout <- as.data.frame(ode(y=c(R=0.06185237*(2.01*0.51/12.01/4/deb_k2_all_fixYukfVmaxKm[[2]]$pars[["fds"]]), 
-            S=2.01*0.51/12.01/4/deb_k2_all_fixYukfVmaxKm[[2]]$pars[["fds"]], 
-            C=25), 
-        parms=deb_k2_all_fixYukfVmaxKm[[2]]$pars, func=DEB, times=sort(Bpred$Time))) %>% select(time, Cmic)
-###Merge
-Gpred<-merge(Gpred, Gout, by.x = c("Time"), by.y = c("time"))
-
-Pd<-rbind(Bpred, Mpred, Gpred)
-
-######################################################
-no_cors<-detectCores()-1
-cl<-makeCluster(no_cors)
-registerDoParallel(cl)
-######################################################
-#Use monod model to fit the data
-source("./Rscripts/monod_test.R")
-monod_test_out<-monod_test(data = Pd, FACT = 2)
-monod_test_out$goodness
-
-#Use monod model to fit the data
-source("../Rscripts/mend_test.R")
-mend_test_out<-mend_test(data = Pd, FACT = 2)
-mend_test_out$goodness
-
-stopImplicitCluster()
+  ggtitle("d)")+
+  theme_min+xlab("16S and ITS Distance")+
+  ylab("Proteins abundance (relative difference)")+
+  theme(plot.margin = unit(c(0.5,1,0.5,2), "in"))
+
+grid.arrange(heat1, heat2, corplot1, corplot2, nrow=2)
+
 ```
@@
 ```{r Monod and Monod plus model parameters, echo=FALSE, fig.height=24, fig.width=20, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Fig.
-S1: Monod (a) and Monod plus (b) model parameters ($V_{MAX}$ - maximum velocity constant, $K_{M}$ - affinity constant, CUE - carbon use efficiency, $k_{MB}$ - microbial biomass decay rate constant, $m_{R}$ - specific maintenance rate constant) estimated separately for three different types of physical barriers presented in incubation bottle (BROTH - black symbols, GLASS - grey symbols and WOOL - empty symbols).
+S1: Monod (a) and Monod plus (b) model parameters (v - maximum velocity constant, k - affinity constant, $Y_{CUE}$ - carbon use efficiency, m - microbial biomass decay rate constant, h - specific maintenance rate constant) estimated separately for three different types of physical barriers presented in incubation bottle (BROTH - black symbols, GLASS - grey symbols and WOOL - empty symbols).
 Symbols represent best parameters values estimated by the Differential Evolution Algorithm and error bars represent standard deviation of posterior parameters distribution estimated by Constrained Markov Chain Monte Carlo simulation.
@@
 #add proper parameter labels
-mlab<-c('Vmax' = 'V[MAX]',
-        'Km' = 'K[M]',
-        'CUE' = 'CUE',
-        'k' = 'k[MB]')
+mlab<-c('Vmax' = 'v',
+        'Km' = 'k',
+        'CUE' = 'Y[CUE]',
+        'k' = 'm')
 
@@
 #add proper parameter labels
-mlab2<-c('Vmax' = 'V[MAX]',
-        'Km' = 'K[M]',
-        'CUE' = 'CUE',
-        'mr' = 'm[R]',
-        'k' = 'k[MB]')
+mlab2<-c('Vmax' = 'v',
+        'Km' = 'k',
+        'CUE' = 'Y[CUE]',
+        'mr' = 'h',
+        'k' = 'm')
 

### 2019-11-07 Community_analysis/community_analysis.R

@@ -37,6 +37,31 @@ theme_min<-theme(axis.text.x=element_text(vjust=0.2, size=18, colour="black"),
                  legend.spacing=unit(0.5, "cm"),
                  plot.title=element_text(size=18, face="bold", hjust=-0.05))
 ###############################################################################################
+#Cell size
+size<-read.xlsx(xlsxFile = c("/home/petacapek/Dokumenty/pracovni/data_statistika/minT/Mint_CSub3_cellsort_size_dnaRFU.xlsx"),
+                2)
+summary(size)
+#Broth
+mean(c(size$B1, size$B2, size$B3, size$B4), na.rm=T)
+#Mixed glass
+mean(c(size$MG0, size$MG1, size$MG2, size$MG3, size$MG4), na.rm=T)
+#Glass wool
+mean(c(size$GW0, size$GW1, size$GW2, size$GW3, size$GW4), na.rm=T)
+
+hist(c(size$B1, size$B2, size$B3, size$B4))
+hist(c(size$MG0, size$MG1, size$MG2, size$MG3, size$MG4))
+hist(c(size$GW0, size$GW1, size$GW2, size$GW3, size$GW4))
+
+dsize<-data.frame(size=c(size$B1, size$B2, size$B3, size$B4,
+                         size$MG0, size$MG1, size$MG2, size$MG3, size$MG4,
+                         size$GW0, size$GW1, size$GW2, size$GW3, size$GW4),
+                  Structure = c(rep("BROTH", times=length(c(size$B1, size$B2, size$B3, size$B4))),
+                                rep("GLASS", times=length(c(size$MG0, size$MG1, size$MG2, size$MG3, size$MG4))),
+                                rep("WOOL", times=length(c(size$GW0, size$GW1, size$GW2, size$GW3, size$GW4)))))
+
+ggplot(dsize, aes(x=size)) + geom_histogram(aes(fill=Structure), alpha=0.5,
+                                            position = position_dodge())
+###############################################################################################
 #DATA
[…]

### 2019-11-07 Rscripts/two_pool_all_fixYukfVmaxKm.R

@@ -1,252 +1,0 @@
-two_pool_all_fixYukfVmaxKm<-function(data, FACT){
-  
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-  
-  
-  #if factor needs to be included, its levels are defined
-  if(FACT==1){
-    
-    dat<-data
-    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
-    
-  }else{
-    
-    if(FACT==2){
-      
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-      
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-      
-    }else{
-      
-      if(FACT==3){
-        
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-        
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-        
-        
-      }else{
-        
-        dat<-data
-        dat$id<-c(1)
-        
-      }
-    }
-  }
-  
-  #deb model
-  deriv<-function(time, state, pars){
-    
-    with(as.list(c(state, pars)),{
-      
-      #Carbon uptake
-      Cu=0.2761818*C*S/(6.3136+C)
-      #Respiration
-      r=m0*S+0.2373511*R*(1-0.3565256)
-      
-      Protc=fpr*R+fps*S
-      DNAc=fds*S
-      
-      dR<-Cu-0.2373511*R
-      dS<-0.2373511*R*0.3565256 - m0*S - 0.004915818*S
-      dC<--Cu + 0.004915818*S
-      
-      
-      return(list(c(dR, dS, dC), r=r, Protc=Protc, DNAc=DNAc))
-      
-    })
-  }
-  #define names of parameters
-  parnames<-c("m0", "fpr", "fps", "fds")
-  
-  #parameters estimation function
-  estim<-function(odeset){
-    
-    #defining cost function
-    cost<-function(x){
-      
-      par<-x[1:length(parnames)]
-      
-      names(par)<-parnames
-      
-      Sinit=2.01*0.51/12.01/4/par[["fds"]]
-      Rinit=0.06185237*Sinit
-      
[…]

### 2019-11-20 SI_code_revised.Rmd
Model renamed

@@
 ---
-title: The effect of habitat structure on microbial community function and composition
-author: "Petr Capek, Katherine Todd-Brown, Natalie Sadler, Jianqiu Zheng, Nancy Hess, Kirsten Hofmockel"
+title: Towards more realistic model representation of microbial population dynamics in complex environment
+author: "Petr Capek, Katherine Todd-Brown, Natalie Sadler, Michal Choma, Jianqiu Zheng, Nancy Hess, Kirsten Hofmockel"
 date: "`r Sys.Date()`"
@@
 
-####Two pool model
-```{r Two pool model respiration calibration, eval=F, echo=T}
+####Sub-microbial model
+```{r Sub-microbial model respiration calibration, eval=F, echo=T}
 #Loading the function 
@@
 
-```{r Two pool model respiration calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
+```{r Sub-microbial model respiration calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
 
@@
 
-kable(t.s3.1, digits = c(2, 2, 3, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('v', 'k', 'h', 'f', 'm', 'Y'), row.names=T, caption = "_**Table S7:** Two pool model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s3.1, digits = c(2, 2, 3, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('v', 'k', 'h', 'f', 'm', 'Y'), row.names=T, caption = "_**Table S7:** Sub-microbial model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(two_pool_r_results_good[, c(3:6)], digits = c(2, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S8:** Goodness of fit of Two pool model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(two_pool_r_results_good[, c(3:6)], digits = c(2, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S8:** Goodness of fit of Sub-microbial model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 For Monod and Monod plus models, initial concentration of microbial biomass is defined as an initial concentration of the DNA in the incubation vial divided by the conversion factor between the DNA concentration and microbial biomass.
-For Two pool model, initial concentration of Structures is defined similarly - i.e.
+For Sub-microbial model, initial concentration of Structures is defined similarly - i.e.
 initial concentration of the DNA in the incubation vial divided by the conversion factor between the DNA concentration and Structures.
@@
 
-####Two pool model
+####Sub-microbial model
 Not all model parameters vary significantly across treatments. Parameters that doesn't show any variability across treatments are fixed.
-```{r Two pool model calibration, eval=F, echo=T}
+```{r Sub-microbial model calibration, eval=F, echo=T}
 #Loading the function 
@@
 
-```{r Two pool model overall calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
+```{r Sub-microbial model overall calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
 
@@
 
-kable(t.s6.1, digits = c(0, 2, 2, 3, 3, 3, 2, 3, 3, 3, 2), align = 'c', "html", booktabs=T, col.names = c(" ","v", 'k', 'h', 'f', 'm', 'Y', '$\\B_{R-Proteins}$', '$\\B_{S-Proteins}$', '$\\B_{S-DNA}$', '$\\RS_{0}$'), row.names=F, caption = "_**Table S21:** Two pool model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s6.1, digits = c(0, 2, 2, 3, 3, 3, 2, 3, 3, 3, 2), align = 'c', "html", booktabs=T, col.names = c(" ","v", 'k', 'h', 'f', 'm', 'Y', '$\\B_{R-Proteins}$', '$\\B_{S-Proteins}$', '$\\B_{S-DNA}$', '$\\RS_{0}$'), row.names=F, caption = "_**Table S21:** Sub-microbial model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(two_pool_all_results_good[, c(1, 4:7)], digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S21:** Goodness of fit of Two pool model calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(two_pool_all_results_good[, c(1, 4:7)], digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S21:** Goodness of fit of Sub-microbial model calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%

### 2019-11-20 Community_analysis/community_analysis.R

@@ -372,9 +372,10 @@ ftax_env$IDs<-colnames(ftaxt)
 ##Genus
 ftaxtg<-ftaxt
 ftaxtg$Genus<-f.norm_taxt$Genus
+ftaxtg$Order<-f.norm_taxt$Order
 
-Ftaxtg<-melt(ftaxtg, id.vars = c("Genus"))
-colnames(Ftaxtg)<-c("Genus", "IDs","Abundance")
+Ftaxtg<-melt(ftaxtg, id.vars = c("Genus", "Order"))
+colnames(Ftaxtg)<-c("Genus", "Order","IDs","Abundance")
 
 Ftaxtg<-merge(Ftaxtg, ftax_env, by.x = "IDs", by.y = "IDs", all.x = TRUE)
 Ftaxtg$Total<-numeric(length = nrow(Ftaxtg))
@@ -382,7 +383,7 @@ for(i in unique(Ftaxtg$IDs)){
   Ftaxtg[Ftaxtg$IDs==i, "Total"]<-sum(as.numeric(Ftaxtg[Ftaxtg$IDs==i, "Abundance"]))
 }
 
-f_gen_f<-Ftaxtg %>% group_by(Genus, Structure, Substrate) %>% 
+f_gen_f<-Ftaxtg %>% group_by(Genus, Order, Structure, Substrate) %>% 
   summarize(Abundance=mean(Abundance/Total, na.rm=T))
 
 
@@ -403,10 +404,21 @@ ggplot(f_gen_f[f_gen_f$Abundance>0.0005, ], aes(Structure, Genus)) +
   theme_min + theme(legend.position = "bottom",
                     plot.margin = unit(c(0.01, 0, 0.01, 0), "in"))
 
+unique(f_gen_f$Order)
+
 f_gen_f %>% filter(Genus=="Gibberella") %>% group_by(Structure) %>% 
   summarize(Abundance=mean(Abundance))
 f_gen_f %>% filter(Genus=="Verticillium") %>% group_by(Structure) %>% 
   summarize(Abundance=mean(Abundance, na.rm=T))
+
+f_gen_f %>% filter(Order=="Saccharomycetales" |
+                     Order=="Cystofilobasidiales" |
+                     Order=="Sporidiobolales" |
+                     Order=="Tremellales" |
+                     Order=="Chaetothyriales") %>% group_by(Structure) %>% 
+  summarize(Abundance=mean(Abundance*100, na.rm=T))
+
+
[…]

### 2020-03-30 SI_code_revised.Rmd
Code revision - removing community data and Monod plus model, simplyfing Sub-microbial model

@@
 DNA and protein concentrations are in $\mu$grams per incubation vial.
-Both values are recalculated to unit of carbon per one ml of media in the vial (4 ml) using the conversion factors reported in Vrede et al.
+Both values are converted to unit of carbon per one ml of media in the vial (4 ml) using the conversion factors reported in Vrede et al.
 (2004).
@@
 Linear Mixed-Effect models are used to evaluate the effect of substrate and the type of physical barrier on measured respiration rate, DNA and protein concentration.
-Becuase all three variables display significant variability over time, the time of measurement is treated as a random effect.
+Because all three variables display significant variability over time, the time of measurement is treated as a random effect factor.
 Type of the barrier is nested within the substrate.
@@
 $~$
-Next, the effect of the substrate and the presence of a physical barrier on microbial community structure and diversity is evaluated.
-First, the bray distance indices are calculated on a normalized data.
-Then, Permutational Multivariate Analysis of Variance is used to estimate the treatment effects.
-Analysis permutations are constrained to an individual times of incubation to avoid the potential confounding effect of time.
-
-```{r vegan analysis, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-#Bacteria
-bac.norm<-decostand(bac2, method=c("normalize"))
-bac.dist<-vegdist(bac.norm, method = "bray")
-bac.ad<-adonis(bac.dist~Substrate+Structure/Substrate, bac_env, strata = bac_env$Day)
-
-#Fungi
-fungi.norm<-decostand(fungi2, method=c("normalize"))
-fungi.dist<-vegdist(fungi.norm, method = "bray")
-fungi.ad<-adonis(fungi.dist~Substrate+Structure/Substrate, fungi_env, strata = fungi_env$Day)
-
-t.s01<-cbind(as.data.frame(bac.ad$aov[c(4, 1, 6)])[c(1:3), ], as.data.frame(fungi.ad$aov[c(4, 1, 6)])[c(1:3), ])
-
-row.names(t.s01)<-c("Substrate", "Structure", "Substrate*Structure")
-
-kable(t.s01, digits = c(2, 1, 3, 2, 1, 3), align = 'c', "html", booktabs=T, col.names = c("F", "df", "p", "F", "df", "p"), row.names=T, caption = "_**Table S2:** Results of permutational multivariate analysis of variance.
-Effect of substrate, the type of physical barrier (Structure) and their interaction was evaluated.
-Statistical analysis was conducted for bacterial and fungal part of the microbial community separately.
-F values, degrees of freedom (df) and the probability (p) are reported._") %>%
-  kable_styling() %>% 
-  add_header_above(c(" " = 1, "Bacteria" = 3, "Fungi" = 3),
-                   italic=T)
-
-```
-$~$
-$~$
-$~$
-$~$
-$~$
-$~$
-```{r community figure, echo=FALSE, fig.height=5, fig.width=10, fig.cap="Fig.
-S2: Multivariate distances between six experimental groups that combine two different substrate treatments (Glucose or Cellobiose) and three different types of physical barrier (BROTH, WOOL, GLASS - black, grey and empty symbols respectively) in respect to structure of bacterial ((a); 16S) and fungal ((b); ITS) part of microbial community."}
-#Bacteria
-bacm<-as.matrix(bac.dist)
-#Intercept - Glucose - BROTH
-mGB<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
-       intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Glucose"))]
-mGB[mGB==0]<-NA
-#Glucose - GLASS
-mGG<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
-       intersect(which(bac_env$Structure=="GLASS"),which(bac_env$Substrate=="Glucose"))]
-mGG[mGG==0]<-NA
-#Glucose - WOOL
-mGW<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
-       intersect(which(bac_env$Structure=="WOOL"),which(bac_env$Substrate=="Glucose"))]
-mGW[mGW==0]<-NA
-#Cellobiose - BROTH
-mCB<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
-       intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose"))]
-mCB[mCB==0]<-NA
-#Cellobiose - GLASS
-mCG<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
-       intersect(which(bac_env$Structure=="GLASS"),which(bac_env$Substrate=="Cellobiose"))]
-mCG[mCG==0]<-NA
-#Cellobiose - WOOL
-mCW<-bacm[intersect(which(bac_env$Structure=="BROTH"),which(bac_env$Substrate=="Cellobiose")), 
-       intersect(which(bac_env$Structure=="WOOL"),which(bac_env$Substrate=="Cellobiose"))]
-mCW[mCW==0]<-NA
-
-bacdiff<-data.frame(Dist=c(mean(mGB, na.rm=T), mean(mGG, na.rm=T), mean(mGW, na.rm=T), 0, mean(mCG, na.rm=T), mean(mCW, na.rm=T)),
-                    Dist.sd=c(sd(mGB, na.rm=T), sd(mGG, na.rm=T), sd(mGW, na.rm=T), sd(mCB, na.rm=T), sd(mCG, na.rm=T), sd(mCW, na.rm=T)),
-                    Substrate=c(rep("Glucose", 3), rep("Cellobiose", 3)),
-                    Structure=rep(c("BROTH", "GLASS", "WOOL"), 2))
-
-#Fungi
-fm<-as.matrix(fungi.dist)
-#Intercept - Glucose - BROTH
-fGB<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
-          intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Glucose"))]
-fGB[fGB==0]<-NA
-#Glucose - GLASS
-fGG<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
-          intersect(which(fungi_env$Structure=="GLASS"),which(fungi_env$Substrate=="Glucose"))]
-fGG[fGG==0]<-NA
-#Glucose - WOOL
-fGW<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
-          intersect(which(fungi_env$Structure=="WOOL"),which(fungi_env$Substrate=="Glucose"))]
-fGW[fGW==0]<-NA
-#Cellobiose - BROTH
-fCB<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
-          intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose"))]
-fCB[fCB==0]<-NA
-#Cellobiose - GLASS
-fCG<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
-          intersect(which(fungi_env$Structure=="GLASS"),which(fungi_env$Substrate=="Cellobiose"))]
-fCG[fCG==0]<-NA
-#Cellobiose - WOOL
-fCW<-fm[intersect(which(fungi_env$Structure=="BROTH"),which(fungi_env$Substrate=="Cellobiose")), 
-          intersect(which(fungi_env$Structure=="WOOL"),which(fungi_env$Substrate=="Cellobiose"))]
-fCW[fCW==0]<-NA
-
-fdiff<-data.frame(Dist=c(mean(fGB, na.rm=T), mean(fGG, na.rm=T), mean(fGW, na.rm=T), 0, mean(fCG, na.rm=T), mean(fCW, na.rm=T)),
-                    Dist.sd=c(sd(fGB, na.rm=T), sd(fGG, na.rm=T), sd(fGW, na.rm=T), sd(fCB, na.rm=T), sd(fCG, na.rm=T), sd(fCW, na.rm=T)),
-                    Substrate=c(rep("Glucose", 3), rep("Cellobiose", 3)),
-                    Structure=rep(c("BROTH", "GLASS", "WOOL"), 2))
-
-bacplot<-ggplot(bacdiff, aes(Substrate, Dist))+
-  geom_point(position = position_dodge(width = 0.5), aes(fill=Structure), shape=21, cex=6)+
-  scale_fill_manual(values = c("black", "grey", "white"))+
-  scale_colour_manual(values = c("black", "black", "black"))+
-  geom_errorbar(aes(ymin=Dist-Dist.sd, ymax=Dist+Dist.sd, colour=Structure),
-                position=position_dodge(width = 0.5), width=0.1)+
-  ylab("16S Distance")+
-  ggtitle("a)")+
-  theme_min+theme(axis.title.x = element_blank(),
-                  legend.title = element_blank(),
-                  legend.position = c(0.8, 0.2))
-fplot<-ggplot(fdiff, aes(Substrate, Dist))+
-  geom_point(position = position_dodge(width = 0.5), aes(fill=Structure), shape=21, cex=6, show.legend = F)+
-  scale_fill_manual(values = c("black", "grey", "white"))+
-  scale_colour_manual(values = c("black", "black", "black"))+
-  geom_errorbar(aes(ymin=Dist-Dist.sd, ymax=Dist+Dist.sd, colour=Structure),
-                position=position_dodge(width = 0.5), width=0.1, show.legend = F)+
-  ylab("ITS Distance")+
-  ggtitle("b)")+
-  theme_min+theme(axis.title.x = element_blank(),
-                  legend.title = element_blank())
-
-grid.arrange(bacplot, fplot, nrow=1)
-
-```
-
-$~$
-$~$
-$~$
-
 ##Microbial-explicit models analysis
-###All models are first calibrated against respiration rate
-Calibration of each model is done in four steps.
+###Both models are first calibrated against respiration rate
+Calibration of both models is done in four steps.
 First, the model is calibrated against data across all treatments.
@@
 #########################################################
-no_cors<-detectCores()
+no_cors<-detectCores()-1
 cl<-makeCluster(no_cors)
@@
 kable(monod_r_results_good[c(3:6)], digits = c(2, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S4:** Goodness of fit of monod model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
-See Materials and Methods for symbols explanation._") %>%
-  kable_styling() %>% 
-  group_rows("All", 1, 1) %>%
-  group_rows("Structure", 2, 2) %>%
-  group_rows("Substrate", 3, 4) %>%
-  group_rows("Each", 4, 4)
-
-```
-
-####Monod plus model
-```{r Monod plus respiration calibration, eval=F, echo=T}
-#Loading the function 
-source("../Rscripts/monodplus_r.R")
-source("../Rscripts/monodplus_r_fix.R")
-#across all treatments
-monodplus_r1<-monodplus_r(data=d, FACT = 4)
-##################################################
-no_cors<-detectCores()-1
-cl<-makeCluster(no_cors)
-registerDoParallel(cl)
-##################################################
-#for different structures
-monodplus_r2<-monodplus_r(data=d, FACT = 2)
-monodplus_r2_fix<-monodplus_r_fix(data=d, FACT = 2)
-#for different substrates
-monodplus_r3<-monodplus_r(data=d, FACT = 1)
-monodplus_r3_fix<-monodplus_r_fix(data=d, FACT = 1)
-#for each separately 
-monodplus_r4<-monodplus_r(data=d, FACT = 3)
-monodplus_r4_fix<-monodplus_r_fix(data=d, FACT = 3)
-##################################################
-stopImplicitCluster()
-##################################################
-monodplus_r1$goodness
-monodplus_r2$goodness
-monodplus_r3$goodness
-monodplus_r4$goodness
-
-# mean(c(monodplus_r4[[1]]$pars[["Cmic_0"]], monodplus_r4[[2]]$pars[["Cmic_0"]],
-#        monodplus_r4[[3]]$pars[["Cmic_0"]], monodplus_r4[[4]]$pars[["Cmic_0"]],
-#        monodplus_r4[[5]]$pars[["Cmic_0"]], monodplus_r4[[6]]$pars[["Cmic_0"]]))
-
-monodplus_r1$goodness
-monodplus_r2_fix$goodness
-monodplus_r3_fix$goodness
-monodplus_r4_fix$goodness
-```
-
-```{r Monod plus respiration calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
-
-#results of the previous iteration (not run)
-# monodplus_r_results_par<-as.data.frame(rbind(monodplus_r1[[1]]$pars[c(1:5)],
-#                     monodplus_r2_fix[[1]]$pars, monodplus_r2_fix[[2]]$pars,
-#                     monodplus_r2_fix[[3]]$pars,
-#                     monodplus_r3_fix[[1]]$pars, monodplus_r3_fix[[2]]$pars,
-#                     monodplus_r4_fix[[1]]$pars, monodplus_r4_fix[[2]]$pars,
-#                     monodplus_r4_fix[[3]]$pars, monodplus_r4_fix[[4]]$pars,
-#                     monodplus_r4_fix[[5]]$pars, monodplus_r4_fix[[6]]$pars))
-# monodplus_r_results_good<-as.data.frame(rbind(monodplus_r1$goodness,
-#                                               monodplus_r2_fix$goodness,
-#                                               monodplus_r3_fix$goodness,
-#                                               monodplus_r4_fix$goodness))
-# write.csv(monodplus_r_results_par, file = c("../monodplus_r_results_par.csv"), row.names = F)
-# write.csv(monodplus_r_results_good, file = c("../monodplus_r_results_good.csv"), row.names = F)
-
-monodplus_r_results_par<-read.csv("monodplus_r_results_par.csv")
-monodplus_r_results_good<-read.csv("monodplus_r_results_good.csv")
-
-t.s2.1<-data.frame(Vmax=monodplus_r_results_par[,1],
-                 Km=monodplus_r_results_par[,2],
-                 CUE=monodplus_r_results_par[,3],
-                 mr=monodplus_r_results_par[,4],
-                 k=monodplus_r_results_par[,5])
-row.names(t.s2.1)<-c("All", "Broth", "Glass", "Wool",
-                    "Celluloze", "Glucose",
-                    "Broth - Cellobiose", "Broth - Glucose",
-                    "Glass - Cellobiose", "Glass - Glucose",
-                    "Wool - Cellobiose", "Wool - Glucose")
-
-options(knitr.kable.NA = '')
-
-kable(t.s2.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('v', 'k', '$\\Y_{CUE}$', 'h', 'm'), row.names=T, caption = "_**Table S5:** Monod plus model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
-See Materials and Methods for symbols explanation._") %>%
-  kable_styling() %>% 
-  group_rows("All", 1, 1) %>%
-  group_rows("Structure", 2, 4) %>%
-  group_rows("Substrate", 5, 6) %>%
-  group_rows("Each", 7, 12)
-
-kable(monodplus_r_results_good[,c(4, 1, 2, 3)], digits = c(2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S6:** Goodness of fit of Monnod plus model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 two_pool_r4_fix$goodness
+
 ```
@@
 #results of the previous iteration (not run)
-# two_pool_r_results_par<-rbind(two_pool_r1[[1]]$pars[c(1:6)],
+# two_pool_r_results_par<-rbind(two_pool_r1[[1]]$pars[c(1:5)],
 #                    two_pool_r2_fix[[1]]$pars, two_pool_r2_fix[[2]]$pars,
@@
                  Km=two_pool_r_results_par[,2],
-                 mo=two_pool_r_results_par[,3],
-                 f=two_pool_r_results_par[,4],
-                 k=two_pool_r_results_par[,5],
-                 Yu=two_pool_r_results_par[,6])
+                 f=two_pool_r_results_par[,3],
+                 k=two_pool_r_results_par[,4],
+                 Yu=two_pool_r_results_par[,5])
 row.names(t.s3.1)<-c("All", "Broth", "Glass", "Wool",
@@
 
-kable(t.s3.1, digits = c(2, 2, 3, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('v', 'k', 'h', 'f', 'm', 'Y'), row.names=T, caption = "_**Table S7:** Sub-microbial model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s3.1, digits = c(2, 2, 2, 3, 2), align = 'c', "html", booktabs=T, col.names = c('v', 'k', 'f', 'm', 'Y'), row.names=T, caption = "_**Table S7:** Sub-microbial model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-###All models are calibrated against respiration rate, proteins and DNA concentration in the cells.
-All models are calibrated against all data we measured.
+###Both models are calibrated against respiration rate, proteins and DNA concentration in the cells.
+Both models are calibrated against all data we measured.
 In addition to model parameters reported in tables above, conversion factors between proteins/DNA concentration and microbial biomass are estimated.
-For Monod and Monod plus models, initial concentration of microbial biomass is defined as an initial concentration of the DNA in the incubation vial divided by the conversion factor between the DNA concentration and microbial biomass.
+For Monod model, initial concentration of microbial biomass is defined as an initial concentration of the DNA in the incubation vial divided by the conversion factor between the DNA concentration and microbial biomass.
 For Sub-microbial model, initial concentration of Structures is defined similarly - i.e.
@@
 This parameter has to be fixed across treatments to reflect the reality.
-In addition, Monod and Monod plus models are calibrated against respiration rate and proteins or DNA concentration separately.
+In addition, Monod model is calibrated against respiration rate and proteins or DNA concentration separately.
 When calibrated againnst protein concentration, initial concentration of microbial biomass is estimated and fixed across treatments.
@@
 
-####Monod plus model
-```{r Monod plus model calibartion, eval=F, echo=T}
-#Proteins
+####Sub-microbial model
+Not all model parameters vary significantly across treatments. Parameters that doesn't show any variability across treatments are fixed.
+```{r Sub-microbial model calibration, eval=F, echo=T}
 #Loading the function 
-source("../Rscripts/monodplus_prot.R")
-source("../Rscripts/monodplus_prot_fix.R")
+source("./Rscripts/two_pool_allx.R")
+source("../Rscripts/two_pool_allx_fix.R")
 #across all treatments
-monodplus_prot1<-monodplus_prot(data=d, FACT = 4)
+two_pool_allx_out1<-two_pool_allx(data=d, FACT = 4)
+
 ###########################################################
@@
 #for different structures
-monodplus_prot2<-monodplus_prot(data=d, FACT = 2)
-monodplus_prot2_fix<-monodplus_prot_fix(data=d, FACT = 2)
+two_pool_allx_out2<-two_pool_allx(data=d, FACT = 2)
+
+two_pool_allx_fix_out2<-two_pool_allx_fix(data=d, FACT = 2)
+two_pool_allx_fix_out2$goodness
+
+two_pool_allx_fix_out2[[1]]$pars
+two_pool_allx_fix_out2[[2]]$pars
+two_pool_allx_fix_out2[[3]]$pars
+
+# mean(c(two_pool_allx_fix_out2[[1]]$pars[["Vmax"]],
+#        two_pool_allx_fix_out2[[2]]$pars[["Vmax"]],
+#        two_pool_allx_fix_out2[[3]]$pars[["Vmax"]]))
+# mean(c(two_pool_allx_fix_out2[[1]]$pars[["Yu"]],
+#        two_pool_allx_fix_out2[[2]]$pars[["Yu"]],
+#        two_pool_allx_fix_out2[[3]]$pars[["Yu"]]))
+# mean(c(two_pool_allx_fix_out2[[1]]$pars[["k"]],
+#        two_pool_allx_fix_out2[[2]]$pars[["k"]],
+#        two_pool_allx_fix_out2[[3]]$pars[["k"]]))
+############################################################################
+#in the following code, the Vmax, Ys and k parameters are fixed across treatments
+#becuase its variability across treatments is minimal (data not shown)
+source("./Rscripts/two_pool_all_fixVmaxYuk.R")
+two_pool_all_fixVmaxYuk_out2<-two_pool_all_fixVmaxYuk(data=d, FACT = 2)
+two_pool_all_fixVmaxYuk_out2$goodness
+
+############################################################################
 #for different substrates
-monodplus_prot3<-monodplus_prot(data=d, FACT = 1)
-monodplus_prot3_fix<-monodplus_prot_fix(data=d, FACT = 1)
+two_pool_allx_fix_out3<-two_pool_allx_fix(data=d, FACT = 1)
+two_pool_allx_fix_out3$goodness
 #for each separately 
-monodplus_prot4<-monodplus_prot(data=d, FACT = 3)
-monodplus_prot4_fix<-monodplus_prot_fix(data=d, FACT = 3)
+two_pool_allx_fix_out4<-two_pool_allx_fix(data=d, FACT = 3)
+two_pool_allx_fix_out4$goodness
+#in the following code, the Vmax Ys, and k parameters are fixed across treatments
+two_pool_all_fixVmaxYuk_out4<-two_pool_all_fixVmaxYuk(data=d, FACT = 3)
+two_pool_all_fixVmaxYuk_out4$goodness
 ###########################################################
@@
 ###########################################################
-monodplus_prot1$goodness
-monodplus_prot2_fix$goodness
-monodplus_prot3_fix$goodness
-monodplus_prot4_fix$goodness
-###########################################################
-#DNA
-#Loading the function 
-source("../Rscripts/monodplus_DNA.R")
-#across all treatments
-monodplus_DNA1<-monodplus_DNA(data=d, FACT = 4)
-###########################################################
-no_cors<-detectCores()-2
-cl<-makeCluster(no_cors)
-registerDoParallel(cl)
-###########################################################
-#for different structures
-monodplus_DNA2<-monodplus_DNA(data=d, FACT = 2)
-#for different substrates
-monodplus_DNA3<-monodplus_DNA(data=d, FACT = 1)
-#for each separately 
-monodplus_DNA4<-monodplus_DNA(data=d, FACT = 3)
-###########################################################
-stopImplicitCluster()
-###########################################################
-monodplus_DNA1$goodness
-monodplus_DNA2$goodness
-monodplus_DNA3$goodness
-monodplus_DNA4$goodness
-###########################################################
-#DNA and Proteins
-#Loading the function 
-source("../Rscripts/monodplus_all.R")
-#across all treatments
-monodplus_all1<-monodplus_all(data=d, FACT = 4)
-###########################################################
-no_cors<-detectCores()-1
-cl<-makeCluster(no_cors)
-registerDoParallel(cl)
-###########################################################
-#for different structures
-monodplus_all2<-monodplus_all(data=d, FACT = 2)
-#for different substrates
-monodplus_all3<-monodplus_all(data=d, FACT = 1)
-#for each separately 
-monodplus_all4<-monodplus_all(data=d, FACT = 3)
-###########################################################
-stopImplicitCluster()
-###########################################################
-monodplus_all1$goodness
-monodplus_all2$goodness
-monodplus_all3$goodness
-monodplus_all4$goodness
-```
-
-```{r Monod plus overall calibration results, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
-
-# #results of the previous iteration (not run)
-# ##All measurements together
-# monodplus_all_results_par<-rbind(monodplus_all1[[1]]$pars, 
-#                             monodplus_all2[[1]]$pars, monodplus_all2[[2]]$pars,
-#                             monodplus_all2[[3]]$pars, 
-#                             monodplus_all3[[1]]$pars, monodplus_all3[[2]]$pars,
-#                             monodplus_all4[[1]]$pars, monodplus_all4[[2]]$pars,
-#                             monodplus_all4[[3]]$pars, monodplus_all4[[4]]$pars,
-#                             monodplus_all4[[5]]$pars, monodplus_all4[[6]]$pars)
-# 
-# monodplus_all_results_good<-rbind(monodplus_all1$goodness[c(2, 3, 1), ],
-#                              monodplus_all2$goodness[c(2, 3, 1), ],
-#                              monodplus_all3$goodness[c(2, 3, 1), ],
-#                              monodplus_all4$goodness[c(2, 3, 1), ])
-# 
-# write.csv(monodplus_all_results_par, file = c("../monodplus_all_results_par.csv"), row.names = F)
-# write.csv(monodplus_all_results_good, file = c("../monodplus_all_results_good.csv"), row.names = F)
-# 
-# ##Respiration rate and proteins
-# monodplus_prot_results_par<-rbind(monodplus_prot1[[1]]$pars[c(1:6)], 
-#                             monodplus_prot2_fix[[1]]$pars, monodplus_prot2_fix[[2]]$pars,
-#                             monodplus_prot2_fix[[3]]$pars, 
-#                             monodplus_prot3_fix[[1]]$pars, monodplus_prot3_fix[[2]]$pars,
-#                             monodplus_prot4_fix[[1]]$pars, monodplus_prot4_fix[[2]]$pars,
-#                             monodplus_prot4_fix[[3]]$pars, monodplus_prot4_fix[[4]]$pars,
-#                             monodplus_prot4_fix[[5]]$pars, monodplus_prot4_fix[[6]]$pars)
-# 
-# monodplus_prot_results_good<-rbind(monodplus_prot1$goodness[c(2,1), ],
-#                              monodplus_prot2_fix$goodness[c(2,1), ],
-#                              monodplus_prot3_fix$goodness[c(2,1), ],
-#                              monodplus_prot4_fix$goodness[c(2,1), ])
-# 
-# write.csv(monodplus_prot_results_par, file = c("../monodplus_prot_results_par.csv"), row.names = F)
-# write.csv(monodplus_prot_results_good, file = c("../monodplus_prot_results_good.csv"), row.names = F)
-# 
-# ##Respiration rate and DNA
-# monodplus_DNA_results_par<-rbind(monodplus_DNA1[[1]]$pars, 
-#                             monodplus_DNA2[[1]]$pars, monodplus_DNA2[[2]]$pars,
-#                             monodplus_DNA2[[3]]$pars, 
-#                             monodplus_DNA3[[1]]$pars, monodplus_DNA3[[2]]$pars,
-#                             monodplus_DNA4[[1]]$pars, monodplus_DNA4[[2]]$pars,
-#                             monodplus_DNA4[[3]]$pars, monodplus_DNA4[[4]]$pars,
-#                             monodplus_DNA4[[5]]$pars, monodplus_DNA4[[6]]$pars)
-#  
-# monodplus_DNA_results_good<-rbind(monodplus_DNA1$goodness[c(2, 1), ],
-#                              monodplus_DNA2$goodness[c(2, 1), ],
-#                              monodplus_DNA3$goodness[c(2, 1), ],
-#                              monodplus_DNA4$goodness[c(2, 1), ])
-# 
-# write.csv(monodplus_DNA_results_par, file = c("../monodplus_DNA_results_par.csv"), row.names = F)
-# write.csv(monodplus_DNA_results_good, file = c("../monodplus_DNA_results_good.csv"), row.names = F)
-
-#All measurements
-monodplus_all_results_par<-read.csv("monodplus_all_results_par.csv")
-monodplus_all_results_good<-read.csv("monodplus_all_results_good.csv")
-monodplus_all_results_good$variable<-rep(c("Respiration", "Proteins", "DNA"), 4)
-monodplus_all_results_good[monodplus_all_results_good$R2<0, "R2"]<-0
-#Respiration and proteins
-monodplus_prot_results_par<-read.csv("monodplus_prot_results_par.csv")
-monodplus_prot_results_good<-read.csv("monodplus_prot_results_good.csv")
-monodplus_prot_results_good$variable<-rep(c("Respiration", "Proteins"), 4)
-monodplus_prot_results_good[monodplus_prot_results_good$R2<0, "R2"]<-0
-#Respiration and DNA
-monodplus_DNA_results_par<-read.csv("monodplus_DNA_results_par.csv")
-monodplus_DNA_results_good<-read.csv("monodplus_DNA_results_good.csv")
-monodplus_DNA_results_good$variable<-rep(c("Respiration", "DNA"), 4)
-monodplus_DNA_results_good[monodplus_DNA_results_good$R2<0, "R2"]<-0
-
-t.s5.1<-data.frame(Name=c("All", "Broth", "Glass", "Wool",
-                    "Cellobiose", "Glucose",
-                    "Broth - Cellobiose", "Broth - Glucose",
-                    "Glass - Cellobiose", "Glass - Glucose",
-                    "Wool - Cellobiose", "Wool - Glucose"),
-                   Vmax=monodplus_all_results_par[,1],
-                   Km=monodplus_all_results_par[,2],
-                   CUE=monodplus_all_results_par[,3],
-                   mr=monodplus_all_results_par[,4],
-                   k=monodplus_all_results_par[,5],
-                   fp=monodplus_all_results_par[,6],
-                   fd=monodplus_all_results_par[,7])
-
-options(knitr.kable.NA = '')
-
-kable(t.s5.1, digits = c(0, 2, 2, 2, 3, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'h', 'm', '$\\f_{Proteins}$', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S15:** Monod plus model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
-See Materials and Methods for symbols explanation._") %>%
-  kable_styling() %>% 
-  group_rows("All", 1, 1) %>%
-  group_rows("Structure", 2, 4) %>%
-  group_rows("Substrate", 5, 6) %>%
-  group_rows("Each", 7, 12)
-
-kable(monodplus_all_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S16:** Goodness of fit of Monod plus model calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
-See Materials and Methods for symbols explanation._") %>%
-  kable_styling() %>% 
-  group_rows("All", 1, 3) %>%
-  group_rows("Structure", 4, 6) %>%
-  group_rows("Substrate", 7, 9) %>%
-  group_rows("Each", 10, 12)
-
-t.s5.2<-data.frame(Name=c("All", "Broth", "Glass", "Wool",
-                    "Cellobiose", "Glucose",
-                    "Broth - Cellobiose", "Broth - Glucose",
-                    "Glass - Cellobiose", "Glass - Glucose",
-                    "Wool - Cellobiose", "Wool - Glucose"),
-                   Vmax=monodplus_prot_results_par[,1],
-                   Km=monodplus_prot_results_par[,2],
-                   CUE=monodplus_prot_results_par[,3],
-                   mr=monodplus_prot_results_par[,4],
-                   k=monodplus_prot_results_par[,5],
-                   fp=monodplus_prot_results_par[,6])
-
-options(knitr.kable.NA = '')
-
-kable(t.s5.2, digits = c(0, 2, 2, 2, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'h', 'm', '$\\f_{Proteins}$'), row.names=F, caption = "_**Table S17:** Monod plus model parameters calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
-See Materials and Methods for symbols explanation._") %>%
-  kable_styling() %>% 
-  group_rows("All", 1, 1) %>%
-  group_rows("Structure", 2, 4) %>%
-  group_rows("Substrate", 5, 6) %>%
-  group_rows("Each", 7, 12)
-
-kable(monodplus_prot_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S18:** Goodness of fit of Monod plus model calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
-See Materials and Methods for symbols explanation._") %>%
-  kable_styling() %>% 
-  group_rows("All", 1, 1) %>%
-  group_rows("Structure", 3, 4) %>%
-  group_rows("Substrate", 5, 6) %>%
-  group_rows("Each", 7, 8)
-
-t.s5.3<-data.frame(Name=c("All", "Broth", "Glass", "Wool",
-                    "Cellobiose", "Glucose",
-                    "Broth - Cellobiose", "Broth - Glucose",
-                    "Glass - Cellobiose", "Glass - Glucose",
-                    "Wool - Cellobiose", "Wool - Glucose"),
-                   Vmax=monodplus_DNA_results_par[,1],
-                   Km=monodplus_DNA_results_par[,2],
-                   CUE=monodplus_DNA_results_par[,3],
-                   mr=monodplus_DNA_results_par[,4],
-                   k=monodplus_DNA_results_par[,5],
-                   fd=monodplus_DNA_results_par[,6])
-
-options(knitr.kable.NA = '')
-
-kable(t.s5.3, digits = c(0, 2, 2, 2, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'h', 'm', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S19:** Monod plus model parameters calibrated against respiration rate and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
-See Materials and Methods for symbols explanation._") %>%
-  kable_styling() %>% 
-  group_rows("All", 1, 1) %>%
-  group_rows("Structure", 2, 4) %>%
-  group_rows("Substrate", 5, 6) %>%
-  group_rows("Each", 7, 12)
-
-kable(monodplus_DNA_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S20:** Goodness of fit of Monod plus  model calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
-See Materials and Methods for symbols explanation._") %>%
-  kable_styling() %>% 
-  group_rows("All", 1, 2) %>%
-  group_rows("Structure", 3, 4) %>%
-  group_rows("Substrate", 5, 6) %>%
-  group_rows("Each", 7, 8)
-```
-
-####Sub-microbial model
-Not all model parameters vary significantly across treatments. Parameters that doesn't show any variability across treatments are fixed.
-```{r Sub-microbial model calibration, eval=F, echo=T}
-#Loading the function 
-source("../Rscripts/two_pool_all.R")
-source("../Rscripts/two_pool_all_fix.R")
-#across all treatments
-two_pool_all1<-two_pool_all(data=d, FACT = 4)
-###########################################################
-no_cors<-detectCores()-1
-cl<-makeCluster(no_cors)
-registerDoParallel(cl)
-###########################################################
-#for different structures
-two_pool_all2<-two_pool_all(data=d, FACT = 2)
-two_pool_all2_fix<-two_pool_all_fix(data=d, FACT = 2)
-############################################################################
-#in the following code, the Ys and k parameters are fixed across treatments
-#becuase its variability across treatments is minimal (data not shown)
-source("./Rscripts/two_pool_all_fixYuk.R")
-two_pool_all_fixYuk2<-two_pool_all_fixYuk(data=d, FACT = 2)
-#Then, parameter f is fixed as well
-source("./Rscripts/two_pool_all_fixYukf.R")
-two_pool_all_fixYukf2<-two_pool_all_fixYukf(data=d, FACT = 2)
-#Then, parameter Vmax is fixed as well
-source("./Rscripts/two_pool_all_fixYukfVmax.R")
-two_pool_all_fixYukfVmax2<-two_pool_all_fixYukfVmax(data=d, FACT = 2)
-############################################################################
-#for different substrates
-two_pool_all3<-two_pool_all(data=d, FACT = 1)
-two_pool_all3_fix<-two_pool_all_fix(data=d, FACT = 1)
-#for each separately 
-two_pool_all4<-two_pool_all(data=d, FACT = 3)
-two_pool_all4_fix<-two_pool_all_fix(data=d, FACT = 3)
-#in the following code, the Ys, k,f and Vmax parameters are fixed across treatments
-two_pool_all_fixYukfVmax4<-two_pool_all_fixYukfVmax(data=d, FACT = 3)
-###########################################################
-stopImplicitCluster()
-###########################################################
-two_pool_all1$goodness
-two_pool_all_fixYukfVmax2$goodness
-two_pool_all3_fix$goodness
-two_pool_all_fixYukfVmax4$goodness
+two_pool_allx_out1$goodness
+two_pool_all_fixVmaxYuk_out2$goodness
+two_pool_allx_fix_out3$goodness
+two_pool_all_fixVmaxYuk_out4$goodness
+
+# mean(c(two_pool_allx_out4[[1]]$pars[["RS_0"]],
+#        two_pool_allx_out4[[2]]$pars[["RS_0"]],
+#        two_pool_allx_out4[[3]]$pars[["RS_0"]],
+#        #two_pool_allx_out4[[4]]$pars[["RS_0"]],
+#        two_pool_allx_out4[[5]]$pars[["RS_0"]],
+#        two_pool_allx_out4[[6]]$pars[["RS_0"]]))
+
 ```
@@
 
-# #results of the previous iteration (not run)
-# two_pool_all_results_par<-rbind(c(two_pool_all1[[1]]$pars[c(1:9)], RS_0 = 0.06185237),
-#                          c(Vmax = 0.2761818, 
-#                            two_pool_all_fixYukfVmax2[[1]]$pars[1],
-#                            two_pool_all_fixYukfVmax2[[1]]$pars[2],
-#                            f = 0.2373511,
-#                            k = 0.004915818,
-#                            Yu = 0.3565256,
-#                            two_pool_all_fixYukfVmax2[[1]]$pars[c(3:5)],
-#                            RS_0 = 0.06185237),
-#                          c(Vmax = 0.2761818, 
-#                            two_pool_all_fixYukfVmax2[[2]]$pars[1],
-#                            two_pool_all_fixYukfVmax2[[2]]$pars[2],
-#                            f = 0.2373511,
-#                            k = 0.004915818,
-#                            Yu = 0.3565256,
-#                            two_pool_all_fixYukfVmax2[[2]]$pars[c(3:5)],
-#                            RS_0 = 0.06185237),
-#                          c(Vmax = 0.2761818, 
-#                            two_pool_all_fixYukfVmax2[[3]]$pars[1],
-#                            two_pool_all_fixYukfVmax2[[3]]$pars[2],
-#                            f = 0.2373511,
-#                            k = 0.004915818,
-#                            Yu = 0.3565256,
-#                            two_pool_all_fixYukfVmax2[[3]]$pars[c(3:5)],
-#                            RS_0 = 0.06185237),
-#                          c(two_pool_all3_fix[[1]]$pars[c(1:9)], RS_0 = 0.06185237),
-#                          c(two_pool_all3_fix[[2]]$pars[c(1:9)], RS_0 = 0.06185237),
-#                          c(Vmax = 0.2761818, 
-#                            two_pool_all_fixYukfVmax4[[1]]$pars[c(1:2)],
-#                            f = 0.2373511,
-#                            k = 0.004915818,
-#                            Yu = 0.3565256,
-#                            two_pool_all_fixYukfVmax4[[1]]$pars[c(3:5)],
-#                            RS_0 = 0.06185237),
-#                          c(Vmax = 0.2761818, 
-#                            two_pool_all_fixYukfVmax4[[2]]$pars[c(1:2)],
-#                            f = 0.2373511,
-#                            k = 0.004915818,
-#                            Yu = 0.3565256,
-#                            two_pool_all_fixYukfVmax4[[2]]$pars[c(3:5)],
-#                            RS_0 = 0.06185237),
-#                          c(Vmax = 0.2761818, 
-#                            two_pool_all_fixYukfVmax4[[3]]$pars[c(1:2)],
-#                            f = 0.2373511,
-#                            k = 0.004915818,
-#                            Yu = 0.3565256,
-#                            two_pool_all_fixYukfVmax4[[3]]$pars[c(3:5)],
-#                            RS_0 = 0.06185237),
-#                          c(Vmax = 0.2761818, 
-#                            two_pool_all_fixYukfVmax4[[4]]$pars[c(1:2)],
-#                            f = 0.2373511,
-#                            k = 0.004915818,
-#                            Yu = 0.3565256,
-#                            two_pool_all_fixYukfVmax4[[4]]$pars[c(3:5)],
-#                            RS_0 = 0.06185237),
-#                          c(Vmax = 0.2761818, 
-#                            two_pool_all_fixYukfVmax4[[5]]$pars[c(1:2)],
-#                            f = 0.2373511,
-#                            k = 0.004915818,
-#                            Yu = 0.3565256,
-#                            two_pool_all_fixYukfVmax4[[5]]$pars[c(3:5)],
-#                            RS_0 = 0.06185237),
-#                          c(Vmax = 0.2761818, 
-#                            two_pool_all_fixYukfVmax4[[6]]$pars[c(1:2)],
-#                            f = 0.2373511,
-#                            k = 0.004915818,
-#                            Yu = 0.3565256,
-#                            two_pool_all_fixYukfVmax4[[6]]$pars[c(3:5)],
-#                            RS_0 = 0.06185237)
+#results of the previous iteration (not run)
+# two_pool_all_results_par<-rbind(c(two_pool_allx_out1[[1]]$pars[c(1:8)], RS_0 = 0.04012019),
+#                         c(Vmax = 0.2468173, 
+#                           two_pool_all_fixVmaxYuk_out2[[1]]$pars[1],
+#                           two_pool_all_fixVmaxYuk_out2[[1]]$pars[2],
+#                           k = 0.007046304,
+#                           Yu = 0.3731948,
+#                           two_pool_all_fixVmaxYuk_out2[[1]]$pars[c(3:5)],
+#                           RS_0 = 0.04012019),
+#                         c(Vmax = 0.2468173, 
+#                           two_pool_all_fixVmaxYuk_out2[[2]]$pars[1],
+#                           two_pool_all_fixVmaxYuk_out2[[2]]$pars[2],
+#                           k = 0.007046304,
+#                           Yu = 0.3731948,
+#                           two_pool_all_fixVmaxYuk_out2[[2]]$pars[c(3:5)],
+#                           RS_0 = 0.04012019),
+#                         c(Vmax = 0.2468173, 
+#                           two_pool_all_fixVmaxYuk_out2[[3]]$pars[1],
+#                           two_pool_all_fixVmaxYuk_out2[[3]]$pars[2],
+#                           k = 0.007046304,
+#                           Yu = 0.3731948,
+#                           two_pool_all_fixVmaxYuk_out2[[3]]$pars[c(3:5)],
+#                           RS_0 = 0.04012019),
+#                         c(two_pool_allx_fix_out3[[1]]$pars[c(1:8)], RS_0 = 0.04012019),
+#                         c(two_pool_allx_fix_out3[[2]]$pars[c(1:8)], RS_0 = 0.04012019),
+#                         c(Vmax = 0.2468173, 
+#                           two_pool_all_fixVmaxYuk_out4[[1]]$pars[c(1:2)],
+#                           k = 0.007046304,
+#                           Yu = 0.3731948,
+#                           two_pool_all_fixVmaxYuk_out4[[1]]$pars[c(3:5)],
+#                           RS_0 = 0.04012019),
+#                         c(Vmax = 0.2468173, 
+#                           two_pool_all_fixVmaxYuk_out4[[2]]$pars[c(1:2)],
+#                           k = 0.007046304,
+#                           Yu = 0.3731948,
+#                           two_pool_all_fixVmaxYuk_out4[[2]]$pars[c(3:5)],
+#                           RS_0 = 0.04012019),
+#                         c(Vmax = 0.2468173, 
+#                           two_pool_all_fixVmaxYuk_out4[[3]]$pars[c(1:2)],
+#                           k = 0.007046304,
+#                           Yu = 0.3731948,
+#                           two_pool_all_fixVmaxYuk_out4[[3]]$pars[c(3:5)],
+#                           RS_0 = 0.04012019),
+#                         c(Vmax = 0.2468173, 
+#                           two_pool_all_fixVmaxYuk_out4[[4]]$pars[c(1:2)],
+#                           k = 0.007046304,
+#                           Yu = 0.3731948,
+#                           two_pool_all_fixVmaxYuk_out4[[4]]$pars[c(3:5)],
+#                           RS_0 = 0.04012019),
+#                         c(Vmax = 0.2468173, 
+#                           two_pool_all_fixVmaxYuk_out4[[5]]$pars[c(1:2)],
+#                           k = 0.007046304,
+#                           Yu = 0.3731948,
+#                           two_pool_all_fixVmaxYuk_out4[[5]]$pars[c(3:5)],
+#                           RS_0 = 0.04012019),
+#                         c(Vmax = 0.2468173, 
+#                           two_pool_all_fixVmaxYuk_out4[[6]]$pars[c(1:2)],
+#                           k = 0.007046304,
+#                           Yu = 0.3731948,
+#                           two_pool_all_fixVmaxYuk_out4[[6]]$pars[c(3:5)],
+#                           RS_0 = 0.04012019)
 #                         )
-# two_pool_all_results_good<-rbind(two_pool_all1$goodness,
-#                                 two_pool_all_fixYukfVmax2$goodness,
-#                                 two_pool_all3_fix$goodness,
-#                                 two_pool_all_fixYukfVmax4$goodness)
+# two_pool_all_results_good<-rbind(two_pool_allx_out1$goodness,
+#                                two_pool_all_fixVmaxYuk_out2$goodness,
+#                                two_pool_allx_fix_out3$goodness,
+#                                two_pool_all_fixVmaxYuk_out4$goodness)
 # 
@@
                    Km=two_pool_all_results_par[,2],
-                   m0=two_pool_all_results_par[,3],
-                   f=two_pool_all_results_par[,4],
-                   k=two_pool_all_results_par[,5],
-                   Yu=two_pool_all_results_par[,6],
-                   fpr=two_pool_all_results_par[,7],
-                   fps=two_pool_all_results_par[,8],
-                   fds=two_pool_all_results_par[,9],
-                   RS_0=two_pool_all_results_par[,10])
+                   f=two_pool_all_results_par[,3],
+                   k=two_pool_all_results_par[,4],
+                   Yu=two_pool_all_results_par[,5],
+                   fpr=two_pool_all_results_par[,6],
+                   fps=two_pool_all_results_par[,7],
+                   fds=two_pool_all_results_par[,8],
+                   RS_0=two_pool_all_results_par[,9])
 
@@
 
-kable(t.s6.1, digits = c(0, 2, 2, 3, 3, 3, 2, 3, 3, 3, 2), align = 'c', "html", booktabs=T, col.names = c(" ","v", 'k', 'h', 'f', 'm', 'Y', '$\\B_{R-Proteins}$', '$\\B_{S-Proteins}$', '$\\B_{S-DNA}$', '$\\RS_{0}$'), row.names=F, caption = "_**Table S21:** Sub-microbial model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s6.1, digits = c(0, 2, 2, 2, 3, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ","v", 'k', 'f', 'm', 'Y', '$\\B_{R-Proteins}$', '$\\B_{S-Proteins}$', '$\\B_{S-DNA}$', '$\\RS_{0}$'), row.names=F, caption = "_**Table S21:** Sub-microbial model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 ```
-
-$~$
-$~$
-$~$
-$~$
-$~$
-$~$
-```{r correlation figure, echo=FALSE, fig.height=12, fig.width=16, fig.cap="Fig.
-S3: Relative abundance of bacterial (a) and fungal (b) species in three experimental treatments differing in presence/absence of physical structure.
-Species with relative abundance above 0.5% are plotted.
-Their abundance is visualized by colour scale.
-Effect of bacterial and fungal composition of microbial community (expressed as multivariate mean distance between groups; see Material and Methods section) on overall microbial community affinity to substrate (i.e.
-parameter k - (a)) and abundance of proteins in microbial biomass (i.e.
-sum of protein abundance in biomass reserve and struscture - (b)).
-The values on y-axes are expressed as a mean difference between groups in order to allow unbiased correlation with microbial community structure metric (see Matrial and Methods section)."}
-#Heatmap
-bac.tax<-as.data.frame(read.csv(file=c("./Community_analysis/OTU_bacteria.txt"), header=T,
-                             sep="\t"))
-bac.tax<-bac.tax[, c(1, 175)]
-bac.tax$X.OTU.ID<-gsub("_","", bac.tax$X.OTU.ID)
-
-#Add taxonomy to bac.norm dataset
-bac.norm_tax<-bac
-#Transform the dataset
-bac.norm_taxt<-as.data.frame(t(bac.norm_tax))
-taxlab<-character()
-for(i in 1:nrow(bac.norm_taxt)){
-  nd<-which(bac.tax$X.OTU.ID==rownames(bac.norm_taxt)[i])
-  taxlab<-append(taxlab, as.character(bac.tax$taxonomy[nd]))
-}
-
-bac.norm_taxt$taxonomy<-taxlab
-
-#Extract kingdom
-pattern<-"k__(.+?),"
-regmatches(bac.norm_taxt$taxonomy[1], regexec(pattern, bac.norm_taxt$taxonomy[1]))[[1]][2]
-kingdoms<-character()
-for(i in 1:nrow(bac.norm_taxt)){
-  kingdoms<-append(kingdoms, 
-                   regmatches(bac.norm_taxt$taxonomy[i], regexec(pattern, bac.norm_taxt$taxonomy[i]))[[1]][2])
-}
-kingdoms[which(kingdoms=="?")]<-c("Unknown")
-bac.norm_taxt$Kingdom<-kingdoms
-
-#Extract phylum
-pattern<-"p__(.+?),"
-
-phylum<-character()
-for(i in 1:nrow(bac.norm_taxt)){
-  phylum<-append(phylum, 
-                   regmatches(bac.norm_taxt$taxonomy[i], regexec(pattern, bac.norm_taxt$taxonomy[i]))[[1]][2])
-}
-phylum[which(phylum=="?")]<-c("Unknown")
-bac.norm_taxt$Phylum<-phylum
-
-#Extract class
-pattern<-"c__(.+?),"
-
-cs<-character()
-for(i in 1:nrow(bac.norm_taxt)){
-  cs<-append(cs, 
-                 regmatches(bac.norm_taxt$taxonomy[i], regexec(pattern, bac.norm_taxt$taxonomy[i]))[[1]][2])
-}
-cs[which(cs=="?")]<-c("Unknown")
-bac.norm_taxt$Class<-cs
-
-#Extract order
-pattern<-"o__(.+?),"
-
-os<-character()
-for(i in 1:nrow(bac.norm_taxt)){
-  os<-append(os, 
-             regmatches(bac.norm_taxt$taxonomy[i], regexec(pattern, bac.norm_taxt$taxonomy[i]))[[1]][2])
-}
-os[which(os=="?")]<-c("Unknown")
-bac.norm_taxt$Order<-os
-
-#Extract family
-pattern<-"f__(.+?),"
-
-fs<-character()
-for(i in 1:nrow(bac.norm_taxt)){
-  fs<-append(fs, 
-             regmatches(bac.norm_taxt$taxonomy[i], regexec(pattern, bac.norm_taxt$taxonomy[i]))[[1]][2])
-}
-
-fs[which(fs=="?")]<-c("Unknown")
-bac.norm_taxt$Family<-fs
-
-#Extract genus
-pattern<-"g__(.+?),"
-
-gs<-character()
-for(i in 1:nrow(bac.norm_taxt)){
-  gs<-append(gs, 
-             regmatches(bac.norm_taxt$taxonomy[i], regexec(pattern, bac.norm_taxt$taxonomy[i]))[[1]][2])
-}
-gs[which(gs=="?")]<-c("Unknown")
-bac.norm_taxt$Genus<-gs
-
-#Extract species
-pattern<-"s__(.+?),"
-
-ss<-character()
-for(i in 1:nrow(bac.norm_taxt)){
-  ss<-append(ss, 
-             regmatches(bac.norm_taxt$taxonomy[i], regexec(pattern, bac.norm_taxt$taxonomy[i]))[[1]][2])
-}
-ss[which(is.na(ss))]<-c("Unknown")
-bac.norm_taxt$Specie<-ss
-
-#Plots
-tax_env<-bac_env
-btaxt<-bac.norm_taxt[, 1:71]
-tax_env$IDs<-colnames(btaxt)
-
-##Genus
-btaxtg<-btaxt
-btaxtg$Genus<-bac.norm_taxt$Genus
-
-Btaxtg<-melt(btaxtg, id.vars = c("Genus"))
-colnames(Btaxtg)<-c("Genus", "IDs","Abundance")
-
-Btaxtg<-merge(Btaxtg, tax_env, by.x = "IDs", by.y = "IDs", all.x = TRUE)
-Btaxtg$Total<-numeric(length = nrow(Btaxtg))
-for(i in unique(Btaxtg$IDs)){
-  Btaxtg[Btaxtg$IDs==i, "Total"]<-sum(as.numeric(Btaxtg[Btaxtg$IDs==i, "Abundance"]))
-}
-
-bac_gen_f<-Btaxtg %>% group_by(Genus, Structure, Substrate) %>% 
-  summarize(Abundance=mean(Abundance/Total, na.rm=T))
-
-#Only genus with abundance above 0.05%
-heat1<-ggplot(bac_gen_f[bac_gen_f$Abundance>0.0005, ], aes(Structure, Genus)) +
-  geom_tile(aes(fill = log10(Abundance*100))) +
-  facet_grid(~ Substrate) +
-  scale_fill_gradientn(colours = c("white","darkorange2","grey60"),
-                       breaks = c(-2, -1, 0, 1, 2),
-                       labels=c("0.01", "0.1", "0", "10", "100"),
-                       name = "Relative abundance (%)",
-                       guide = guide_colourbar(direction = "horizontal", title.position = "bottom",
-                                               title.hjust = 0.5, barwidth = 12)) +
-  scale_x_discrete(expand = c(0,0)) + 
-  scale_y_discrete(expand = c(0,0)) +  
-  xlab("Structure") + ylab("")+
-  ggtitle("a)")+
-  theme_min + theme(legend.position = "bottom",
-                    axis.text = element_text(size=14),
-                    strip.text = element_text(size=14))
-
-##Fungi
-f.tax<-as.data.frame(read.csv(file=c("./Community_analysis/OTU_fungi.txt"), header=T,
-                             sep="\t"))
-#Add taxonomy to fungi dataset
-f.norm_tax<-fungi
-#Transform the dataset
-f.norm_taxt<-as.data.frame(t(f.norm_tax))
-ftaxlab<-character()
-for(i in 1:nrow(f.norm_taxt)){
-  nd<-which(f.tax$X.OTU.ID==rownames(f.norm_taxt)[i])
-  ftaxlab<-append(ftaxlab, as.character(f.tax$taxonomy[nd]))
-}
-
-f.norm_taxt$taxonomy<-ftaxlab
-
-#Extract kingdom
-pattern<-"k__(.+?),"
-regmatches(f.norm_taxt$taxonomy[1], regexec(pattern, f.norm_taxt$taxonomy[1]))[[1]][2]
-kingdoms<-character()
-for(i in 1:nrow(f.norm_taxt)){
-  kingdoms<-append(kingdoms, 
-                   regmatches(f.norm_taxt$taxonomy[i], regexec(pattern, f.norm_taxt$taxonomy[i]))[[1]][2])
-}
-kingdoms[which(kingdoms=="?")]<-c("Unknown")
-f.norm_taxt$Kingdom<-kingdoms
-
-#Extract phylum
-pattern<-"p__(.+?),"
-
-phylum<-character()
-for(i in 1:nrow(f.norm_taxt)){
-  phylum<-append(phylum, 
-                 regmatches(f.norm_taxt$taxonomy[i], regexec(pattern, f.norm_taxt$taxonomy[i]))[[1]][2])
-}
-phylum[which(phylum=="?")]<-c("Unknown")
-f.norm_taxt$Phylum<-phylum
-
-#Extract class
-pattern<-"c__(.+?),"
-
-cs<-character()
-for(i in 1:nrow(f.norm_taxt)){
-  cs<-append(cs, 
-             regmatches(f.norm_taxt$taxonomy[i], regexec(pattern, f.norm_taxt$taxonomy[i]))[[1]][2])
-}
-cs[which(cs=="?")]<-c("Unknown")
-f.norm_taxt$Class<-cs
-
-#Extract order
-pattern<-"o__(.+?),"
-
-os<-character()
-for(i in 1:nrow(f.norm_taxt)){
-  os<-append(os, 
-             regmatches(f.norm_taxt$taxonomy[i], regexec(pattern, f.norm_taxt$taxonomy[i]))[[1]][2])
-}
-os[which(os=="?")]<-c("Unknown")
-f.norm_taxt$Order<-os
-
-#Extract family
-pattern<-"f__(.+?),"
-
-fs<-character()
-for(i in 1:nrow(f.norm_taxt)){
-  fs<-append(fs, 
-             regmatches(f.norm_taxt$taxonomy[i], regexec(pattern, f.norm_taxt$taxonomy[i]))[[1]][2])
-}
-
-fs[which(fs=="?")]<-c("Unknown")
-f.norm_taxt$Family<-fs
-
-#Extract genus
-pattern<-"g__(.+?),"
-
-gs<-character()
-for(i in 1:nrow(f.norm_taxt)){
-  gs<-append(gs, 
-             regmatches(f.norm_taxt$taxonomy[i], regexec(pattern, f.norm_taxt$taxonomy[i]))[[1]][2])
-}
-gs[which(gs=="?")]<-c("Unknown")
-f.norm_taxt$Genus<-gs
-
-#Extract species
-pattern<-"s__(.+?),"
-
-ss<-character()
-for(i in 1:nrow(f.norm_taxt)){
-  ss<-append(ss, 
-             regmatches(f.norm_taxt$taxonomy[i], regexec(pattern, f.norm_taxt$taxonomy[i]))[[1]][2])
-}
-ss[which(is.na(ss))]<-c("Unknown")
-f.norm_taxt$Specie<-ss
-
-#Plots
-ftax_env<-fungi_env
-ftaxt<-f.norm_taxt[, 1:64]
-ftax_env$IDs<-colnames(ftaxt)
-
-##Genus
-ftaxtg<-ftaxt
-ftaxtg$Genus<-f.norm_taxt$Genus
-
-Ftaxtg<-melt(ftaxtg, id.vars = c("Genus"))
-colnames(Ftaxtg)<-c("Genus", "IDs","Abundance")
-
-Ftaxtg<-merge(Ftaxtg, ftax_env, by.x = "IDs", by.y = "IDs", all.x = TRUE)
-Ftaxtg$Total<-numeric(length = nrow(Ftaxtg))
-for(i in unique(Ftaxtg$IDs)){
-  Ftaxtg[Ftaxtg$IDs==i, "Total"]<-sum(as.numeric(Ftaxtg[Ftaxtg$IDs==i, "Abundance"]))
-}
-
-f_gen_f<-Ftaxtg %>% group_by(Genus, Structure, Substrate) %>% 
-  summarize(Abundance=mean(Abundance/Total, na.rm=T))
-
-#Genus with abundance above 0.05%
-heat2<-ggplot(f_gen_f[f_gen_f$Abundance>0.0005, ], aes(Structure, Genus)) +
-  geom_tile(aes(fill = log(Abundance*100))) +
-  facet_grid(~ Substrate) +
-  scale_fill_gradientn(colours = c("white","darkorange2","grey60"),
-                       breaks = c(-2, -1, 0, 1, 2),
-                       labels=c("0.01", "0.1", "0", "10", "100"),
-                       name = "Relative abundance (%)",
-                       guide = guide_colourbar(direction = "horizontal", title.position = "bottom",
-                                               title.hjust = 0.5, barwidth = 12)) +
-  scale_x_discrete(expand = c(0,0)) + 
-  scale_y_discrete(expand = c(0,0)) +  
-  xlab("Structure") + ylab("")+
-  ggtitle("b)")+
-  theme_min + theme(legend.position = "bottom",
-                    axis.text = element_text(size=14),
-                    strip.text = element_text(size=14))
-##########################################################################################
-dpars<-t.s6.1[c(2:4), -1]
-dpars$Structure<-c("BROTH", "GLASS", "WOOL")
-
-#Bacteria
-bacm<-as.matrix(bac.dist)
-#Intercept - BROTH
-mBB<-bacm[which(bac_env$Structure=="BROTH"), 
-       which(bac_env$Structure=="BROTH")]
-mBB[mBB==0]<-NA
-#BROTH - GLASS
-mBG<-bacm[which(bac_env$Structure=="BROTH"), 
-       which(bac_env$Structure=="GLASS")]
-
-#BRTOH - WOOL
-mBW<-bacm[which(bac_env$Structure=="BROTH"), 
-       which(bac_env$Structure=="WOOL")]
-
-bacBGW<-data.frame(Dist=c(0, mean(mBG, na.rm=T), mean(mBW, na.rm=T)),
-                    Dist.sd=c(sd(mBB, na.rm=T), sd(mBG, na.rm=T), sd(mBW, na.rm=T)),
-                    Structure=c("BROTH", "GLASS", "WOOL"))
-
-#Fungi
-#Intercept - BROTH
-fBB<-fm[which(fungi_env$Structure=="BROTH"), 
-       which(fungi_env$Structure=="BROTH")]
-fBB[fBB==0]<-NA
-#BROTH - GLASS
-fBG<-fm[which(fungi_env$Structure=="BROTH"), 
-       which(fungi_env$Structure=="GLASS")]
-
-#BRTOH - WOOL
-fBW<-fm[which(fungi_env$Structure=="BROTH"), 
-       which(fungi_env$Structure=="WOOL")]
-
-fBGW<-data.frame(Dist=c(0, mean(fBG, na.rm=T), mean(fBW, na.rm=T)),
-                    Dist.sd=c(sd(fBB, na.rm=T), sd(fBG, na.rm=T), sd(fBW, na.rm=T)),
-                    Structure=c("BROTH", "GLASS", "WOOL"))
-
-dpars2<-merge(dpars, bacBGW, by.y=c("Structure"), by.x=c("Structure"))
-
-dpars3<-merge(dpars, fBGW, by.y=c("Structure"), by.x=c("Structure"))
-
-for(i in 2:11){
-  dpars2[, i]<-sqrt((dpars2[, i]-as.numeric(dpars2[1, i]))^2)
-}
-
-for(i in 2:11){
-  dpars3[, i]<-sqrt((dpars3[, i]-as.numeric(dpars3[1, i]))^2)
-}
-
-dpars2$group<-c("Bacteria")
-dpars3$group<-c("Fungi")
-dpars4<-rbind(dpars2, dpars3)
-
-corplot1.p<-coefficients(nls(Km~exp(Dist*x)-a, dpars4, start = list(x=10, a=2)))
-
-corplot1<-ggplot(dpars4[dpars4$group=="Bacteria", ], aes(Dist, Km))+
-  geom_point(cex=6, aes(fill=Structure), show.legend=T, shape=21)+
-  geom_point(data=dpars4[dpars4$group=="Fungi", ], cex=6, 
-             aes(x=Dist, y=Km, fill=Structure), show.legend=F, shape=22)+
-  scale_fill_manual(values = c("black", "grey", "white"))+
-  stat_function(fun = function(x){(exp(x*corplot1.p[1])-corplot1.p[2])}, lwd=1, lty=2, colour="grey30")+
-  xlab("16S and ITS Distance")+
-  ggtitle("c)")+
-  ylab("Affinity to substrate (relative difference)")+
-  theme_min+theme(legend.title = element_blank(),
-                  legend.position = c(0.2,0.7),
-                  plot.margin = unit(c(0.5,1,0.5,2), "in"))
-
-corplot2<-ggplot(dpars4[dpars4$group=="Bacteria", ], aes(Dist, fps+fpr))+
-  geom_point(cex=6, aes(fill=Structure), show.legend=F, shape=21)+
-  geom_point(data=dpars4[dpars4$group=="Fungi", ], cex=6, 
-             aes(x=Dist, y=fps+fpr, fill=Structure), show.legend=F, shape=22)+
-  stat_smooth(method = lm, colour="grey30", lty=2, se=F)+
-  scale_fill_manual(values = c("black", "grey", "white"))+
-  ggtitle("d)")+
-  theme_min+xlab("16S and ITS Distance")+
-  ylab("Proteins abundance (relative difference)")+
-  theme(plot.margin = unit(c(0.5,1,0.5,2), "in"))
-
-grid.arrange(heat1, heat2, corplot1, corplot2, nrow=2)
-
-```
-
-##Supplementary figure
-```{r Monod and Monod plus model parameters, echo=FALSE, fig.height=24, fig.width=20, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Fig.
-S1: Monod (a) and Monod plus (b) model parameters (v - maximum velocity constant, k - affinity constant, $Y_{CUE}$ - carbon use efficiency, m - microbial biomass decay rate constant, h - specific maintenance rate constant) estimated separately for three different types of physical barriers presented in incubation bottle (BROTH - black symbols, GLASS - grey symbols and WOOL - empty symbols).
-Symbols represent best parameters values estimated by the Differential Evolution Algorithm and error bars represent standard deviation of posterior parameters distribution estimated by Constrained Markov Chain Monte Carlo simulation.
-Parameters estimation was constrained by either observed respiration rates, observed respiration rates and cellular proteins concentration, observed respiration rates and DNA concentration or by all measurements at the same time."}
-
-# # Monod model calibrated against the respiration rate (not run)
-# #uncertainty of model parameters
-# monod_r_sd<-as.data.frame(rbind(apply(monod_r2_fix[[1]]$par_prof$pars, 2, sd),
-#                                apply(monod_r2_fix[[2]]$par_prof$pars, 2, sd),
-#                                apply(monod_r2_fix[[3]]$par_prof$pars, 2, sd)))[,-5]
-# colnames(monod_r_sd)<-c("Vmax.sd", "Km.sd", "CUE.sd", "k.sd")
-# 
-# #Combine with the estimated parameters
-# monod_rs1<-cbind(monod_r_results_par[c(2:4), c(1:4)], monod_r_sd)
-# monod_rs1$Legend<-rep("Respiration", times=3)
-# 
-# #Monod model calibrated against the respiration rate and the rest (not run)
-# # uncertainty of model parameters
-# monod_rest_sd<-as.data.frame(rbind(apply(monod_prot2_fix[[1]]$par_prof$pars, 2, sd)[c(1:4)],
-#                                  apply(monod_prot2_fix[[2]]$par_prof$pars, 2, sd)[c(1:4)],
-#                                  apply(monod_prot2_fix[[3]]$par_prof$pars, 2, sd)[c(1:4)],
-#                                  apply(monod_DNA2[[1]]$par_prof$pars, 2, sd)[c(1:4)],
-#                                  apply(monod_DNA2[[2]]$par_prof$pars, 2, sd)[c(1:4)],
-#                                  apply(monod_DNA2[[3]]$par_prof$pars, 2, sd)[c(1:4)],
-#                                  apply(monod_all2[[1]]$par_prof$pars, 2, sd)[c(1:4)],
-#                                  apply(monod_all2[[2]]$par_prof$pars, 2, sd)[c(1:4)],
-#                                  apply(monod_all2[[3]]$par_prof$pars, 2, sd)[c(1:4)]))[,-c(5,6)]
-# colnames(monod_rest_sd)<-c("Vmax.sd", "Km.sd", "CUE.sd", "k.sd")
-# #Combine with the estimated parameters
-# monod_alls<-cbind(rbind(monod_prot_results_par[c(2:4), c(1:4)],
-#                         monod_DNA_results_par[c(2:4), c(1:4)],
-#                         monod_all_results_par[c(2:4), c(1:4)]), monod_rest_sd)
-# 
-# 
-# monod_alls$Legend<-c(rep("Respiration & Proteins", times=3),
-#                     rep("Respiration & DNA", times=3),
-#                     rep("All", times=3))
-#  
-# #Combine monod_alls and monod_rs1
-# monod_s<-rbind(monod_rs1, monod_alls)
-# #Add labels
-# monod_s$Structure<-rep(c("BROTH", "GLASS", "WOOL"), times=4)
-# 
-# # Monod plus model calibrated against the respiration rate (not run)
-# # uncertainty of model parameters
-# monodplus_r_sd<-as.data.frame(rbind(apply(monodplus_r2_fix[[1]]$par_prof$pars, 2, sd),
-#                                apply(monodplus_r2_fix[[2]]$par_prof$pars, 2, sd),
-#                                apply(monodplus_r2_fix[[3]]$par_prof$pars, 2, sd)))
-# colnames(monodplus_r_sd)<-c("Vmax.sd", "Km.sd", "CUE.sd", "mr.sd", "k.sd")
-# 
-# #Combine with the estimated parameters
-# monodplus_rs1<-cbind(monodplus_r_results_par[c(2:4), c(1:5)], monodplus_r_sd)
-# monodplus_rs1$Legend<-rep("Respiration", times=3)
-# 
-# #Monod model calibrated against the respiration rate and the rest (not run)
-# # uncertainty of model parameters
-# monodplus_rest_sd<-as.data.frame(rbind(apply(monodplus_prot2_fix[[1]]$par_prof$pars, 2, sd)[c(1:5)],
-#                                  apply(monodplus_prot2_fix[[2]]$par_prof$pars, 2, sd)[c(1:5)],
-#                                  apply(monodplus_prot2_fix[[3]]$par_prof$pars, 2, sd)[c(1:5)],
-#                                  apply(monodplus_DNA2[[1]]$par_prof$pars, 2, sd)[c(1:5)],
-#                                  apply(monodplus_DNA2[[2]]$par_prof$pars, 2, sd)[c(1:5)],
-#                                  apply(monodplus_DNA2[[3]]$par_prof$pars, 2, sd)[c(1:5)],
-#                                  apply(monodplus_all2[[1]]$par_prof$pars, 2, sd)[c(1:5)],
-#                                  apply(monodplus_all2[[2]]$par_prof$pars, 2, sd)[c(1:5)],
-#                                  apply(monodplus_all2[[3]]$par_prof$pars, 2, sd)[c(1:5)]))
-# colnames(monodplus_rest_sd)<-c("Vmax.sd", "Km.sd", "CUE.sd", "mr.sd", "k.sd")
-# #Combine with the estimated parameters
-# monodplus_alls<-cbind(rbind(monodplus_prot_results_par[c(2:4), c(1:5)],
-#                         monodplus_DNA_results_par[c(2:4), c(1:5)],
-#                         monodplus_all_results_par[c(2:4), c(1:5)]), monodplus_rest_sd)
-# 
-# 
-# monodplus_alls$Legend<-c(rep("Respiration & Proteins", times=3),
-#                     rep("Respiration & DNA", times=3),
-#                     rep("All", times=3))
-#  
-# #Combine monod_alls and monod_rs1
-# monodplus_s<-rbind(monodplus_rs1, monodplus_alls)
-# #Add labels
-# monodplus_s$Structure<-rep(c("BROTH", "GLASS", "WOOL"), times=4)
-# 
-# write.csv(monod_s, file = c("../monod_FigS1.csv"), row.names = F)
-# write.csv(monodplus_s, file = c("../monodplus_FigS1.csv"), row.names = F)
-
-monod_s<-read.csv("monod_FigS1.csv")
-monodplus_s<-read.csv("monodplus_FigS1.csv")
-
-Monod_s<-melt(monod_s[, c(1:4, 9, 10)], id.vars=c("Legend", "Structure"))
-Monod_s$sd<-melt(monod_s[, c(5:8, 9, 10)], id.vars=c("Legend", "Structure"))[, "value"]
-
-#add proper parameter labels
-mlab<-c('Vmax' = 'v',
-        'Km' = 'k',
-        'CUE' = 'Y[CUE]',
-        'k' = 'm')
-
-figS1a<-ggplot(Monod_s, aes(Legend, value))+
-  geom_point(shape=21, cex=6, aes(fill=Structure), position = position_dodge(width = 0.6))+
-  geom_errorbar(aes(ymin=value-sd, ymax=value+sd, colour=Structure), 
-                position = position_dodge(width = 0.6))+
-  scale_fill_manual(values = c("black", "grey", "white"))+
-  scale_colour_manual(values = c("black", "black", "black"))+theme_min+
-  theme(axis.title.x = element_blank(),
-        legend.title = element_blank(),
-        axis.text.x=element_text(vjust=0.2, size=18, colour="black", angle = 90),
-        legend.position = c("top"),
-        plot.title=element_text(size=18, face="bold", hjust=-0.01))+
-  facet_wrap(~variable, scale="free", nrow=1, 
-             labeller = labeller(variable = as_labeller(mlab, label_parsed)))+
-  ylab("Parameter value")+ggtitle("a)")
-
-Monodplus_s<-melt(monodplus_s[, c(1:5, 11, 12)], id.vars=c("Legend", "Structure"))
-Monodplus_s$sd<-melt(monodplus_s[, c(6:10, 11, 12)], id.vars=c("Legend", "Structure"))[, "value"]
-
-#add proper parameter labels
-mlab2<-c('Vmax' = 'v',
-        'Km' = 'k',
-        'CUE' = 'Y[CUE]',
-        'mr' = 'h',
-        'k' = 'm')
-
-figS1b<-ggplot(Monodplus_s, aes(Legend, value))+
-  geom_point(shape=21, cex=6, aes(fill=Structure), position = position_dodge(width = 0.6),
-             show.legend = F)+
-  geom_errorbar(aes(ymin=value-sd, ymax=value+sd, colour=Structure), 
-                position = position_dodge(width = 0.6), show.legend = F)+
-  scale_fill_manual(values = c("black", "grey", "white"))+
-  scale_colour_manual(values = c("black", "black", "black"))+theme_min+
-  theme(axis.title.x = element_blank(),
-        legend.title = element_blank(),
-        axis.text.x=element_text(vjust=0.2, size=18, colour="black", angle = 90),
-        plot.title=element_text(size=18, face="bold", hjust=-0.01))+
-  facet_wrap(~variable, scale="free", nrow=1, 
-             labeller = labeller(variable = as_labeller(mlab2, label_parsed)))+
-  ylab("Parameter value")+ggtitle("b)")
-
-grid.arrange(figS1a, figS1b, nrow=2)
-
-```

### 2020-03-30 Rscripts/monod_test.R

@@ -1,257 +1,0 @@
-monod_test<-function(data, FACT){
-  
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-  
-  if(FACT==1){
-    
-    dat<-data
-    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
-    
-  }else{
-    
-    if(FACT==2){
-      
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-      
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-      
-    }else{
-      
-      if(FACT==3){
-        
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-        
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-        
-        
-      }else{
-        
-        dat<-data
-        dat$id<-c(1)
-        
-      }
-    }
-  }
-  
-  
-  #monod growth function
-  deriv<-function(time, state, pars){
-    
-    with(as.list(c(state, pars)),{
-      
-      dCmic<--k*Cmic+CUE*Vmax*Cmic*C/(Km+C)
-      dC<-k*Cmic-Vmax*Cmic*C/(Km+C)
-      
-      return(list(c(dCmic, dC), r=(1-CUE)*Vmax*Cmic*C/(Km+C)))
-      
-    })
-  }
-  
-  #define names of parameters
-  parnames<-c("Vmax", "Km", "CUE", "k")
-  
-  #parameters estimation function
-  estim<-function(odeset){
-    
-    #defining cost function
-    cost<-function(x){
-      
-      par<-x[1:length(parnames)]
-      
-      names(par)<-parnames
-      
-      Cmic0<-numeric()
-      if(odeset$Structure[1]=="Broth"){
-        Cmic0<-0.1287723+2.08193
-      }else{
-        if(odeset$Structure[1]=="Mixed glass"){
-          Cmic0<-0.2931153+4.73895
-        }else{
-          Cmic0<-0.221984+3.588933
-        }
-      }
-      
-      #first, pars dependent output from ode is matched with measured values
[…]

### 2020-03-30 Rscripts/monodplus_DNA.R

@@ -1,244 +1,0 @@
-monodplus_DNA<-function(data, FACT){
-  
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-  
-  if(FACT==1){
-    
-    dat<-data
-    dat$id<-ifelse(d$Substrate=="Glucose", 1, 2)
-    
-  }else{
-    
-    if(FACT==2){
-      
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-      
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-      
-    }else{
-      
-      if(FACT==3){
-        
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-        
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-        
-        
-      }else{
-        
-        dat<-data
-        dat$id<-c(1)
-        
-      }
-    }
-  }
-  
-  #mend model
-  deriv<-function(time, state, pars){
-    
-    with(as.list(c(state, pars)),{
-      
-      #equations
-      #C uptake
-      Cu=Vmax*Cmic*C/(Km+C)
-      #respiration
-      r=mr*Cmic+(1-CUE)*Cu
-      
-      #Protc=Cmic*fp
-      DNAc=Cmic*fd
-      
-      #states
-      dCmic<-CUE*Cu-mr*Cmic - k*Cmic
-      dC<--Cu + k*Cmic
-      
-      return(list(c(dCmic, dC), r=r, DNAc=DNAc))
-      
-    })
-  }
-  #define names of parameters
-  parnames<-c("Vmax", "Km", "CUE", "mr", "k", "fd")
-  
-  #parameters estimation function
-  estim<-function(odeset){
-    
-    #defining cost function
-    cost<-function(x){
-      
-      par<-x[1:length(parnames)]
-      
-      names(par)<-parnames
-      
-      #first, pars dependent output from ode is matched with measured values
-      yhat_all<-as.data.frame(ode(y=c(Cmic=2.01*0.51/12.01/4/par[["fd"]], C=25), parms=par, deriv, times=sort(odeset$Time)))
-      
-      #select time and the measured variables 
[…]

### 2020-03-30 Rscripts/monodplus_r.R

@@ -1,236 +1,0 @@
-monodplus_r<-function(data, FACT){
-  
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-  
-  if(FACT==1){
-    
-    dat<-data
-    dat$id<-ifelse(d$Substrate=="Glucose", 1, 2)
-    
-  }else{
-    
-    if(FACT==2){
-      
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-      
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-      
-    }else{
-      
-      if(FACT==3){
-        
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-        
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-        
-        
-      }else{
-        
-        dat<-data
-        dat$id<-c(1)
-        
-      }
-    }
-  }
-  
-  #mend model
-  deriv<-function(time, state, pars){
-    
-    with(as.list(c(state, pars)),{
-      
-      #equations
-      #C uptake
-      Cu=Vmax*Cmic*C/(Km+C)
-      #respiration
-      r=mr*Cmic+(1-CUE)*Cu
-      
-      #Protc=Cmic*fp
-      #DNAc=Cmic*fd
-      
-      #states
-      dCmic<-CUE*Cu-mr*Cmic-k*Cmic
-      dC<--Cu + k*Cmic
-      
-      return(list(c(dCmic, dC), r=r))
-      
-    })
-  }
-  #define names of parameters
-  parnames<-c("Vmax", "Km", "CUE", "mr", "k", "Cmic_0")
-  
-  #parameters estimation function
-  estim<-function(odeset){
-    
-    #defining cost function
-    cost<-function(x){
-      
-      par<-x[1:length(parnames)]
-      
-      names(par)<-parnames
-      
-      #first, pars dependent output from ode is matched with measured values
-      yhat_all<-as.data.frame(ode(y=c(Cmic=par[["Cmic_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))
-      
-      #select time and the measured variables 
[…]

### 2020-03-30 Rscripts/monodplus_r_fix.R

@@ -1,236 +1,0 @@
-monodplus_r_fix<-function(data, FACT){
-  
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-  
-  if(FACT==1){
-    
-    dat<-data
-    dat$id<-ifelse(d$Substrate=="Glucose", 1, 2)
-    
-  }else{
-    
-    if(FACT==2){
-      
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-      
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-      
-    }else{
-      
-      if(FACT==3){
-        
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-        
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-        
-        
-      }else{
-        
-        dat<-data
-        dat$id<-c(1)
-        
-      }
-    }
-  }
-  
-  #mend model
-  deriv<-function(time, state, pars){
-    
-    with(as.list(c(state, pars)),{
-      
-      #equations
-      #C uptake
-      Cu=Vmax*Cmic*C/(Km+C)
-      #respiration
-      r=mr*Cmic+(1-CUE)*Cu
-      
-      #Protc=Cmic*fp
-      #DNAc=Cmic*fd
-      
-      #states
-      dCmic<-CUE*Cu-mr*Cmic-k*Cmic
-      dC<--Cu + k*Cmic
-      
-      return(list(c(dCmic, dC), r=r))
-      
-    })
-  }
-  #define names of parameters
-  parnames<-c("Vmax", "Km", "CUE", "mr", "k")
-  
-  #parameters estimation function
-  estim<-function(odeset){
-    
-    #defining cost function
-    cost<-function(x){
-      
-      par<-x[1:length(parnames)]
-      
-      names(par)<-parnames
-      
-      #first, pars dependent output from ode is matched with measured values
-      yhat_all<-as.data.frame(ode(y=c(Cmic=0.09125118, C=25), parms=par, deriv, times=sort(odeset$Time)))
-      
-      #select time and the measured variables 
-      yhat<-select(yhat_all, c("time", "r"))
[…]

### 2020-03-30 Rscripts/monoplus_all.R

@@ -1,244 +1,0 @@
-monodplus_all<-function(data, FACT){
-  
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-  
-  if(FACT==1){
-    
-    dat<-data
-    dat$id<-ifelse(d$Substrate=="Glucose", 1, 2)
-    
-  }else{
-    
-    if(FACT==2){
-      
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-      
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-      
-    }else{
-      
-      if(FACT==3){
-        
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-        
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-        
-        
-      }else{
-        
-        dat<-data
-        dat$id<-c(1)
-        
-      }
-    }
-  }
-  
-  #mend model
-  deriv<-function(time, state, pars){
-    
-    with(as.list(c(state, pars)),{
-      
-      #equations
-      #C uptake
-      Cu=Vmax*Cmic*C/(Km+C)
-      #respiration
-      r=mr*Cmic+(1-CUE)*Cu
-      
-      Protc=Cmic*fp
-      DNAc=Cmic*fd
-      
-      #states
-      dCmic<-CUE*Cu-mr*Cmic - k*Cmic
-      dC<--Cu + k*Cmic
-      
-      return(list(c(dCmic, dC), r=r, DNAc=DNAc, Protc=Protc))
-      
-    })
-  }
-  #define names of parameters
-  parnames<-c("Vmax", "Km", "CUE", "mr", "k", "fd", "fp")
-  
-  #parameters estimation function
-  estim<-function(odeset){
-    
-    #defining cost function
-    cost<-function(x){
-      
-      par<-x[1:length(parnames)]
-      
-      names(par)<-parnames
-      
-      #first, pars dependent output from ode is matched with measured values
-      yhat_all<-as.data.frame(ode(y=c(Cmic=2.01*0.51/12.01/4/par[["fd"]], C=25), parms=par, deriv, times=sort(odeset$Time)))
-      
[…]

### 2020-03-30 Rscripts/two_pool_all.R

@@ -1,252 +1,0 @@
-two_pool_all<-function(data, FACT){
-  
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-  
-  
-  #if factor needs to be included, its levels are defined
-  if(FACT==1){
-    
-    dat<-data
-    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
-    
-  }else{
-    
-    if(FACT==2){
-      
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-      
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-      
-    }else{
-      
-      if(FACT==3){
-        
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-        
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-        
-        
-      }else{
-        
-        dat<-data
-        dat$id<-c(1)
-        
-      }
-    }
-  }
-  
-  #deb model
-  deriv<-function(time, state, pars){
-    
-    with(as.list(c(state, pars)),{
-      
-      #Carbon uptake
-      Cu=Vmax*C*S/(Km+C)
-      #Respiration
-      r=m0*S+f*R*(1-Yu)
-      
-      Protc=fpr*R+fps*S
-      DNAc=fds*S
-      
-      dR<-Cu-f*R
-      dS<-f*R*Yu - m0*S - k*S
-      dC<--Cu + k*S
-      
-      
-      return(list(c(dR, dS, dC), r=r, Protc=Protc, DNAc=DNAc))
-      
-    })
-  }
-  #define names of parameters
-  parnames<-c("Vmax", "Km", "m0", "f", "k", "Yu", "fpr", "fps", "fds", "RS_0")
-  
-  #parameters estimation function
-  estim<-function(odeset){
-    
-    #defining cost function
-    cost<-function(x){
-      
-      par<-x[1:length(parnames)]
-      
-      names(par)<-parnames
-      
-      Sinit=2.01*0.51/12.01/4/par[["fds"]]
-      Rinit=par[["RS_0"]]*Sinit
-      
-      #first, pars dependent output from ode is matched with measured values
[…]

### 2020-03-30 Rscripts/two_pool_all_fix.R

@@ -1,252 +1,0 @@
-two_pool_all_fix<-function(data, FACT){
-  
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-  
-  
-  #if factor needs to be included, its levels are defined
-  if(FACT==1){
-    
-    dat<-data
-    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
-    
-  }else{
-    
-    if(FACT==2){
-      
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-      
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-      
-    }else{
-      
-      if(FACT==3){
-        
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-        
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-        
-        
-      }else{
-        
-        dat<-data
-        dat$id<-c(1)
-        
-      }
-    }
-  }
-  
-  #deb model
-  deriv<-function(time, state, pars){
-    
-    with(as.list(c(state, pars)),{
-      
-      #Carbon uptake
-      Cu=Vmax*C*S/(Km+C)
-      #Respiration
-      r=m0*S+f*R*(1-Yu)
-      
-      Protc=fpr*R+fps*S
-      DNAc=fds*S
-      
-      dR<-Cu-f*R
-      dS<-f*R*Yu - m0*S - k*S
-      dC<--Cu + k*S
-      
-      
-      return(list(c(dR, dS, dC), r=r, Protc=Protc, DNAc=DNAc))
-      
-    })
-  }
-  #define names of parameters
-  parnames<-c("Vmax", "Km", "m0", "f", "k", "Yu", "fpr", "fps", "fds")
-  
-  #parameters estimation function
-  estim<-function(odeset){
-    
-    #defining cost function
-    cost<-function(x){
-      
-      par<-x[1:length(parnames)]
-      
-      names(par)<-parnames
-      
-      Sinit=2.01*0.51/12.01/4/par[["fds"]]
-      Rinit=0.06185237*Sinit
-      
-      #first, pars dependent output from ode is matched with measured values
[…]

### 2020-03-30 Rscripts/two_pool_all_fixVmaxYuk.R

@@ -1,11 +1,13 @@
-monodplus_prot_fix<-function(data, FACT){
+two_pool_all_fixVmaxYuk<-function(data, FACT){
   
   #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
   
+  
+  #if factor needs to be included, its levels are defined
   if(FACT==1){
     
     dat<-data
-    dat$id<-ifelse(d$Substrate=="Glucose", 1, 2)
+    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
     
   }else{
     
@@ -38,31 +40,30 @@ monodplus_prot_fix<-function(data, FACT){
     }
   }
   
-  #mend model
+  #deb model
   deriv<-function(time, state, pars){
     
     with(as.list(c(state, pars)),{
       
-      #equations
-      #C uptake
-      Cu=Vmax*Cmic*C/(Km+C)
-      #respiration
-      r=mr*Cmic+(1-CUE)*Cu
+      #Carbon uptake
+      Cu=0.2468173*C*S/(Km+C)
+      #Respiration
+      r=f*R*(1-0.3731948)
       
-      Protc=Cmic*fp
-      #DNAc=Cmic*fd
+      Protc=fpr*R+fps*S
+      DNAc=fds*S
       
-      #states
-      dCmic<-CUE*Cu-mr*Cmic-k*Cmic
-      dC<--Cu + k*Cmic
+      dR<-Cu-f*R
+      dS<-f*R*0.3731948 - 0.007046304*S
+      dC<--Cu + 0.007046304*S
       
-      return(list(c(dCmic, dC), r=r, Protc=Protc))
       
+      return(list(c(dR, dS, dC), r=r, Protc=Protc, DNAc=DNAc))
       
     })
   }
   #define names of parameters
-  parnames<-c("Vmax", "Km", "CUE", "mr", "k","fp")
+  parnames<-c("Km", "f", "fpr", "fps", "fds")
   
   #parameters estimation function
   estim<-function(odeset){
@@ -74,55 +75,62 @@ monodplus_prot_fix<-function(data, FACT){
       
       names(par)<-parnames
       
-      #first, pars dependent output from ode is matched with measured values
-      yhat_all<-as.data.frame(ode(y=c(Cmic=0.09125118, C=25), parms=par, deriv, times=sort(odeset$Time)))
+      Sinit=2.01*0.51/12.01/4/par[["fds"]]
+      Rinit=0.04012019*Sinit
       
[…]

### 2020-03-30 Rscripts/two_pool_all_fixYuk.R

@@ -1,252 +1,0 @@
-two_pool_all_fixYuk<-function(data, FACT){
-  
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-  
-  
-  #if factor needs to be included, its levels are defined
-  if(FACT==1){
-    
-    dat<-data
-    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
-    
-  }else{
-    
-    if(FACT==2){
-      
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-      
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-      
-    }else{
-      
-      if(FACT==3){
-        
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-        
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-        
-        
-      }else{
-        
-        dat<-data
-        dat$id<-c(1)
-        
-      }
-    }
-  }
-  
-  #deb model
-  deriv<-function(time, state, pars){
-    
-    with(as.list(c(state, pars)),{
-      
-      #Carbon uptake
-      Cu=Vmax*C*S/(Km+C)
-      #Respiration
-      r=m0*S+f*R*(1-0.3565256)
-      
-      Protc=fpr*R+fps*S
-      DNAc=fds*S
-      
-      dR<-Cu-f*R
-      dS<-f*R*0.3565256 - m0*S - 0.004915818*S
-      dC<--Cu + 0.004915818*S
-      
-      
-      return(list(c(dR, dS, dC), r=r, Protc=Protc, DNAc=DNAc))
-      
-    })
-  }
-  #define names of parameters
-  parnames<-c("Vmax", "Km", "m0", "f", "fpr", "fps", "fds")
-  
-  #parameters estimation function
-  estim<-function(odeset){
-    
-    #defining cost function
-    cost<-function(x){
-      
-      par<-x[1:length(parnames)]
-      
-      names(par)<-parnames
-      
-      Sinit=2.01*0.51/12.01/4/par[["fds"]]
-      Rinit=0.06185237*Sinit
-      
[…]

### 2020-03-30 Rscripts/two_pool_all_fixYukf.R

@@ -1,252 +1,0 @@
-two_pool_all_fixYukf<-function(data, FACT){
-  
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-  
-  
-  #if factor needs to be included, its levels are defined
-  if(FACT==1){
-    
-    dat<-data
-    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
-    
-  }else{
-    
-    if(FACT==2){
-      
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-      
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-      
-    }else{
-      
-      if(FACT==3){
-        
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-        
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-        
-        
-      }else{
-        
-        dat<-data
-        dat$id<-c(1)
-        
-      }
-    }
-  }
-  
-  #deb model
-  deriv<-function(time, state, pars){
-    
-    with(as.list(c(state, pars)),{
-      
-      #Carbon uptake
-      Cu=Vmax*C*S/(Km+C)
-      #Respiration
-      r=m0*S+0.2373511*R*(1-0.3565256)
-      
-      Protc=fpr*R+fps*S
-      DNAc=fds*S
-      
-      dR<-Cu-0.2373511*R
-      dS<-0.2373511*R*0.3565256 - m0*S - 0.004915818*S
-      dC<--Cu + 0.004915818*S
-      
-      
-      return(list(c(dR, dS, dC), r=r, Protc=Protc, DNAc=DNAc))
-      
-    })
-  }
-  #define names of parameters
-  parnames<-c("Vmax", "Km", "m0", "fpr", "fps", "fds")
-  
-  #parameters estimation function
-  estim<-function(odeset){
-    
-    #defining cost function
-    cost<-function(x){
-      
-      par<-x[1:length(parnames)]
-      
-      names(par)<-parnames
-      
-      Sinit=2.01*0.51/12.01/4/par[["fds"]]
-      Rinit=0.06185237*Sinit
-      
-      #first, pars dependent output from ode is matched with measured values
[…]

### 2020-03-30 Rscripts/two_pool_all_fixYukfVmax.R

@@ -1,252 +1,0 @@
-two_pool_all_fixYukfVmax<-function(data, FACT){
-  
-  #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
-  
-  
-  #if factor needs to be included, its levels are defined
-  if(FACT==1){
-    
-    dat<-data
-    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
-    
-  }else{
-    
-    if(FACT==2){
-      
-      dat<-data
-      dat<-data
-      ids<-dat %>% group_by(Structure) %>% summarise(id=n())
-      ids$id<-seq(1:nrow(ids))
-      
-      dat<-merge(dat, ids, by.x=c("Structure"), by.y=c("Structure"))
-      
-    }else{
-      
-      if(FACT==3){
-        
-        dat<-data
-        ids<-dat %>% group_by(Substrate, Structure) %>% summarise(id=n())
-        ids$id<-seq(1:nrow(ids))
-        
-        dat<-merge(dat, ids, by.x=c("Substrate","Structure"), by.y=c("Substrate", "Structure"))
-        
-        
-      }else{
-        
-        dat<-data
-        dat$id<-c(1)
-        
-      }
-    }
-  }
-  
-  #deb model
-  deriv<-function(time, state, pars){
-    
-    with(as.list(c(state, pars)),{
-      
-      #Carbon uptake
-      Cu=0.2761818*C*S/(Km+C)
-      #Respiration
-      r=m0*S+0.2373511*R*(1-0.3565256)
-      
-      Protc=fpr*R+fps*S
-      DNAc=fds*S
-      
-      dR<-Cu-0.2373511*R
-      dS<-0.2373511*R*0.3565256 - m0*S - 0.004915818*S
-      dC<--Cu + 0.004915818*S
-      
-      
-      return(list(c(dR, dS, dC), r=r, Protc=Protc, DNAc=DNAc))
-      
-    })
-  }
-  #define names of parameters
-  parnames<-c("Km", "m0", "fpr", "fps", "fds")
-  
-  #parameters estimation function
-  estim<-function(odeset){
-    
-    #defining cost function
-    cost<-function(x){
-      
-      par<-x[1:length(parnames)]
-      
-      names(par)<-parnames
-      
-      Sinit=2.01*0.51/12.01/4/par[["fds"]]
-      Rinit=0.06185237*Sinit
-      
[…]

### 2020-03-30 Rscripts/two_pool_allx.R

@@ -1,11 +1,13 @@
-mend_test<-function(data, FACT){
+two_pool_allx<-function(data, FACT){
   
   #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
   
+  
+  #if factor needs to be included, its levels are defined
   if(FACT==1){
     
     dat<-data
-    dat$id<-ifelse(d$Substrate=="Glucose", 1, 2)
+    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
     
   }else{
     
@@ -38,26 +40,30 @@ mend_test<-function(data, FACT){
     }
   }
   
-  #mend model
+  #deb model
   deriv<-function(time, state, pars){
     
     with(as.list(c(state, pars)),{
       
-      #equations
-      #C uptake
-      Cu=Vmax*Cmic*C/(Km+C)
-      #respiration
-      r=mr*Cmic+(1-CUE)*Cu
+      #Carbon uptake
+      Cu=Vmax*C*S/(Km+C)
+      #Respiration
+      r=f*R*(1-Yu)
       
-      #states
-      dCmic<-CUE*Cu-mr*Cmic-k*Cmic
-      dC<--Cu + k*Cmic
+      Protc=fpr*R+fps*S
+      DNAc=fds*S
       
-      return(list(c(dCmic, dC), r=r))
+      dR<-Cu-f*R
+      dS<-f*R*Yu - k*S
+      dC<--Cu + k*S
+      
+      
+      return(list(c(dR, dS, dC), r=r, Protc=Protc, DNAc=DNAc))
+      
     })
   }
   #define names of parameters
-  parnames<-c("Vmax", "Km", "CUE", "mr", "k")
+  parnames<-c("Vmax", "Km", "f", "k", "Yu", "fpr", "fps", "fds", "RS_0")
   
   #parameters estimation function
   estim<-function(odeset){
@@ -69,76 +75,62 @@ mend_test<-function(data, FACT){
       
       names(par)<-parnames
       
-      Cmic0<-numeric()
-      if(odeset$Structure[1]=="Broth"){
-        Cmic0<-0.1287723+2.08193
-      }else{
-        if(odeset$Structure[1]=="Mixed glass"){
-          Cmic0<-0.2931153+4.73895
-        }else{
-          Cmic0<-0.221984+3.588933
-        }
-      }
+      Sinit=2.01*0.51/12.01/4/par[["fds"]]
+      Rinit=par[["RS_0"]]*Sinit
       
[…]

### 2020-03-30 Rscripts/two_pool_allx_fix.R

@@ -1,11 +1,13 @@
-monodplus_prot<-function(data, FACT){
+two_pool_allx_fix<-function(data, FACT){
   
   #FACT 1 = Substrate, 2=Structure, 3=Both, 4=No
   
+  
+  #if factor needs to be included, its levels are defined
   if(FACT==1){
     
     dat<-data
-    dat$id<-ifelse(d$Substrate=="Glucose", 1, 2)
+    dat$id<-ifelse(data$Substrate=="Glucose", 1, 2)
     
   }else{
     
@@ -38,31 +40,30 @@ monodplus_prot<-function(data, FACT){
     }
   }
   
-  #mend model
+  #deb model
   deriv<-function(time, state, pars){
     
     with(as.list(c(state, pars)),{
       
-      #equations
-      #C uptake
-      Cu=Vmax*Cmic*C/(Km+C)
-      #respiration
-      r=mr*Cmic+(1-CUE)*Cu
+      #Carbon uptake
+      Cu=Vmax*C*S/(Km+C)
+      #Respiration
+      r=f*R*(1-Yu)
       
-      Protc=Cmic*fp
-      #DNAc=Cmic*fd
+      Protc=fpr*R+fps*S
+      DNAc=fds*S
       
-      #states
-      dCmic<-CUE*Cu-mr*Cmic-k*Cmic
-      dC<--Cu + k*Cmic
+      dR<-Cu-f*R
+      dS<-f*R*Yu - k*S
+      dC<--Cu + k*S
       
-      return(list(c(dCmic, dC), r=r, Protc=Protc))
       
+      return(list(c(dR, dS, dC), r=r, Protc=Protc, DNAc=DNAc))
       
     })
   }
   #define names of parameters
-  parnames<-c("Vmax", "Km", "CUE", "mr", "k","fp", "Cmic_0")
+  parnames<-c("Vmax", "Km", "f", "k", "Yu", "fpr", "fps", "fds")
   
   #parameters estimation function
   estim<-function(odeset){
@@ -74,55 +75,62 @@ monodplus_prot<-function(data, FACT){
       
       names(par)<-parnames
       
-      #first, pars dependent output from ode is matched with measured values
-      yhat_all<-as.data.frame(ode(y=c(Cmic=par[["Cmic_0"]], C=25), parms=par, deriv, times=sort(odeset$Time)))
+      Sinit=2.01*0.51/12.01/4/par[["fds"]]
+      Rinit=0.04012019*Sinit
       
[…]

### 2020-03-30 Rscripts/two_pool_r.R

@@ -48,10 +48,10 @@ two_pool_r<-function(data, FACT){
       #Carbon uptake
       Cu=Vmax*C*S/(Km+C)
       #Respiration
-      r=m0*S + f*R*(1-Yu) 
+      r=f*R*(1-Yu) 
       
       dR<-Cu-f*R
-      dS<-f*R*Yu - k*S - m0*S
+      dS<-f*R*Yu - k*S
       dC<--Cu + k*S
       
       return(list(c(dR, dS, dC), r=r))
@@ -59,7 +59,7 @@ two_pool_r<-function(data, FACT){
     })
   }
   #define names of parameters
-  parnames<-c("Vmax", "Km", "m0", "f", "k", "Yu", "R_0", "S_0")
+  parnames<-c("Vmax", "Km", "f", "k", "Yu", "R_0", "S_0")
   
   #parameters estimation function
   estim<-function(odeset){
@@ -120,9 +120,9 @@ two_pool_r<-function(data, FACT){
     }
     
     #approximate parameter estimation is done by MCMC method
-    par_mcmc<-modMCMC(f=cost, p=c(Vmax=0.1, Km=3, m0=0.01, f=4, k=1e-2, Yu=0.8,R_0=0.03483481, S_0=0.07661063),
-                      lower=c(Vmax=1e-5, Km=1e-2, m0=1e-5, f=1e-3, k=1e-6, Yu=0.01, R_0=1e-4, S_0=1e-4),
-                      upper=c(Vmax=1, Km=20, m0=1, f=100, k=1, Yu=0.8, R_0=2, S_0=2), niter=10000)
+    par_mcmc<-modMCMC(f=cost, p=c(Vmax=0.1, Km=3, f=4, k=1e-2, Yu=0.8,R_0=0.03483481, S_0=0.07661063),
+                      lower=c(Vmax=1e-5, Km=1e-2, f=1e-3, k=1e-6, Yu=0.01, R_0=1e-4, S_0=1e-4),
+                      upper=c(Vmax=1, Km=20, f=100, k=1, Yu=0.8, R_0=2, S_0=2), niter=10000)
     
     #lower and upper limits for parameters are extracted
     pl<-summary(par_mcmc)["min",]

### 2020-03-30 Rscripts/two_pool_r_fix.R

@@ -48,10 +48,10 @@ two_pool_r_fix<-function(data, FACT){
       #Carbon uptake
       Cu=Vmax*C*S/(Km+C)
       #Respiration
-      r=m0*S + f*R*(1-Yu) 
+      r=f*R*(1-Yu) 
       
       dR<-Cu-f*R
-      dS<-f*R*Yu - k*S - m0*S
+      dS<-f*R*Yu - k*S
       dC<--Cu + k*S
       
       return(list(c(dR, dS, dC), r=r))
@@ -59,7 +59,7 @@ two_pool_r_fix<-function(data, FACT){
     })
   }
   #define names of parameters
-  parnames<-c("Vmax", "Km", "m0", "f", "k", "Yu")
+  parnames<-c("Vmax", "Km", "f", "k", "Yu")
   
   #parameters estimation function
   estim<-function(odeset){
@@ -72,7 +72,7 @@ two_pool_r_fix<-function(data, FACT){
       names(par)<-parnames
       
       #first, pars dependent output from ode is matched with measured values
-      yhat_all<-as.data.frame(ode(y=c(R=0.01908309, S=0.2687341, C=25), parms=par, deriv, times=sort(odeset$Time)))
+      yhat_all<-as.data.frame(ode(y=c(R=0.08814891, S=0.4807859, C=25), parms=par, deriv, times=sort(odeset$Time)))
       
       #select time and the measured variables
       yhat<-select(yhat_all, c("time", "r"))
@@ -95,7 +95,7 @@ two_pool_r_fix<-function(data, FACT){
       names(par)<-parnames
       
       #first, pars dependent output from ode is matched with measured values
-      yhat_all<-as.data.frame(ode(y=c(R=0.01908309, S=0.2687341, C=25), parms=par, deriv, times=sort(odeset$Time)))
+      yhat_all<-as.data.frame(ode(y=c(R=0.08814891, S=0.4807859, C=25), parms=par, deriv, times=sort(odeset$Time)))
       
       #select time and the measured variables
       yhat<-select(yhat_all, c("time", "r"))
@@ -120,9 +120,9 @@ two_pool_r_fix<-function(data, FACT){
     }
     
     #approximate parameter estimation is done by MCMC method
[…]

### 2020-03-31 SI_code_revised.Rmd
Tables numbering corrected

@@
 
-#Bacteria
-bac<-as.data.frame(t(read.csv(file=c("bacteria.csv"), skip = 4)[,-1]))
-colnames(bac)<-t(read.csv(file=c("bacteria.csv"), skip = 4)[,1])
-##remove OTUs with all zeros
-bac2<-bac[, colSums(bac!=0)>0]
-
-bac_env<-as.data.frame(t(read.csv(file=c("bacteria.csv"), nrows = 4, header = F)[,-1]))
-colnames(bac_env)<-t(read.csv(file=c("bacteria.csv"), nrows = 4, header = F)[,1])
-
-bac_env$Structure<-recode(bac_env$Structure, 'Broth'="BROTH", 'Glass wool'="WOOL",
-                          'Mixed glass'="GLASS")
-
-#Fungi
-fungi<-as.data.frame(t(read.csv(file=c("fungi.csv"), skip = 4)[,-1]))
-colnames(fungi)<-t(read.csv(file=c("fungi.csv"), skip = 4)[,1])
-
-##remove OTUs with all zeros
-fungi2<-fungi[, colSums(fungi!=0)>0]
-
-fungi_env<-as.data.frame(t(read.csv(file=c("fungi.csv"), nrows = 4, header = F)[,-1]))
-colnames(fungi_env)<-t(read.csv(file=c("fungi.csv"), nrows = 4, header = F)[,1])
-
-fungi_env$Structure<-recode(fungi_env$Structure, 'Broth'="BROTH", 'Glass wool'="WOOL",
-                          'Mixed glass'="GLASS")
-
 ```
@@
 ```
-##Statistical analysis
+
+## Statistical analysis
+
 Linear Mixed-Effect models are used to evaluate the effect of substrate and the type of physical barrier on measured respiration rate, DNA and protein concentration.
@@
 $~$
-##Microbial-explicit models analysis
-###Both models are first calibrated against respiration rate
+
+## Microbial-explicit models analysis
+
+### Both models are first calibrated against respiration rate
+
 Calibration of both models is done in four steps.
@@
 
-####Monod model
+#### Monod model
+
 ```{r monod respiration calibration, eval=F, echo=T}
@@
 
-kable(t.s1.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('v', 'k', '$\\Y_{CUE}$', 'm'), row.names=T, caption = "_**Table S3:** Monod model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s1.1, digits = c(2, 2, 2, 3), align = 'c', "html", booktabs=T, col.names = c('v', 'k', '$\\Y_{CUE}$', 'm'), row.names=T, caption = "_**Table S2:** Monod model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(monod_r_results_good[c(3:6)], digits = c(2, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S4:** Goodness of fit of monod model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(monod_r_results_good[c(3:6)], digits = c(2, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S3:** Goodness of fit of monod model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-####Sub-microbial model
+#### Sub-microbial model
+
 ```{r Sub-microbial model respiration calibration, eval=F, echo=T}
@@
 #                               two_pool_r4$goodness)
-# write.csv(two_pool_r_results_par, file = c("../two_pool_r_results_par.csv"), row.names = F)
-# write.csv(two_pool_r_results_good, file = c("../two_pool_r_results_good.csv"), row.names = F)
+# write.csv(two_pool_r_results_par, file = c("./two_pool_r_results_par.csv"), row.names = F)
+# write.csv(two_pool_r_results_good, file = c("./two_pool_r_results_good.csv"), row.names = F)
 
@@
 
-kable(t.s3.1, digits = c(2, 2, 2, 3, 2), align = 'c', "html", booktabs=T, col.names = c('v', 'k', 'f', 'm', 'Y'), row.names=T, caption = "_**Table S7:** Sub-microbial model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s3.1, digits = c(2, 2, 2, 3, 2), align = 'c', "html", booktabs=T, col.names = c('v', 'k', 'f', 'm', 'Y'), row.names=T, caption = "_**Table S4:** Sub-microbial model parameters calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(two_pool_r_results_good[, c(3:6)], digits = c(2, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S8:** Goodness of fit of Sub-microbial model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(two_pool_r_results_good[, c(3:6)], digits = c(2, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c('Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S5:** Goodness of fit of Sub-microbial model calibrated against respiration rate across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-###Both models are calibrated against respiration rate, proteins and DNA concentration in the cells.
+### Both models are calibrated against respiration rate, proteins and DNA concentration in the cells.
+
 Both models are calibrated against all data we measured.
@@
 
-####Monod model
+#### Monod model
+
 ```{r Monod model - biomass components, eval=F, echo=T}
@@
 
-kable(t.s4.1, digits = c(0, 2, 2, 2, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'm', '$\\f_{Proteins}$', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S9:** Monod model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s4.1, digits = c(0, 2, 2, 2, 3, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'm', '$\\f_{Proteins}$', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S6:** Monod model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(monod_all_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S10:** Goodness of fit of Monod model calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(monod_all_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S7:** Goodness of fit of Monod model calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(t.s4.2, digits = c(0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'm', '$\\f_{Proteins}$'), row.names=F, caption = "_**Table S11:** Monod model parameters calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s4.2, digits = c(0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'm', '$\\f_{Proteins}$'), row.names=F, caption = "_**Table S8:** Monod model parameters calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(monod_prot_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S12:** Goodness of fit of Monod model calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(monod_prot_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S9:** Goodness of fit of Monod model calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(t.s4.3, digits = c(0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'm', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S13:** Monod model parameters calibrated against respiration rate and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s4.3, digits = c(0, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ", 'v', 'k', '$\\Y_{CUE}$', 'm', '$\\f_{DNA}$'), row.names=F, caption = "_**Table S10:** Monod model parameters calibrated against respiration rate and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(monod_DNA_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S14:** Goodness of fit of Monod model calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(monod_DNA_results_good[,c(1, 4:7)], digits = c(0, 2, 2, 0, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S11:** Goodness of fit of Monod model calibrated against respiration rate and protein concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-####Sub-microbial model
+#### Sub-microbial model
+
 Not all model parameters vary significantly across treatments. Parameters that doesn't show any variability across treatments are fixed.
+
 ```{r Sub-microbial model calibration, eval=F, echo=T}
@@
 # 
-# write.csv(two_pool_all_results_par, file = c("../two_pool_all_results_par.csv"), row.names = F)
-# write.csv(two_pool_all_results_good, file = c("../two_pool_all_results_good.csv"), row.names = F)
+# write.csv(two_pool_all_results_par, file = c("./two_pool_all_results_par.csv"), row.names = F)
+# write.csv(two_pool_all_results_good, file = c("./two_pool_all_results_good.csv"), row.names = F)
 
@@
 
-kable(t.s6.1, digits = c(0, 2, 2, 2, 3, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ","v", 'k', 'f', 'm', 'Y', '$\\B_{R-Proteins}$', '$\\B_{S-Proteins}$', '$\\B_{S-DNA}$', '$\\RS_{0}$'), row.names=F, caption = "_**Table S21:** Sub-microbial model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(t.s6.1, digits = c(0, 2, 2, 2, 3, 2, 2, 2, 3, 3), align = 'c', "html", booktabs=T, col.names = c(" ","v", 'k', 'f', 'm', 'Y', '$\\B_{R-Proteins}$', '$\\B_{S-Proteins}$', '$\\B_{S-DNA}$', '$\\RS_{0}$'), row.names=F, caption = "_**Table S12:** Sub-microbial model parameters calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%
@@
 
-kable(two_pool_all_results_good[, c(1, 4:7)], digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S21:** Goodness of fit of Sub-microbial model calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
+kable(two_pool_all_results_good[, c(1, 4:7)], digits = c(0, 2, 2, 2, 2), align = 'c', "html", booktabs=T, col.names = c(" " , 'Log Likelihood', '$\\R^{2}$', 'Number of parameters', 'AIC'), row.names=F, caption = "_**Table S13:** Goodness of fit of Sub-microbial model calibrated against respiration rate, protein and DNA concentration across all treatments (All), for different levels of incubation system complexity separately (Structure), for Glucose and Celluloze treatments separately (Substrate), and for all experimental treatments separatelly (Each).
 See Materials and Methods for symbols explanation._") %>%

### 2020-04-01 SI_code_revised.Rmd
Updated figures

@@
 ```
+
+```{r}
+
+fpars<-t.s6.1[c(2:4), ]
+fpars$SA<-c(0, 332, 400)
+
+fpars$Km.sd<-c(summary(two_pool_all_fixVmaxYuk_out2[[1]]$par_prof)[2,1],
+               summary(two_pool_all_fixVmaxYuk_out2[[2]]$par_prof)[2,1],
+               summary(two_pool_all_fixVmaxYuk_out2[[3]]$par_prof)[2,1])
+fpars$fr.sd<-c(summary(two_pool_all_fixVmaxYuk_out2[[1]]$par_prof)[2,4]+
+                 summary(two_pool_all_fixVmaxYuk_out2[[1]]$par_prof)[2,5],
+               summary(two_pool_all_fixVmaxYuk_out2[[2]]$par_prof)[2,4]+
+                 summary(two_pool_all_fixVmaxYuk_out2[[2]]$par_prof)[2,5],
+               summary(two_pool_all_fixVmaxYuk_out2[[3]]$par_prof)[2,4]+
+                 summary(two_pool_all_fixVmaxYuk_out2[[3]]$par_prof)[2,5])
+
+grid.arrange(ggplot(fpars, aes(SA, Km)) + 
+  geom_point(cex=6, pch=21, aes(fill = Name)) + theme_min + stat_smooth(method=lm, se = F, colour="grey30") +
+  geom_errorbar(aes(ymin=Km-Km.sd, ymax=Km+Km.sd)) +
+  scale_fill_manual(values=c("black", "grey", "white")) +
+  xlab(expression(paste("Total surface area (", cm^{2}, ")"))) +
+  ylab(expression(paste("Affinity to substrate (", mu, "mol ", L^{-1}, ")"))) +
+  theme(legend.title = element_blank(),
+        legend.position = c(0.75, 0.75))+
+  ggtitle("A)"), 
+  ggplot(fpars, aes(SA, fpr+fps)) + 
+  geom_point(cex=6, pch=21, aes(fill = Name), show.legend = F) + theme_min + stat_smooth(method=lm, se = F, colour="grey30") +
+  geom_errorbar(aes(ymin=fpr+fps-fr.sd, ymax=fpr+fps+fr.sd)) +
+  scale_fill_manual(values=c("black", "grey", "white")) +
+  xlab(expression(paste("Total surface area (", cm^{2}, ")"))) +
+  ylab("Abundance of proteins in cells") + ggtitle("B)"),
+  nrow=1)
+
+```
+
+```{r}
+source("./Rscripts/two_pool_plot.R")
+
+sim_out<-two_pool_plot(fpars)
+
+bplot<-m0 %>% filter(Substrate=="Glucose" | Substrate=="Cellobiose") %>% 
+  group_by(Structure, Substrate, Day) %>%
+  summarise(P = mean(Prot.in*0.46/12.01/4, na.rm=T),
+            P.sd = sd(Prot.in*0.46/12.01/4, na.rm=T),
+            D = mean(DNA*0.51/12.01/4, na.rm=T),
+            D.sd = sd(DNA*0.51/12.01/4, na.rm=T)) 
+bplot$Struc<-NA
+
+for(i in 1:nrow(bplot)){
+  if(bplot$Structure[i]=="Broth"){
+    bplot$Struc[i]<-"BROTH"
+  }else{
+    if(bplot$Structure[i]=="Glass wool"){
+      bplot$Struc[i]<-"WOOL"
+    }else{
+      bplot$Struc[i]<-"GLASS"
+    }
+  }
+}
+
+bplot$Struc<-factor(bplot$Struc, levels=c("BROTH", "GLASS", "WOOL"))
+
+ggplot(bplot, aes(x=Day*24))+
+    geom_point(cex=6, aes(y=P, colour="Cellular Proteins", shape=Substrate)) +
+  geom_errorbar(aes(ymin=P-P.sd, ymax=P+P.sd, color="Cellular Proteins"), width=0.0001, lwd=1)+
+  geom_line(data = sim_out, aes(x=time, y=Protc, colour="Cellular Proteins"), lwd=1.5)+
+    geom_point(cex=6, aes(y=D*28, colour="DNA", shape=Substrate))+
+  geom_errorbar(aes(ymin=(D-D.sd)*28, ymax=(D+D.sd)*28, color="DNA"), width=0.0001, lwd=1)+
+  geom_line(data = sim_out, aes(x=time, y=DNAc*28, colour="DNA"), lwd=1.5)+
+    scale_y_continuous(sec.axis = sec_axis(~./28, 
+                       name = expression(paste("DNA (", mu ,"mol(",C[DNA],")" ,ml^{-1}, ")"))), limits = c(0, 3))+
+    scale_colour_manual(values = c( "black", "grey60"))+
+    labs(y = expression(paste("Cellular Proteins (", mu ,"mol(",C[Proteins],")" ,ml^{-1}, ")")),
+         x = "Time (h)")+
+    theme_min + facet_wrap(~Struc, ncol=1) + expand_limits(x = 0, y = 0) +
+  theme(legend.title = element_blank(),
+        legend.direction = "horizontal",
+        legend.spacing.y = unit(2.6, "in"),
+        legend.spacing.x = unit(0.05, "in"),
+        legend.position = c(0.65, 0.45),
+        legend.background = element_blank(),
+        )
+
+source("./Rscripts/monod_plot.R")
+sim_monod_out<-monod_plot(t.s4.1)
+
+bplot$Struc<-factor(bplot$Struc, levels=c("BROTH", "GLASS", "WOOL"))
+
+ggplot(bplot, aes(x=Day*24))+
+    geom_point(cex=6, aes(y=P, colour="Cellular Proteins", shape=Substrate)) +
+  geom_errorbar(aes(ymin=P-P.sd, ymax=P+P.sd, color="Cellular Proteins"), width=0.0001, lwd=1)+
+  geom_line(data = sim_out, aes(x=time, y=Protc, colour="Cellular Proteins"), lwd=1.5)+
+  geom_line(data = sim_monod_out, aes(x=time, y=Protc, colour="Cellular Proteins"), lwd=1, linetype="dashed")+
+    geom_point(cex=6, aes(y=D*28, colour="DNA", shape=Substrate))+
+  geom_errorbar(aes(ymin=(D-D.sd)*28, ymax=(D+D.sd)*28, color="DNA"), width=0.0001, lwd=1)+
+  geom_line(data = sim_out, aes(x=time, y=DNAc*28, colour="DNA"), lwd=1.5)+
+  geom_line(data = sim_monod_out, aes(x=time, y=DNAc*28, colour="DNA"), lwd=1, linetype="dashed")+
+    scale_y_continuous(sec.axis = sec_axis(~./28, 
+                       name = expression(paste("DNA (", mu ,"mol(",C[DNA],")" ,ml^{-1}, ")"))), limits = c(0, 3))+
+    scale_colour_manual(values = c( "black", "grey60"))+
+    labs(y = expression(paste("Cellular Proteins (", mu ,"mol(",C[Proteins],")" ,ml^{-1}, ")")),
+         x = "Time (h)")+
+    theme_min + facet_wrap(~Struc, ncol=1) + expand_limits(x = 0, y = 0) +
+  theme(legend.title = element_blank(),
+        legend.direction = "horizontal",
+        legend.spacing.y = unit(2.6, "in"),
+        legend.spacing.x = unit(0.05, "in"),
+        legend.position = c(0.65, 0.45),
+        legend.background = element_blank(),
+        )
+
+ggplot(sim_out, aes(x=time))+
+    geom_line(lwd=1.5, aes(y=Protc/(R+S)*100, colour="Cellular Proteins")) +
+    geom_line(lwd=1.5, aes(y=DNAc/(R+S)*15*100, colour="DNA"))+
+    scale_y_continuous(sec.axis = sec_axis(~./15, 
+                       name = expression(paste(frac(C[DNA], Biomass)%*%100))))+
+    scale_colour_manual(values = c( "black", "grey60"))+
+    labs(y = expression(paste(frac(C[Proteins], Biomass)%*%100)),
+         x = "Time (h)")+
+    theme_min + facet_wrap(~Struc, ncol=1) + expand_limits(x = 0, y = 0) +
+  theme(legend.title = element_blank(),
+        legend.direction = "horizontal",
+        legend.spacing.y = unit(2.6, "in"),
+        legend.spacing.x = unit(0.05, "in"),
+        legend.position = c(0.55, 0.45),
+        legend.background = element_blank(),
+        )
+
+```

### 2020-04-01 Rscripts/monod_plot.R

@@ -1,0 +1,40 @@
+monod_plot<-function(ep){
+  #model
+  deriv<-function(time, state, pars){
+    
+    with(as.list(c(state, pars)),{
+      
+      Protc=fp*Cmic
+      DNAc=fd*Cmic
+      
+      dCmic<--k*Cmic+CUE*Vmax*Cmic*C/(Km+C)
+      dC<-k*Cmic-Vmax*Cmic*C/(Km+C)
+      
+      return(list(c(dCmic, dC), r=(1-CUE)*Vmax*Cmic*C/(Km+C), Protc=Protc, DNAc=DNAc))
+      
+    })
+  }
+  #define names of parameters
+  parnames<-c("Vmax", "Km", "CUE", "k", "fp", "fd")
+  
+  #Broth
+  par<-as.numeric(ep[ep$Name=="Broth", c(2:7)])
+  names(par)<-parnames
+  yhatB<-as.data.frame(ode(y=c(Cmic=2.01*0.51/12.01/4/par[["fd"]], C=25), 
+                           parms=par, deriv, times=seq(0, 120)))
+  yhatB$Struc<-"BROTH"
+  #Glass
+  par<-as.numeric(ep[ep$Name=="Glass", c(2:7)])
+  names(par)<-parnames
+  yhatG<-as.data.frame(ode(y=c(Cmic=2.01*0.51/12.01/4/par[["fd"]], C=25), 
+                             parms=par, deriv, times=seq(0, 120)))
+  yhatG$Struc<-"GLASS"
+  #Wool
+  par<-as.numeric(ep[ep$Name=="Wool", c(2:7)])
+  names(par)<-parnames
+  yhatW<-as.data.frame(ode(y=c(Cmic=2.01*0.51/12.01/4/par[["fd"]], C=25), 
+                           parms=par, deriv, times=seq(0, 120)))
+  yhatW$Struc<-"WOOL"
+  
+  return(rbind(yhatB, yhatG, yhatW))
+}

### 2021-03-19 SI_code_2019.Rmd
reorg files for 2021 reanalysis

--- 2020-04-01 SI_code_revised.Rmd
@@
 ---
-title: Towards more realistic model representation of microbial population dynamics in complex environment
+title: Physical structure of an environment and microbial biomass composition variability affect the performance of microbially explicit model
 author: "Petr Capek, Katherine Todd-Brown, Natalie Sadler, Michal Choma, Jianqiu Zheng, Nancy Hess, Kirsten Hofmockel"
